### 2010-12-13 faces_system.tex
Initial import

%% bare_jrnl.tex
%% V1.1
%% 2002/08/13
%% by Michael Shell
%% mshell@ece.gatech.edu
%%
%% NOTE: This text file uses MS Windows line feed conventions. When (human)
%% reading this file on other platforms, you may have to use a text
%% editor that can handle lines terminated by the MS Windows line feed
%% characters (0x0D 0x0A).
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.6 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.ieee.org
%% and/or
%% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
%%
%% This code is offered as-is - no warranty - user assumes all risk.
%% Free to use, distribute and modify.

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% Testflow can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/testflow

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print. Authors are encouraged to use U.S. letter paper when
% submitting to IEEE. Use the testflow package mentioned above to verify
% correct handling of both paper sizes by the author's LaTeX system.
%
% Also note that the "draftcls", not "draft", option should be used if
% it is desired that the figures are to be displayed in draft mode.

% *************************************************************************
% You can now use this template both for submitting to peer review and,
% if your paper is accepted, for sending in final publication materials.
% To change from peer review format (single column, double-spaced) to
% final publication format (double column, single-spaced), just move the
% comment-line sign (%) from one \documentclass line to the other.
% The first is for peer review format(single column), the second is for final publication(double column).

\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
%\documentclass[9.5pt,journal,final,finalsubmission,twocolumn]{IEEEtran}
% *************************************************************************

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{pdfsync}
\usepackage{helvet}
\usepackage{url}
\usepackage{color}

% Include other packages here, before hyperref.
\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\subj}{\textup{subj to}}

\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}

\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}

\begin{document}
%
% paper title
\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
Illumination by Sparse Representation}
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
        John~Wright,~\IEEEmembership{Member,~IEEE,}
        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
        Hossein~Mobahi,
        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{A. Wagner, A. Ganesh, Z. Zhou, and Y. Ma are with the Dept. of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign. H. Mobahi is with the Computer Science Dept. at the University of Illinois at Urbana-Champaign.  J. Wright and Y. Ma are with Microsoft Research Asia.  Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W. Main st. Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.

% If you want to put a publisher's ID mark on the page
% (can leave text blank if you just want to see how the
% text height on the first page will be reduced by IEEE)
%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}

% use only for invited papers
%\specialpapernotice{(Invited Paper)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%% ABSTRACT
\IEEEcompsoctitleabstractindextext{
\vspace{-15mm}
\begin{abstract}
Many classic and contemporary face recognition algorithms work well on public data sets, but degrade sharply when they are  used in a real recognition system. This is mostly due to the difficulty of simultaneously handling variations in illumination, image misalignment, and occlusion in the test image. We consider a scenario where the training images are well controlled, and test images are only loosely controlled. We propose a conceptually simple face recognition system that achieves a high degree of robustness and stability to illumination variation, image misalignment, and partial occlusion. The system uses tools from sparse representation to align a test face image to a set of frontal training images. The region of attraction of our alignment algorithm is computed empirically for public face datasets such as Multi-PIE. We demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under uncontrolled illumination. In order to evaluate how our algorithms work under practical testing conditions, we have implemented a complete face recognition system, including a projector-based training acquisition system. Our system can efficiently and effectively recognize faces under a variety of realistic conditions, using only frontal images under the proposed illuminations as training.\vspace{0mm}
\end{abstract}

% make the title area

\begin{IEEEkeywords}
Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
Sparse Representation,  Error Correction, Validation and Outlier Rejection.
\end{IEEEkeywords}}

\maketitle

\IEEEdisplaynotcompsoctitleabstractindextext

\IEEEpeerreviewmaketitle

%%%%%%%%% BODY TEXT
\section{Introduction}
%Automatic face recognition remains one of the most active areas in computer vision.

\PARstart{T}{here} is a historical tendency for face recognition algorithms to work well under laboratory conditions but degrade when tested in less-controlled environments.  While there have been several high profile failed trials of face recognition technology for extremely difficult mass surveillance / watch-list applications, face recognition has not even seen widespread applications in more tractable settings, such as access control for buildings, computer systems, automobiles, or even automatic teller machines.\footnote{Face recognition scenarios where both gallery and test data are uncontrolled is an active research area as well \cite{LFW}.}  These applications are very interesting due to their potential sociological impact.  They are also qualitatively different from surveillance applications because they demand extremely reliable systems, and because the gallery subjects are allies, rather than opponents, of the recognition system.  This creates the possibility of carefully controlling the acquisition of the training data, even if the testing data may be less controlled for usability reasons.

Classical holistic subspace based methods
\cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for
their speed and simplicity, as well as for their natural
extension to linear illumination models.  However, their
performance has been shown to be extremely brittle not only to
alignment variation, but to occlusions. One promising recent
direction, set forth in \cite{Wright2009-PAMI}, casts the
recognition problem as one of finding a sparse representation
of the test image in terms of the training set as a whole, up
to some sparse error due to occlusion. A \emph{sparse
representation-based classification} (SRC) method is then proposed for
recognition. The main idea is that the sparse nonzero
coefficients should concentrate on the training samples with
the same class label as the test sample. SRC has demonstrated
striking recognition performance despite severe occlusion or
corruption by solving a simple convex program.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}& \hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png} \\
\includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}& \hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png} \\
\includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} & \hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
\end{tabular}
\caption{{\bf Compound effect of registration and illumination}. The task is to identify the girl among 20 subjects, by computing the sparse representation of her input face with respect to the entire training set. The absolute sum of the coefficients associated with each subject is plotted on the right. We also show the faces reconstructed with each subject's training images weighted by the associated sparse coefficients. The red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The input face is from Viola and Jones' face detector (the black box) and all 38 illuminations specified in Section \ref{sec:illumination} are used in the training.  {\bf Middle:} The input face is well-aligned (the white box) with the training by our algorithm specified in Section \ref{sec:registration} but only 24 frontal illuminations are used in the training for recognition (see Section \ref{sec:illumination}). {\bf Bottom:} Informative representation obtained by using both well-aligned input face and sufficient (all 38) illuminations in the training. }
\label{fig:promo}
%\vspace{-.5in}
\end{figure}

%\subsection{Compound Effect of Registration and Illumination}
Unfortunately, while SRC achieves impressive results, it does not deal with misalignment
between the test and training images.  Furthermore, the experiments were carried out on public
datasets
taken under controlled laboratory conditions
such as
Extended Yale B \cite{Georghiades2001-PAMI}, which do not contain realistic test images.
%two critical aspects of real world face recognition must be addressed:  significant variations in both the {\em image domain} and in
%the {\em image value}.
We illustrate this with an example in Figure \ref{fig:promo}.
The task is to identify the girl among 20 subjects. If the test
face image, say obtained from an off-the-shelf face detector,
has even a small amount of registration error against the
training images (caused by mild pose, scale, or misalignment),
the representation is no longer informative, even if sufficient
illuminations are present in the training, as shown in Figure
\ref{fig:promo}(top). In addition, in order to linearly span
the illuminations of a typical indoor (or outdoor) environment,
illuminations from behind the subject are also needed in the
training. Otherwise, even for perfectly aligned test images,
the representation will not necessarily be sparse or
informative, as shown by the example in Figure
\ref{fig:promo}(middle). Clearly, both good alignment, as well
as sufficient training images are needed.
%Unfortunately, most public face databases lack images with a
%significant component of rear (more than 90 degrees from
%frontal) illumination, either for training or testing.

\subsection{Related Work}

We briefly review existing techniques for recognition, image registration, and handling of illumination variation.
Our system is based purely on 2D techniques.  This fact immediately distinguishes our technique from systems that either require a 3D data as an input, or attempt to estimate a 3D model from input 2D data \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these techniques have been shown to achieve better robustness to pose variation given a sufficiently accurate 3D model, for access control applications where the user is cooperative in aligning their head to the camera, robustness to moderate misalignment is sufficient.  Furthermore, many 3D techniques rely on multiple views to estimate their model, and may have trouble if all of the training images are captured from one pose.  Note that images of people's faces under varying illumination contain shape-related discriminative information, and this information can be leveraged by 2D algorithms even if shape is not reconstructed explicitly.

% Alignment
In holistic recognition algorithms, correspondence between points in the test image and in the training must be achieved.  A long line of research exists on using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related Active Shape Models \cite{cootes1992active} to register images against a relatively high-dimensional model of plausible face appearances, often leveraging face-specific contours.  While these model-based techniques have advantages in dealing with variations in expression and pose, they may add unnecessary complexity to a system where subjects are willing to present a neutral expression.  We prefer to focus on warpings with far fewer parameters, and to use the training images themselves as the appearance model. Iterative registration in this spirit dates at least back to the Lucas-Kanade algorithm \cite{lucas1981iterative}.

However, whereas much of the early work on image registration is aimed at the problem of registering nearly identical images, say by minimizing a sum of squared distances, here we must confront several physical factors simultaneously: misalignment, illumination variations, and certain amount of occlusion. As we discuss further below, illumination variation can be dealt by expressing the test image as a linear combination of an appropriate set of training images. Similar representations have been exploited in illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}). For robustness to gross errors, the $\ell^1$-norm of the residual is a more appropriate objective function than the classical $\ell^2$-norm. Its use here is loosely motivated by theoretical results due to Candes and Tao  \cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations lead us to pose the registration problem as the search for a set of transformations and illumination coefficients that minimize the $\ell^1$-norm of the representation error. We solve this problem using a generalized Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of these problems can also be solved efficiently using recently developed first-order techniques for $\ell^1$-minimization, which are reviewed in \cite{YangA2010-pp}.

% Illumination
Researchers have tried various techniques to deal with illumination variation.  In almost all recognition algorithms where only a single gallery image is available per individual, illumination effects are regarded as a nuisance that must be removed before the algorithm can continue.  This is typically done by making statistical assumptions about how illumination affects the image, and using those assumptions to extract a new representation that is claimed to be illumination invariant.  Recent examples include \cite{chen2006total} and \cite{zhou2007appearance}.
However, despite these efforts, truly illumination-invariant features are difficult to obtain from a single input image.
So, if one's primary concern in designing the recognition system is achieving a high recognition rate, it makes no sense to limit the gallery to a single image per person.  We therefore take the strategy of sampling many gallery images of each individual under varying illuminations.  These images are used as the basis for either a convex cone model \cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model \cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct projector based light stage.  Most light stages used for face recognition have been constructed for the purpose of creating public data sets to study illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other light stages have been used for computer graphics purposes \cite{debevec2000acquiring, jones2005performance}. The light source can be moved around manually \cite{masselus2002free}, but this may result in poor consistency between users.  Structured light applications use projectors to directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D reconstruction, but this is very disturbing to the user.  To our knowledge, we are the first to use projectors to indirectly illuminate a subject's face for the purpose of face recognition.

\subsection{Contributions} In this paper, we show how the two
{\em strongly coupled} issues of registration and illumination
can be naturally addressed within the sparse representation
framework. We show that face registration, a challenging
nonlinear problem, can be solved by a series of linear programs
that iteratively minimize the sparsity of the registration
error. This leads to an efficient and effective alignment
algorithm for face images that works for a large range of
variation in translation, rotation, and scale, even when
the face is only partially visible due to eyeglasses, closed
eyes and open mouth, sensor saturation, etc.  We also propose a
sufficient set of training illuminations
that is capable of interpolating typical indoor and outdoor
lighting, along with a practical hardware system for capturing
them.

% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.  While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.

% We use tools from sparse representation
% to align a test face image with a set of frontal training
% images in the presence of registration error, illumination variation, and partial occlusion. The region
% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions. We show
% that our system can efficiently and effectively recognize faces under
% a variety of realistic conditions, using only
% frontal images under the proposed illuminations as training.

%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.  We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.  Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.  We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.  We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.

Finally, we demonstrate the effectiveness of the proposed new
methods with a complete face recognition system that is {\em
simple, stable, and scalable}. The proposed system performs
robust automatic recognition of subjects from loosely
controlled input images taken both indoors and outdoors, using
labeled frontal views of the subjects' faces under the proposed
illuminations for training and an off-the-shelf face
detector\footnote{We use the OpenCV
implementation of the Viola and Jones' face detector
\cite{Viola2004-IJCV}.} to detect faces in images.

We conduct extensive experiments on the proposed system with
both public databases and a face database that is collected by
our own acquisition system. Our experimental results on
large-scale public face databases show that our algorithm
indeed achieves very good performance on these databases,
exceeding or competing with the state-of-the-art algorithms. 
Additionally, our experimental results on our own database
clearly demonstrate that our system not only works well with
images taken under controlled laboratory conditions, but is
capable of handling practical indoor and outdoor illuminations as well.

\noindent{\bf Organization of this paper.} In Section
\ref{sec:registration}, we derive our robust registration and
recognition algorithm within the sparse representation
framework. We elaborate on the implementation issues of the
algorithm, conduct region of attraction experiments with
respect to both 2D in-plane deformation and 3D pose variation,
and discuss its relationship to existing work in the
literature. Section \ref{sec:illumination} is dedicated to our
training acquisition system. Using this system, we investigate empirically the set of training illuminations required to handle practical illumination variations, and suggest a sufficient set of 38 training illuminations. Extensive experiments on
a large-scale public database and on our own database are conducted
in Section \ref{sec:multipie} and Section \ref{sec:own-data},
respectively, to verify the proposed system. Section
\ref{sec:conclusion} concludes our work with discussion of
promising future directions.

\section{Handling Practical Registration Error}\label{sec:registration}
As demonstrated in Figure \ref{fig:promo}(top), the main
limitation of the {\em sparse representation and
classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
the assumption of pixel-accurate alignment between the test
image and the training set. This leads to brittleness under
pose and misalignment, making it inappropriate for deployment
outside a laboratory setting. In this section, we show how this
weakness can be rectified while still preserving the conceptual
simplicity and good recognition performance of SRC.

SRC assumes access to a database of multiple registered
training images per subject, taken under varying illuminations.
The images of subject $i$, stacked as vectors, form a matrix
$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
well-aligned test image $\y_0$ can be represented as a sparse
linear combination $A \x_0$ of all of the images in the
database,\footnote{We assume the illuminations in the training
set are sufficient. We will address how to ensure illumination
sufficiency in the next section.} plus a sparse error $\e_0$
due to occlusion. The sparse representation can be recovered by
minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
$\x$ and $\e$:
\begin{equation}
\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
\label{eqn:robust-l1}
\end{equation}
Now suppose that $\y_0$ is subject to some pose or
misalignment, so that instead of observing $\y_0$, we observe
the warped image $\y = \y_0 \circ \tau^{-1}$, for some
transformation $\tau \in T$ where $T$ is a finite-dimensional
group of transformations acting on the image domain.  The
transformed image $\y$ no longer has a sparse representation of
the form $\y = A \x_0 + \e_0$, and naively applying the
algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
as seen in Figure \ref{fig:promo}(top).

\subsection{Batch and Individual Alignment} Notice that if the
true deformation $\tau^{-1}$ can be found, then we can apply
its inverse $\tau$ to the test image and it again becomes
possible to find a sparse representation of the resulting
image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
  This sparsity
provides a strong cue for finding the correct deformation
$\tau$: conceptually, one would like to seek a transformation
$\tau$ that allows the sparsest representation, by solving
\begin{equation} \label{eqn:L1-L1-conceptual}
\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
\end{equation}
For fixed $\tau$, this problem is jointly convex in $\x$ and
$\e$. However, as a simultaneous optimization over the
coefficients $\x$, error representation $\e$, and
transformation $\tau$, it is a difficult, nonconvex
optimization problem. One source of difficulty is the presence
of multiple faces in the matrix $A$:
\eqref{eqn:L1-L1-conceptual} has many local minima that
correspond to aligning $\y$ to different subjects. In this
sense, the misaligned recognition problem differs from the
well-aligned version studied in \cite{Wright2009-PAMI}. For the
well-aligned case, it is possible to directly solve for a
global representation, with no concern for local minima. With
possible misalignment, it is more appropriate to seek the best
alignment of the test face with each subject $i$:
\begin{equation} \label{eqn:per-subject-L1}
\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
\end{equation}
We no longer penalize $\| \x \|_1$, since $A_i$ consists of
only images of subject $i$ and so $\x$ is no longer expected to
be sparse.

\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
Jacobian of $\y \circ \tau$ with respect to the transformation
parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
above equation is underdetermined if we allow the registration
error $\e$ to be arbitrary. Near the correct alignment we
expect the aligned testing image to differ from $A_i \x$ only
for the minority of the pixels corrupted by occlusions. Thus,
we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
\label{eqn:L1-align}
\end{equation}
This is different from the popular choice that
minimizes the $\ell^2$-norm of the registration error:
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step
$\Delta  \tau$ by solving the least-square problem:
$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
\|_2$. Empirically, we find that if there is only small noise
between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
$\eqref{eqn:L2-align}$ have similar performance.  However, if
there are occlusions in $\y_0$, sequential
$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
better than sequential $\ell^2$-minimization
\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
example.

The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images. Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}. While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm. In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution. We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.

\begin{figure}
\centering
\begin{tabular}{cccc}
\includegraphics[height=1.2in]{figures_cvpr/L1_cropped} &
\includegraphics[height=1.2in]{figures_cvpr/y_warp_L1} &
\includegraphics[height=1.2in]{figures_cvpr/y_hat_L1} &
\includegraphics[height=1.2in]{figures_cvpr/e_L1} \\
\includegraphics[height=1.2in]{figures_cvpr/L2_cropped} &
\includegraphics[height=1.2in]{figures_cvpr/y_warp_L2} &
\includegraphics[height=1.2in]{figures_cvpr/y_hat_L2} &
\includegraphics[height=1.2in]{figures_cvpr/e_L2} \\
(a) & (b) & (c) & (d)
\end{tabular}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by
$\ell^1$ and $\ell^2$ minimization.}
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$;
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
\vspace{-.3in}
\end{figure}

In addition to normalizing the training images (which is done
once), it is important to normalize the warped testing image
$\y \circ \tau$ as the algorithm runs.  Without normalization,
the algorithm may fall into a degenerate global minimum
corresponding to zooming in on a dark region of the test
image.  Normalization is done by replacing the linearization of
$\y \circ \tau$ with a linearization of the normalized version
$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
The proposed alignment algorithm can be easily extended to work
in a {\em multiscale} fashion, with benefits both in
convergence behavior and computational cost.  The alignment
algorithm is simply run to completion on progressively less
downsampled versions of the training and testing images, using
the result of one level to initialize the next.

\subsection{Robust Recognition by Sparse Representation} Once
the best transformation $\tau_i$ has been computed for each
subject $i$, the training sets $A_i$ can be aligned to $\y$,
and a global sparse representation problem of the form
\eqref{eqn:robust-l1} can be solved to obtain a discriminative
representation in terms of the entire training set. Moreover,
the per-subject alignment residuals $\| \e \|_1$ can be used to
prune unpromising candidates from the global optimization,
leaving a much smaller and more efficiently solvable problem.
The complete optimization procedure is summarized as Algorithm
\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
considered together to provide a sparse representation for the
test image. If $S = 1$, the algorithm reduces to classification
by registration error; but considering the test image might be
an invalid subject, we typically choose $S = 10$. Since valid
images have a sparse representation in terms of this larger
set, we can reject invalid test images using the {\em sparsity
concentration index} proposed in \cite{Wright2009-PAMI}.
The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
selects coefficients from the vector $\x$ corresponding to subject $i$.

Another important free parameter in Algorithm
\ref{alg:deformable-src} is the class of deformations $T$. In
our experiments, we typically use 2D similarity
transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{For
readers unfamiliar with this terminology, SE stands for Special
Euclidean, a class of transformations representing image
rotation and translation.  The $\Re_+$ accounts for the scaling
variation that can be represented by a 2D similarity
transformation.}, for removing alignment error incurred by face
detector, or 2D projective transformations, $T =
\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
This class of transformations is able to represent distortion
in a perspective image of a planar object.}$, for handling some
pose variation.

In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
performance of our system, based on the observation that the face detector output may be poorly centered on
the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
instead of aligning the training sets to the
original $\y$ directly obtained from the face detector, we
compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
is believed to be better centered, and update $\y$ according to
$\bar{\tau}$. For the 2D similarity transformations, which are
used in our system when initialized by the face detector, a
transformation $\tau$ can be parameterized as $\tau = (\tau^1,
\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
the translations in $x$- and $y$-axis, $\tau^3$ represents the
rotation angle and $\tau^4$ represents the scale. Then the
average transformation is simply obtained by taking the
component-wise mean:
\begin{displaymath}
\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
\tau_{k_S}^i) / S, i = 1,2,3,4.
\end{displaymath}
Finally, the training sets are aligned to the new $\y$.

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
$\y\in\Re^m$ and a deformation group $T$. 
\STATE
{\bf for} each subject $i$, 
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$. 
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
\STATE \hspace{6mm}
$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$ 
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$; 
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem:
$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$$
\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
\end{algorithmic}
\end{algorithm}

So far, these transformations have been expressed in a very
general form.  The transformation defines a mapping between the
coordinates of pixels in the large original image and a smaller
(un)warped image. The pixels of the small image get stacked into a vector.
For a simple implementation, a rectangular window with
regular sampling can used, but in general, the small image need
not be regularly sampled in pixel coordinates.  For example,
the sample locations could be arbitrarily selected from within a
``face shaped'' area. We will discuss how choosing
different windows can affect the performance of our algorithm
in Section \ref{sec:multipie}. Note that to prevent aliasing
artifacts in the downsampled image, it is necessary to apply a
smoothing filter to the original image.
%The strength of the
%filter depends on how much the warping downsamples the image.
%Although the scale component of the warping may vary with each
%step of the iterative alignment, it is safe in practice to
%choose a fixed conservative smoothing filter (erring slightly
%on the side of more smoothing). Whenever a transformation
%operator is not shown for a vector in our notation (i.e. for
%the columns of $A$ or $J$), it means that a suitable
%transformation has already been applied to a larger image
%(otherwise, the dimensions of the arrays would not agree).

\subsection{System Implementation}
In this section, we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
%The most classical
%solution to these $\ell^1$-minimization problems is the so
%called \emph{primal-dual interior-point} method
%\cite{BoydS2004}. Conceptually, it operates in an iterative
%fashion, and a linear system is solved at each iteration.
%Therefore, its complexity per iteration is at least $O(n^3)$,
%where $n$ is the total number of training images for the face
%recognition problem. Meanwhile, it is also observed in
%\cite{YangA2010-pp} that its computational cost with respect to
%the dimension of each sample, $m$, grows much fast than other
%alternative methods. Hence for real-world applications, as the
%image resolution and number of classes grow, this method will
%soon become too slow to be useful.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
%We refer the interested readers to
%\cite{YangA2010-pp} for a comprehensive survey in the
%literature and a detailed study of the performance of five
%representative approaches on solving the two $\ell^1$-minimization
%problems in Step 6 and 14 of
%Algorithm~\ref{alg:deformable-src}.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.
%method, is always
%among the most accurate and fastest algorithms and achieves
%better performance in both effectiveness and efficiency
%compared to other first-order methods.
%Most importantly, as first-order methods, they require only
%lightweight vector operations and matrix-vector
%multiplications, contrast to many other methods, such as
%primal-dual interior-point method and the homotopy method
%\cite{DonohoD2006-pp}, which all involve expensive operations
%such as solving a linear system. For the rest of this section,
%we briefly describe the ALM method.

We use Step 14 as an example to illustrate the ALM
method, since solving Step 6 is very similar. Recall that in
Step 14 the problem we are interested in is:
\begin{equation}
\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
A \x + \e.
\end{equation}
Its corresponding augmented Lagrangian function is
\begin{equation}
L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
\end{equation}
where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
(\x,\e,\blamda)$ by alternating between optimizing with respect
to the primal variables $\x, \e$ and updating the dual variable
$\blamda$, with the other fixed, as follows:
\begin{equation}
\left \{
\begin{array}{lll}
(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
\end{array}
\right .
\label{eq:alm}
\end{equation}
As one can see, although updating $\blamda$ is trivial,
minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
$\x$ and $\e$ could still be costly. To further reduce the
complexity of the problem, we adopt an approach used in
\cite{YangJ2009-pp}, called \emph{alternating direction
method of multipliers} \cite{Glowinski1975-TR}, which proposes to minimize $L_{\mu} (\x,\e,\blamda_k)$
with respect to $\x$ and $\e$ separately. Then the multiplier
$\blamda$ is updated after just one iteration of minimization
with respect to the two primal variables, giving:
\begin{equation}
\left \{
\begin{array}{lll}
\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
\end{array}
\right .
\label{eq:adm}
\end{equation}
In order to discuss the solution to the above sub-problems, we
need to define the following soft-thresholding operator for a
vector $\x$ and a scalar $\alpha \geq 0$:
\begin{equation}
\textup{shrink}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
\end{equation}
where all the operations are performed component-wise. It is
easy to show that the sub-problem with respect to $\e$ has a
closed-form solution given by the soft-thresholding operator:
\begin{equation}
\e_{k+1} = \textup{shrink}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
\end{equation}
To solve the subproblem associated with $\x$, we
apply a first-order $\ell^1$-minimization method,
called \emph{fast iterative shrinkage-threshold algorithm}
(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
iteratively minimize a quadratic approximation $Q(\x, \z)$ to
$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
carefully chosen in order to achieve a good convergence
rate. We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method for Global
Recognition)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$;
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alm}
\end{algorithm}

%
%Note
%that, however, Algorithm~\ref{alg:alm} differs from
%\cite{YangJ2009-pp} mainly in the way that the sub-problem
%associated with $\x$ is solved. In \cite{YangJ2009-pp}, only an
%approximation to the solution is obtained by just computing one
%iteration of the Conjugate Gradient algorithm. As pointed out
%by \cite{YangA2010-pp}, although this further reduces the
%complexity of the algorithm, it converges very slowly when the
%condition number of the matrix $A^TA$ is large, which is indeed
%the case for the face recognition problem.

We have selected this algorithm because it strikes the best
balance between accuracy and scalability for our problem out of
all that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory, 
running on our database containing images size $80\times 60$ 
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is 
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

\subsection{Simulations and Experiments on Region of
Attraction} We will now present three experimental results
demonstrating the effectiveness of the individual alignment
procedure outlined in the previous section. They show the sufficiency of the
region of attraction, verify effectiveness of the multiscale extension,
and show stability to small pose variations.  We delay large-scale recognition experiments
to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
have discussed the issue of illumination in the next section.
\begin{enumerate}

\item{\em 2D Deformation.}  We first verify the
    effectiveness of our alignment algorithm with images
    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
    We select all the subjects in Session 1, use 7
    illuminations per person from Session 1 for training,
    and test on one new illumination from Session
    2.\footnote{The training are illuminations $\{0, 1, 7,
    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
    testing is the illumination 10. } We manually select
    eye corners in both training and testing as the ground
    truth for registration. We downsample the images to
    $80\times 60$ pixels\footnote{Unless otherwise stated,
    this will be the default resolution at which we prepare
    all our training and testing datasets and run all our
    experiments.} and the distance between the two outer
    eye corners is normalized to be 50 pixels for each
    person. We introduce artificial deformation to the
    testing image with a combination of translation,
    rotation and scaling. We further use the alignment
    error $\|\e\|_1$ as an indicator of success. Let $r_0$
    be the alignment error obtained by aligning a test
    image to the training images without any artificial
    perturbation. When the test image is artificially
    perturbed and aligned, resulting in an alignment error
    $r$, we consider the alignment successful if $|r - r_0 | \leq
    0.01r_0$. Figure \ref{fig:attraction} shows the
    percentage of successful registrations for all subjects
    for each artificial deformation. The results suggest
    that our algorithm works extremely well with
    translation up to 20\% of the eye distance (or 10
    pixels) in all directions and up to $30^\circ$ in-plane
    rotation. We have also tested our alignment algorithm
    with scale variation and it can handle up to 15\%
    change in scale.

We have gathered the statistics of the Viola and Jones'
 face detector on the Multi-PIE dataset. For 4,600 frontal
 images of 230 subjects under 20 different illuminations,
 using manual registration as the ground truth, the average
 misalignment error of the detected faces is about 6 pixels
 and the average variation in scale is 8\%. This falls
 safely inside the region of attraction for our alignment
 algorithm.
\begin{figure}
\centering
\begin{tabular}{ccc}
\includegraphics[height=1.4in]{figures_pami/x_y_roa.png} &
\includegraphics[height=1.4in]{figures_pami/y_theta_roa.png} &
\includegraphics[height=1.4in]{figures_pami/y_s_roa.png}\\
(a)&(b)&(c)
\end{tabular}
\begin{tabular}{cccc}
\includegraphics[height=1.1in]{figures_pami/x_tr.png} &
\includegraphics[height=1.1in]{figures_pami/y_tr.png} &
\includegraphics[height=1.1in]{figures_pami/theta.png} &
\includegraphics[height=1.1in]{figures_pami/scale.png}\\
(d)&(e)&(f)& (g)
\end{tabular}
\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.  The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees. {\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions. (b) Simultaneous translation in $y$ direction and in-plane rotation. (c) Simultaneous translation in $y$ direction and scale variation. {\bf Bottom row:} (d) Translation in $x$ direction only. (e) Translation in $y$ direction only. (f) In-plane rotation only. (g) Scale variation only.}
\label{fig:attraction}
\vspace{-.3in}
\end{figure}
\item{\em Multiscale Implementation.}
        %The
    %gain here is a reduction in the computational cost. We
    %briefly explain how this works here. In our system, a
    %truncated Gaussian kernel is used for blurring. In
    %order to ensure accurate operation, we change the
    %window size of the Gaussian kernel in proportion to its
    %standard deviation. Consequently, achieving a large
    %amount of blur requires a large convolution
    %mask\footnote{It is possible to apply a low-bandwidth
    %Gaussian kernel multiple times to achieve the same
    %effect of a single high-bandwidth kernel. However, the
    %time required will be at least as great as performing
    %the single large blur.} and hence is computationally
    %expensive.
    %To avoid using a large window, we observe that, by
    %down-sampling the image, a smaller blur kernel can
    %achieve similar blurring effect. However, working at a
    %lower resolution degrades the alignment precision.
    %Therefore, we use a multi-scale framework to achieve
    %computational efficiency and high precision at the same
    %time. It starts from the smallest scale with largest
%blur. The strong blur can handle large misalignment, but
%the low resolution results in poor precision. We then use
%the estimated alignment to initialize a similar process
%that works at a higher resolution and with smaller blur.
%This process repeats and thus the solution is computed
%progressively.
Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost. Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
%To see the relationship between number of scales and the
%standard deviation $\sigma$ of the Gaussian kernel applied
%between scales,
We conduct an experiment similar to the previous 2D
deformation experiment. We use the same 7 illuminations in
Session 1 as training, and all 20 illuminations in the same
session as testing. We introduce artificial deformation in
both $x$ and $y$ directions up to 16 pixels in the
$80\times 60$ frame, with a step size of 4 pixels, i.e.,
$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
\{-16,-12,\ldots,12,16\}$. We consider an alignment
successful if the estimated coordinates of the eye-corners
are within 1 pixel from the ground truth in the original
image.  In Figure \ref{fig:multiscale}, we report the
alignment success rate, averaged over the artificially
perturbed initial deformations, as a function of the
standard deviation of the Gaussian kernel $\sigma$, for
three choices of the number of scales. As one can see,
using multiscale indeed improves the performance, and when
3 scales are used, a smaller convolution kernel can achieve
a similar performance compared to a much larger kernel when
only 2 scales are used.
\begin{figure}
\centering
\includegraphics[height=1.8in]{figures_pami/multiscale.png}
\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations. A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
\label{fig:multiscale}
\vspace{-.5in}
\end{figure}

\item{\em 3D Pose Variation.} As densely sampled pose and
    illumination face images are not available in any of
    the public databases, including Multi-PIE, we have
    collected our own dataset using our own system (to be
    introduced in the next section). We use frontal face
    images of a subject under the 38 illuminations proposed
    in the next section as training. For testing, we
    collect images of the subject under a typical indoor
    lighting condition at pose ranging from $-90^\circ$ to
    $+90^\circ$ with step size 5.625$^\circ$, a total of 33
    poses. We use Viola and Jones' face detector to
    initialize our alignment algorithm.
%whenever it succeeds. For extreme poses where the detector fails (usually for poses $> 45^\circ$), we use manual initialization by putting a reasonable square window around the face.
Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
to $\pm 45^\circ$.
Note that this level of out-of-plane
 pose variation is beyond what we intend to handle with our formulation.
%While we don't intend to perform 2D recognition with this much pose,
%the alignment is at least stable.
\begin{figure}
\centering
\begin{tabular}{ccccc}
\includegraphics[height=1in]{figures_cvpr/5} &
\includegraphics[height=1in]{figures_cvpr/7} &
\includegraphics[height=1in]{figures_cvpr/09} &
\includegraphics[height=1in]{figures_cvpr/11} &
\includegraphics[height=1in]{figures_cvpr/13} \\ \vspace{-.1in}
(a) & (b) & (c) & (d) & (e)\vspace{3mm}\\
\includegraphics[height=1in]{figures_cvpr/15} &
\includegraphics[height=1in]{figures_cvpr/17} &
\includegraphics[height=1in]{figures_cvpr/19} &
\includegraphics[height=1in]{figures_cvpr/21} &
\includegraphics[height=1in]{figures_cvpr/3} \\ \vspace{-.1in}
(f) & (g) & (h) & (i) & (j)
\end{tabular}
\caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
}\label{fig:pose-alignment} \vspace{-.3in}
\end{figure}
\end{enumerate}

\subsection{Comparison with Related Work}
Our modification to SRC roots solidly in the tradition of
adding deformation-robustness to face recognition algorithms
\cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}.
However, the only previous work to investigate face alignment
in the context of sparse signal representation and SRC is the
work of \cite{Huang2008-CVPR}. They consider the case where the
training images themselves are misaligned and allow one
deformation per training image. They linearize the training
rather than the test, which is computationally more costly as
it effectively triples the size of the training set. In
addition, as they align the test image to all subjects
simultaneously, it potentially is more prone to local minima
when the number of subjects increases, as we will see in the
following experimental comparisons.
\begin{enumerate}
\item {\em Extended Yale B.} In this experiment, we have
    used the same experimental settings as in
    \cite{Huang2008-CVPR}. 20 subjects are selected and
    each has 32 frontal images (selected at random) as
    training and another 32 for testing. An artificial
    translation of 10 pixels (in both $x$ and $y$
    directions) is introduced to the test image. For our
    algorithm we downsample all the images to $88\times
    80$ for memory reasons, whereas the work of
    \cite{Huang2008-CVPR} uses random projections.
Note that the use of cropped images in this experiment introduces image boundary effects.
    Our
    algorithm achieves the recognition rate 93.7\%,
    compared to 89.1\% recognition rate reported in
    \cite{Huang2008-CVPR}.
\item {\em CMU Multi-PIE.} In this experiment, we choose
    all subjects from the CMU Multi-PIE database, 7
    training images from Session 1 and 1 test image from
    Session 2 per person. The setting is exactly the same
    as the previous experiment on 2D deformation. We again
    work with downsampled images of size $80\times 60$ pixels. An
    artificial translation of 5 pixels (in both $x$ and $y$
    directions) was induced in the test image. The
    algorithm of \cite{Huang2008-CVPR} achieves a
    recognition rate of 67.5\%,\footnote{That algorithm has
    two free parameters - $l$ and $d$, which govern the tradeoff between
	accuracy and run-time. For this experiment
    we chose $l = 1$ and $d = 514$.} while ours achieves 92.2\%.
    \end{enumerate}

\section{Handling Practical Illumination Variation}\label{sec:illumination}
In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.  Under what conditions is this a reasonable assumption to make?  What can we say from first principles about how the training images should be chosen?

\subsection{The Illumination Model}

% % META: Most of this is about understanding the Basi Result.  For the Paper, don't try to derive the result; just state it.
% % Derivation of linearity using strong assumptions
% First, let us assume that
% the illumination if the subject's face is distant.  This model
% will be a good approximation as long as distance to the nearest
% light source is much larger than the person's face.  If we
% further assume that the object is convex so that there is no
% self-shadowing, the illumination incident on a surface patch
% will depend only on the surface normal, and not on the location on the surface.
% Furthermore, if the object is Lambertian, the image intensity will not depend on the angle between patch surface normal and the sensor, or on the depth of the patch.  Since each pixel in the image corresponds to
% some patch on the object, the vector of image intensities is a
% linear function of the radiance of the corresponding portions
% of the object.
% % Motivation of low dimensionality from spherical harmonics, low-pass interpretation
% Consider the vector space of all illuminations
% of the object.  These are just positive functions (or more
% generally, distributions) defined on the sphere, where each
% point on the sphere corresponds to a direction of incoming
% light.  Consider the vector space formed by the light exiting
% the object.  This too is a positive function defined on the
% sphere, where now each point on the sphere corresponds to the
% normal of a patch on the object (or any other patch with the
% same normal).  Under the Lambertian assumption, the radiance of
% a patch depends on the cosine of the incident angle of incoming
% light, and is integrated over the half-sphere of illumination
% that the patch can "see". Thus, the surface of the object acts
% on the sphere in a manner very similar to a half-cosine
% approximation to a low-pass filter for a one-dimensional
% signal.  Thus the energy leaving the object is
% disproportionately concentrated in the subspace corresponding
% to low spatial frequencies.  Since the image is a linear
% function of this, the space of images of an object will also
% tend to fall on (the positive portion of) a subspace.   In
% \cite{Basri2003-PAMI}, Basri showed using spherical harmonic
% basis functions that nine basis illuminations (corresponding to
% the lowest frequency spherical harmonics) result in training
% images that do a good job of linearly interpolating all other
% training images.  It should be noted, however, that these basis
% illuminations are not strictly positive, and thus neither are
% the training images they generate (when rendered on a computer
% that can handle negative illumination).  \footnote{Note that
% while the spherical harmonic basis functions have regions where
% they are negative, this does no necessary preclude this
% theoretical result from being put to use in practice.  If the
% geometry of the training illumination system were well
% calibrated, it would be possible to generate approximations of
% the positive and (rectified) negative components of the basis
% illuminations separately, taking the difference of the
% resulting two images, and storing it using a signed datatype.
% Combinations of these images may not be positive, but they may
% still be good enough to be useful.}

The strongest theoretical results so far regarding the relationship
between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
That paper assumes that convex, Lambertian objects are photographed under distant illuminations with a fixed pose.
Under those assumptions, the incident and reflected light are represented using distributions on a sphere and can thus be studied in a spherical harmonic basis.
The Lambertian nature of the object acts as a low-pass filter between the incident and reflected light distributions,
and as a result, the set of images of the object end up lying very
close to the subspace corresponding to low frequency spherical harmonics.\footnote{This is highly condensed; please refer to \cite{Basri2003-PAMI} for a  complete description of this result.}
Indeed, only nine (properly chosen) basis illuminations are sufficient to generate basis images
that span all possible images of the object to a good approximation.  While this is a very important result for understanding the image
formation process, the direct application of this result in most practical
systems is misguided for several reasons.
% Non-physicality
Neither the basis illuminations, nor the basis images used
in the Basri analysis are physical.  Like Fourier bases, they have negative components,
and are thus not, at least directly, physically realizable.
% This alone is not
% a showstopper, however: one could, for example, decompose the basis images into
% their positive and negative parts and use them separately.  If the images
% are being rendered from a 3D model, ``negative'' light can be permitted.
There are ways to engineer around this difficulty\footnote{such as allowing negative light while rendering from a 3D model, decomposing mixed-sign illuminations and images into positive negative parts, and enforcing positivity only in the output image}, but the photometric and geometric assumptions behind this model limit its applicability even more severely.

%Studio photographers and movie-makers
%go to great lengths to control their effects through makeup and lighting.
% Not only does specularity allow some of the high angular frequency components of the
% incident illumination to affect the outgoing illumination, it even prevents
% the relationship between incoming and outgoing light from even being expressable as a convolution.
% Indeed, in the extreme case of an image of a mirror-surface sphere, the image is basically a warped
% version of the sphere's incident illumination, and there is no reason to believe the set of
% images is any lower dimensional than the set of illuminations.
Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
and they all do so in a way that violates the modeling assumptions of the Basri analysis.
Self-shadowing and inter-reflection break the simple relationship between the outgoing light
and the resulting images, and specularity even invalidates the representation of outgoing light itself,
since it is no longer independent of viewing angle.
%If the surface has a specular component, the relationship between incoming and outgoing light
%is no longer expressible as a convolution, and if there is self-shadowing or inter-reflection the illumination
%of each patch will depend not only on its location on the surface, but on the illumination of other pixels in the image.

Fortunately, even with these effects, the relationship between the space of illuminations and the space of images is still linear; all we have to assume is that the image is dominated by scattered light.\footnote{Phenomena that break this assumption include fluorescence, the photochromic (``Transition'') lenses in some eyeglasses, and, the infrared light radiated by all warm objects.}.  Note that while the relationship between illuminations and images is linear, only positive weights are allowed; the space of all images of an object with fixed pose and varying illumination is a convex cone lying in the positive orthant.  The question becomes, how many images does it take to do a good job of representing images sampled from this cone?

It has been observed in various empirical studies that
one can get away with using a small number of frontal
illuminations to linearly represent a wide range of new frontal
illuminations, when they are all taken under the same laboratory conditions
\cite{Georghiades2001-PAMI}. This is the case for many public
face datasets, including AR, ORL, PIE, and Multi-PIE.
Unfortunately, we have found that in practice, a training
database consisting purely of frontal illuminations is not
sufficient to linearly represent images of a faces taken
under typical indoor or outdoor conditions (see the experiment
conducted in Section \ref{sec:own-data}). As illustrated by the
example in Figure \ref{fig:promo}, an insufficient number of
training illuminations can result in recognition failure. To
ensure our algorithm works in practice, we need to find a set
of training illuminations that are indeed {\em sufficient} to
linearly represent a wide variety of practical indoor and
outdoor illuminations.

\subsection{Capturing a Sufficient Set of Training
Illuminations} To this end, we have designed a system that can
acquire frontal images of a subject while simultaneously illuminating the subject from all directions above horizontal.
A sketch of the
system is shown in Figure \ref{fig:system}: The illumination
system consists of four projectors that display various bright
patterns onto the three white walls in the corner of a dark
room.  The light reflects off of the walls and illuminates the
user's head indirectly.  After taking the frontal illuminations
we rotate the chair by 180 degrees and take pictures from the
opposite direction.  Having two cameras speeds the process
since only the chair needs to be moved in between frontal and
rear illuminations. Our projector-based system has several
advantages over flash-based illumination systems:
\begin{itemize}
\item The illuminations can be modified in software, rather than hardware.
\item It is easy to capture many different illuminations quickly.
\item Good coverage and distant illumination can be achieved simultaneously.
\item There is no need to mount anything on the walls or construct a large dome.
\item The system can be assembled from off-the-shelf hardware.
\end{itemize}
\begin{figure}
\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}}
\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
\vspace{-.5in}
\label{fig:system}
\end{figure}
With our projector system, our choice of illuminations is
constrained only by the need to achieve a good
SNR\footnote{Since illuminations with more pixels illuminated
will have a better SNR (provided they don't saturate), there is
an engineering tradeoff between the SNR and the number of
training images.}, avoid saturation, and achieve
a reasonably short acquisition time.  Two simplifying
assumptions that we make are that every pixel is either turned
fully on or off in every illumination, and that the illuminated
regions do not overlap.

Assuming that each pixel is fully on or off enables us to guarantee
that each illumination image has the same overall intensity, merely
by guaranteeing that we illuminate the same number of pixels in each image.\footnote{Since DLP projectors may have dramatically different response
curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
Since our algorithm depends only the
linearity between the illuminations and the images, and not on the
relative intensities of the illuminations, the designer has the freedom to choose the overall intensity of the illuminations
to prevent saturation or low SNR, in a sort of offline exposure control.

%Assuming that no pixels are shared by multiple illuminations is important;
%for example, any other non-negative basis illuminations formed from positive linear
%combinations of our basis illuminations would span a cone of lower
%volume, and thus the resulting training images might not be
%able to represent as wide a variety of testing images without using
%negative coefficients in $\x$.
%\footnote{This is true if $\x>0$ is enforced;
%the effect of negative coefficients in $\x$ is still under study.}

Assuming that the sequentially illuminated regions do not overlap
results in a set of training images that span a larger
cone than a similar number of overlapping regions.  
This results in training images that require fewer negative
coefficients in $\x$ to represent test images under natural illuminations.
The effect of negative coefficients in $\x$ appears to depend partly
on how the test images are taken and is still under study.

We ran two experiments to guide our choice of illuminations for
our large-scale experiments:
\begin{itemize}
\item {\em Coverage Experiment.} In the first experiment we
    attempt to determine what coverage of the sphere is
    required to achieve good interpolation for test images.
    The subject was illuminated by 100 (50 front, 50 back)
    illuminations arranged in concentric rings centered at
    the front camera.  Subsets of the training images were
    chosen, starting at the front camera and adding a ring
    at a time.  Each time a ring was added to the training
    illumination set, the average $\ell^1$ registration
    error (residual) for a set of test images taken under
    sunlight was computed and plotted in Figure
    \ref{fig:illumination-patterns}(a).  The more rings of
    training illuminations are added, the lower the
    representation error becomes, with diminishing returns.
%\end{itemize}
%\begin{itemize}
\item {\em Granularity Experiment.} In the second
    experiment we attempt to determine how finely divided
    the illumination sphere should be.  At the first
    granularity level, the projectors  illuminate the
    covered area uniformly.  At each subsequent granularity
    level each illuminated cell is divided in two along its
    longer side but intensity doubled.  For each
    granularity level the average $\ell^1$ registration
    error is computed as in the coverage experiment and
    shown in Figure \ref{fig:illumination-sufficiency}(b).
    Again, diminishing returns are observed as more
    illuminations are added.
\end{itemize}
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_cvpr/coverage_experiment_asplode.png} &
\includegraphics[height=1.5in]{figures_cvpr/final_cvpr_illuminations_asplode.png} \\
(a) Coverage Experiment & (b) Chosen Illumination Patterns
\end{tabular}\vspace{2mm}
\caption{{\bf Illumination patterns.}   The cells are illuminated in sequence.  For rear illuminations the sequence is reversed.  In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations. The four rectangular regions correspond to the four projectors.}
\label{fig:illumination-patterns}
%\vspace{-.5in}
\end{figure}
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
\includegraphics[height=1.5in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
(a) Coverage & (b) Granularity
\end{tabular}\vspace{-2mm}
\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
\label{fig:illumination-sufficiency}
%\vspace{-.5in}
\end{figure}

\subsection{Chosen Illumination Patterns} In the plot for the
coverage experiment, Figure \ref{fig:illumination-sufficiency}(a),
 we clearly see two plateau regions: one is after 4 rings
and one is after 10 rings. The first four rings represent the
typical frontal illuminations, which are present in most public
face datasets; however, we see that the residual stabilizes
after 10 rings which include some illuminations from the back
of the subject. This suggests that although the frontal
illuminations account for most of the illumination on the face,
some illuminations from the back are needed in the training set to
represent images with illumination coming from all directions.
In the plot for the granularity experiment, Figure
\ref{fig:illumination-sufficiency}(b), we observe that the
residual reaches a plateau after four divisions, corresponding
to a total of 32 illuminations. Based on the results from both
experiments, we decide to partition the area covered by the
first 10 rings into a total of 38 cells, whose layout is
explained in Figure \ref{fig:illumination-patterns}(b). For
our large-scale experiments, we have collected those
illuminations for all our subjects.\footnote{It is possible
that with further experimentation a reduced set of illuminations
can be found that performs as well or better.}

See below for the 38 training images of one subject:
\begin{figure}[h]
\centering
\includegraphics[width=6.3in]{figures_cvpr/training.png}
\vspace{-.3in}
\end{figure}

%\subsection{Should Positivity in the Representation Coefficients
%be Enforced?} One critical issue in linear illumination models
%is whether to enforce nonnegativity in the coefficients $\x$,
%i.e. whether to model illumination using a cone or a subspace.
%Some authors, i.e. \cite{Basri2003-PAMI}, have chosen to allow
%negative components in the coefficient vector $\x$ when
%representing one image as a sum of other images, and only
%enforce that the linear combination $A\x$ be positive
%everywhere.  Unfortunately, for non-convex objects such as
%faces, enforcing non-negativity of the representation is not
%sufficient to guarantee that the resulting image of the object
%is physical.  For example, consider a Lambertian planar object
%with a cylindrical hole drilled in it that is aligned with the
%optical axis of an orthographic camera.  Consider two images:
%one under illumination from a distant point source aligned with
%the hole, and one off of the hole axis.  Positively weighting
%the on-axis image and negatively weighting the on-axis image
%can result in an image where the bottom of the hole is more
%strongly illuminated than the surrounding plane.  The image is
%positive, but it is clearly non-physical.  It is worth
%emphasizing that this phenomenon is a result of some points on
%an object seeing a restricted subset of the illumination sphere
%(a violation of convexity),  and not a result of the
%directionality of the light chosen in the thought experiment.
%While the number of coefficients that actually end up negative
%after optimization is usually very small, we have observed
%pathological cases where the shadows next to the user's nose in
%the reconstructed image invert in value.

%Instead of using a smaller set of training images, allowing the
%coefficients to go negative to try to represent more test
%illuminations, and risking over-fitting, we decided to enforce
%non-negativity of the coefficients.  Nonnegative combinations
%of images are guaranteed to correspond to physically plausible
%images.  all physical illuminations unless the training images
%actually span the boundary of the illumination cone. Because we
%have a flexible acquisition system, we can directly generate a
%set of illuminations that span most of the illumination cone,
%without resorting to negative coefficients and risking
%overfitting.  Thus, in Algorithm 1, we have enforced that the
%coefficients $\x$ be non-negative.

%One of the main factors that complicates face recognition, and computer vision in general, is that two images of the same person's face can be very different, even if the pose is carefully controlled and all points on the person's face are visible in both images.  The most safe, reliable, and general assumption that can be made about the imaging process is that it is linear with respect to the intensity of different light sources.  The set of images of a given object is therefore a cone in the non-negative orthant of the pixel basis.  All recognition algorithms that rely on vision for training rely on images sampled from this cone.

\subsection{Compensating for Gamma Compression} Any algorithm
based on representing a testing image as a weighted sum of
training images must make sure that the image intensity has
been coded linearly with image intensity. While most modern machine vision
cameras capture linearly in intensity, older analog video
cameras often used NTSC defined gamma compression, while most
modern digital consumer and SLR cameras record in sRGB defined
gamma compression.  There are several public face recognition
databases that do not specify the gamma compression of their
images. Similarly most of the software
tools used for computer vision research do not provide any facilities
for gamma decompression, and several common image file formats fail to specify the gamma encoding of
the data they contain.  Because of this situation, the burden of handling
gamma correctly falls entirely on the practitioner.  If gamma correction is not applied properly,
most algorithms (including ours)
will degrade gracefully, but the performance of our recognition algorithm
benefits significantly if gamma is handled properly for both
the training and testing images.

\section{Tests on Public Databases}\label{sec:multipie}
In this section and the next section, we conduct comprehensive
experiments on large-scale face databases to verify the performance
of our algorithm and system. We
first test on the largest public face database available that
is suitable for testing our algorithm, the CMU Multi-PIE.
One shortcoming of the CMU Multi-PIE database for our purposes
is that there is no separate set of test images taken under natural illuminations;
we are left to choose which sets of images to use for testing and training.
To challenge our algorithm, we choose only a small set of illuminations for the training set,
yet we include all illuminations in the testing set. In the following section, we will test our
algorithm on a face dataset that is collected by our own
system. The goal for that experiment will be to show that with a sufficient set of
training illuminations for each subject, our algorithm indeed
works stably and robustly with practical illumination,
misalignment, pose, and occlusion, as already indicated by our
experiment shown in Figure \ref{fig:promo}(bottom).

CMU Multi-PIE provides the most extensive test set among public
datasets. This database contains images of 337 subjects across
simultaneous variation in pose, expression, and illumination.
Of these 337 subjects, we use all of the 249 subjects present
in Session 1 as the training set. The remaining 88 subjects are
treated as ``imposters'', or invalid images. For each of the
249 training subjects, we include frontal images of 7 frontal
illuminations,\footnote{They are illuminations
$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
directional illumination, we subtract the ambient-illuminated
image 0.} taken with neutral expression. As suggested by the
work of \cite{Georghiades2001-PAMI}, these extreme frontal
illuminations would be sufficient to linearly represent other
frontal illuminations, as will also be corroborated by the next
experiment on our own dataset. For the test set, we use all 20
illuminations from Sessions 2-4, which were recorded at
distinct times over a period of several months. The dataset is
challenging due to the large number of subjects, and due to
natural variation in subject appearance over time. Table
\ref{tab:MultiPIE-recognition} shows the result of our
algorithm on each of the 3 testing sessions. Our algorithm
achieves recognition rates above $90\%$ for all three sessions,
with input {\em directly} obtained from the Viola and Jones'
face detector; there was no manual intervention for the test images.
\begin{table}
\caption{Recognition rates on large-scale Multi-PIE database.}
\centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & Session 2 & Session 3 & Session 4  \\
\hline
\hline
LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
\hline
NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
\hline
NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
\hline
\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
\hline
{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
\hline
{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
\hline
\end{tabular}
\label{tab:MultiPIE-recognition} }
\end{table}
\begin{figure}
\centering
\begin{tabular}{cccccc}
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
(a) & (b) & (c) & (d) & (e) & (f)
\end{tabular}
\caption{{\bf Representative examples of failed Multi-PIE subjects}. {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2. Due to change of hair style, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
\label{fig:failed-examples}
%\vspace{-.8in}
\end{figure}
\subsection{Improving the Performance by Choosing Different Sampling Windows}

Our algorithm's errors are mostly caused by a few subjects who
significantly change their appearances between sessions (such
as hair, facial hair, and eyeglasses). Some representative
examples are shown in Figure \ref{fig:failed-examples}. For those subjects, alignment and recognition fail on
almost all test illuminations.

Meanwhile, this observation also suggests that we might be able
to improve the performance of our method by carefully choosing
a face region which is less affected by the above factors for
recognition. In particular, since the forehead region is likely
to be affected by the change of hair style, we try replacing
the previous $80 \times 60$ canonical frame with a new
window that better excludes the forehead. We adjust the
resolution of the window to keep $m$ approximately constant. In addition,
we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
the observation that in many cases the corners
actually contain background. An example of the new window
is shown in Figure~\ref{fig:new-mask}.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_pami/example.png} &
\includegraphics[height=1.5in]{figures_pami/example_new.png}\\
(a) & (b)
\end{tabular}
\caption{{\bf Choosing different sampling windows.} (a) The default window. (b) The newly proposed window.}
\label{fig:new-mask}
%\vspace{-.6in}
\end{figure}

It is shown in Table \ref{tab:MultiPIE-recognition} that the
recognition rates on Multi-PIE indeed increase with this new
window. In addition, Figure \ref{fig:failed-examples}(a), (b),
and (c) illustrate three representative subjects for which the
recognition rates of our algorithm are significantly boosted
with the new window.

However, we should mention that the best choice of the window
is very problem-specific and there is
not a simple guideline to follow. For example, although the new
window performs better on Multi-PIE, the same window doesn't
help at all on our own database, which will be introduced in
the next section. This is because most of the training and
testing images in our database are taken on the same day so the
variation in hair style is very small. Hence, excluding the
forehead part may actually result in loss of useful
discriminative information.

\subsection{Subject Validation}

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS and LDA.

Figure \ref{fig:roc-multipie} plots the receiver operating
characteristic (ROC) curves, which are generated by sweeping
the threshold $t$ through the entire range of possible values
for each algorithm.\footnote{Rejecting invalid images not in
the entire database is much more difficult than deciding if two
face images are the same subject. Figure 8 should not be
confused with typical ROC curves for face similarity, e.g.,
\cite{PhillipsP2007}.} We can see that SCI based approach
significantly outperforms the other three algorithms. Similar
contrasts between our algorithm and baseline algorithms were
also observed for SRC in \cite{Wright2009-PAMI}, though on much
smaller datasets.

\begin{figure}
\centerline{
\includegraphics[width=3in]{figures_pami/pami_roc2.png}
}
\caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
%\vspace{-.6in}
\end{figure}

\subsection{Comparison to Existing Work}
We compare our result to baseline linear-projection-based
algorithms, such as Nearest Neighbor (NN), Nearest Subspace
(NS) \cite{Lee2005-PAMI}, and Linear Discriminant Analysis
(LDA) \cite{Belhumeur1997-PAMI}.\footnote{We do not list
results on PCA \cite{Turk1991-CVPR} as its performance is
always below that of Nearest Subspace.} Since these algorithms
assume pixel-accurate alignment, they are not expected to work
well if the test is not well aligned with the training. In
Table~\ref{tab:MultiPIE-recognition}, we report the results of
those algorithms with two types of input: 1.\ the output of the
Viola and Jones' detector, indicated by a subscript ``$d$'';
2.\ the input face is aligned to the training with manually
selected outer eye corners, indicated by a subscript ``$m$''.
Notice that even with careful manual registration, these
baseline algorithms perform significantly worse than our
algorithm, which uses input directly from the face detector.
The performance drop of the LDA algorithm on Multi-PIE reported here
seems to agree with that reported already in
\cite{Gross2008-FGR}.

We also compare our result to the most recent work
\cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
initial registration is again obtained from manually selected
outer eye corners. Then, a supervised hierarchical sparse
coding model based on local image descriptors is trained, which
enjoys certain translation invariant properties. With the same
training and testing sets, \cite{Yang2010-CVPR} is able to
handle any remaining misalignment and achieves state-of-the-art
performance on the CMU Multi-PIE database.
Table~\ref{tab:MultiPIE-recognition} shows that our algorithm
achieves similar or better performance on different sessions of
Multi-PIE.

\subsection{Recognition with Synthetic Random Block Occlusion}

We further test the robustness of our $\ell^1$-norm based
algorithm to synthetic occlusion. We simulate various levels of
occlusion from 10\% to 50\% by replacing a randomly located
block of the face image with an image of a baboon, as shown in
Figure~\ref{fig:multipie-occ-rec}. In this experiment, to avoid
any other factors that may contribute to extra occlusion of the
face (such as the change of hair style), we choose illumination
10 from Session 1\footnote{This is the same session as the
training set.} as testing. The rest of the experimental setting
remains unchanged. The table in
Figure~\ref{fig:multipie-occ-rec} shows that our algorithm is
indeed capable of handling a moderate amount of occlusion. For
example, at 20\% occlusion, our algorithm still achieves 94.9\%
recognition rate.

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ10.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ20.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ30.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ40.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
\caption{{\bf Recognition under varying level of
random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion (see table below). }
\begin{tabular}{|c|c|c|c|c|c| }
\hline
Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
\hline
\hline
Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
\hline
\end{tabular}
\label{fig:multipie-occ-rec}
\vspace{-.5in}
\end{figure}

\subsection{Recognition with Pose and Expression} We now run tests of
our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
Using the same training set as above, we test our algorithm on
images in Session 2 with $15^\circ$ pose, for all 20
illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
algorithm on images in Session 3 with smile, again for all 20
illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
the performance of our method will be significantly improved if pose and expression data
are available in the training.

\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
system we described in Section \ref{sec:illumination}, and shown in Figure
\ref{fig:system}, we have collected the frontal view of 109
subjects {\em without eyeglasses} under 38 illuminations shown
in Figure \ref{fig:illumination-patterns}. For testing our
algorithm, we have also taken 935 images of these subjects with
a different camera under a variety of practical conditions.

\subsection{Necessity of Rear Illuminations} To see how
training illuminations affect the performance of our algorithm
in practice, we now compare how well a few frontal
illuminations can linearly represent: 1. other frontal illuminations
taken under the same laboratory conditions, and 2. typical
indoor and outdoor illuminations. To this end, we use the face
database acquired by our system and use 7 illuminations per
subject as training. The illuminations are chosen to be similar
to the 7 illuminations used in the previous experiment on
Multi-PIE.\footnote{We use the illumination set $\{6, 9, 12,
13, 18, 21, 22\}$ shown in Figure
\ref{fig:illumination-patterns}(b) to mimic the illumination
set $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
our algorithm on the remaining $24 - 7 = 17$ frontal
illuminations for all the subjects. The recognition rate is
$99.8\%$, nearly perfect. We also test our algorithm on 310
indoor images and 168 outdoor images of these subjects taken
under a variety of lighting conditions (category 1 and 2
specified below), similar to the one shown in Figure
\ref{fig:promo}, and the recognition rates for indoor and
outdoor images drop down to $94.2\%$ and $89.2\%$,
respectively. This is a strong indication that 
frontal illuminations taken under laboratory conditions
are insufficient for representing test images under typical indoor and
outdoor illuminations.\vspace{-0.2in}

\begin{figure}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} 
%\vspace{-.25in}
 \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
%\vspace{-.25in}
 \caption{{\bf Representative examples of category C4}. Top row: successful examples with our method using overlapping blocks. Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
\vspace{-.35in}
\end{figure}

\subsection{Large-Scale Test with Sufficient Training
Illuminations} Now we use all 109 subjects and 38 illuminations
in the training and test on 935 images taken under a variety of
practical illuminations and conditions. We have manually partitioned the test images into four main
categories:
\begin{description}
\item[C1:] 310 \emph{indoor} images of 72 subjects without
    eyeglasses, frontal view
    (Fig.~\ref{fig:examples1-3}, row 1).
\item[C2:] 168 \emph{outdoor} images of 48 subjects without
    eyeglasses, frontal view
    (Fig.~\ref{fig:examples1-3}, row 2).
\item[C3:] 211 images of 32 subjects with \emph{eyeglasses}
    (Fig.~\ref{fig:examples1-3}, row 3).
\item[C4:] 246 images of 56 subjects with \emph{sunglasses}
    (Fig.~\ref{fig:examples4}).
\end{description}
We apply Viola and Jones' face detector on these images and
directly use the detected faces as the input to our algorithm.
Table~\ref{tab:UIUC-recognition} reports the performance of our
algorithm on each category.
\begin{table}	
\centering \caption{Recognition rates on our own database.}
\begin{tabular}{|c|c|c|c|c| }
\hline
Test Category & C1 & C2 & C3 & C4  \\
\hline
\hline
Recognition rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
\hline
\end{tabular}
\label{tab:UIUC-recognition}
\vspace{-.25in}
\end{table}
Since our focus is on face
recognition, the errors do not include failures of the face
detector on some of the more challenging images.
As one can see, our algorithm achieves higher than 95\%
recognition rates on categories 1-3. Furthermore, using the
full set of 38 illuminations indeed improves the performance of
our system under practical illumination conditions compared to
only using a small subset of 7 illuminations. However, the
performance dramatically drops when the faces are occluded by
various types of sunglasses, which could cover up to 40\% of
the entire face. Given the previous experimental results on
synthetic random block occlusions, and given that the
illuminations are more challenging, the result is not
surprising. In the next subsection, we will show how additional
assumptions can be used to improve the recognition performance.

\subsection{Improving the Performance with Occlusion using Overlapping Blocks}
A traditional approach to improve the performance of face
recognition under severe occlusion is to use subregions instead
the entire face as a whole. This idea has been explored in many
earlier works; see \cite{Pentland1994-CVPR, Wright2009-PAMI}
for examples. Since in most real world cases the occlusion is contiguous, it is reasonable to argue that a minority of the
subregions are likely to be affected by the occlusion. In this
paper, we adopt the same idea and partition the face into four
overlapping blocks to better handle sunglasses. This
scheme is illustrated in Figure~\ref{fig:occ-block}. Notice
that in this example three out of the four blocks are partially
or almost completely occluded. In our experiment, each block is
of size $90\times 48$ and covers about two-fifths of the entire
face. The testing and training sets are partitioned in the same
way. We then independently apply
Algorithm~\ref{alg:deformable-src} and compute a sparse
representation after registration for each block independently
with respect to the training set. The recognition
results for individual blocks are then aggregated by voting.

\begin{figure}
\centerline{
\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{-0.2in}}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{-.5in}
\end{figure}

In this experiment, we found that the using the \emph{sparsity
concentration index} (SCI) scores for voting achieves higher
recognition rate than the residual measure used in Algorithm~\ref{alg:deformable-src}, on
category 4 (sunglasses) of our database. The recognition rate
is increased to 78.3\%, compared to 40.9\% obtained without
this partition scheme. This is another evidence of the superior
ability of SCI on subject validation, since a heavily occluded
block can be regarded as an outlier for recognition and should
be rejected while voting.

However, we should point out that a major problem with this
approach is that occlusion cannot always be expected to fall within
 any fixed partition of the face image. Therefore, the
proposed scheme should only be viewed as an example which shows
that the performance under occlusion can be boosted by
leveraging local information of a face as well as global information. Meanwhile, we will
leave the further investigation of more general models (such as Markov
random fields \cite{ZhouZ2009}) for face
recognition with both misalignment and occlusion as an
interesting future work.\vspace{-0.2in}

\section{Conclusion}\label{sec:conclusion}
We have proposed a new algorithm and system for recognizing
human faces from images taken under practical conditions. The
proposed system is very {\em simple}, {\em scalable} both in
terms of computational complexity and recognition performance.
The system is compatible with off-the-shelf face
detectors.
The system achieves extremely {\em stable} performance under
a wide range of variations in illumination, misalignment, and even under small amounts of
pose and occlusion. We achieve very good recognition performance on
large-scale tests with public datasets as well as our practical face
images, while using only frontal 2D images in the training without
any explicit 3D face model.
Our system could potentially be extended to better handle large pose
and expression, either by incorporating training images with different poses or
expressions or by explicitly modeling and compensating the associated deformations
in the alignment stage.

Another important direction for future
investigation is to extend the alignment algorithm to better
tackle contiguous occlusion. We have demonstrated in this
work that misalignment can be naturally handled within the
sparse representation framework. More complicated models for
spatial continuity, such as Markov random fields, have also
been successfully integrated into the computation of a sparse
representation of well-aligned test images
\cite{Cevher2008-NIPS, ZhouZ2009}. A unified approach
for face alignment and recognition in the presence of
contiguous occlusion remains an open problem. \vspace{-0.2in}

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%
% Use this command to get the appendices' numbers in "A", "B" instead of the
% default capitalized Roman numerals ("I", "II", etc.).
% However, the capital letter form may result in awkward subsection numbers
% (such as "A-A"). Capitalized Roman numerals are the default.
%\useRomanappendicesfalse
%
%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgement
\section*{Acknowledgment}
% optional entry into table of contents (if used)
%\addcontentsline{toc}{section}{Acknowledgment}
This work was supported by grants NSF IIS 08-49292, NSF ECCS
07-01676, and ONR N00014-09-1-0230. John Wright would like to
thank Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
Mathematics for discussions related to this work. He would like
to acknowledge support from a Microsoft Research Fellowship and
the Lemelson-Illinois Student Prize.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section
% NOTE: BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/

% can use a bibliography generated by BibTeX as a .bbl file
% standard IEEE bibliography style from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,faces}
\bibliography{faces}

% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% biography section
%
% If you had an eps/pdf photo file (graphicx package needed)
% the extra braces prevent the LaTeX parser from getting confused
% when it sees the complicated \includegraphics command within an
% optional argument. You can create your own macro to make things
% simpler here.
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell.eps}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%
%\begin{biography}{Andrew Wagner}
%Biography text here.
%\end{biography}
%
%\begin{biography}{John Wright}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Arvind Ganesh}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Zihan Zhou}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Yi Ma}
%Biography text here.
%\end{biography}
%

% if you will not have a photo
%\begin{biographynophoto}{John Doe}
%Biography text here.
%\end{biographynophoto}

% insert where needed to balance the two columns on the last page
%\newpage

%\begin{biographynophoto}{Jane Doe}
%Biography text here.
%$\end{biographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2010-12-13 improvements.tex
Initial import

\documentclass{article}
\usepackage{amsmath,amssymb}

\begin{document}

\begin{center}
{\bf \large Summary of Improvements upon our CVPR '09 paper}
\end{center}

\begin{enumerate}

\item We have proposed a faster algorithm, based on the method of Augmented Lagrange Multipliers, for solving the $\ell^1$-norm minimization problems at the alignment and recognition stages of our system. Compared to the highly customized interior-point method used in the conference version of this paper, the new algorithm is slightly faster at the alignment step and is more than 10 times faster for global recognition.

\item Our alignment algorithm has been discussed in more detail in the current version of our paper. For example, the effectiveness of using multiple scales for alignment is empirically justified, and the local convergence of the iterative alignment algorithm is briefly discussed.

\item The discussion on illumination models motivating our acquisition system has been expanded, and a small section on gamma correction, aimed at practitioners, has been added.

\item We have improved the performance of our system by carefully studying the failure cases in the experiments. Observing that the output of a face detector may be poorly centered on the face, and may contain a significant amount of background, we now compute an average transformation based on the per subject alignment results and apply it to the test image before global recognition. In our experiments on CMU Multi-PIE dataset, the effect of using different sampling windows is studied and a new window is proposed to further improve the performance of our system. Higher recognition rates on both public Multi-PIE dataset and our own dataset are reported.

\item We have conducted additional experiments on the Multi-PIE database to study the effect of random block occlusion. We have also compared our approach with the recently proposed Supervised Translation-Invariant Sparse Coding model (Yang et al).

\item We have expanded our own dataset by including more subjects, and by adding many more challenging test images. The number of subjects in our dataset has been increased from 74 to 109, and the number of test images has been increased from 593 to 935.

\item We have proposed a simple partition scheme to improve the performance of system with contiguous occlusion, i.e., sunglasses.

\end{enumerate}

\end{document}

### 2010-12-14 rebuttal.txt
Knocked a couple tasks off of the list

We thank the reviewers for their constructive comments, and in particular
reviewer 1 for an extra in-depth review.  While we have philosophical
differences on several points, we feel the valuable constructive criticism has
helped us strengthen the paper considerably.  We will now respond to the
reviews in order.

Reviewer 1 makes several specific recommendations for reducing the length and
generally improving the presentation of the paper which we have heeded.  Please
refer to the change list for details.

The core of Reviewer 1's objections to our paper is that face recognition in
the scenario we are targetting is solved, and thus, unworthy of further study.
Although it is impossible to prove that no practical face recognition system
exists for access control, to help address this we have added 
experimental comparison to demonstrate both
suboptimal performance of a state-of-the-art general recognition algorithm
on our data, and to maintain ballance, suboptimal performance of our algorithm
on the NIST-style data sets. MAY NEED TO BE TWEAKED PENDING EXPERIMENT RESULTS

We will now respond the objections enumerated by Reviewer 1:

1. Reviewer 1 asserts that in applications where users are cooperative, the
importance of handling occlusions is diminished.  While it is true that
cooperative users will not deliberately obscure their face, it is still
desirable to be robust to things like changes in eyewear and partial occclusion
from hats and scarves.  Furthermore, we have found that the use of the L1-norm
as a robust error function improves performance even when the test image does
not contain occlusions.  Any boundaries where the intensity changes in a
misaligned image can result in a large localised error component.  While
unoccluded faces are mostly smooth, sharp boundaries still occur where there
are cast shadows, especially at depth discontinuities such as the nostrils, the
sides of the nose, and partially open mouths.

Reviewer 1 further compares face recognition to other biometrics such as iris
recognition and fingerprints, and claims that these technogolies are far more
reliable and convenient for access control applications.  We consider this is a
false dichotomy, as multi-modal systems are often used for high-stakes
applications.  Furthermore, it should not be under-emphasized that of all of
these biometrics, face recognition has the potential to be the least intrusive
to the user on a daily basis.  Ideally, face recognition will take place
automatically as you walk up to a door.

The reviewer further takes issue with the practicality of the acquisition
system.  Part of this may be due to a misunderstanding of the system.  The
"illumination pattern" is just a block of white pixels that moves around, and
the purpose is simply to ensure that we illuminate the subject from a wide
variety of illumination directions.  The geometry of the projectors is not
calibrated; they are simply hand positioned on a table so that the projected
images touch without much overlap.  A purpose-build illumination system could
largely do away with even this requirement.

2. Reviewer 1's interest in face recognition for terrorist watchlist and law
enforcement applications is of limited relevance to the evaluation of this
article.  By claiming that "the field of face recognition has set itself the
goal of recognition with as few gallery images as possible", the reviewer is
conflating the constraints of one application of face recognition
(surveillance) with those of face recognition as a whole.  A huge part of the
message of our paper is that for application that allow it, we can and should
be gathering more and higher quality data at enrolment.  This idea is certainly
not new or even recent.  What ~is new is the confluence of: 1. The availability
of hardware to efficiently acquire such data (high intensity, high contrast
ratio DLP projectors) 2. Robust computational tools to use such data in a
principled fashion 3. Computer hardware fast enough to even contemplate using
entire images as features.

3. "It is conceivable that SRC is applicable to any feature as well - which
underscores the fact that it is simply a pattern classification toolbox - on
par with any found in a textbook on pattern recognition. What is it that makes
it special for face recognition?"

Most general pattern classification algorithms do not take into account the
physically motivated linear model that is central (though by no means unique)
to the problem of illumination variation.

4. Our goal is not to achieve the best recognition rates on the NIST family of
benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance the state of
the art in face recognition given cooperative subjects.  The motivating
application of the NIST data sets is government law enforcement (originally for
enforcement of drug prohibition, and more recently shifting towards border
security).  This requires backwards compatibility with old image data, some of
which was gathered before computers were even in widespread use.  Due to these
historical reasons, none of these data sets contain multiple controlled
illuminations.  For example, FRGC contains at most two illuminations per person
in the gallery.  (Two studio lights were kept on, and a third light was
switched either on or off.) The techniques used to gather these data sets are a
poor choice for applications where users are cooperative, and are needlessly
restrictive.  While large public data sets are critical to the rigorous
comparison of recognition algorithms, their existence should not be used to
argue against the development of new data acquisition techniques and algorithms
appropriate for them.

5. We have added comparisons with XYZ systems that are unrelated to our framework.  
Please see the change list for details. UNFINISHED!!!

6. Although we claim to have a minimalist system that works in a realistic
access control setting, we do not claim to have solved all problems in face
recognition.  We will tone down the rhetoric in XYZ places. UNFINISHED!!

7. We will add further discussion contrasting the advantages and disadvantages of our
system vs. systems incorporating 3D models.

8. The issues of (mis)alignment and illumination variance are coupled in the
sense that if one problem is solved, the other problem becomes a lot easier.
WE HAVE CLARIFIED THIS IN THE PAPER.

9. If the data fits our model perfectly, at the correct alignment the only
corruptions are due to occlusions.  Until the images are perfectly aligned,
there is indeed a component of dense error caused by misalignment.  This dense
error component is indeed largest where the image gradient is high (or even
discontinuous), and the derivative of this error component is what drives the
algorithm to its solution.  We explicitly do blur and downsample the images to
mollify registration effects; multiscale alignment is a well-known technique,
and is not new to our paper.  HOSSEIN-PERHAPS YOU HAVE A REFERENCE WE COULD ADD
HERE?

10. Scattered light is light that has had its direction modified by interaction
with materials.  DREW-FIND A REFERENCE OR ARGUE FROM MAXWELL"S EQUATIONS AND
MATERIAL PROPERTIES

11. Rigorous justification of the choice of illuminations is a subject of 
continuing research.  Our current choice is motivated empirically by the
experiments described in the paper.

12. To our knowlege, we are the first to use projectors for illumination in
face recognition applications.  Section 1A of the previous submission already
cites several other applications that use projectors as illumination sources.
WE HAVE ADDED THE REQUESTED REFERENCES TO THE CURRENT SUBMISSION.

13.  It is true that one or both stages of the algorithm could be substituted
with other techniques.  That said, the registration and recognition stages are
coherent in the sense that they are motivated by same modeling assumptions
(illumination model, sparse occlusion, etc.), they use the same robust penalty
on image error (the L1-norm), and they use the same numerical optimization
techniques (ALM).  While conceptual simplicity is a secondary concern in
algorithm design, starting with a minimalist foundation benefits both
practitioners as well as researchers.  Other researchers have already published
more complex extensions to our work.

14. Since we perform alignment separately for each class in the gallery, the
computation for the alignment stage is linear in the size of the gallery.  The
variation in alignment across classes is indeed small, but still changes the
objective enough to affect recognition results.  Aligning to a generic face (or
a gallery based on "aligned" images of many people) is a good idea, and one we
have considered as well.  If nothing else, it could be used to improve
alignment initialization instead of using raw face detector output.  It may
also be useful to use the roughly aligned test image to quicky eliminate some
of the classes from further consideration.  Again, we have deliberately tried
to keep the algorithm as minimalist as possible for this publication.

15. The adaptation of the Lucas-Kanade algorithm to both the l1 norm cost
function and a linear illumination model is novel to our knowlege.  As
discussed earlier, we disagree with Reviewer 1's contention that data
acquisition is not a critical part of the face recognition problem.  We believe
we have demonstrated that using a well-though-out combination of existing ideas
(iterative image alignment, l1-error function, SRC, using projectors for
illumination), one can build a recognition that is conceptually simple, well
motivated, and competitive with state-of-the-art recognition systems for access
control scenarios.

Reviewer 2 raises 5 questions, which we will now address.  Since the
recognition optimization is convex and has no difficultly converging, we assume
that (i)-(iii) are asking about the iterative alignment.

(i) "How often an infinite loop is encountered, and how it is handled."

We place a hard limit on the maximal number of alignment iterations, so the
algorithm is guaranteed to terminate in finite time.  In practice, the
alignment algorithm tends to converge within 4-10 iterations for each
resolution of the multi-scale implementation.  We also place hard bounds on how
large the alignment can be, and terminate alignment early if these bounds are
reached.  In the cases where the alignment "slips off" the face, this is how
alignment terminates.

(ii) "How often the starting point happens to be located in a wrong basin of
attraction, i.e. a wrong face. (If there is any way to know this.)"
(iii) "How often the solution gets drifted away into a wrong basin of
attraction. (If there is any way to know this.)"

Since we align the test image to the training users individually, we avoid the
potential problem of the test image converging to a local minimum corresponding
to a non-match face.

(iv) "What is the criteria or threshold for convergence? If the threshold is
stringent, the algorithms are likely to get stuck in infinite loops, and if the
threshold is liberal, premature convergence would occur. How are such
thresholds picked and how universal are they? Does every new dataset need its
own threshold?"

For the l1 minimizations we threshold the l1-norm of the difference between
successive values of the state (this includes the both the explicit
optimization variables and the lagrangle multipliers).  Similarly, for the
alignment iteration, we threshold the l1-norm of the update to the
transformation parameters.  We used the same set convergence thresholds for all
of the data sets we ran against.  In general, the l1-minimzations are
guaranteed to converge eventually, although convergence may be very slow if the
optimization parameters are chosen poorly.  ARVIND-DISCUSS ANY FURTHER RESULTS
ON CONVERGENCE YOU KNOW OF HERE.

(v) "The problem, however, is that if we minimize a function one variable at
a time, we may actually end up increasing the function (L_{mu}(x,e,lambda_{k})) value."

ARVIND, PLEASE CITE CONVERGENCE RESULTS FOR THIS

Reviewer 3 makes the following requests:

"I would have liked to see the experiment for the case S=1, without
impostors in the dataset, to judge the effect of the "Sparse
Concentration Index"

While we observe improved identification performance when using SCI, the main
purpose of SCI is imposter rejection.  
WE HAVE ADDED EXPERIMENTS TO THE PAPER DEMONSTRATING THE UTILITY OF THE SCI-BASED
RECOGNITION STAGE FOR BOTH IDENTIFICATION AND IMPOSTER REJECTION.

"I would also retitle the paper to include "robust measure"
instead of "sparse representation", as the L1 Cost is just a robust
cost, and the system would probably work as well or better with other
(even non-convex) robust cost functions."

We hope reviewer will find our supplemental justification for the (Sparse
Representation based) recognition convincing.  If so, the reviewer may
therefore also agree that the current title is indeed appropriate.

"It might be possible to streamline the introduction a bit more, such that it
clearly describes the algorithm before going into the details."

WE HAVE ATTEMPTED TO MAKE THE OVERALL STRUCTURE OF OUR SYSTEM CLEAR EARLIER IN THE
PAPER TO AVOID CONFUSION.

"Using and requiring several input images of a person asks for a
comment on the advantage or difference to a system that uses
a 3D representation of a face. Please comment on this!"

WE HAVE ENHANCED OUR DISCUSSION OF SYSTEMS THAT USE 3D MODELS IN THE PAPER.  PLEASE SEE THE
CHANGE LIST FOR DETAILS.

### 2010-12-14 faces_system.tex
Knocked a couple tasks off of the list

@@
 % Alignment
-In holistic recognition algorithms, correspondence between points in the test image and in the training must be achieved.
-A long line of research exists on using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related Active Shape Models \cite{cootes1992active} to register images against a relatively high-dimensional model of plausible face appearances, often leveraging face-specific contours.
-While these model-based techniques have advantages in dealing with variations in expression and pose, they may add unnecessary complexity to a system where subjects are willing to present a neutral expression.
-We prefer to focus on warpings with far fewer parameters, and to use the training images themselves as the appearance model.
-Iterative registration in this spirit dates at least back to the Lucas-Kanade algorithm \cite{lucas1981iterative}.
-
-However, whereas much of the early work on image registration is aimed at the problem of registering nearly identical images, say by minimizing a sum of squared distances, here we must confront several physical factors simultaneously: misalignment, illumination variations, and certain amount of occlusion.
-As we discuss further below, illumination variation can be dealt by expressing the test image as a linear combination of an appropriate set of training images.
-Similar representations have been exploited in illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
-For robustness to gross errors, the $\ell^1$-norm of the residual is a more appropriate objective function than the classical $\ell^2$-norm.
-Its use here is loosely motivated by theoretical results due to Candes and Tao  \cite{CandesE2005-IT} (see also \cite{Wright2008-IT}).
-These two observations lead us to pose the registration problem as the search for a set of transformations and illumination coefficients that minimize the $\ell^1$-norm of the representation error.
-We solve this problem using a generalized Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}.
-Each of these problems can also be solved efficiently using recently developed first-order techniques for $\ell^1$-minimization, which are reviewed in \cite{YangA2010-pp}.
+In holistic recognition algorithms, correspondence between points in the test
+image and in the training must be achieved.  A long line of research exists on
+using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
+Active Shape Models \cite{cootes1992active} to register images against a
+relatively high-dimensional model of plausible face appearances, often
+leveraging face-specific contours.  While these model-based techniques have
+advantages in dealing with variations in expression and pose, they may add
+unnecessary complexity to a system where subjects are willing to present a
+neutral expression.  We prefer to focus on warpings with far fewer parameters,
+and to use the training images themselves as the appearance model. Iterative
+registration in this spirit dates at least back to the Lucas-Kanade algorithm
+\cite{lucas1981iterative}.
+
+However, whereas much of the early work on image registration is aimed at the
+problem of registering nearly identical images, say by minimizing a sum of
+squared distances, here we must confront several physical factors
+simultaneously: misalignment, illumination variations, and certain amount of
+occlusion. As we discuss further below, illumination variation can be dealt by
+expressing the test image as a linear combination of an appropriate set of
+training images. Similar representations have been exploited in
+illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
+For robustness to gross errors, the $\ell^1$-norm of the residual is a more
+appropriate objective function than the classical $\ell^2$-norm. Its use here
+is loosely motivated by theoretical results due to Candes and Tao
+\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
+lead us to pose the registration problem as the search for a set of
+transformations and illumination coefficients that minimize the $\ell^1$-norm
+of the representation error. We solve this problem using a generalized
+Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
+minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
+these problems can also be solved efficiently using recently developed
+first-order techniques for $\ell^1$-minimization, which are reviewed in
+\cite{YangA2010-pp}.
 
@@
 Researchers have tried various techniques to deal with illumination variation.
-In almost all recognition algorithms where only a single gallery image is available per individual, illumination effects are regarded as a nuisance that must be removed before the algorithm can continue.
-This is typically done by making statistical assumptions about how illumination affects the image, and using those assumptions to extract a new representation that is claimed to be illumination invariant.
-Recent examples include \cite{chen2006total} and \cite{zhou2007appearance}.
-However, despite these efforts, truly illumination-invariant features are difficult to obtain from a single input image.
-So, if one's primary concern in designing the recognition system is achieving a high recognition rate, it makes no sense to limit the gallery to a single image per person.
-We therefore take the strategy of sampling many gallery images of each individual under varying illuminations.
-These images are used as the basis for either a convex cone model \cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model \cite{Basri2003-PAMI}.
-Images are captured using a simple-to-construct projector based light stage.
-Most light stages used for face recognition have been constructed for the purpose of creating public data sets to study illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.
-Many other light stages have been used for computer graphics purposes \cite{debevec2000acquiring, jones2005performance}.
-The light source can be moved around manually \cite{masselus2002free}, but this may result in poor consistency between users.
-Structured light applications use projectors to directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D reconstruction, but this is very disturbing to the user.
-To our knowledge, we are the first to use projectors to indirectly illuminate a subject's face for the purpose of face recognition.
+In almost all recognition algorithms where only a single gallery image is
+available per individual, illumination effects are regarded as a nuisance that
+must be removed before the algorithm can continue.  This is typically done by
+making statistical assumptions about how illumination affects the image, and
+using those assumptions to extract a new representation that is claimed to be
+illumination invariant.  Recent examples include \cite{chen2006total} and
+\cite{zhou2007appearance}.  However, despite these efforts, truly
+illumination-invariant features are difficult to obtain from a single input
+image.  So, if one's primary concern in designing the recognition system is
+achieving a high recognition rate, it makes no sense to limit the gallery to a
+single image per person.  We therefore take the strategy of sampling many
+gallery images of each individual under varying illuminations.  These images
+are used as the basis for either a convex cone model
+\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
+\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
+projector based light stage.  To our knowledge, we
+are the first to use projectors to indirectly illuminate a subject's face for
+the purpose of face recognition.
 
@@
 between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
-That paper assumes that convex, Lambertian objects are photographed under distant illuminations with a fixed pose.
-Under those assumptions, the incident and reflected light are represented using distributions on a sphere and can thus be studied in a spherical harmonic basis.
-The Lambertian nature of the object acts as a low-pass filter between the incident and reflected light distributions,
-and as a result, the set of images of the object end up lying very
-close to the subspace corresponding to low frequency spherical harmonics.\footnote{This is highly condensed; please refer to \cite{Basri2003-PAMI} for a  complete description of this result.}
-Indeed, only nine (properly chosen) basis illuminations are sufficient to generate basis images
-that span all possible images of the object to a good approximation.  While this is a very important result for understanding the image
+The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
+%Under those assumptions, the incident and reflected light are represented using distributions on a sphere and can thus be studied in a spherical harmonic basis.
+%The Lambertian nature of the object acts as a low-pass filter between the incident and reflected light distributions,
+%and as a result, the set of images of the object end up lying very
+%close to the subspace corresponding to low frequency spherical harmonics.\footnote{This is highly condensed; please refer to \cite{Basri2003-PAMI} for a  complete description of this result.}
+all images of the object can be well approximated by linear combinations of
+nine (properly chosen) basis images.  The basis images have mixed sign, and
+their illuminations consist of the lowest frequency spherical harmonics.
+While this is a very important result for understanding the image
 formation process, the direct application of this result in most practical
@@
 % Non-physicality
-Neither the basis illuminations, nor the basis images used
-in the Basri analysis are physical.  Like Fourier bases, they have negative components,
-and are thus not, at least directly, physically realizable.
+%Neither the basis illuminations, nor the basis images used
+%in the Basri analysis are physical.  Like Fourier bases, they have negative components,
+%and are thus not, at least directly, physically realizable.
 % This alone is not
@@
 % are being rendered from a 3D model, ``negative'' light can be permitted.
-There are ways to engineer around this difficulty\footnote{such as allowing negative light while rendering from a 3D model, decomposing mixed-sign illuminations and images into positive negative parts, and enforcing positivity only in the output image}, but the photometric and geometric assumptions behind this model limit its applicability even more severely.
-
+%There are ways to engineer around this difficulty\footnote{such as allowing negative light while rendering from a 3D model, decomposing mixed-sign illuminations and images into positive negative parts, and enforcing positivity only in the output image}, but the photometric and geometric assumptions behind this model limit its applicability even more severely.
 %Studio photographers and movie-makers
@@
 and they all do so in a way that violates the modeling assumptions of the Basri analysis.
-Self-shadowing and inter-reflection break the simple relationship between the outgoing light
-and the resulting images, and specularity even invalidates the representation of outgoing light itself,
-since it is no longer independent of viewing angle.
-%If the surface has a specular component, the relationship between incoming and outgoing light
+%Specifically, self-shadowing and inter-reflection break the simple relationship between the outgoing light
+%and the resulting images, and specularity even invalidates the representation of outgoing light that is used.
+% itself,
+%since it is no longer independent of viewing angle.
+%%If the surface has a specular component, the relationship between incoming and outgoing light
 %is no longer expressible as a convolution, and if there is self-shadowing or inter-reflection the illumination
@@
 
-Fortunately, even with these effects, the relationship between the space of illuminations and the space of images is still linear; all we have to assume is that the image is dominated by scattered light.\footnote{Phenomena that break this assumption include fluorescence, the photochromic (``Transition'') lenses in some eyeglasses, and, the infrared light radiated by all warm objects.}.
-Note that while the relationship between illuminations and images is linear, only positive weights are allowed; the space of all images of an object with fixed pose and varying illumination is a convex cone lying in the positive orthant.
-The question becomes, how many images does it take to do a good job of representing images sampled from this cone?
+Fortunately, even with these effects, the relationship between the space of
+illuminations and the space of images is still linear
+\footnote{Materials that break
+this assumption include fluorescent materials and the photochromic (``Transition'') lenses
+in some eyeglasses.  Most materials emmit light in proportion to their
+incident light.}, 
+and that the sensor has a linear response curve. 
+\footnote{ Proper handling of gamma encoding is an important consideration for
+practitioners.  Most cameras apply a non-linear and often undocumented response
+curve to captured images.  A slight degradation of performance will occur if
+gamma compressed images are treated as if they were linear.  We recommend the use
+of cameras with well documented response curves that can be inverted when the 
+image file is loaded.}
+For a more indepth study 
+of the relationship between illumination and images, we refer the reader to
+\cite{belhumeur1998set}.
+While the relationship between illuminations and images is linear,
+only positive weights are allowed; the space of all images of an object with
+fixed pose and varying illumination is a convex cone lying in the positive
+orthant. The question becomes, how many images does it take to do a good job
+of representing images sampled from this cone?
 
@@
 
-\subsection{Capturing a Sufficient Set of Training
-Illuminations} To this end, we have designed a system that can
-acquire frontal images of a subject while simultaneously illuminating the subject from all directions above horizontal.
+\subsection{Capturing a Sufficient Set of Training Illuminations} To this end,
+we have designed a system that can acquire frontal images of a subject while
+simultaneously illuminating the subject from all directions above horizontal.
+
+Most light stages used for face recognition have
+been constructed for the purpose of creating public data sets to study
+illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
+light stages have been used for computer graphics purposes
+\cite{debevec2000acquiring, jones2005performance}. 
+In particular, both LED's
+and projectors have been used for the purpose of acquiring BRDF's for various materials.
+
+" Y. Y. Schechner, S. K. Nayar, P. N. Belhumeur " A Theory of multiplexed illumination", ICCV 2003.  
+Y. Y. Schechner, S. K. Nayar, P. N. Belhumeur "Multiplexing for Optimal Lighting", PAMI 2007.  "
+
+The light source can be
+moved around manually \cite{masselus2002free}, but this may result in poor
+consistency of illuminations between users.  Structured light applications use projectors to
+directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
+reconstruction, but this is very disturbing to the user.
+
 A sketch of the
@@
 rear illuminations. Our projector-based system has several
-advantages over flash-based illumination systems:
+advantages over flash-based illumination systems for face recognition:
 \begin{itemize}
@@
 All recognition algorithms that rely on vision for training rely on images sampled from this cone.
-
-\subsection{Compensating for Gamma Compression} Any algorithm
-based on representing a testing image as a weighted sum of
-training images must make sure that the image intensity has
-been coded linearly with image intensity. While most modern machine vision
-cameras capture linearly in intensity, older analog video
-cameras often used NTSC defined gamma compression, while most
-modern digital consumer and SLR cameras record in sRGB defined
-gamma compression.  There are several public face recognition
-databases that do not specify the gamma compression of their
-images. Similarly most of the software
-tools used for computer vision research do not provide any facilities
-for gamma decompression, and several common image file formats fail to specify the gamma encoding of
-the data they contain.  Because of this situation, the burden of handling
-gamma correctly falls entirely on the practitioner.  If gamma correction is not applied properly,
-most algorithms (including ours)
-will degrade gracefully, but the performance of our recognition algorithm
-benefits significantly if gamma is handled properly for both
-the training and testing images.
 

### 2010-12-21 rebuttal.txt
First almost complete version

@@
 
-Reviewer 1 makes several specific recommendations for reducing the length and
-generally improving the presentation of the paper which we have heeded.  Please
-refer to the change list for details.
+Reviewer 1 makes several specific recommendations for improving the
+presentation of the paper which we have heeded.  Please refer to the change
+list for details.
 
@@
 Although it is impossible to prove that no practical face recognition system
-exists for access control, to help address this we have added 
-experimental comparison to demonstrate both
-suboptimal performance of a state-of-the-art general recognition algorithm
-on our data, and to maintain ballance, suboptimal performance of our algorithm
-on the NIST-style data sets. MAY NEED TO BE TWEAKED PENDING EXPERIMENT RESULTS
+exists for access control, to help address this we have added experimental
+comparisons of our system with LBP, one of the leading techniques for single
+image face recognition, on the FERET dataset, as well as on the Multi-PIE
+dataset.  This demonstrates both suboptimal performance of LBP for applications
+where many training images are available, and conversely, suboptimal
+performance of our algorithm on the FERET dataset.
 
@@
 
-5. We have added comparisons with XYZ systems that are unrelated to our framework.  
-Please see the change list for details. UNFINISHED!!!
+5. We have added comparison with a Local Binary Patterns based recognition system, and
+have shown that both our system and LBP perform better in the application domains
+for which they were designed.  Please see the change list for details.
 
@@
 sense that if one problem is solved, the other problem becomes a lot easier.
-WE HAVE CLARIFIED THIS IN THE PAPER.
 
@@
 mollify registration effects; multiscale alignment is a well-known technique,
-and is not new to our paper.  HOSSEIN-PERHAPS YOU HAVE A REFERENCE WE COULD ADD
-HERE?
+and is not new to our paper.
 
-10. Scattered light is light that has had its direction modified by interaction
-with materials.  DREW-FIND A REFERENCE OR ARGUE FROM MAXWELL"S EQUATIONS AND
-MATERIAL PROPERTIES
+10. We have removed the reference to scattered light, as there appear to
+be several common uses of the term.  That said, it is true that linear illumination
+models are physically motivated for scenes far more general than for convex 
+lambertian objects.  Maxwell's equations in free space are linear, and materials
+that create non-linear boundary conditions are the exception
+rather than the rule.  Without proportionality between radiance and irradiance 
+the concept of a Bidirectional Reflectance Distribution Function and its
+generalizations are not even meaningful.
 
@@
 cites several other applications that use projectors as illumination sources.
-WE HAVE ADDED THE REQUESTED REFERENCES TO THE CURRENT SUBMISSION.
+The requested references propose a technique for reducing the SNR of projector
+based illumination sysems for certain classes of objects.  We have added
+discussion of this to the paper.
 
@@
 
-While we observe improved identification performance when using SCI, the main
-purpose of SCI is imposter rejection.  
-WE HAVE ADDED EXPERIMENTS TO THE PAPER DEMONSTRATING THE UTILITY OF THE SCI-BASED
-RECOGNITION STAGE FOR BOTH IDENTIFICATION AND IMPOSTER REJECTION.
+We have added experiments to the paper demonstrating the utility of the
+sci-based recognition stage for both identification and imposter rejection.  In
+general, we see a few percentage points of improvement in recognition rate when
+using SCI over recognition based purely on alignment residual.  
+It is worth emphasizing that while this effect is desirable, 
+the primary value of SCI is its use for imposter rejection.
 
@@
 
-WE HAVE ATTEMPTED TO MAKE THE OVERALL STRUCTURE OF OUR SYSTEM CLEAR EARLIER IN THE
-PAPER TO AVOID CONFUSION.
+We have attempted to make the overall structure of our system clear earlier in the
+paper, by clarifying early on that SRC forms the recognition stage of our system
+in the introduction, as well as by making it clear in the caption for the our motivating example
+is not the final pipeline.
 
@@
 
-WE HAVE ENHANCED OUR DISCUSSION OF SYSTEMS THAT USE 3D MODELS IN THE PAPER.  PLEASE SEE THE
-CHANGE LIST FOR DETAILS.
+The primary advantage of systems that internally use a 3D model of human faces
+is that they have the potential to work over a wider range of pose variation in
+the test image.  In general, systems using 2D models are simpler and easier to
+engineer.  As far as data acquisition is concerned, 2D cameras still have a far
+higher resolution and lower noise levels than 2D cameras.  Our set of 2D images
+may very well encode more useful shape information about a user's face than a
+single low-resolution depth image, especially if recognition is to be performed
+on 2D images.
+
+YI- ON THIS LAST ONE, WOULD IT BE BETTER STRATEGICALLY TO ADMIT NOT HAVING ANY
+EXPERIENCE WITH SYSTEMS THAT USE 3D MODELS FOR FACES?

### 2010-12-21 faces_system.tex
First almost complete version

@@
 Classical holistic subspace based methods
-\cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for
-their speed and simplicity, as well as for their natural
-extension to linear illumination models.  However, their
-performance has been shown to be extremely brittle not only to
-alignment variation, but to occlusions. One promising recent
-direction, set forth in \cite{Wright2009-PAMI}, casts the
-recognition problem as one of finding a sparse representation
-of the test image in terms of the training set as a whole, up
-to some sparse error due to occlusion. A \emph{sparse
-representation-based classification} (SRC) method is then proposed for
-recognition. The main idea is that the sparse nonzero
-coefficients should concentrate on the training samples with
-the same class label as the test sample. SRC has demonstrated
-striking recognition performance despite severe occlusion or
-corruption by solving a simple convex program.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}& \hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png} \\
-\includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}& \hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png} \\
-\includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} & \hspace{3mm}
+\cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and
+simplicity, as well as for their natural extension to linear illumination
+models.  However, their performance has been shown to be extremely brittle not
+only to alignment variation, but to even minor occlusions caused by, say, a
+wisp of hair, a blinked eye, or mouth that is slightly open. One promising
+recent direction, set forth in \cite{Wright2009-PAMI}, casts the recognition
+problem as one of finding a sparse representation of the test image in terms of
+the training set as a whole, up to some sparse error due to occlusion. A
+\emph{sparse representation-based classification} (SRC) method is then proposed
+for recognition. The main idea is that the sparse nonzero coefficients should
+concentrate on the training samples with the same class label as the test
+sample. SRC has demonstrated striking recognition performance despite severe
+occlusion or corruption by solving a simple convex program.  For this
+reason, the final recognition stage of our proposed system will consist of SRC,
+and the alignment stage that precedes it will be based on similar techniques.
+
+\begin{figure} \centering \begin{tabular}{cc}
+\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
+\hspace{3mm}
+\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
+\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
+\hspace{3mm}
+\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
+\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
+\hspace{3mm}
 \includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
-\end{tabular}
-\caption{{\bf Compound effect of registration and illumination}.
-The task is to identify the girl among 20 subjects, by computing the sparse representation of her input face with respect to the entire training set.
-The absolute sum of the coefficients associated with each subject is plotted on the right.
-We also show the faces reconstructed with each subject's training images weighted by the associated sparse coefficients.
-The red line (cross) corresponds to her true identity, subject 12.
-{\bf Top:} The input face is from Viola and Jones' face detector (the black box) and all 38 illuminations specified in Section \ref{sec:illumination} are used in the training.
-{\bf Middle:} The input face is well-aligned (the white box) with the training by our algorithm specified in Section \ref{sec:registration} but only 24 frontal illuminations are used in the training for recognition (see Section \ref{sec:illumination}).
-{\bf Bottom:} Informative representation obtained by using both well-aligned input face and sufficient (all 38) illuminations in the training.
-}
-\label{fig:promo}
+\end{tabular} \caption{{\bf Effects of registration and illumination on
+Recognition}. In this example we identify the girl among 20 subjects, by
+computing the sparse representation of her input face with respect to the
+entire training set. The absolute sum of the coefficients associated with each
+subject is plotted on the right. We also show the faces reconstructed with each
+subject's training images weighted by the associated sparse coefficients. The
+red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
+input face is from Viola and Jones' face detector (the black box) and all 38
+illuminations specified in Section \ref{sec:illumination} are used in the
+training.  {\bf Middle:} The input face is well-aligned (the white box) with
+the training by our algorithm specified in Section \ref{sec:registration} but
+only 24 frontal illuminations are used in the training for recognition (see
+Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and 
+a sufficient set (all 38) of
+illuminations are used in the training. Both are necessary for correct recognition
+using SRC.} \label{fig:promo}
 %\vspace{-.5in}
 \end{figure}
-
 %\subsection{Compound Effect of Registration and Illumination}
+
 Unfortunately, while SRC achieves impressive results, it does not deal with misalignment
-between the test and training images.  Furthermore, the experiments were carried out on public
-datasets
-taken under controlled laboratory conditions
-such as
-Extended Yale B \cite{Georghiades2001-PAMI}, which do not contain realistic test images.
+between the test and training images, and it requires a rich set of illuminations
+in the gallery images for good performance.  
+%Furthermore, the experiments were carried out on public datasets taken under
+%controlled laboratory conditions such as Extended Yale B
+%\cite{Georghiades2001-PAMI}, which do not contain realistic test images.
 %two critical aspects of real world face recognition must be addressed:  significant variations in both the {\em image domain} and in
@@
 
-We briefly review existing techniques for recognition, image registration, and handling of illumination variation.
-Our system is based purely on 2D techniques.
-This fact immediately distinguishes our technique from systems that either require a 3D data as an input, or attempt to estimate a 3D model from input 2D data \cite{amberg2007reconstructing,Blanz2003-PAMI}.
-While these techniques have been shown to achieve better robustness to pose variation given a sufficiently accurate 3D model, for access control applications where the user is cooperative in aligning their head to the camera, robustness to moderate misalignment is sufficient.
-Furthermore, many 3D techniques rely on multiple views to estimate their model, and may have trouble if all of the training images are captured from one pose.
-Note that images of people's faces under varying illumination contain shape-related discriminative information, and this information can be leveraged by 2D algorithms even if shape is not reconstructed explicitly.
+We briefly review existing techniques for recognition, image registration, and
+handling of illumination variation.  Our system is based purely on 2D
+techniques.  This fact immediately distinguishes our technique from systems
+that either require a 3D data as an input, or attempt to estimate a 3D model
+from input 2D data \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
+techniques have been shown to achieve better robustness to pose variation given
+a sufficiently accurate 3D model, for access control applications where the
+user is cooperative in aligning their head to the camera, robustness to
+moderate misalignment is sufficient.  Furthermore, many 3D techniques rely on
+multiple views to estimate their model, and may have trouble if all of the
+training images are captured from one pose.  Note that images of people's faces
+under varying illumination contain shape-related discriminative information,
+and this information can be leveraged by 2D algorithms even if shape is not
+reconstructed explicitly.
 
@@
 squared distances, here we must confront several physical factors
-simultaneously: misalignment, illumination variations, and certain amount of
-occlusion. As we discuss further below, illumination variation can be dealt by
+simultaneously: misalignment, illumination variations, and corrupted pixels.
+As we discuss further below, illumination variation can be dealt by
 expressing the test image as a linear combination of an appropriate set of
@@
 illumination-invariant features are difficult to obtain from a single input
-image.  So, if one's primary concern in designing the recognition system is
-achieving a high recognition rate, it makes no sense to limit the gallery to a
+image.  We argue that if one has the luxury of designinig the recognition algorithm,
+but also the acquisition system, and the application demands a high recognition rate,
+it is inadvisable to limit the gallery to a
 single image per person.  We therefore take the strategy of sampling many
@@
 \cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
-projector based light stage.  To our knowledge, we
+projector based light stage.  While similar systems have been used for
+other applications, to our knowledge, we
 are the first to use projectors to indirectly illuminate a subject's face for
@@
 
-\subsection{Contributions} In this paper, we show how the two
-{\em strongly coupled} issues of registration and illumination
+\subsection{Contributions} In this paper, we show how the 
+mutually confounding issues of registration and illumination
 can be naturally addressed within the sparse representation
@@
 sufficiency in the next section.} plus a sparse error $\e_0$
-due to occlusion. The sparse representation can be recovered by
+due to corrupted pixels. The sparse representation can be recovered by
 minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
@@
 \cite{debevec2000acquiring, jones2005performance}.
-In particular, both LED's
-and projectors have been used for the purpose of acquiring BRDF's for various materials.
-
-" Y. Y. Schechner, S. K. Nayar, P. N. Belhumeur " A Theory of multiplexed illumination", ICCV 2003.  
-Y. Y. Schechner, S. K. Nayar, P. N. Belhumeur "Multiplexing for Optimal Lighting", PAMI 2007.  "
-
 The light source can be
@@
 reconstruction, but this is very disturbing to the user. 
+Y. Schechner \cite{schechner2007multiplexing}
+studies techniques for multiplexing illumination that can dramatically reduce the noise 
+of the demultiplexed images for certain classes of objects and cameras.  
+While these techniques have not been incorporated into the current
+system, they fit elegantly into our framework and will likely be used
+in future implementations.  We stress that use of this multiplexing technique
+is independent from the choice of original (directional) illuminations.
 
@@
 
-Assuming that the sequentially illuminated regions do not overlap
-results in a set of training images that span a larger
-cone than a similar number of overlapping regions.  
-This results in training images that require fewer negative
-coefficients in $\x$ to represent test images under natural illuminations.
-The effect of negative coefficients in $\x$ appears to depend partly
-on how the test images are taken and is still under study.
+Assuming that the sequentially illuminated regions do not overlap results in a
+set of training images that span a larger cone than a similar number of
+overlapping regions.  This results in training images that require fewer
+negative coefficients in $\x$ to represent test images under natural
+illuminations.  The effect of negative coefficients in $\x$ appears to depend
+partly on how the test images are taken and is still under study.
 
@@
 face detector; there was no manual intervention for the test images.
+To demonstrate that the SCI-based recognition step is indeed beneficial
+even when there are no imposters, we include results for recognition
+based only on the alignment error residuals, shown in row 6.
+{\bf DISCUSSION OF ZIHAN'S LBP IMPLEMENTATION GOES HERE}
 \begin{table}
@@
 \cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
+\hline
+{LBP} & ?\% & ?\% & ?\% \\
+\hline
+{Algorithm 1 w/o SCI} & 90.7\% & 89.6\% & 87.5\% \\
 \hline
@@
 \end{figure}
+Our algorithm is not designed for recognition when there is only a single
+gallery image per user, and will not achieve competitive performance,
+especially if the illumination in the test image is significantly different
+from the gallery image. For applications that are constrained to a single
+gallery image per user, algorithms such as LBP are much more appropriate.  We
+demonstrate this by running Algorithm 1, as well as our LBP implementation, on
+three standard subsets of the FERET \cite{phillips1998feret} dataset,
+summarized in Table \ref{tab:FERET-recognition}.  
+\begin{table}
+\caption{Recognition rates on single gallery image FERET dataset}
+\centerline{
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Recognition rate & fb & fc & ? \\
+\hline
+\hline
+{LBP} & ?\% & ?\% & ?\% \\
+\hline
+{Algorithm 1} & ?\% & ?\% & ?\% \\
+\hline
+\end{tabular}
+\label{tab:FERET-recognition} }
+\end{table}
+
 \subsection{Improving the Performance by Choosing Different Sampling Windows}
-
 Our algorithm's errors are mostly caused by a few subjects who
@@
 \hline
-Recognition rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
+Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
 \hline
@@
 \section{Conclusion}\label{sec:conclusion}
-We have proposed a new algorithm and system for recognizing
-human faces from images taken under practical conditions. The
-proposed system is very {\em simple}, {\em scalable} both in
-terms of computational complexity and recognition performance.
-The system is compatible with off-the-shelf face
-detectors.
+Using a well-though-out combination of existing ideas
+(iterative image alignment, l1-error function, SRC, using projectors for
+illumination), one can build a system for recognizing human faces
+from images taken under practical conditions that is conceptually simple, well
+motivated, and competitive with state-of-the-art recognition systems for access
+control scenarios.
+
 The system achieves extremely {\em stable} performance under

### 2010-12-25 rebuttal.tex
Added Yi's revised rebuttal

\documentclass[11pt]{article}

\usepackage{epsfig}
\usepackage{latexsym, amsmath, amsfonts}
%\usepackage{pdfsync}

\renewcommand{\baselinestretch}{1.0}
\parskip 2.2mm
\parindent 0mm
\topmargin -0.60in \oddsidemargin 0.0625in \textheight 9.00in
\textwidth 6.50in

\renewcommand{\Re}{{\mathbb R}}
\newcommand{\Ze}{{\mathbb Z}}
\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}
\newenvironment{proof}{{\bf Proof: \ }}{ \hfill \QED}

\newcommand{\ie}{{\it i.e., }}
\newcommand{\D}{\displaystyle}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{fact}{Fact}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}

\begin{document}

   \bibliographystyle{plain}

   \title{\Large {\bf Manuscript  Revision Statement}}

       \author{{Andrew Wagner, John Wright, Arvind Ganesh, Zihan Zhou, Hossein Mobahi, and Yi Ma}}

   \date{}
   \maketitle

We thank the reviewers for their constructive comments, and in particular
reviewer 1 for an extra in-depth review.  While we have philosophical
differences on several points, we deeply appreciate that the valuable constructive 
criticisms raised by the reviewers have helped us strengthen the paper considerably.  
Below we will respond to comments of the reviewers in order as well as indicate 
specific changes that we have made to address their concerns.

\section{Respond to Reviewer \#1}
Reviewer 1 makes several specific recommendations for improving the
presentation of the paper, many of which we have heeded.  Please refer to the change
list for details.

\paragraph{About Presentation:} The reviewer has suggested we can reduce the description
of the ALM algorithm in Section II-D. We agree ALM has been extensively studied in the optimization 
literature. We do not claim any novelty here. However, we believe a complete description of the algorithm 
here is important for the TPAMI audience: There are
many algorithms for solving the associated optimization problem, with drastically different complexity
and speed. Typical vision researchers might not be so familiar with some of the recent
developments in these algorithms. In particular some of the more effective improvements to ALM are fairly 
recent and scattered in different literatures. In fact, many of the questions raised by reviewer \#2
ask for even more detailed explanation about the algorithms. So we believe, a complete but brief 
description of the ALM-based algorithm is important for the audience of TPAMI to implement and 
test the ideas efficiently and correctly.

We find other suggestions on presentation useful and have made changes accordingly.

\paragraph{About Positioning and Novelty:}
As we have tried to make it clear in the very beginning, the proposed method does not aim to solve face 
recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, hence it
may not have any advantages over other face recognition techniques/systems specifically designed for such scenarios. 
However, there are many practical scenarios where the proposed setting is very relevant: access control
or recognizing refugees, where cooperative training images can (and should) be obtained in advance but no cooperative test
images are needed or available. No need of cooperative test images is arguably the main (or only) advantage that 
face recognition has over other biometric measures for security, such as fingerprint, iris, and DNA. 
If the subjects can cooperate for both training and testing, as suggested by the reviewers, then face 
recognition may no longer be the best choice, in terms of cost and accuracy. The proposed system is an attempt
to build a system that is not only accurate and robust enough to compete with other biometric measures but also 
discreet and non-intrusive when in use. To the best of our knowledge (probably the reviewer may agree), even 
for this admittedly restricted scenario, there is no practical face-recognition system whose performance can meet
the standards of most access control and security purposes. Existing systems have found success  mostly in applications
that are less-demanding in performance such as online image search or photo album organization etc.

\paragraph{About Summary Recommendation:} While we agree with the reviewer that careful evaluation of our
system is needed by comparing it with other face recognition techniques/systems. We must point out that most face recognition systems are not designed for the scenario considered by this paper, hence a fair and meaningful comparison can be difficult. Nevertheless, to better evaluate both the strength and limitation of our method, we have added experimental comparison of our system with LBP, one of the leading techniques used for face verification for most commercial face recognition systems (in addition to other techniques already compared in the paper). We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
datasets.  The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is not controlled nor sufficient.

\paragraph{Detailed Concerns or Comments by Reviewer 1:}
\begin{enumerate}
\item As we have mentioned earlier, we completely agree with the reviewer that if the users are completely cooperative at
both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
and iris are arguably better choices. However, our method does NOT require the users to be cooperative at the test
stage. The purpose of having a good training set and a robust recognition system is precisely to alleviate the users 
from being so cooperative at the test stage. This will allow the system to be less intrusive and much more user-friendly 
as it can handle typically daily change of faces such as occlusion by glasses, hats, or beard and variation in lighting and shadow. Ideally, face recognition will take place automatically and unknowingly as a user walks through a door or works in front a desktop.

In addition, if the accuracy and reliability of a face recognition system can truly get close to that of fingerprint and iris, it can serve as a complementary measure in a multi-modal system and improve the overall security for high-stake applications. To the best of our knowledge, no existing face recognition system, with or without cooperative training, comes close to meet the standards of high-stake security applications.

Furthermore, we have found that the use of the L1-norm as a robust error function improves performance 
even when the test image does not contain occlusions.  Any boundaries where the intensity changes in a
misaligned image can result in a large localised error component.  While
unoccluded faces are mostly smooth, sharp boundaries still occur where there
are cast shadows, especially at depth discontinuities such as the nostrils, the
sides of the nose, and partially open mouths.

\item Reviewer 1's interest in face recognition for terrorist watchlist and law
enforcement applications is of limited relevance to the evaluation of this
article.  By claiming that ``the field of face recognition has set itself the
goal of recognition with as few gallery images as possible,'' the reviewer is
conflating the constraints of one particular type of applications of face recognition
(surveillance) with those of face recognition as a whole. This is a (noble) goal probably easy to
set than to meet. Although progresses have been made towards this goal, the reviewer
probably has to agree that after several decades (and several high-profile failures of
face recognition systems in real-world testing), people are still looking for techniques 
that can achieve the performance required by high-end applications such as homeland 
security. We believe that, before such techniques can be found or such a goal can be
achieved at all, we should not preventing ourselves from making progress in developing systems that 
can already provide good guarantee for success, even if the conditions are somewhat limited.

A huge part of the message of our paper is that for applications that allow it, we can and should
be gathering more and higher quality data at enrollment to our advantage. In fact, majority 
of the crime watch-list consists of criminals who have been committed before. Today,
the acquisition standards for fingerprints are far more strict than that for
capturing face images of criminals and foreigners -- precisely because they are shown to
help fingerprint identification systems. Of course, the idea of better training helps is certainly
not new or even recent.  What is new is the confluence of: 1. The availability
of hardware to efficiently acquire such data (high intensity, high contrast
ratio DLP projectors), 2. Robust and efficient computational tools to use such data in a
principled fashion, 3. Computer hardware fast enough to even contemplate using
entire images as features.

\item ``It is conceivable that SRC is applicable to any feature as well - which
underscores the fact that it is simply a pattern classification toolbox - on
par with any found in a textbook on pattern recognition. What is it that makes
it special for face recognition?''

Most general pattern classification algorithms do not take into account the
physically motivated linear model that is central (though by no means unique)
to the problem of illumination variation.

\item Our goal is not to achieve the best recognition rates on the NIST family of
benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
the art in face recognition given cooperative subjects.  The motivating
application of the NIST data sets is government law enforcement (originally for
enforcement of drug prohibition, and more recently shifting towards border
security).  This requires backwards compatibility with old image data, some of
which was gathered before computers and digital cameras were even in widespread use.  
Due to these historical reasons, none of these data sets contain multiple controlled
illuminations.  For example, FRGC contains at most two illuminations per person
in the gallery.  (Two studio lights were kept on, and a third light was
switched either on or off.) The techniques used to gather these data sets are not a wise 
choice for applications where users are cooperative (foreigners applying for a visa)
or forced to be so (suspects or criminals in custody), and are needlessly
restrictive.  While large public data sets are critical to the rigorous
comparison of recognition algorithms, their existence should not be used to
argue against the development of new data acquisition techniques and algorithms
appropriate for them.

\item We have added comparison with a Local Binary Patterns based recognition system, and
have shown that both our system and LBP perform better in the application domains or settings
for which they were designed.  Please see the change list for details.

\item Although we claim to have a preliminary system that works in a realistic
access control setting, we have never claimed that our method solves all problems in face
recognition. We do not know where the reviewer gets that wrong impression. 
From the very beginning and then at multiple places later, we actually emphasize
the importance of having the more restrictive training data for guaranteeing the success 
of our method.

If it helps, we will try to tone down the rhetoric in XYZ places. UNFINISHED!!

\item We will add further discussion contrasting the advantages and disadvantages of our
system vs. systems incorporating 3D models. This paper does not intend to deal with large 3D pose
although the proposed techniques are directly applicable to the cases when images of more poses or 
even a full 3D models are available.

\item The issues of (mis)alignment and illumination variance are coupled in the
sense that if one problem is solved, the other problem becomes a lot easier. But solving them
separately is not sufficient for solving the face recognition problem.

\item If the data fits our model perfectly, at the correct alignment the only
corruptions are due to occlusions.  Until the images are perfectly aligned,
there is indeed a component of dense error caused by misalignment.  This dense
error component is indeed largest where the image gradient is high (or even
discontinuous), and the derivative of this error component is what drives the
algorithm to its solution.  We explicitly do blur and downsample the images to
mollify registration effects; multiscale alignment is a well-known technique,
and is not new to our paper.

\item We have removed the reference to scattered light, as there appear to
be several common uses of the term.  That said, it is true that linear illumination
models are physically motivated for scenes far more general than for convex 
lambertian objects.  Maxwell's equations in free space are linear, and materials
that create non-linear boundary conditions are the exception
rather than the rule.  Without proportionality between radiance and irradiance 
the concept of a Bidirectional Reflectance Distribution Function and its
generalizations are not even meaningful.

\item Rigorous justification of the choice of illuminations is a subject of 
continuing research.  Our current choice is motivated empirically by the
experiments described in the paper.

\item To our knowlege, we are the first to use projectors for illumination in
face recognition applications.  Section 1A of the previous submission already
cites several other applications that use projectors as illumination sources.
The requested references propose a technique for reducing the SNR of projector
based illumination sysems for certain classes of objects.  We have added
discussion of this to the paper.

\item  It is true that one or both stages of the algorithm could be substituted
with other techniques.  That said, the registration and recognition stages are
coherent in the sense that they are motivated by same modeling assumptions
(illumination model, sparse occlusion, etc.), they use the same robust penalty
on image error (the L1-norm), and they use the same numerical optimization
techniques (ALM).  While conceptual simplicity is a secondary concern in
algorithm design, starting with a minimalist foundation benefits both
practitioners as well as researchers.  Other researchers have already published
more complex extensions to our work.

\item Since we perform alignment separately for each class in the gallery, the
computation for the alignment stage is linear in the size of the gallery.  The
variation in alignment across classes is indeed small, but still changes the
objective enough to affect recognition results.  Aligning to a generic face (or
a gallery based on ``aligned'' images of many people) is a good idea, and one we
have considered as well.  If nothing else, it could be used to improve
alignment initialization instead of using raw face detector output.  It may
also be useful to use the roughly aligned test image to quicky eliminate some
of the classes from further consideration.  Again, we have deliberately tried
to keep the algorithm as minimalist as possible for this publication.

\item The adaptation of the Lucas-Kanade algorithm to both the L1 norm cost
function and a linear illumination model is novel to our knowlege.  As
discussed earlier, we disagree with Reviewer 1's contention that data
acquisition is not a critical part of the face recognition problem -- careful (training) 
data acquisition has been a critical part for almost all other biometric identification 
systems (e.g. fingerprint and iris and DNA)! Why face recognition needs to be so different?

We believe we have demonstrated that using a well-though-out combination of existing ideas
(iterative image alignment, L1-error function, SRC, using projectors for
illumination), one can build a recognition system that is conceptually simple, well
motivated, and competitive with state-of-the-art recognition systems for access
control scenarios.

\end{enumerate}

\section{Respond to Reviewer \#2}
Reviewer 2 raises 5 questions, which we will now address.  Since the
recognition optimization is convex and has no difficultly converging, we assume
that (1)-(3) are asking about the iterative alignment scheme.

\begin{enumerate}
\item (i) ``How often an infinite loop is encountered, and how it is handled."

We place a hard limit on the maximal number of alignment iterations, so the
algorithm is guaranteed to terminate in finite time.  In practice, the
alignment algorithm tends to converge within 4-10 iterations for each
resolution of the multi-scale implementation.  We also place hard bounds on how
large the alignment can be, and terminate alignment early if these bounds are
reached.  In the cases where the alignment ``slips off'' the face, this is how
alignment terminates.

\item (ii) ``How often the starting point happens to be located in a wrong basin of
attraction, i.e. a wrong face. (If there is any way to know this.)''

Since we align the test image to the training users individually, we avoid the
potential problem of the test image converging to a local minimum corresponding
to a non-match face.

\item (iii) ``How often the solution gets drifted away into a wrong basin of
attraction. (If there is any way to know this.)''

In Section II E, we have conducted careful experiments to measure the region of attraction
for the proposed alignment method.

\item (iv) ``What is the criteria or threshold for convergence? If the threshold is
stringent, the algorithms are likely to get stuck in infinite loops, and if the
threshold is liberal, premature convergence would occur. How are such
thresholds picked and how universal are they? Does every new dataset need its
own threshold?''

For the l1 minimizations we threshold the l1-norm of the difference between
successive values of the state (this includes the both the explicit
optimization variables and the Largrange multipliers).  Similarly, for the
alignment iteration, we threshold the l1-norm of the update to the
transformation parameters.  We used the same set convergence thresholds for all
of the data sets we ran against.  With the suggested thresholds, the l1-minimzations are
guaranteed to converge eventually, although convergence may be very slow if they
are chosen poorly.

\item (v) ``The problem, however, is that if we minimize a function one variable at
a time, we may actually end up increasing the function $(L_{\mu}(x,e,\lambda_{k}))$ value.''

This may be true for general nonlinear optimization problem. However for convex problems
we are solving here, the alternating direction methods have been proven to converge, see
[27] and related work by the same authors.

\end{enumerate}

\section{Respond to Reviewer \#3}
Reviewer 3 makes the following requests:

\begin{enumerate}
\item ``I would have liked to see the experiment for the case S=1, without
impostors in the dataset, to judge the effect of the ``Sparse
Concentration Index''.

We have added experiments to the paper demonstrating the utility of the
sci-based recognition stage for both identification and imposter rejection.  In
general, we see a few percentage points of improvement in recognition rate when
using SCI over recognition based purely on alignment residual.  
It is worth emphasizing that while this effect is desirable, 
the primary value of SCI is its use for imposter rejection.

\item ``I would also retitle the paper to include ``robust measure''
instead of ``sparse representation", as the L1 Cost is just a robust
cost, and the system would probably work as well or better with other
(even non-convex) robust cost functions.''

We hope reviewer will find our supplemental justification for the (Sparse
Representation based) recognition convincing.  If so, the reviewer may
therefore also agree that the current title is indeed appropriate.

\item ``It might be possible to streamline the introduction a bit more, such that it
clearly describes the algorithm before going into the details.''

We have attempted to make the overall structure of our system clear earlier in the
paper, by clarifying early on that SRC forms the recognition stage of our system
in the introduction, as well as by making it clear in the caption of the motivating example
is not the final pipeline.

\item ``Using and requiring several input images of a person asks for a
comment on the advantage or difference to a system that uses
a 3D representation of a face. Please comment on this!"

The primary advantage of systems that internally use a 3D model of human faces
is that they have the potential to work over a wider range of pose variation in
the test image.  In general, systems using 2D models are simpler and easier to
engineer.  As far as data acquisition is concerned, 2D cameras still have a far
higher resolution and lower noise levels than 3D cameras.  Based on the principles
of photometric stereos, our set of 2D images may very well encode more useful shape 
information about a user's face than a single low-resolution depth image, especially if 
recognition is to be performed on 2D input images.

\end{enumerate}

\end{document}

### 2010-12-28 rebuttal.tex
Changes from Yi, handing off to John.

@@
 
+Thanks to the reviewer's comments, we have modified the introduction to make this point more clear, especially clarifying the specific scenario where our method can claim novelty.
+
 \paragraph{About Summary Recommendation:} While we agree with the reviewer that careful evaluation of our
@@
 both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
-and iris are arguably better choices. However, our method does NOT require the users to be cooperative at the test
-stage. The purpose of having a good training set and a robust recognition system is precisely to alleviate the users 
+and iris are arguably better choices.
+However, our method does NOT require the users to be completely cooperative at the test stage.
+The purpose of having a good training set and a robust recognition system is precisely to alleviate the users
 from being so cooperative at the test stage. This will allow the system to be less intrusive and much more user-friendly 
@@
 potential problem of the test image converging to a local minimum corresponding
-to a non-match face.
+to a wrong face.
 
@@
 In Section II E, we have conducted careful experiments to measure the region of attraction
-for the proposed alignment method.
+for the proposed alignment method. We have studied the region of attraction of the proposed
+method for both 2D and 3D pose variation, as well as how it can be improved with a proper
+multi-scale implementation.
 
@@
 
+In addition, the sparse representation techniques are not limited to frontal 2D images.
+They can be easily applied to 2D images from multiple poses or to establish alignment
+between a 2D image and a 3D face model (if available). These are all good directions
+for future research.
+
+We have modified Section I.A and the Conclusion to address these issues.
+
 \end{enumerate}

### 2010-12-28 faces_system.tex
Changes from Yi, handing off to John.

@@
 
-\PARstart{T}{here} is a historical tendency for face recognition algorithms to work well under laboratory conditions but degrade when tested in less-controlled environments.
-While there have been several high profile failed trials of face recognition technology for extremely difficult mass surveillance / watch-list applications, face recognition has not even seen widespread applications in more tractable settings, such as access control for buildings, computer systems, automobiles, or even automatic teller machines.\footnote{Face recognition scenarios where both gallery and test data are uncontrolled is an active research area as well \cite{LFW}.}  These applications are very interesting due to their potential sociological impact.
-They are also qualitatively different from surveillance applications because they demand extremely reliable systems, and because the gallery subjects are allies, rather than opponents, of the recognition system.
-This creates the possibility of carefully controlling the acquisition of the training data, even if the testing data may be less controlled for usability reasons.
-
-Classical holistic subspace based methods
-\cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and
-simplicity, as well as for their natural extension to linear illumination
+\PARstart{T}{here} has been a historical tendency for face recognition algorithms to work well under laboratory conditions  (well aligned images, similar lighting conditions, and little change in appearances) but degrade drastically when tested in less-controlled real-world environments.
+Although achieving high recognition performance under uncontrolled conditions and with limited gallery images has been the traditional goal for face recognition, existing techniques have not yet been able to meet the performance requirements of high-stake security-related applications.
+While there have been many successes on less-demanding applications such as online image search and family photo album organization (e.g.
+the freely available Google's Picassa, Microsoft's Photo Gallery, and Apple's iPhoto), there have been many high profile failed trials of face recognition technology worldwide for extremely challenging mass surveillance/watch-list applications.
+
+Somewhat surprisingly, face recognition has not seen widespread applications or success in more tractable security settings, such as access control for secure facilities (e.g., prisons, office buildings), computer systems, automobiles, or even automatic teller machines,  where controlled gallery images can be obtained in advance.\footnote{Face recognition scenarios where both gallery and test data are uncontrolled continue to be an active research area \cite{LFW}, and becomes more popular recently due to aforementioned online applications.}  These applications are very interesting due to their potential sociological and technological impact.
+They are also qualitatively different from mass surveillance or online image applications because they demand extremely reliable systems whose performance must compete with other biometric measures such as fingerprint and iris.
+Now because the gallery subjects are allies, rather than opponents, of the recognition system, this creates the possibility of carefully controlling the acquisition of the training data (as we normally do for all other biometrics).
+However, unlike other biometrics, face recognition has the potential in working with much less controlled testing data so that the access control system can be made more discreet and much less intrusive than using other biometrics.
+To some extent, the goal of this paper is to show how a reliable face recognition system can be built for this admittedly restrictive but important scenario.
+
+Classical holistic subspace based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and simplicity, as well as for their natural extension to linear illumination
 models.  However, their performance has been shown to be extremely brittle not
@@
 sample. SRC has demonstrated striking recognition performance despite severe
-occlusion or corruption by solving a simple convex program.  For this
-reason, the final recognition stage of our proposed system will consist of SRC,
-and the alignment stage that precedes it will be based on similar techniques.
+occlusion or corruption by solving a simple convex program.  
+%For this reason, the final recognition stage of our proposed system will consist of SRC,
+%and the alignment stage that precedes it will be based on similar techniques.
 
@@
 
-Unfortunately, while SRC achieves impressive results, it does not deal with misalignment
+Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
 between the test and training images, and it requires a rich set of illuminations
@@
 %the {\em image value}.
-We illustrate this with an example in Figure \ref{fig:promo}.
+We illustrate the compounded effect of both alignment and illumination 
+with an example in Figure \ref{fig:promo}.
 The task is to identify the girl among 20 subjects. If the test
@@
 training images (caused by mild pose, scale, or misalignment),
-the representation is no longer informative, even if sufficient
-illuminations are present in the training, as shown in Figure
-\ref{fig:promo}(top). In addition, in order to linearly span
+the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative, 
+even if sufficient illuminations are present in the training, as shown in Figure
+\ref{fig:promo}(top). In addition, in order to span
 the illuminations of a typical indoor (or outdoor) environment,
@@
 training. Otherwise, even for perfectly aligned test images,
-the representation will not necessarily be sparse or
+the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
 informative, as shown by the example in Figure
 \ref{fig:promo}(middle). Clearly, both good alignment, as well
-as sufficient training images are needed.  
+as sufficient training images are needed to ensure success of the sparsity-based 
+recognition method proposed by \cite{Wright2009-PAMI}.  
 %Unfortunately, most public face databases lack images with a
@@
 %frontal) illumination, either for training or testing.
+In this paper, we examine how to handle alignment and illumination simultaneously
+in the sparse representation framework, hence bring the method proposed in \cite{Wright2009-PAMI}
+closer to practical use.
 
@@
 techniques have been shown to achieve better robustness to pose variation given
-a sufficiently accurate 3D model, for access control applications where the
-user is cooperative in aligning their head to the camera, robustness to
-moderate misalignment is sufficient.  Furthermore, many 3D techniques rely on
-multiple views to estimate their model, and may have trouble if all of the
-training images are captured from one pose.  Note that images of people's faces
-under varying illumination contain shape-related discriminative information,
-and this information can be leveraged by 2D algorithms even if shape is not
-reconstructed explicitly.
+a sufficiently accurate 3D model, for access control applications where only 
+moderate misalignment and pose are present, the proposed method will be more
+than sufficient.   Note that according to the principle of photometric stereo, images of faces
+under varying illuminations already contain 3D shape-related information,\footnote{In principle,
+one can recover the 3D shape of the face from multiple illuminations using photometric stereo.} 
+and this information can be leveraged by 2D algorithms for alignment and recognition 
+even if shape is not reconstructed explicitly. In addition, the proposed sparsity-based robust
+alignment and recognition techniques are not limited to 2D data and can be easily extended
+to 3D data if necessary.
 
@@
 advantages in dealing with variations in expression and pose, they may add
-unnecessary complexity to a system where subjects are willing to present a
-neutral expression.  We prefer to focus on warpings with far fewer parameters,
-and to use the training images themselves as the appearance model. Iterative
-registration in this spirit dates at least back to the Lucas-Kanade algorithm
-\cite{lucas1981iterative}.
+unnecessary complexity to applications where subjects normally present a neutral face
+or only have moderate expression. We prefer to focus on deformations with far fewer 
+degrees of freedom (say an affine transform), and to use the training images themselves 
+as the appearance model. Iterative registration in this spirit dates at least back to the 
+Lucas-Kanade algorithm \cite{lucas1981iterative}.
 
@@
 problem of registering nearly identical images, say by minimizing a sum of
-squared distances, here we must confront several physical factors
+squared distances or maximize normalized correlation, here we must confront several physical factors
 simultaneously: misalignment, illumination variations, and corrupted pixels.
@@
 illumination-invariant features are difficult to obtain from a single input
-image.  We argue that if one has the luxury of designinig the recognition algorithm,
-but also the acquisition system, and the application demands a high recognition rate,
-it is inadvisable to limit the gallery to a
+image.  We argue that if one has the luxury of designing the acquisition system 
+and the application demands a high recognition rate,
+it is then unwise to limit the gallery to a
 single image per person.  We therefore take the strategy of sampling many
@@
 \subsection{Contributions} In this paper, we show how the 
-mutually confounding issues of registration and illumination
-can be naturally addressed within the sparse representation
+mutually compounding issues of registration and illumination
+can be naturally addressed within a robust sparse representation
 framework. We show that face registration, a challenging
@@
 
-\noindent{\bf Organization of this paper.} In Section
+\noindent{\em Organization of this paper:} In Section
 \ref{sec:registration}, we derive our robust registration and
@@
 
-\section{Handling Practical Registration Error}\label{sec:registration}
+\section{Robust Alignment}\label{sec:registration}
 As demonstrated in Figure \ref{fig:promo}(top), the main
@@
 \begin{algorithmic}[1]
-\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
+\begin{small}
+\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
 $\y\in\Re^m$ and a deformation group $T$. 
@@
 \STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
+\end{small}
 \end{algorithmic}
@@
 \cite{YangJ2009-pp}, called \emph{alternating direction
-method of multipliers} \cite{Glowinski1975-TR}, which proposes to minimize $L_{\mu} (\x,\e,\blamda_k)$
-with respect to $\x$ and $\e$ separately. Then the multiplier
-$\blamda$ is updated after just one iteration of minimization
+method of multipliers} (ADM) \cite{Glowinski1975-TR}, which proposes to minimize $L_{\mu} (\x,\e,\blamda_k)$
+with respect to $\x$ and $\e$ separately. As the objective function is convex and alternation is between two
+terms, this procedure has been proven to converge to the global optima (see \cite{YangJ2009-pp} and references therein).
+
+Then the multiplier $\blamda$ is updated after just one iteration of minimization
 with respect to the two primal variables, giving:
@@
 in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
-\begin{algorithm}[h]
+\begin{algorithm}[t]
 \caption{\bf (Augmented Lagrange Multiplier Method for Global
@@
 \begin{algorithmic}[1]
+\begin{small}
 \STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
@@
 \ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$;
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
 \ENDWHILE \STATE
 {\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{small}
 \end{algorithmic}
@@
 We have selected this algorithm because it strikes the best
-balance between accuracy and scalability for our problem out of
-all that we have tested. We refer the interested reader to
+balance between speed, accuracy, and scalability for our problem out of
+many algorithms that we have tested. We refer the interested reader to
 \cite{YangA2010-pp} for a more detailed discussion of competing
@@
 have discussed the issue of illumination in the next section.
-\begin{enumerate}
-
-\item{\em 2D Deformation.}  We first verify the
+
+\noindent {1) \em 2D Deformation.}  We first verify the
     effectiveness of our alignment algorithm with images
@@
 \end{figure}
-\item{\em Multiscale Implementation.}
+
+\noindent{2) \em Multiscale Implementation.}
         %The
@@
 \label{fig:multiscale}
-\vspace{-.5in}
+\vspace{-.2in}
 \end{figure}
 
-\item{\em 3D Pose Variation.} As densely sampled pose and
+\noindent{3) \em 3D Pose Variation.} As densely sampled pose and
     illumination face images are not available in any of
@@
 \end{figure}
-\end{enumerate}
 
@@
 curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
-Since our algorithm depends only the
+Since our algorithm depends only on  the
 linearity between the illuminations and the images, and not on the
@@
 our large-scale experiments:
-\begin{itemize}
+\begin{enumerate}
 \item {\em Coverage Experiment.} In the first experiment we
@@
     illuminations are added.
-\end{itemize}
+\end{enumerate}
 \begin{figure}
@@
 \end{tabular}\vspace{2mm}
-\caption{{\bf Illumination patterns.}   The cells are illuminated in sequence.
+\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
 For rear illuminations the sequence is reversed.
 In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations.
-The four rectangular regions correspond to the four projectors.}
-\label{fig:illumination-patterns}
+The four rectangular regions correspond to the four projectors.
+}
+\label{fig:illumination-patterns}\label{fig:illumination-sufficiency}
 %\vspace{-.5in}
 \end{figure}
+
 \begin{figure}
@@
 based only on the alignment error residuals, shown in row 6.
-{\bf DISCUSSION OF ZIHAN'S LBP IMPLEMENTATION GOES HERE}
+
 \begin{table}
@@
 \hline
+LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
+\hline
 \cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-{LBP} & ?\% & ?\% & ?\% \\
 \hline
@@
 \end{figure}
-Our algorithm is not designed for recognition when there is only a single
-gallery image per user, and will not achieve competitive performance,
-especially if the illumination in the test image is significantly different
-from the gallery image. For applications that are constrained to a single
-gallery image per user, algorithms such as LBP are much more appropriate.  We
-demonstrate this by running Algorithm 1, as well as our LBP implementation, on
-three standard subsets of the FERET \cite{phillips1998feret} dataset,
-summarized in Table \ref{tab:FERET-recognition}.  
-\begin{table}
-\caption{Recognition rates on single gallery image FERET dataset}
-\centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & fb & fc & ? \\
-\hline
-\hline
-{LBP} & ?\% & ?\% & ?\% \\
-\hline
-{Algorithm 1} & ?\% & ?\% & ?\% \\
-\hline
-\end{tabular}
-\label{tab:FERET-recognition} }
-\end{table}
 
@@
 Multi-PIE.
+
+Finally, we compare our result to Local Binary Patterns (LBP)
+\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
+able to capture final details of facial appearance and texture.
+Due to its robustness to variations in illumination, facial
+expression, aging and other changes, LBP has achieved the
+state-of-the-art face recognition performance in the scenario
+when only one sample per person is used for training
+\cite{Tan06facerecognition}. In this paper, we follow the same
+steps as in \cite{Ahonen2006-PAMI} to construct a LBP
+descriptor for each training and testing sample. The $80\times
+60$ face region is first divided into a regular $10\times 10$
+grid of cells, each of size $8\times 6$ pixels. Within each
+cell, the histogram of 59 uniform binary patterns is then
+computed, where the patterns are generated by thresholding 8
+neighboring pixels in a circle of radius 2 using the central
+pixel value. Finally, the local histograms are concatenated to
+produce the global descriptor vector. As suggested in
+\cite{Ahonen2006-PAMI}, the recognition is performed using a
+nearest neighbor classifier with Chi square distance as the
+distance measure and we report the recognition rates with the
+same two types of input as before. As shown in
+Table~\ref{tab:MultiPIE-recognition}, LBP achieves competitive
+recognition rates given manually aligned training and testing
+samples, which demonstrates its robustness to illumination
+changes. However, its performance is very sensitive to
+misalignment.
+
+Meanwhile, although our algorithm is not designed for
+recognition when there is only a single gallery image per user,
+and will not achieve competitive performance, we compare its
+performance with LBP within such setting for the purpose of
+better evaluation of the strength and limitation of both
+algorithms. For this experiment, we use the FERET dataset
+\cite{phillips1998feret}, which contains five standard
+partitions: `fa' is the gallery containing 1196 frontal images
+of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
+sets of probe images. The testing sets differ from the training
+in facial expression (`fb'), illumination (`fc'), aging (`dup1'
+) and long aging (`dup2'). In fact, except for `fb', we notice
+significant changes of illumination in all the other three test
+sets. For the training, we again crop and normalize the face
+region from each original image to an $80\times 60$ window
+using manually marked eye coordinates \cite{Deng2010-PR}. In
+Table~\ref{tab:FERET-recognition}, we report the performance of
+our algorithm on the four test sets, with input directly
+obtained from the Viola and Jones' detector, and the
+performance of LBP with input obtained from both face detector
+and manual alignment using the eye coordinates.
+
+\begin{table}
+\caption{Recognition rates on single gallery image FERET
+dataset} \centerline{
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Recognition rate & fb & fc & dup1 & dup2 \\
+\hline
+\hline
+LBP$_d$ (LBP$_m$) & 54.8 (96.6)\% & 10.3 (58.8)\% & 29.8 (71.6)\% & 19.8 (61.5)\% \\
+\hline
+{Algorithm 1} & 95.2\% & 28.4\% & 46.1\% & 20.3\% \\
+\hline
+\end{tabular}
+\label{tab:FERET-recognition} }
+\end{table}
+
+As expected, our algorithm doesn't perform well except for
+`fb', in which case the illumination is similar to the training and 
+the mere variation in facial expression is handled well by the sparse error 
+model. For the other three test sets, our algorithm fails because the illumination 
+changes and other variations seriously violate the assumptions of our method. 
+On the other hand, while LBP achieves the best recognition rates given
+manually aligned training and testing samples, its performance
+degrades drastically when the input is obtained directly from
+the face detector. It is also worth noting that similar poor
+performance of LBP, as well as other descriptors, has been
+observed on the Labeled Face in the Wild (LFW) database, where
+the training is uncontrolled and limited and the input is
+directly obtained from the face detector \cite{Wolf2008-ECCV}.
+
+All these experimental results re-emphasize the importance that
+both illumination and alignment need to be simultaneously handled
+well in order to obtain accurate face recognition, even when there is
+no obvious occlusion or corruption in the test.
 
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.35in}
+\vspace{-.15in}
 \end{figure}
@@
 \label{fig:occ-block}
-\vspace{-.5in}
+\vspace{-.2in}
 \end{figure}

### 2010-12-28 rebuttal.tex
Fixed a small english mistake

--- 2010-12-25 rebuttal.tex
@@
 reviewer 1 for an extra in-depth review.  While we have philosophical
-differences on several points, we deeply appreciate that the valuable constructive 
-criticisms raised by the reviewers have helped us strengthen the paper considerably.  
-Below we will respond to comments of the reviewers in order as well as indicate 
-specific changes that we have made to address their concerns.
+differences on several points, we deeply appreciate that the valuable
+constructive criticisms raised by the reviewers have helped us strengthen the
+paper considerably.  Below we will respond to comments of the reviewers in
+order as well as indicate specific changes that we have made to address their
+concerns.
 
-\section{Respond to Reviewer \#1}
-Reviewer 1 makes several specific recommendations for improving the
-presentation of the paper, many of which we have heeded.  Please refer to the change
-list for details.
+\section{Respond to Reviewer \#1} Reviewer 1 makes several specific
+recommendations for improving the presentation of the paper, many of which we
+have heeded.  Please refer to the change list for details.
 
-\paragraph{About Presentation:} The reviewer has suggested we can reduce the description
-of the ALM algorithm in Section II-D. We agree ALM has been extensively studied in the optimization 
-literature. We do not claim any novelty here. However, we believe a complete description of the algorithm 
-here is important for the TPAMI audience: There are
-many algorithms for solving the associated optimization problem, with drastically different complexity
-and speed. Typical vision researchers might not be so familiar with some of the recent
-developments in these algorithms. In particular some of the more effective improvements to ALM are fairly 
-recent and scattered in different literatures. In fact, many of the questions raised by reviewer \#2
-ask for even more detailed explanation about the algorithms. So we believe, a complete but brief 
-description of the ALM-based algorithm is important for the audience of TPAMI to implement and 
-test the ideas efficiently and correctly.
+\paragraph{About Presentation:} The reviewer has suggested we can reduce the
+description of the ALM algorithm in Section II-D. We agree ALM has been
+extensively studied in the optimization literature. We do not claim any novelty
+here. However, we believe a complete description of the algorithm here is
+important for the TPAMI audience: There are many algorithms for solving the
+associated optimization problem, with drastically different complexity and
+speed. Typical vision researchers might not be so familiar with some of the
+recent developments in these algorithms. In particular some of the more
+effective improvements to ALM are fairly recent and scattered accross the
+litterature.  In fact, many of the questions raised by reviewer \#2 ask for
+even more detailed explanation about the algorithms. We feel that a complete
+but brief description of the ALM-based algorithm is important for the audience
+of TPAMI to implement and test the ideas efficiently and correctly.
 
-We find other suggestions on presentation useful and have made changes accordingly.
+We find other suggestions on presentation useful and have made changes
+accordingly.
 
-\paragraph{About Positioning and Novelty:}
-As we have tried to make it clear in the very beginning, the proposed method does not aim to solve face 
-recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, hence it
-may not have any advantages over other face recognition techniques/systems specifically designed for such scenarios. 
-However, there are many practical scenarios where the proposed setting is very relevant: access control
-or recognizing refugees, where cooperative training images can (and should) be obtained in advance but no cooperative test
-images are needed or available. No need of cooperative test images is arguably the main (or only) advantage that 
-face recognition has over other biometric measures for security, such as fingerprint, iris, and DNA. 
-If the subjects can cooperate for both training and testing, as suggested by the reviewers, then face 
-recognition may no longer be the best choice, in terms of cost and accuracy. The proposed system is an attempt
-to build a system that is not only accurate and robust enough to compete with other biometric measures but also 
-discreet and non-intrusive when in use. To the best of our knowledge (probably the reviewer may agree), even 
-for this admittedly restricted scenario, there is no practical face-recognition system whose performance can meet
-the standards of most access control and security purposes. Existing systems have found success  mostly in applications
-that are less-demanding in performance such as online image search or photo album organization etc.
+\paragraph{About Positioning and Novelty:} As we have tried to make it clear in
+the very beginning, the proposed method does not aim to solve face recognition
+for all scenarios! It is not designed for the case where the training is
+uncontrolled and limited, hence it may not have any advantages over other face
+recognition techniques/systems specifically designed for such scenarios.
+However, there are many practical scenarios where the proposed system is very
+relevant: access control or recognizing refugees, where cooperative training
+images can (and should) be obtained in advance but NO COOPERATIVE TEST IMAGES
+ARE NEEDED OR AVAILABLE. No need of cooperative test images is arguably the
+main (or only) advantage that face recognition has over other biometric
+measures for security, such as fingerprint, iris, and DNA.  If the subjects can
+cooperate for both training and testing, as suggested by the reviewers, then
+face recognition may no longer be the best choice, in terms of cost and
+accuracy. The proposed system is an attempt to build a system that is not only
+accurate and robust enough to compete with other biometric measures but also
+discreet and non-intrusive when in use. To the best of our knowledge (probably
+the reviewer may agree), even for this admittedly restricted scenario, there is
+no practical face-recognition system whose performance can meet the standards
+of most access control and security purposes. Existing systems have found
+success  mostly in applications that are less-demanding in performance such as
+online image search or photo album organization etc.
 
-\paragraph{About Summary Recommendation:} While we agree with the reviewer that careful evaluation of our
-system is needed by comparing it with other face recognition techniques/systems.
-We must point out that most face recognition systems are not designed for the scenario considered by this paper, hence a fair and meaningful comparison can be difficult.
-Nevertheless, to better evaluate both the strength and limitation of our method, we have added experimental comparison of our system with LBP, one of the leading techniques used for face verification for most commercial face recognition systems (in addition to other techniques already compared in the paper).
-We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
-datasets.
-The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is not controlled nor sufficient.
+\paragraph{About Summary Recommendation:} While we agree with the reviewer that
+careful evaluation of our system is needed by comparing it with other face
+recognition techniques/systems, we must point out that most face recognition
+systems are not designed for the scenario considered by this paper, hence a
+fair and meaningful comparison can be difficult. Nevertheless, to better
+evaluate both the strengths and limitations of our method, we have added an
+experimental comparison of our system with LBP, one of the leading techniques
+used for face verification for most commercial face recognition systems (in
+addition to other techniques already compared in the paper). We compare both
+methods with the same settings on two datasets that are representative of
+controlled and uncontrolled training: the Multi-PIE and the FERET datasets.
+The results indicate both suboptimal performance of LBP for applications where
+many controlled training images (under different illuminations) are available,
+and conversely, suboptimal performance of our algorithm on the FERET dataset
+where the training is not controlled nor sufficient.
 
-\paragraph{Detailed Concerns or Comments by Reviewer 1:}
-\begin{enumerate}
-\item As we have mentioned earlier, we completely agree with the reviewer that if the users are completely cooperative at
-both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
-and iris are arguably better choices. However, our method does NOT require the users to be cooperative at the test
-stage. The purpose of having a good training set and a robust recognition system is precisely to alleviate the users 
-from being so cooperative at the test stage. This will allow the system to be less intrusive and much more user-friendly 
-as it can handle typically daily change of faces such as occlusion by glasses, hats, or beard and variation in lighting and shadow.
-Ideally, face recognition will take place automatically and unknowingly as a user walks through a door or works in front a desktop.
+\paragraph{Detailed Concerns or Comments by Reviewer 1:} \begin{enumerate}
+\item As we have mentioned earlier, we completely agree with the reviewer that
+if the users are completely cooperative at both training and testing phases,
+face recognition may not necessarily be the best choice of biometric measure --
+fingerprint and iris are arguably better choices. However, our method does NOT
+require the users to be cooperative at the test stage. The purpose of having a
+good training set and a robust recognition system is precisely to alleviate the
+users from being so cooperative at the test stage. This will allow the system
+to be less intrusive and much more user-friendly as it can handle typically
+daily change of faces such as occlusion by glasses, hats, or beard and
+variation in lighting and shadow. Ideally, face recognition will take place
+automatically and unknowingly as a user walks through a door or works in front
+a desktop.
 
-In addition, if the accuracy and reliability of a face recognition system can truly get close to that of fingerprint and iris, it can serve as a complementary measure in a multi-modal system and improve the overall security for high-stake applications.
-To the best of our knowledge, no existing face recognition system, with or without cooperative training, comes close to meet the standards of high-stake security applications.
+In addition, if the accuracy and reliability of a face recognition system can
+truly get close to that of fingerprint and iris, it can serve as a
+complementary measure in a multi-modal system and improve the overall security
+for high-stake applications. To the best of our knowledge, no existing face
+recognition system, with or without cooperative training, comes close to meet
+the standards of high-stake security applications.
 
-Furthermore, we have found that the use of the L1-norm as a robust error function improves performance 
-even when the test image does not contain occlusions.  Any boundaries where the intensity changes in a
-misaligned image can result in a large localised error component.  While
-unoccluded faces are mostly smooth, sharp boundaries still occur where there
-are cast shadows, especially at depth discontinuities such as the nostrils, the
-sides of the nose, and partially open mouths.
+Furthermore, we have found that the use of the L1-norm as a robust error
+function improves performance even when the test image does not contain
+occlusions.  Any boundaries where the intensity changes in a misaligned image
+can result in a large localised error component.  While unoccluded faces are
+mostly smooth, sharp boundaries still occur where there are cast shadows,
+especially at depth discontinuities such as the nostrils, the sides of the
+nose, and partially open mouths.
 
@@
 goal of recognition with as few gallery images as possible,'' the reviewer is
-conflating the constraints of one particular type of applications of face recognition
-(surveillance) with those of face recognition as a whole. This is a (noble) goal probably easy to
-set than to meet. Although progresses have been made towards this goal, the reviewer
-probably has to agree that after several decades (and several high-profile failures of
-face recognition systems in real-world testing), people are still looking for techniques 
-that can achieve the performance required by high-end applications such as homeland 
-security. We believe that, before such techniques can be found or such a goal can be
-achieved at all, we should not preventing ourselves from making progress in developing systems that 
-can already provide good guarantee for success, even if the conditions are somewhat limited.
+conflating the constraints of one particular type of applications of face
+recognition (surveillance) with those of face recognition as a whole. This is a
+(noble) goal probably easy to set than to meet. Although progresses have been
+made towards this goal, the reviewer probably has to agree that after several
+decades (and several high-profile failures of face recognition systems in
+real-world testing), people are still looking for techniques that can achieve
+the performance required by high-end applications such as homeland security. We
+believe that, before such techniques can be found or such a goal can be
+achieved at all, we should not preventing ourselves from making progress in
+developing systems that can already provide good guarantee for success, even if
+the conditions are somewhat limited.
 
-A huge part of the message of our paper is that for applications that allow it, we can and should
-be gathering more and higher quality data at enrollment to our advantage. In fact, majority 
-of the crime watch-list consists of criminals who have been committed before. Today,
-the acquisition standards for fingerprints are far more strict than that for
-capturing face images of criminals and foreigners -- precisely because they are shown to
-help fingerprint identification systems. Of course, the idea of better training helps is certainly
-not new or even recent.  What is new is the confluence of: 1. The availability
-of hardware to efficiently acquire such data (high intensity, high contrast
-ratio DLP projectors), 2. Robust and efficient computational tools to use such data in a
-principled fashion, 3. Computer hardware fast enough to even contemplate using
-entire images as features.
+A huge part of the message of our paper is that for applications that allow it,
+we can and should be gathering more and higher quality data at enrollment to
+our advantage. In fact, majority of the crime watch-list consists of criminals
+who have been committed before. Today, the acquisition standards for
+fingerprints are far more strict than that for capturing face images of
+criminals and foreigners -- precisely because they are shown to help
+fingerprint identification systems. Of course, the idea of better training
+helps is certainly not new or even recent.  What is new is the confluence of:
+1. The availability of hardware to efficiently acquire such data (high
+intensity, high contrast ratio DLP projectors), 2. Robust and efficient
+computational tools to use such data in a principled fashion, 3. Computer
+hardware fast enough to even contemplate using entire images as features.
 
@@
 
-\item Our goal is not to achieve the best recognition rates on the NIST family of
-benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
+\item Our goal is not to achieve the best recognition rates on the NIST family
+of benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
 the art in face recognition given cooperative subjects.  The motivating
@@
 security).  This requires backwards compatibility with old image data, some of
-which was gathered before computers and digital cameras were even in widespread use.  
-Due to these historical reasons, none of these data sets contain multiple controlled
-illuminations.  For example, FRGC contains at most two illuminations per person
-in the gallery.  (Two studio lights were kept on, and a third light was
-switched either on or off.) The techniques used to gather these data sets are not a wise 
-choice for applications where users are cooperative (foreigners applying for a visa)
-or forced to be so (suspects or criminals in custody), and are needlessly
-restrictive.  While large public data sets are critical to the rigorous
-comparison of recognition algorithms, their existence should not be used to
-argue against the development of new data acquisition techniques and algorithms
-appropriate for them.
+which was gathered before computers and digital cameras were even in widespread
+use.  Due to these historical reasons, none of these data sets contain multiple
+controlled illuminations.  For example, FRGC contains at most two illuminations
+per person in the gallery.  (Two studio lights were kept on, and a third light
+was switched either on or off.) The techniques used to gather these data sets
+are not a wise choice for applications where users are cooperative (foreigners
+applying for a visa) or forced to be so (suspects or criminals in custody), and
+are needlessly restrictive.  While large public data sets are critical to the
+rigorous comparison of recognition algorithms, their existence should not be
+used to argue against the development of new data acquisition techniques and
+algorithms appropriate for them.
 
-\item We have added comparison with a Local Binary Patterns based recognition system, and
-have shown that both our system and LBP perform better in the application domains or settings
-for which they were designed.  Please see the change list for details.
+\item We have added comparison with a Local Binary Patterns based recognition
+system, and have shown that both our system and LBP perform better in the
+application domains or settings for which they were designed.  Please see the
+change list for details.
 
@@
 
-\item We will add further discussion contrasting the advantages and disadvantages of our
-system vs. systems incorporating 3D models. This paper does not intend to deal with large 3D pose
-although the proposed techniques are directly applicable to the cases when images of more poses or 
-even a full 3D models are available.
+\item We will add further discussion contrasting the advantages and
+disadvantages of our system vs. systems incorporating 3D models. This paper
+does not intend to deal with large 3D pose although the proposed techniques are
+directly applicable to the cases when images of more poses or even a full 3D
+models are available.
 
 \item The issues of (mis)alignment and illumination variance are coupled in the
-sense that if one problem is solved, the other problem becomes a lot easier. But solving them
-separately is not sufficient for solving the face recognition problem.
+sense that if one problem is solved, the other problem becomes a lot easier.
+But solving them separately is not sufficient for solving the face recognition
+problem.
 
@@
 
-\item We have removed the reference to scattered light, as there appear to
-be several common uses of the term.  That said, it is true that linear illumination
-models are physically motivated for scenes far more general than for convex 
-lambertian objects.  Maxwell's equations in free space are linear, and materials
-that create non-linear boundary conditions are the exception
+\item We have removed the reference to scattered light, as there appear to be
+several common uses of the term.  That said, it is true that linear
+illumination models are physically motivated for scenes far more general than
+for convex lambertian objects.  Maxwell's equations in free space are linear,
+and materials that create non-linear boundary conditions are the exception
 rather than the rule.  Without proportionality between radiance and irradiance
@@
 objective enough to affect recognition results.  Aligning to a generic face (or
-a gallery based on ``aligned'' images of many people) is a good idea, and one we
-have considered as well.  If nothing else, it could be used to improve
+a gallery based on ``aligned'' images of many people) is a good idea, and one
+we have considered as well.  If nothing else, it could be used to improve
 alignment initialization instead of using raw face detector output.  It may

### 2010-12-31 rebuttal.tex
Changes from John.

--- 2010-12-28 rebuttal.tex
@@
 reviewer 1 for an extra in-depth review.  While we have philosophical
-differences on several points, we deeply appreciate that the valuable constructive 
-criticisms raised by the reviewers have helped us strengthen the paper considerably.  
-Below we will respond to comments of the reviewers in order as well as indicate 
+differences on several points, we deeply appreciate the constructive 
+criticisms raised by the reviewers, which have helped us strengthen the paper considerably.  
+Below, we respond to comments of the reviewers in order, indicating
 specific changes that we have made to address their concerns.
 
-\section{Respond to Reviewer \#1}
-Reviewer 1 makes several specific recommendations for improving the
+\section{Response to Reviewer \#1}
+Reviewer 1 makes specific recommendations for improving the
 presentation of the paper, many of which we have heeded.  Please refer to the change
@@
 
-\paragraph{About Presentation:} The reviewer has suggested we can reduce the description
-of the ALM algorithm in Section II-D. We agree ALM has been extensively studied in the optimization 
+\paragraph{About Presentation:} The reviewer suggests that we can reduce the description
+of the ALM algorithm in Section II-D. We agree that ALM has been extensively studied in the optimization 
 literature. We do not claim any novelty here. However, we believe a complete description of the algorithm 
@@
 recent and scattered in different literatures. In fact, many of the questions raised by reviewer \#2
-ask for even more detailed explanation about the algorithms. So we believe, a complete but brief 
+ask for even more detailed explanation about the algorithms. So, we believe that a complete but brief 
 description of the ALM-based algorithm is important for the audience of TPAMI to implement and 
@@
 
-We find other suggestions on presentation useful and have made changes accordingly.
+We find the reviewer's other suggestions on presentation useful and have made changes accordingly.
 
 \paragraph{About Positioning and Novelty:}
-As we have tried to make it clear in the very beginning, the proposed method does not aim to solve face 
-recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, hence it
+As we have tried to make clear, the proposed method does not aim to solve face 
+recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, and hence it
 may not have any advantages over other face recognition techniques/systems specifically designed for such scenarios. 
 However, there are many practical scenarios where the proposed setting is very relevant: access control
-or recognizing refugees, where cooperative training images can (and should) be obtained in advance but no cooperative test
-images are needed or available. No need of cooperative test images is arguably the main (or only) advantage that 
-face recognition has over other biometric measures for security, such as fingerprint, iris, and DNA. 
+or recognizing fugitives, where cooperative training images can (and should) be obtained in advance but no cooperative test
+images are needed or available. The possibility of obtaining reliable identification without cooperative test images is arguably the main (or only) advantage of 
+face recognition over other biometric measures such as fingerprint, iris, and DNA. 
 If the subjects can cooperate for both training and testing, as suggested by the reviewers, then face 
@@
 discreet and non-intrusive when in use. To the best of our knowledge (probably the reviewer may agree), even 
-for this admittedly restricted scenario, there is no practical face-recognition system whose performance can meet
-the standards of most access control and security purposes. Existing systems have found success  mostly in applications
-that are less-demanding in performance such as online image search or photo album organization etc.
+for this admittedly restricted scenario, there is no practical face recognition system whose performance can meet
+the standards of most access control and security applications. Existing systems have found success  mostly in applications
+that are less demanding in performance, such as online image search or photo album organization.
 
@@
 
-\paragraph{About Summary Recommendation:} While we agree with the reviewer that careful evaluation of our
-system is needed by comparing it with other face recognition techniques/systems.
-We must point out that most face recognition systems are not designed for the scenario considered by this paper, hence a fair and meaningful comparison can be difficult.
-Nevertheless, to better evaluate both the strength and limitation of our method, we have added experimental comparison of our system with LBP, one of the leading techniques used for face verification for most commercial face recognition systems (in addition to other techniques already compared in the paper).
+\paragraph{About Summary Recommendation:} We agree with the reviewer that there is a need for a careful evaluation of our system and comparison to other techniques or systems for face recognition.
+However, because most face recognition systems are not designed for the scenario considered by this paper, a fair and meaningful comparison is difficult.
+Nevertheless, to clarify both the strengths and limitations of our approach, we have augmented the experimental section with comparisons to Local Binary Patterns (LBP), in addition to the comparisons already performed.
+LBP is one of the most popular tools for face verification, and is used in many commercial face recognition systems.
+
 We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
 datasets.
-The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is not controlled nor sufficient.
+The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is neither controlled nor sufficient.
 
@@
 \begin{enumerate}
-\item As we have mentioned earlier, we completely agree with the reviewer that if the users are completely cooperative at
+\item As mentioned above, we agree that if the users are completely cooperative at
 both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
@@
 However, our method does NOT require the users to be completely cooperative at the test stage.
-The purpose of having a good training set and a robust recognition system is precisely to alleviate the users
-from being so cooperative at the test stage. This will allow the system to be less intrusive and much more user-friendly 
-as it can handle typically daily change of faces such as occlusion by glasses, hats, or beard and variation in lighting and shadow.
+The purpose of having a good training set and a robust recognition system is precisely to alleviate this requirement.
+This will allow the system to be less intrusive and much more user-friendly,
+as it can handle typical daily changes in facial appearance, such as occlusion by glasses, hats, or facial hair and variations in lighting.
 Ideally, face recognition will take place automatically and unknowingly as a user walks through a door or works in front a desktop.
 
-In addition, if the accuracy and reliability of a face recognition system can truly get close to that of fingerprint and iris, it can serve as a complementary measure in a multi-modal system and improve the overall security for high-stake applications.
+In addition, if the accuracy and reliability of a face recognition system can truly get close to that of fingerprint and iris, it can serve as a complementary measure in a multi-modal system and improve the overall security for high-stakes applications.
 To the best of our knowledge, no existing face recognition system, with or without cooperative training, comes close to meet the standards of high-stake security applications.
@@
 even when the test image does not contain occlusions.  Any boundaries where the intensity changes in a
-misaligned image can result in a large localised error component.  While
+misaligned image can result in a large, localized error component.  While
 unoccluded faces are mostly smooth, sharp boundaries still occur where there
@@
 enforcement applications is of limited relevance to the evaluation of this
-article.  By claiming that ``the field of face recognition has set itself the
+manuscript.  By claiming that ``the field of face recognition has set itself the
 goal of recognition with as few gallery images as possible,'' the reviewer is
 conflating the constraints of one particular type of applications of face recognition
-(surveillance) with those of face recognition as a whole. This is a (noble) goal probably easy to
-set than to meet. Although progresses have been made towards this goal, the reviewer
+(surveillance) with those of face recognition as a whole. This (noble) goal is probably easier to
+set than to meet. Although progress has been made towards this goal, the reviewer
 probably has to agree that after several decades (and several high-profile failures of
@@
 that can achieve the performance required by high-end applications such as homeland 
-security. We believe that, before such techniques can be found or such a goal can be
-achieved at all, we should not preventing ourselves from making progress in developing systems that 
-can already provide good guarantee for success, even if the conditions are somewhat limited.
+security. Having such a high, long-term goal should not prevent us from making progress in developing systems that 
+can already provide good guarantees of success, even if the conditions are somewhat limited.
 
 A huge part of the message of our paper is that for applications that allow it, we can and should
-be gathering more and higher quality data at enrollment to our advantage. In fact, majority 
-of the crime watch-list consists of criminals who have been committed before. Today,
-the acquisition standards for fingerprints are far more strict than that for
-capturing face images of criminals and foreigners -- precisely because they are shown to
-help fingerprint identification systems. Of course, the idea of better training helps is certainly
-not new or even recent.  What is new is the confluence of: 1. The availability
+be gathering more and higher quality data at enrollment. In fact, criminal watchlists typically contain many criminals who have been incarcerated before. Today,
+the acquisition standards for fingerprints are far more strict than those for
+capturing face images of criminals -- precisely because the availability of better data has been shown to 
+help fingerprint identification systems. Of course, the idea that better training helps is certainly
+not new or even recent.  What is new is: 1. The availability
 of hardware to efficiently acquire such data (high intensity, high contrast
@@
 benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
-the art in face recognition given cooperative subjects.  The motivating
+the art in face recognition given cooperative training images.  The motivating
 application of the NIST data sets is government law enforcement (originally for
@@
 recognition. We do not know where the reviewer gets that wrong impression. 
-From the very beginning and then at multiple places later, we actually emphasize
-the importance of having the more restrictive training data for guaranteeing the success 
+In the introduction, and at multiple places later in the manuscript, we emphasize
+the importance of having more restrictive training data for guaranteeing the success 
 of our method.
 
-If it helps, we will try to tone down the rhetoric in XYZ places. UNFINISHED!!
+%If it helps, we will try to tone down the rhetoric in XYZ places. UNFINISHED!!
 
 \item We will add further discussion contrasting the advantages and disadvantages of our
-system vs. systems incorporating 3D models. This paper does not intend to deal with large 3D pose
+system vs.\ systems incorporating 3D models. This paper does not intend to deal with large 3D pose
 although the proposed techniques are directly applicable to the cases when images of more poses or 
@@
 
-\item The issues of (mis)alignment and illumination variance are coupled in the
+\item The issues of (mis)alignment and illumination variation are coupled in the
 sense that if one problem is solved, the other problem becomes a lot easier. But solving them
@@
 discontinuous), and the derivative of this error component is what drives the
-algorithm to its solution.  We explicitly do blur and downsample the images to
+algorithm to its solution.  We do explicitly  blur and downsample the images to
 mollify registration effects; multiscale alignment is a well-known technique,
@@
 models are physically motivated for scenes far more general than for convex 
-lambertian objects.  Maxwell's equations in free space are linear, and materials
+Lambertian objects.  Maxwell's equations in free space are linear, and materials
 that create non-linear boundary conditions are the exception
@@
 
-\item To our knowlege, we are the first to use projectors for illumination in
+\item To our knowledge, we are the first to use projectors for illumination in
 face recognition applications.  Section 1A of the previous submission already
 cites several other applications that use projectors as illumination sources.
-The requested references propose a technique for reducing the SNR of projector
-based illumination sysems for certain classes of objects.  We have added
+The requested references propose a technique for improving the SNR of projector-based 
+illumination systems for certain classes of objects.  We have added
 discussion of this to the paper.
@@
 alignment initialization instead of using raw face detector output.  It may
-also be useful to use the roughly aligned test image to quicky eliminate some
+also be useful to use the roughly aligned test image to quickly eliminate some
 of the classes from further consideration.  Again, we have deliberately tried
@@
 \item The adaptation of the Lucas-Kanade algorithm to both the L1 norm cost
-function and a linear illumination model is novel to our knowlege.  As
+function and a linear illumination model is novel to our knowledge.  As
 discussed earlier, we disagree with Reviewer 1's contention that data
@@
 data acquisition has been a critical part for almost all other biometric identification 
-systems (e.g. fingerprint and iris and DNA)! Why face recognition needs to be so different?
+systems (e.g.\ fingerprint and iris and DNA)! Why should face recognition needs be different?
 
-We believe we have demonstrated that using a well-though-out combination of existing ideas
+We believe we have demonstrated that using a well-thought-out combination of existing ideas
 (iterative image alignment, L1-error function, SRC, using projectors for
@@
 
-\section{Respond to Reviewer \#2}
-Reviewer 2 raises 5 questions, which we will now address.  Since the
-recognition optimization is convex and has no difficultly converging, we assume
+\section{Response to Reviewer \#2}
+Reviewer 2 raises five questions, which we will now address.  Since the
+recognition optimization is convex, and convergence is therefore not an issue, we assume
 that (1)-(3) are asking about the iterative alignment scheme.
@@
 
-We place a hard limit on the maximal number of alignment iterations, so the
+We place a hard limit on the maximum number of alignment iterations, so the
 algorithm is guaranteed to terminate in finite time.  In practice, the
 alignment algorithm tends to converge within 4-10 iterations for each
-resolution of the multi-scale implementation.  We also place hard bounds on how
-large the alignment can be, and terminate alignment early if these bounds are
+resolution of the multi-scale implementation. This low number of iterations is expected, since the iterative alignment scheme is locally quadratically convergent. We also place hard bounds on how
+large the deformation can be, and terminate alignment early if these bounds are
 reached.  In the cases where the alignment ``slips off'' the face, this is how
@@
 transformation parameters.  We used the same set convergence thresholds for all
-of the data sets we ran against.  With the suggested thresholds, the l1-minimzations are
+of the data sets we ran against.  With the suggested thresholds, the l1-minimizations are
 guaranteed to converge eventually, although convergence may be very slow if they
@@
 
-\section{Respond to Reviewer \#3}
+\section{Response to Reviewer \#3}
 Reviewer 3 makes the following requests:
@@
 We have added experiments to the paper demonstrating the utility of the
-sci-based recognition stage for both identification and imposter rejection.  In
+sci-based recognition stage for both identification and impostor rejection.  In
 general, we see a few percentage points of improvement in recognition rate when
@@
 It is worth emphasizing that while this effect is desirable, 
-the primary value of SCI is its use for imposter rejection.
+the primary value of SCI is its use for impostor rejection.
 
@@
 higher resolution and lower noise levels than 3D cameras.  Based on the principles
-of photometric stereos, our set of 2D images may very well encode more useful shape 
+of photometric stereo, our set of 2D images may very well encode more useful shape 
 information about a user's face than a single low-resolution depth image, especially if 

### 2010-12-31 faces_system.tex
Changes from John.

@@
 
-\PARstart{T}{here} has been a historical tendency for face recognition algorithms to work well under laboratory conditions  (well aligned images, similar lighting conditions, and little change in appearances) but degrade drastically when tested in less-controlled real-world environments.
-Although achieving high recognition performance under uncontrolled conditions and with limited gallery images has been the traditional goal for face recognition, existing techniques have not yet been able to meet the performance requirements of high-stake security-related applications.
-While there have been many successes on less-demanding applications such as online image search and family photo album organization (e.g.
-the freely available Google's Picassa, Microsoft's Photo Gallery, and Apple's iPhoto), there have been many high profile failed trials of face recognition technology worldwide for extremely challenging mass surveillance/watch-list applications.
-
-Somewhat surprisingly, face recognition has not seen widespread applications or success in more tractable security settings, such as access control for secure facilities (e.g., prisons, office buildings), computer systems, automobiles, or even automatic teller machines,  where controlled gallery images can be obtained in advance.\footnote{Face recognition scenarios where both gallery and test data are uncontrolled continue to be an active research area \cite{LFW}, and becomes more popular recently due to aforementioned online applications.}  These applications are very interesting due to their potential sociological and technological impact.
+\PARstart{T}{here} has been a historical tendency for face recognition algorithms to work well under laboratory conditions  (well aligned images, similar lighting conditions, and little change in appearance) but degrade drastically when tested in less-controlled real-world environments.
+Although achieving high recognition performance under uncontrolled conditions and with limited gallery images has been the traditional goal of face recognition research, existing techniques are not yet able to meet the performance requirements of high-stakes security-related applications.
+While there have been many successes in less-demanding applications such as online image search and family photo album organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto), there have been a number of high-profile failed trials of face recognition technology in extremely challenging mass surveillance/watch-list applications.
+
+Somewhat surprisingly, face recognition has not seen widespread applications or success in more tractable security settings, such as access control for secure facilities (e.g., prisons, office buildings), computer systems, automobiles, or automatic teller machines, where controlled gallery images can be obtained in advance.
+%\footnote{Face recognition scenarios where both gallery and test data are uncontrolled continue to be an active research area \cite{LFW}, and has become more popular recently due to the aforementioned online applications.}
+These applications are very interesting due to their potential sociological and technological impact.
 They are also qualitatively different from mass surveillance or online image applications because they demand extremely reliable systems whose performance must compete with other biometric measures such as fingerprint and iris.
-Now because the gallery subjects are allies, rather than opponents, of the recognition system, this creates the possibility of carefully controlling the acquisition of the training data (as we normally do for all other biometrics).
-However, unlike other biometrics, face recognition has the potential in working with much less controlled testing data so that the access control system can be made more discreet and much less intrusive than using other biometrics.
-To some extent, the goal of this paper is to show how a reliable face recognition system can be built for this admittedly restrictive but important scenario.
-
-Classical holistic subspace based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and simplicity, as well as for their natural extension to linear illumination
+Because the gallery subjects are allies, rather than opponents, of the recognition system, this creates the possibility of carefully controlling the acquisition of the training data (as is common practice for other biometrics).
+However, unlike other biometrics, face recognition has the potential of working with test data that is much less controlled, allowing the access control system to be made more discreet and much less intrusive.
+To some extent, the goal of this paper is to show how a reliable face recognition system can be built for this admittedly restrictive, but still important, scenario.
+
+Classical holistic subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and simplicity, as well as for their natural extension to linear illumination
 models.  However, their performance has been shown to be extremely brittle not
@@
 using SRC.} \label{fig:promo}
-%\vspace{-.5in}
+\vspace{-.5in}
 \end{figure}
@@
 handling of illumination variation.  Our system is based purely on 2D
-techniques.  This fact immediately distinguishes our technique from systems
+techniques.  This fact immediately distinguishes our approach from systems
 that either require a 3D data as an input, or attempt to estimate a 3D model
-from input 2D data \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
-techniques have been shown to achieve better robustness to pose variation given
+from 2D input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
+techniques can achieve better robustness to pose variation given
 a sufficiently accurate 3D model, for access control applications where only 
@@
 problem of registering nearly identical images, say by minimizing a sum of
-squared distances or maximize normalized correlation, here we must confront several physical factors
+squared distances or maximizing normalized correlation, here we must confront several physical factors
 simultaneously: misalignment, illumination variations, and corrupted pixels.
@@
 appropriate objective function than the classical $\ell^2$-norm. Its use here
-is loosely motivated by theoretical results due to Candes and Tao
+is loosely motivated by theoretical results due to Cand\`{e}s and Tao
 \cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
@@
 recognition algorithm within the sparse representation
-framework. We elaborate on the implementation issues of the
-algorithm, conduct region of attraction experiments with
+framework. We elaborate on implementation issues with our algorithm, conduct region of attraction experiments with
 respect to both 2D in-plane deformation and 3D pose variation,
-and discuss its relationship to existing work in the
-literature. Section \ref{sec:illumination} is dedicated to our
+and discuss its relationship to existing work. Section \ref{sec:illumination} is dedicated to our
 training acquisition system.
@@
 linear combination $A \x_0$ of all of the images in the
-database,\footnote{We assume the illuminations in the training
-set are sufficient. We will address how to ensure illumination
+database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
 sufficiency in the next section.} plus a sparse error $\e_0$
@@
 
-\subsection{Batch and Individual Alignment} Notice that if the
+\subsection{Batch and Individual Alignment} If the
 true deformation $\tau^{-1}$ can be found, then we can apply
@@
 $\tau$: conceptually, one would like to seek a transformation
-$\tau$ that allows the sparsest representation, by solving
+$\tau$ that allows the sparsest representation, via
 \begin{equation} \label{eqn:L1-L1-conceptual}
@@
 \centering
+{\small
 \begin{tabular}{cccc}
@@
 \end{tabular}
+\vspace{-.1in}}
 \caption{{\bf Comparing alignment of a subject wearing sunglasses by
@@
 (c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-.3in}
+\vspace{-.4in}
 \end{figure}
@@
 our experiments, we typically use 2D similarity
-transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{For
-readers unfamiliar with this terminology, SE stands for Special
-Euclidean, a class of transformations representing image
-rotation and translation.  The $\Re_+$ accounts for the scaling
-variation that can be represented by a 2D similarity
-transformation.}, for removing alignment error incurred by face
+transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
+Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
 detector, or 2D projective transformations, $T =
@@
 \mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem:
-$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$$
+$\ell^1$-minimization problem: \hspace{2mm}
+$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
 \STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
@@
 
-So far, these transformations have been expressed in a very
-general form.  The transformation defines a mapping between the
+The transformation $\tau$ defines a mapping between the
 coordinates of pixels in the large original image and a smaller
-(un)warped image. The pixels of the small image get stacked into a vector.
-For a simple implementation, a rectangular window with
+(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
+artifacts in the downsampled image, one should apply a
+smoothing filter to the original image. For a simple implementation, a rectangular window with
 regular sampling can used, but in general, the small image need
@@
 different windows can affect the performance of our algorithm
-in Section \ref{sec:multipie}. Note that to prevent aliasing
-artifacts in the downsampled image, it is necessary to apply a
-smoothing filter to the original image.
+in Section \ref{sec:multipie}. 
 %The strength of the
@@
 \end{equation}
-As one can see, although updating $\blamda$ is trivial,
+Although updating $\blamda$ is trivial,
 minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
@@
 \cite{YangJ2009-pp}, called \emph{alternating direction
-method of multipliers} (ADM) \cite{Glowinski1975-TR}, which proposes to minimize $L_{\mu} (\x,\e,\blamda_k)$
-with respect to $\x$ and $\e$ separately. As the objective function is convex and alternation is between two
-terms, this procedure has been proven to converge to the global optima (see \cite{YangJ2009-pp} and references therein).
-
-Then the multiplier $\blamda$ is updated after just one iteration of minimization
-with respect to the two primal variables, giving:
+method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
+over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
 \begin{equation}
@@
 \end{equation}
-In order to discuss the solution to the above sub-problems, we
+As the objective function is convex and alternation is between two
+terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).
+
+In order to discuss the solution to the above subproblems, we
 need to define the following soft-thresholding operator for a
@@
 where all the operations are performed component-wise. It is
-easy to show that the sub-problem with respect to $\e$ has a
+easy to show that the subproblem with respect to $\e$ has a
 closed-form solution given by the soft-thresholding operator:
@@
     this will be the default resolution at which we prepare
-    all our training and testing datasets and run all our
-    experiments.} and the distance between the two outer
+    all our training and testing datasets.} and the distance between the two outer
     eye corners is normalized to be 50 pixels for each
@@
 \centering
+{\small
 \begin{tabular}{ccc}
@@
 \end{tabular}
+\vspace{-.2in}
+}
 \caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.
@@
 \label{fig:attraction}
-\vspace{-.3in}
+\vspace{-.4in}
 \end{figure}
@@
 %between scales,
-We conduct an experiment similar to the previous 2D
-deformation experiment. We use the same 7 illuminations in
+We use the same 7 illuminations in
 Session 1 as training, and all 20 illuminations in the same
@@
 \includegraphics[height=1.8in]{figures_pami/multiscale.png}
+\vspace{-.3in}
 \caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
@@
 \label{fig:multiscale}
-\vspace{-.2in}
+\vspace{-.5in}
 \end{figure}
@@
 \centering
+{\small
 \begin{tabular}{ccccc}
@@
 \includegraphics[height=1in]{figures_cvpr/11} &
-\includegraphics[height=1in]{figures_cvpr/13} \\ \vspace{-.1in}
-(a) & (b) & (c) & (d) & (e)\vspace{3mm}\\
+\includegraphics[height=1in]{figures_cvpr/13} \vspace{-.1in} \\ 
+(a) & (b) & (c) & (d) & (e)\\
 \includegraphics[height=1in]{figures_cvpr/15} &
@@
 \includegraphics[height=1in]{figures_cvpr/21} &
-\includegraphics[height=1in]{figures_cvpr/3} \\ \vspace{-.1in}
-(f) & (g) & (h) & (i) & (j)
+\includegraphics[height=1in]{figures_cvpr/3} \vspace{-.1in} \\ 
+(f) & (g) & (h) & (i) & (j) \vspace{-.1in}
 \end{tabular}
+}
 \caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
 to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-.3in}
+}\label{fig:pose-alignment} \vspace{-.2in}
 \end{figure}
@@
 
-Fortunately, even with these effects, the relationship between the space of
-illuminations and the space of images is still linear
-\footnote{Materials that break
+Fortunately, even with these effects, for most materials the relationship between
+illumination and image is still linear,\footnote{Materials that break
 this assumption include fluorescent materials and the photochromic (``Transition'') lenses
-in some eyeglasses.  Most materials emmit light in proportion to their
-incident light.}, 
-and that the sensor has a linear response curve. 
-\footnote{ Proper handling of gamma encoding is an important consideration for
+in some eyeglasses.  Most materials emit light in proportion to their
+incident light.} provided the sensor has a linear response curve.\footnote{Proper handling of gamma encoding is an important consideration for
 practitioners.  Most cameras apply a non-linear and often undocumented response
@@
 image file is loaded.}
-For a more indepth study 
+For a more in-depth study 
 of the relationship between illumination and images, we refer the reader to
@@
 
-\subsection{Capturing a Sufficient Set of Training Illuminations} To this end,
-we have designed a system that can acquire frontal images of a subject while
-simultaneously illuminating the subject from all directions above horizontal.
-
-Most light stages used for face recognition have
-been constructed for the purpose of creating public data sets to study
-illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
-light stages have been used for computer graphics purposes
-\cite{debevec2000acquiring, jones2005performance}.
-The light source can be
-moved around manually \cite{masselus2002free}, but this may result in poor
-consistency of illuminations between users.  Structured light applications use projectors to
-directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
-reconstruction, but this is very disturbing to the user. 
-Y. Schechner \cite{schechner2007multiplexing}
-studies techniques for multiplexing illumination that can dramatically reduce the noise 
-of the demultiplexed images for certain classes of objects and cameras.  
-While these techniques have not been incorporated into the current
-system, they fit elegantly into our framework and will likely be used
-in future implementations.  We stress that use of this multiplexing technique
-is independent from the choice of original (directional) illuminations.
-
-A sketch of the
+\subsection{Projector-based Illumination System}
+
+We have designed a system that can acquire frontal images of a subject while
+simultaneously illuminating the subject from all directions above horizontal. A sketch of the
 system is shown in Figure \ref{fig:system}: The illumination
@@
 \begin{figure}
-\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}}
+\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}\vspace{-.2in}}
 \caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
-\vspace{-.5in}
+\vspace{-.3in}
 \label{fig:system}
@@
 partly on how the test images are taken and is still under study.
+
+{\em Relationship to existing work:} Most light stages used for face recognition have
+been constructed for the purpose of creating public data sets to study
+illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
+light stages have been used for computer graphics purposes
+\cite{debevec2000acquiring, jones2005performance}.
+The light source can be
+moved around manually \cite{masselus2002free}, but this may result in poor
+consistency of illuminations between users.  Structured light applications use projectors to
+directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
+reconstruction, but this is very disturbing to the user. 
+Y.\ Schechner \cite{schechner2007multiplexing}
+studies techniques for multiplexing illumination that can dramatically reduce the noise 
+of the demultiplexed images for certain classes of objects and cameras.  
+While these techniques have not been incorporated into the current
+system, they fit elegantly into our framework and will likely be used
+in future implementations.  We stress that use of this multiplexing technique
+is independent from the choice of original (directional) illuminations.
+
+\subsection{Choice of Illumination Patterns}
 
@@
 (a) Coverage Experiment & (b) Chosen Illumination Patterns 
-\end{tabular}\vspace{2mm}
+\end{tabular}\vspace{-2mm}
 \caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
@@
 \label{fig:illumination-patterns}\label{fig:illumination-sufficiency}
-%\vspace{-.5in}
+\vspace{-.2in}
 \end{figure}
@@
 \label{fig:illumination-sufficiency}
-%\vspace{-.5in}
+\vspace{-.5in}
 \end{figure}
 
-\subsection{Chosen Illumination Patterns} In the plot for the
+ In the plot for the
 coverage experiment, Figure \ref{fig:illumination-sufficiency}(a),
@@
 in Session 1 as the training set. The remaining 88 subjects are
-treated as ``imposters'', or invalid images. For each of the
+treated as ``impostors'', or invalid images. For each of the
 249 training subjects, we include frontal images of 7 frontal
@@
 To demonstrate that the SCI-based recognition step is indeed beneficial
-even when there are no imposters, we include results for recognition
+even when there are no impostors, we include results for recognition
 based only on the alignment error residuals, shown in row 6.
@@
 \caption{Recognition rates on large-scale Multi-PIE database.}
+\vspace{-.2in}
 \centerline{
@@
 \centering
+{\small
 \begin{tabular}{cccccc}
@@
 \includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-(a) & (b) & (c) & (d) & (e) & (f)
+\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{-.1in} \\
+(a) & (b) & (c) & (d) & (e) & (f) \vspace{-.1in}
 \end{tabular}
-\caption{{\bf Representative examples of failed Multi-PIE subjects}.
+}
+\caption{{\bf Representative failures from Multi-PIE}.
 {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2.
-Due to change of hair style, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
+Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
 \label{fig:failed-examples}
-%\vspace{-.8in}
+\vspace{-.5in}
 \end{figure}
@@
 \centering
+{\small
 \begin{tabular}{cc}
 \includegraphics[height=1.5in]{figures_pami/example.png} &
-\includegraphics[height=1.5in]{figures_pami/example_new.png}\\
-(a) & (b)
+\includegraphics[height=1.5in]{figures_pami/example_new.png} \vspace{-.1in} \\
+(a) & (b) \vspace{-.1in}
 \end{tabular}
+}
 \caption{{\bf Choosing different sampling windows.} (a) The default window. (b) The newly proposed window.}
 \label{fig:new-mask}
-%\vspace{-.6in}
+\vspace{-.3in}
 \end{figure}
 
-It is shown in Table \ref{tab:MultiPIE-recognition} that the
+Table \ref{tab:MultiPIE-recognition} shows that the
 recognition rates on Multi-PIE indeed increase with this new
@@
 recognition rates of our algorithm are significantly boosted
-with the new window.
-
-However, we should mention that the best choice of the window
-is very problem-specific and there is
+with the new window. However, we should mention that the best choice of the window
+is problem-specific and there is
 not a simple guideline to follow. For example, although the new
-window performs better on Multi-PIE, the same window doesn't
+window performs better on Multi-PIE, the same window does not
 help at all on our own database, which will be introduced in
@@
 \includegraphics[width=3in]{figures_pami/pami_roc2.png}
-}
+}\vspace{-.2in}
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
-%\vspace{-.6in}
+\vspace{-.35in}
 \end{figure}
@@
 
-We also compare our result to the most recent work
+We also compare our result to the very recent work
 \cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
@@
 \cite{Ahonen2006-PAMI}, a local appearance descriptor which is
-able to capture final details of facial appearance and texture.
+able to capture fine details of facial appearance and texture.
 Due to its robustness to variations in illumination, facial
@@
 \cite{Tan06facerecognition}. In this paper, we follow the same
-steps as in \cite{Ahonen2006-PAMI} to construct a LBP
+steps as in \cite{Ahonen2006-PAMI} to construct an LBP
 descriptor for each training and testing sample. The $80\times
@@
 \caption{Recognition rates on single gallery image FERET
-dataset} \centerline{
+dataset} \vspace{-.1in} \centerline{
 \begin{tabular}{|c|c|c|c|c| }
@@
 \end{tabular}
-\label{tab:FERET-recognition} }
+\label{tab:FERET-recognition} } \vspace{-.4in}
 \end{table}
 
-As expected, our algorithm doesn't perform well except for
-`fb', in which case the illumination is similar to the training and 
+As expected, our algorithm does not perform well except for
+`fb', in which the illumination is similar to the training and 
 the mere variation in facial expression is handled well by the sparse error 
@@
 example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.
+recognition rate.\vspace{-.1in}
 
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
+\vspace{-.5in}
 \caption{{\bf Recognition under varying level of
-random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%.
-Our method maintains high recognition rates up to 30\% occlusion (see table below).
+random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:} \vspace{1mm}
+{\small
+\begin{tabular}{|c|c|c|c|c|c| }
+\hline
+Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
+\hline
+\hline
+Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
+\hline
+\end{tabular}
 }
-\begin{tabular}{|c|c|c|c|c|c| }
-\hline
-Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
-\hline
-\hline
-Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
-\hline
-\end{tabular}
 \label{fig:multipie-occ-rec}
@@
 to the 7 illuminations used in the previous experiment on
-Multi-PIE.\footnote{We use the illumination set $\{6, 9, 12,
+Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
 13, 18, 21, 22\}$ shown in Figure
-\ref{fig:illumination-patterns}(b) to mimic the illumination
-set $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
+\ref{fig:illumination-patterns}(b) to mimic the illuminations
+ $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
 our algorithm on the remaining $24 - 7 = 17$ frontal
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} 
-%\vspace{-.25in}
+\vspace{-.2in}
  \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
+ \vspace{-.1in}
 \end{figure}
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
-%\vspace{-.25in}
+\vspace{-.2in}
  \caption{{\bf Representative examples of category C4}.
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.15in}
+\vspace{-.5in}
 \end{figure}
@@
 \begin{table}
+\vspace{-.1in}	
 \centering \caption{Recognition rates on our own database.}
+\vspace{-.2in}
 \begin{tabular}{|c|c|c|c|c| }
@@
 \label{tab:UIUC-recognition}
-\vspace{-.25in}
+\vspace{-.4in}
 \end{table}
@@
 \label{fig:occ-block}
-\vspace{-.2in}
+\vspace{-.5in}
 \end{figure}
@@
 that the performance under occlusion can be boosted by
-leveraging local information of a face as well as global information. Meanwhile, we will
-leave the further investigation of more general models (such as Markov
-random fields \cite{ZhouZ2009}) for face
+leveraging local information of a face as well as global information. We 
+leave the investigation of more general models (e.g., MRF \cite{ZhouZ2009}) for face
 recognition with both misalignment and occlusion as an
-interesting future work.\vspace{-0.2in}
+interesting future work.
 
@@
 investigation is to extend the alignment algorithm to better
-tackle contiguous occlusion. We have demonstrated in this
-work that misalignment can be naturally handled within the
+tackle contiguous occlusion. We have demonstrated that misalignment can be naturally handled within the
 sparse representation framework. More complicated models for
@@
 for face alignment and recognition in the presence of
-contiguous occlusion remains an open problem. \vspace{-0.2in}
+contiguous occlusion remains an open problem.
 
@@
 This work was supported by grants NSF IIS 08-49292, NSF ECCS
-07-01676, and ONR N00014-09-1-0230. John Wright would like to
-thank Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
-Mathematics for discussions related to this work. He would like
-to acknowledge support from a Microsoft Research Fellowship and
-the Lemelson-Illinois Student Prize.
+07-01676, and ONR N00014-09-1-0230. JW thanks Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
+Mathematics for discussions related to this work, and 
+acknowledges support from a Microsoft Fellowship and
+the Lemelson-Illinois Student Prize. %} %\vspace{-0.2in}
 

### 2010-12-31 rebuttal.tex
Changes from Zihan

@@
 \newtheorem{definition}{Definition}
+
+\newcommand{\e}{\mathbf{e}}
+\newcommand{\x}{\mathbf{x}}
 
@@
 
-\paragraph{About Positioning and Novelty:}
-As we have tried to make clear, the proposed method does not aim to solve face 
-recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, and hence it
-may not have any advantages over other face recognition techniques/systems specifically designed for such scenarios. 
-However, there are many practical scenarios where the proposed setting is very relevant: access control
-or recognizing fugitives, where cooperative training images can (and should) be obtained in advance but no cooperative test
-images are needed or available. The possibility of obtaining reliable identification without cooperative test images is arguably the main (or only) advantage of 
-face recognition over other biometric measures such as fingerprint, iris, and DNA. 
-If the subjects can cooperate for both training and testing, as suggested by the reviewers, then face 
-recognition may no longer be the best choice, in terms of cost and accuracy. The proposed system is an attempt
-to build a system that is not only accurate and robust enough to compete with other biometric measures but also 
-discreet and non-intrusive when in use. To the best of our knowledge (probably the reviewer may agree), even 
-for this admittedly restricted scenario, there is no practical face recognition system whose performance can meet
-the standards of most access control and security applications. Existing systems have found success  mostly in applications
-that are less demanding in performance, such as online image search or photo album organization.
+\paragraph{About Positioning and Novelty:} As we have tried to make clear, the
+proposed method does not aim to solve face recognition for all scenarios! It is
+not designed for the case where the training is uncontrolled and limited, and
+hence it may not have any advantages over other face recognition
+techniques/systems specifically designed for such scenarios.  The proposed
+system is an attempt to build a system that is not only accurate and robust
+enough to compete with other biometric measures but also discreet and
+non-intrusive when in use. There are many practical scenarios where the
+proposed setting is very relevant, such as access control, where well
+controlled training images can (and should) be obtained in advance but no well
+controlled test images are needed or available. Particularly, the system should
+be able to handle typical daily changes in facial appearance, such as occlusion
+by glasses, hats, or facial hair and variations in lighting and moderate
+expression. The possibility of achieving fully automated (at recognition time),
+touch-free identification is arguably the main advantage of face recognition
+over other biometric measures such as fingerprint and iris recognition.  To the
+best of our knowledge even for this admittedly restricted scenario, there is no
+practical face recognition system whose performance can meet the standards of
+most access control and security applications. Existing systems have found
+success  mostly in applications that are less demanding in performance, such as
+online image search or photo album organization.
 
@@
 
-\paragraph{Detailed Concerns or Comments by Reviewer 1:}
-\begin{enumerate}
-\item As mentioned above, we agree that if the users are completely cooperative at
-both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
-and iris are arguably better choices.
-However, our method does NOT require the users to be completely cooperative at the test stage.
-The purpose of having a good training set and a robust recognition system is precisely to alleviate this requirement.
-This will allow the system to be less intrusive and much more user-friendly,
-as it can handle typical daily changes in facial appearance, such as occlusion by glasses, hats, or facial hair and variations in lighting.
-Ideally, face recognition will take place automatically and unknowingly as a user walks through a door or works in front a desktop.
+\paragraph{Detailed Concerns or Comments by Reviewer 1:} \begin{enumerate}
+\item Reviewer 1 is correct that use of our system is largely restricted to
+users who are cooperative (in the sense of non-adversarial). Again, the main
+point of our paper is that we can make identification more reliable and
+convenient, at the relatively small cost of doing a better job of gathering
+data at enrollment time.  Ideally, face recognition will take place
+automatically as a user walks through a door or works in front a desktop.
 
-In addition, if the accuracy and reliability of a face recognition system can truly get close to that of fingerprint and iris, it can serve as a complementary measure in a multi-modal system and improve the overall security for high-stakes applications.
-To the best of our knowledge, no existing face recognition system, with or without cooperative training, comes close to meet the standards of high-stake security applications.
+In addition, if the accuracy and reliability of a face recognition system can
+truly get close to that of fingerprint and iris, it can serve as a
+complementary measure in a multi-modal system and improve the overall security
+for high-stakes applications. To the best of our knowledge, no existing face
+recognition system, with or without cooperative training, comes close to meet
+the standards of high-stake security applications.
 
-Furthermore, we have found that the use of the L1-norm as a robust error function improves performance 
-even when the test image does not contain occlusions.  Any boundaries where the intensity changes in a
-misaligned image can result in a large, localized error component.  While
+In response to the reviewer's question "is occlusion really a problem?", we have found that the use of
+the L1-norm as a robust error function improves performance even when the test
+image does not contain occlusions.  Any boundaries where the intensity changes
+in a misaligned image can result in a large, localized error component.  While
 unoccluded faces are mostly smooth, sharp boundaries still occur where there
@@
 
-\item Since we perform alignment separately for each class in the gallery, the
-computation for the alignment stage is linear in the size of the gallery.  The
-variation in alignment across classes is indeed small, but still changes the
-objective enough to affect recognition results.  Aligning to a generic face (or
-a gallery based on ``aligned'' images of many people) is a good idea, and one we
-have considered as well.  If nothing else, it could be used to improve
-alignment initialization instead of using raw face detector output.  It may
-also be useful to use the roughly aligned test image to quickly eliminate some
-of the classes from further consideration.  Again, we have deliberately tried
-to keep the algorithm as minimalist as possible for this publication.
+\item  {\bf There is no extensive computation involved in
+    Step 13 of Algorithm 1. In fact, for each candidate
+    subject $i\in {k_1, \ldots, k_s}$ for recognition, Step
+    13 simply uses the transformation $\tau_i$ obtained
+    from the iterative alignment stage (Step 2-9) to
+    re-sample each original training
+    image of that subject, so that this new set of training
+    samples is aligned with the test sample. Nevertheless,
+    it worth mentioning that, since we perform alignment
+    separately for each class in the gallery, the
+    computation for the alignment stage (Step 2-9) is
+    linear in the size of the gallery. However, thanks to
+    its separability, the alignment stage can be easily
+    parallelized.}
+
+    The variation in alignment across classes is indeed
+    small, but still changes the objective enough to affect
+    recognition results. Aligning to a generic face (or a
+    gallery based on ``aligned'' images of many people) is
+    a good idea, and one we have considered as well.  If
+    nothing else, it could be used to improve alignment
+    initialization instead of using raw face detector
+    output.  It may also be useful to use the roughly
+    aligned test image to quickly eliminate some of the
+    classes from further consideration.  Again, we have
+    deliberately tried to keep the algorithm as minimalistic
+    as possible for this publication.
 
@@
 
-For the l1 minimizations we threshold the l1-norm of the difference between
-successive values of the state (this includes the both the explicit
-optimization variables and the Largrange multipliers).  Similarly, for the
-alignment iteration, we threshold the l1-norm of the update to the
-transformation parameters.  We used the same set convergence thresholds for all
-of the data sets we ran against.  With the suggested thresholds, the l1-minimizations are
-guaranteed to converge eventually, although convergence may be very slow if they
-are chosen poorly.
+{\bf For the $\ell_1$ minimizations we terminate the ALM
+algorithm when the relative changes of primal variables in
+two consecutive iterations become small, as suggested by
+[27]. That is,
+\begin{displaymath}
+\frac{\|\x_{k+1}-\x_k\|_2}{\|\x_k\|_2} < \epsilon \quad
+\textup{and} \quad \frac{\|\e_{k+1}-\e_k\|_2}{\|\e_k\|_2} <
+\epsilon,
+\end{displaymath}
+where $\epsilon>0$ is the tolerance. With the suggested
+criteria, the l1-minimizations are guaranteed to converge
+eventually, although convergence may be very slow if they
+are chosen poorly. For the alignment
+iteration, we terminate when the change of objective value
+(\ie $\|\e\|_1$) in two consecutive iterations becomes
+small.} We use the same set of convergence thresholds for
+all the data sets we ran against.
 
@@
 
-We have added experiments to the paper demonstrating the utility of the
-sci-based recognition stage for both identification and impostor rejection.  In
-general, we see a few percentage points of improvement in recognition rate when
-using SCI over recognition based purely on alignment residual.  
-It is worth emphasizing that while this effect is desirable, 
-the primary value of SCI is its use for impostor rejection.
+{\bf We have added experiments to the paper demonstrating
+the utility of the sparse representation(SR)-based
+recognition stage for classification.  In general, we see a
+few percentage points of improvement in recognition rate
+when using the representation residual obtained from the
+SR-based recognition stage (with $S=10$) over the
+recognition based purely on alignment error (with $S$=1).
+While this effect is desirable, as discussed in the paper,
+it is worthy emphasizing that the SR-based recognition also
+enables us to compute SCI, which has demonstrated its power
+for impostor rejection.}
 
@@
 \end{document}
+ a user's face than a single low-resolution depth image, especially if
+recognition is to be performed on 2D input images.
+
+In addition, the sparse representation techniques are not limited to frontal 2D images.
+They can be

### 2010-12-31 faces_system.tex
Changes from Zihan

@@
 a sufficiently accurate 3D model, for access control applications where only 
-moderate misalignment and pose are present, the proposed method will be more
+moderate pose variation is present, the proposed method will be more
 than sufficient.   Note that according to the principle of photometric stereo, images of faces
@@
 above equation is underdetermined if we allow the registration
-error $\e$ to be arbitrary. Near the correct alignment we
-expect the aligned testing image to differ from $A_i \x$ only
+error $\e$ to be arbitrary. At the correct alignment we
+expect the test image to differ from $A_i \x$ only
 for the minority of the pixels corrupted by occlusions. Thus,
@@
 face detector; there was no manual intervention for the test images.
-To demonstrate that the SCI-based recognition step is indeed beneficial
+To demonstrate that the sparse representation based recognition step is indeed beneficial
 even when there are no impostors, we include results for recognition
-based only on the alignment error residuals, shown in row 6.
+based only on the alignment error residuals(i.e. S=1), shown in row 6.
 
@@
 \hline
-{Algorithm 1 w/o SCI} & 90.7\% & 89.6\% & 87.5\% \\
+{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
 \hline
@@
 
-\subsection{Subject Validation}
-
-We test the algorithms' ability to reject invalid images of the
-88 subjects not appearing in the training database. As
-mentioned before, the \emph{sparsity concentration index} (SCI)
-is used as the outlier rejection rule. Given the sparse
-representation $\x$ of a test image with respect to $K$
-training classes, the SCI measures how concentrated the
-coefficients are on a single class in the dataset and is
-defined as in \cite{Wright2009-PAMI}:
-\begin{displaymath}
-\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
-\|\x\|_1 - 1}{K - 1} \in [0,1] .
-\end{displaymath}
-It is easy to see that if $\textup{SCI}(\x) = 1$, the test
-image is represented using images from one single subject
-class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
-evenly over all classes. Thus, we can choose a threshold $t \in
-[0,1]$ for the proposed method and accept a test image as valid
-if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
-invalid. We compare this classifier to classifiers based on
-thresholding the error residuals of NN, NS and LDA.
-
-Figure \ref{fig:roc-multipie} plots the receiver operating
-characteristic (ROC) curves, which are generated by sweeping
-the threshold $t$ through the entire range of possible values
-for each algorithm.\footnote{Rejecting invalid images not in
-the entire database is much more difficult than deciding if two
-face images are the same subject. Figure 8 should not be
-confused with typical ROC curves for face similarity, e.g.,
-\cite{PhillipsP2007}.} We can see that SCI based approach
-significantly outperforms the other three algorithms. Similar
-contrasts between our algorithm and baseline algorithms were
-also observed for SRC in \cite{Wright2009-PAMI}, though on much
-smaller datasets.
-
-\begin{figure}
-\centerline{
-\includegraphics[width=3in]{figures_pami/pami_roc2.png}
-}\vspace{-.2in}
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
-\vspace{-.35in}
-\end{figure}
-
 \subsection{Comparison to Existing Work}
@@
 no obvious occlusion or corruption in the test.
+
+\subsection{Subject Validation}
+
+We test the algorithms' ability to reject invalid images of the
+88 subjects not appearing in the training database. As
+mentioned before, the \emph{sparsity concentration index} (SCI)
+is used as the outlier rejection rule. Given the sparse
+representation $\x$ of a test image with respect to $K$
+training classes, the SCI measures how concentrated the
+coefficients are on a single class in the dataset and is
+defined as in \cite{Wright2009-PAMI}:
+\begin{displaymath}
+\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
+\|\x\|_1 - 1}{K - 1} \in [0,1] .
+\end{displaymath}
+It is easy to see that if $\textup{SCI}(\x) = 1$, the test
+image is represented using images from one single subject
+class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
+evenly over all classes. Thus, we can choose a threshold $t \in
+[0,1]$ for the proposed method and accept a test image as valid
+if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
+invalid. We compare this classifier to classifiers based on
+thresholding the error residuals of NN, NS, LDA, and LBP.
+
+Figure \ref{fig:roc-multipie} plots the receiver operating
+characteristic (ROC) curves, which are generated by sweeping
+the threshold $t$ through the entire range of possible values
+for each algorithm.\footnote{Rejecting invalid images not in
+the entire database is much more difficult than deciding if two
+face images are the same subject. Figure 8 should not be
+confused with typical ROC curves for face similarity, e.g.,
+\cite{PhillipsP2007}.} We can see that SCI based approach
+significantly outperforms the other three algorithms. Similar
+contrasts between our algorithm and baseline algorithms were
+also observed for SRC in \cite{Wright2009-PAMI}, though on much
+smaller datasets.
+
+\begin{figure}
+\centerline{
+\includegraphics[width=3in]{figures_pami/pami_roc2.png}
+}\vspace{-.2in}
+\caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
+\vspace{-.35in}
+\end{figure}
 

### 2010-12-31 rebuttal.tex
Removed bold stuff from Zihan's version

@@
 
-\item  {\bf There is no extensive computation involved in
+\item  There is no extensive computation involved in
     Step 13 of Algorithm 1. In fact, for each candidate
@@
     its separability, the alignment stage can be easily
-    parallelized.}
+    parallelized.
 
@@
 
-{\bf For the $\ell_1$ minimizations we terminate the ALM
+For the $\ell_1$ minimizations we terminate the ALM
 algorithm when the relative changes of primal variables in
@@
 (\ie $\|\e\|_1$) in two consecutive iterations becomes
-small.} We use the same set of convergence thresholds for
+small. We use the same set of convergence thresholds for
 all the data sets we ran against.
@@
 
-{\bf We have added experiments to the paper demonstrating
+We have added experiments to the paper demonstrating
 the utility of the sparse representation(SR)-based
@@
 enables us to compute SCI, which has demonstrated its power
-for impostor rejection.}
+for impostor rejection.
 

### 2011-01-04 faces_system.tex
Worked on introduction

@@
 
-\PARstart{T}{here} has been a historical tendency for face recognition algorithms to work well under laboratory conditions  (well aligned images, similar lighting conditions, and little change in appearance) but degrade drastically when tested in less-controlled real-world environments.
-Although achieving high recognition performance under uncontrolled conditions and with limited gallery images has been the traditional goal of face recognition research, existing techniques are not yet able to meet the performance requirements of high-stakes security-related applications.
-While there have been many successes in less-demanding applications such as online image search and family photo album organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto), there have been a number of high-profile failed trials of face recognition technology in extremely challenging mass surveillance/watch-list applications.
-
-Somewhat surprisingly, face recognition has not seen widespread applications or success in more tractable security settings, such as access control for secure facilities (e.g., prisons, office buildings), computer systems, automobiles, or automatic teller machines, where controlled gallery images can be obtained in advance.
-%\footnote{Face recognition scenarios where both gallery and test data are uncontrolled continue to be an active research area \cite{LFW}, and has become more popular recently due to the aforementioned online applications.}
-These applications are very interesting due to their potential sociological and technological impact.
-They are also qualitatively different from mass surveillance or online image applications because they demand extremely reliable systems whose performance must compete with other biometric measures such as fingerprint and iris.
-Because the gallery subjects are allies, rather than opponents, of the recognition system, this creates the possibility of carefully controlling the acquisition of the training data (as is common practice for other biometrics).
-However, unlike other biometrics, face recognition has the potential of working with test data that is much less controlled, allowing the access control system to be made more discreet and much less intrusive.
-To some extent, the goal of this paper is to show how a reliable face recognition system can be built for this admittedly restrictive, but still important, scenario.
-
-Classical holistic subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and simplicity, as well as for their natural extension to linear illumination
-models.  However, their performance has been shown to be extremely brittle not
-only to alignment variation, but to even minor occlusions caused by, say, a
-wisp of hair, a blinked eye, or mouth that is slightly open. One promising
-recent direction, set forth in \cite{Wright2009-PAMI}, casts the recognition
-problem as one of finding a sparse representation of the test image in terms of
-the training set as a whole, up to some sparse error due to occlusion. A
-\emph{sparse representation-based classification} (SRC) method is then proposed
-for recognition. The main idea is that the sparse nonzero coefficients should
+\PARstart{F}{ace} recognition applications to date have fallen into roughly two
+categories.  Face recognition has recently seen a lot of success in a family of
+less-demanding applications such as online image search and family photo album
+organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto).
+At the other end of the tractability spectrum there are the terrorist watchlist
+and mass surveillance applications that have for the most part dominated the
+field of face recognition research.  However, there are many face recognition
+that fall roughly between these extremes, where very high recognition
+performance is desired, but the users in the gallery are still allies of the
+system rather than adversaries.  These applications include access control for
+secure facilities (e.g., prisons, office buildings), computer systems,
+automobiles, or automatic teller machines, where controlled gallery images can
+be obtained in advance.  These applications are very interesting due to their
+potential sociological impact.  Since the gallery subjects are allies, rather
+than opponents, of the recognition system, this creates the possibility of
+carefully controlling the acquisition of the training data. While the same can
+be said for other biometrics such as fingerprints and iris recognition, face
+recognition has the potential of working with test data that is much less
+controlled, allowing the access control system to be made less intrusive to the
+users of the system. To some extent, the goal of this paper is to show how a
+reliable face recognition system can be built for this admittedly restricted,
+but still important, scenario.
+
+Very few recognition systems specifically target applications where many well-
+controlled training images are available.  Of these, the classical holistic
+subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
+are well known for their speed and simplicity, as well as for their natural
+extension to linear illumination models.  However, their performance has been
+shown to be extremely brittle not only to alignment variation, but to even
+minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
+is slightly open. One promising recent direction, set forth in
+\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
+sparse representation of the test image in terms of the training set as a
+whole, up to some sparse error due to occlusion. A \emph{sparse
+representation-based classification} (SRC) method is then proposed for
+recognition. The main idea is that the sparse nonzero coefficients should
 concentrate on the training samples with the same class label as the test
@@
 occlusion or corruption by solving a simple convex program.  
-%For this reason, the final recognition stage of our proposed system will consist of SRC,
-%and the alignment stage that precedes it will be based on similar techniques.
+%For this reason, the final recognition stage of our proposed system will
+%consist of SRC, and the alignment stage that precedes it will be based on
+%similar techniques.
 

### 2011-01-07 faces_system.tex
Some language fixes from Drew.

@@
 users of the system. To some extent, the goal of this paper is to show how a
-reliable face recognition system can be built for this admittedly restricted,
+reliable face recognition system can be built for this restricted,
 but still important, scenario.
@@
 even if sufficient illuminations are present in the training, as shown in Figure
-\ref{fig:promo}(top). In addition, in order to span
+\ref{fig:promo}(top). Additionally, in order to span
 the illuminations of a typical indoor (or outdoor) environment,
-illuminations from behind the subject are also needed in the
-training. Otherwise, even for perfectly aligned test images,
+illuminations from behind the subject are needed in the
+training set. Otherwise, even for perfectly aligned test images,
 the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
@@
 In this paper, we examine how to handle alignment and illumination simultaneously
-in the sparse representation framework, hence bring the method proposed in \cite{Wright2009-PAMI}
+in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
 closer to practical use.
@@
 moderate pose variation is present, the proposed method will be more
-than sufficient.   Note that according to the principle of photometric stereo, images of faces
+than sufficient.   Note that 2D images of faces
 under varying illuminations already contain 3D shape-related information,\footnote{In principle,
@@
 and this information can be leveraged by 2D algorithms for alignment and recognition 
-even if shape is not reconstructed explicitly. In addition, the proposed sparsity-based robust
-alignment and recognition techniques are not limited to 2D data and can be easily extended
-to 3D data if necessary.
+even if shape is not reconstructed explicitly. 
+%In addition, the proposed sparsity-based robust
+%alignment and recognition techniques are not limited to 2D data and can be easily extended
+%to 3D data if necessary.
 
@@
 or only have moderate expression. We prefer to focus on deformations with far fewer 
-degrees of freedom (say an affine transform), and to use the training images themselves 
+degrees of freedom, i.e. similarity transformations, and to use the training images themselves 
 as the appearance model. Iterative registration in this spirit dates at least back to the 
@@
 
-However, whereas much of the early work on image registration is aimed at the
+Whereas much of the early work on image registration is aimed at the
 problem of registering nearly identical images, say by minimizing a sum of
@@
 
-\subsection{Contributions} In this paper, we show how the 
-mutually compounding issues of registration and illumination
-can be naturally addressed within a robust sparse representation
+\subsection{Contributions} In this paper, we show how 
+registration and illumination
+can be simultaneously addressed within a robust sparse representation
 framework. We show that face registration, a challenging
@@
 sufficient set of training illuminations
-that is capable of interpolating typical indoor and outdoor
+that is capable of linearly representing typical indoor and outdoor
 lighting, along with a practical hardware system for capturing
@@
 
-Finally, we demonstrate the effectiveness of the proposed new
+We then demonstrate the effectiveness of the proposed new
 methods with a complete face recognition system that is {\em
@@
 robust automatic recognition of subjects from loosely
-controlled input images taken both indoors and outdoors, using
-labeled frontal views of the subjects' faces under the proposed
-illuminations for training and an off-the-shelf face
+controlled probe images taken both indoors and outdoors, 
+using a gallery of 
+frontal views of the subjects' faces under the proposed
+illuminations. An off-the-shelf face
 detector\footnote{We use the OpenCV
 implementation of the Viola and Jones' face detector
-\cite{Viola2004-IJCV}.} to detect faces in images.
+\cite{Viola2004-IJCV}.} is used to detect faces in the test images.
 
@@
 
-\noindent{\em Organization of this paper:} In Section
-\ref{sec:registration}, we derive our robust registration and
-recognition algorithm within the sparse representation
-framework. We elaborate on implementation issues with our algorithm, conduct region of attraction experiments with
-respect to both 2D in-plane deformation and 3D pose variation,
-and discuss its relationship to existing work. Section \ref{sec:illumination} is dedicated to our
-training acquisition system.
-Using this system, we investigate empirically the set of training illuminations required to handle practical illumination variations, and suggest a sufficient set of 38 training illuminations.
-Extensive experiments on
-a large-scale public database and on our own database are conducted
-in Section \ref{sec:multipie} and Section \ref{sec:own-data},
-respectively, to verify the proposed system. Section
-\ref{sec:conclusion} concludes our work with discussion of
-promising future directions.
+\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
+we derive our robust registration and recognition algorithm within the sparse
+representation framework. We elaborate on algorithmic implementation issues,
+conduct region of attraction experiments with respect to both 2D in-plane
+deformation and 3D pose variation, and discuss its relationship to existing
+work. Section \ref{sec:illumination} is dedicated to our training acquisition
+system. Using this system, we investigate empirically how many training
+illuminations are required to handle practical illumination variations, and
+suggest a sufficient set of 38 training illuminations. Extensive experiments on
+a large-scale public database and on our own database are conducted in Section
+\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
+proposed system. Section \ref{sec:conclusion} concludes our work with
+discussion of promising future directions.
 
@@
 
-Meanwhile, although our algorithm is not designed for
+Although our algorithm is not designed for
 recognition when there is only a single gallery image per user,
-and will not achieve competitive performance, we compare its
-performance with LBP within such setting for the purpose of
-better evaluation of the strength and limitation of both
-algorithms. For this experiment, we use the FERET dataset
+we compare its
+performance with LBP within this setting for completeness.
+For this experiment, we use the FERET dataset
 \cite{phillips1998feret}, which contains five standard
@@
 
-All these experimental results re-emphasize the importance that
+All of these experimental results confirm that
 both illumination and alignment need to be simultaneously handled
-well in order to obtain accurate face recognition, even when there is
+well in order to achieve accurate face recognition, even when there is
 no obvious occlusion or corruption in the test.
@@
 (iterative image alignment, l1-error function, SRC, using projectors for
-illumination), one can build a system for recognizing human faces
+illumination), we have proposed a system for recognizing human faces
 from images taken under practical conditions that is conceptually simple, well
@@
 
-The system achieves extremely {\em stable} performance under
+The system achieves extremely stable performance under
 a wide range of variations in illumination, misalignment, and even under small amounts of
@@
 large-scale tests with public datasets as well as our practical face
-images, while using only frontal 2D images in the training without
-any explicit 3D face model.
+images, while using only frontal 2D images in the gallery and no
+explicit 3D face model.
 Our system could potentially be extended to better handle large pose

### 2011-01-14 rebuttal.tex
Final formatting changes in the paper, and updated references in the rebuttal.

@@
 \section{Response to Reviewer \#1}
-Reviewer 1 makes specific recommendations for improving the
-presentation of the paper, many of which we have heeded.  Please refer to the change
-list for details.
 
@@
 We find the reviewer's other suggestions on presentation useful and have made changes accordingly. 
+In particular, the discussions of gamma and the illumination model have been made more succinct.  As
+requested, the discussion of prior projector-based systems has been expanded.
 
 \paragraph{About Positioning and Novelty:} As we have tried to make clear, the
-proposed method does not aim to solve face recognition for all scenarios! It is
+proposed method does not aim to solve face recognition for all scenarios. It is
 not designed for the case where the training is uncontrolled and limited, and
@@
 
-\paragraph{About Summary Recommendation:} We agree with the reviewer that there is a need for a careful evaluation of our system and comparison to other techniques or systems for face recognition.
-However, because most face recognition systems are not designed for the scenario considered by this paper, a fair and meaningful comparison is difficult.
-Nevertheless, to clarify both the strengths and limitations of our approach, we have augmented the experimental section with comparisons to Local Binary Patterns (LBP), in addition to the comparisons already performed.
-LBP is one of the most popular tools for face verification, and is used in many commercial face recognition systems.
+\paragraph{About Summary Recommendation:} We agree with the reviewer that there
+is a need for a careful evaluation of our system and comparison to other
+techniques or systems for face recognition. However, because most face
+recognition systems are not designed for the scenario considered by this paper,
+a fair and meaningful comparison is difficult. Nevertheless, to clarify both
+the strengths and limitations of our approach, we have augmented the
+comparisons already performed with comparisons to Local Binary Patterns
+(LBP).  LBP is one of the most popular tools for face verification, and is used
+in many commercial face recognition systems.
 
-We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
+We compare both methods on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
 datasets.
@@
 for high-stakes applications. To the best of our knowledge, no existing face
-recognition system, with or without cooperative training, comes close to meet
-the standards of high-stake security applications.
+recognition system, with or without cooperative training, comes close to meeting
+the standards of high-stakes access control applications.
 
@@
 goal of recognition with as few gallery images as possible,'' the reviewer is
-conflating the constraints of one particular type of applications of face recognition
+conflating the constraints of one particular application of face recognition
 (surveillance) with those of face recognition as a whole. This (noble) goal is probably easier to
@@
 
-A huge part of the message of our paper is that for applications that allow it, we can and should
-be gathering more and higher quality data at enrollment. In fact, criminal watchlists typically contain many criminals who have been incarcerated before. Today,
-the acquisition standards for fingerprints are far more strict than those for
-capturing face images of criminals -- precisely because the availability of better data has been shown to
-help fingerprint identification systems. Of course, the idea that better training helps is certainly
-not new or even recent.  What is new is: 1. The availability
-of hardware to efficiently acquire such data (high intensity, high contrast
-ratio DLP projectors), 2. Robust and efficient computational tools to use such data in a
+A huge part of the message of our paper is that for applications that allow it,
+we can and should be gathering more and higher quality data at enrollment.
+Note that the acquisition of fingerprints has become increasingly complete and
+accurate over years, yet for some reason the same has not happened for face
+data.  Of course, the idea that better training data helps is certainly not new
+or even recent.  What is new is: 1.  The availability of hardware to
+efficiently acquire such data (high intensity, high contrast ratio DLP
+projectors), 2.  Robust and efficient computational tools to use such data in a
 principled fashion, 3. Computer hardware fast enough to even contemplate using
@@
 
-\item Our goal is not to achieve the best recognition rates on the NIST family of
-benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
+\item Our goal is not to achieve the best recognition rates on the NIST family
+of benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
 the art in face recognition given cooperative training images.  The motivating
@@
 security).  This requires backwards compatibility with old image data, some of
-which was gathered before computers and digital cameras were even in widespread use.
-Due to these historical reasons, none of these data sets contain multiple controlled
-illuminations.  For example, FRGC contains at most two illuminations per person
-in the gallery.  (Two studio lights were kept on, and a third light was
-switched either on or off.) The techniques used to gather these data sets are not a wise
-choice for applications where users are cooperative (foreigners applying for a visa)
-or forced to be so (suspects or criminals in custody), and are needlessly
-restrictive.  While large public data sets are critical to the rigorous
-comparison of recognition algorithms, their existence should not be used to
-argue against the development of new data acquisition techniques and algorithms
-appropriate for them.
+which was gathered before computers and digital cameras were even in widespread
+use.  Due to these historical reasons, none of these data sets contain multiple
+controlled illuminations.  For example, FRGC contains at most two illuminations
+per person in the gallery.  (Two studio lights were kept on, and a third light
+was switched either on or off.) The techniques used to gather these data sets
+are not a wise choice for applications where users are cooperative, for which
+the constraints on the acquisition are needlessly restrictive.  Furthermore,
+while large public data sets are critical to the rigorous comparison of
+recognition algorithms, their existence should not be used to argue against the
+development of new data acquisition techniques and associated algorithms.
 
-\item We have added comparison with a Local Binary Patterns based recognition system, and
-have shown that both our system and LBP perform better in the application domains or settings
-for which they were designed.  Please see the change list for details.
+\item We have added comparison with a Local Binary Patterns based recognition system.  Please
+see section IV-B for details.  As can be seen in Tables I and II, and Figure 11, 
+both our system and LBP perform better in the application domains 
+for which they were designed.
 
@@
 
-\item We will add further discussion contrasting the advantages and disadvantages of our
-system vs.\ systems incorporating 3D models. This paper does not intend to deal with large 3D pose
-although the proposed techniques are directly applicable to the cases when images of more poses or
-even a full 3D models are available.
+\item Section I-A (Related Work) contains discussion contrasting the
+advantages and disadvantages of our system vs.\ systems incorporating 3D
+models.  Although the proposed techniques have the potential to be extended to
+cases when images of more poses or even a full 3D models are available, this
+paper does not claim to deal with large 3D pose variation.
 
@@
 
-\item We have removed the reference to scattered light, as there appear to
+\item We have removed the reference to scattered light from Section III-A, as there appear to
 be several common uses of the term.  That said, it is true that linear illumination
@@
 \item To our knowledge, we are the first to use projectors for illumination in
-face recognition applications.  Section 1A of the previous submission already
+face recognition applications.  Section I-A of the previous submission already
 cites several other applications that use projectors as illumination sources.
@@
 illumination systems for certain classes of objects.  We have added
-discussion of this to the paper.
+discussion of this to the paper in Section III-B.
 
@@
 acquisition is not a critical part of the face recognition problem -- careful (training)
-data acquisition has been a critical part for almost all other biometric identification
-systems (e.g.\ fingerprint and iris and DNA)! Why should face recognition needs be different?
+data acquisition has been critical for almost all other biometric identification
+systems (e.g.\ fingerprint and iris)! Why should face recognition be different?
 
-We believe we have demonstrated that using a well-thought-out combination of existing ideas
+We believe we have demonstrated that by using a well-thought-out combination of existing ideas
 (iterative image alignment, L1-error function, SRC, using projectors for
@@
 
-In Section II E, we have conducted careful experiments to measure the region of attraction
+In Section II-E, we have conducted careful experiments to measure the region of attraction
 for the proposed alignment method. We have studied the region of attraction of the proposed
@@
 where $\epsilon>0$ is the tolerance. With the suggested
-criteria, the l1-minimizations are guaranteed to converge
-eventually, although convergence may be very slow if they
-are chosen poorly. For the alignment
+criteria, the $\ell_1$ minimizations are guaranteed to converge
+eventually, although convergence may be very slow if the constants
+are poorly tuned. For the alignment
 iteration, we terminate when the change of objective value
@@
 
-We have added experiments to the paper demonstrating
-the utility of the sparse representation(SR)-based
-recognition stage for classification.  In general, we see a
-few percentage points of improvement in recognition rate
-when using the representation residual obtained from the
-SR-based recognition stage (with $S=10$) over the
-recognition based purely on alignment error (with $S$=1).
-While this effect is desirable, as discussed in the paper,
-it is worthy emphasizing that the SR-based recognition also
-enables us to compute SCI, which has demonstrated its power
-for impostor rejection.
+We have added experiment results to Table I demonstrating the utility of the
+sparse representation(SR) based recognition stage for classification.  In
+general, we see a few percentage points of improvement in recognition rate when
+using the representation residual obtained from the SR-based recognition stage
+(with $S=10$) over the recognition based purely on alignment error (with
+$S$=1).  While this effect is desirable, as discussed in the paper, it is
+worthy emphasizing that the SR-based recognition also enables us to compute
+SCI, which has demonstrated its power for impostor rejection.
 
@@
 
-We hope reviewer will find our supplemental justification for the (Sparse
+We hope reviewer will find our above justification for the (Sparse
 Representation based) recognition convincing.  If so, the reviewer may
@@
 They can be easily applied to 2D images from multiple poses or to establish alignment
-between a 2D image and a 3D face model (if available). These are all good directions
+between a 2D image and a renderings from a 3D face model (if available). These are all good directions
 for future research.
 
-We have modified Section I.A and the Conclusion to address these issues.
+We have modified Section I-A and the Conclusion to address these issues.
 

### 2011-01-14 faces_system.tex
Final formatting changes in the paper, and updated references in the rebuttal.

@@
 using SRC.} \label{fig:promo}
-\vspace{-.5in}
+\vspace{-.25in}
 \end{figure}
@@
 experiment on our own dataset. For the test set, we use all 20
-illuminations from Sessions 2-4, which were recorded at
-distinct times over a period of several months. The dataset is
+illuminations from Sessions 2-4, which were recorded 
+over a period of several months. The dataset is
 challenging due to the large number of subjects, and due to
@@
 algorithm on each of the 3 testing sessions. Our algorithm
-achieves recognition rates above $90\%$ for all three sessions,
-with input {\em directly} obtained from the Viola and Jones'
-face detector; there was no manual intervention for the test images.
+achieves recognition rates above $90\%$ for all three sessions.
+For the test images, 
+alignment was initialized automatically via the Viola and Jones'
+face detector. 
 To demonstrate that the sparse representation based recognition step is indeed beneficial
@@
 \label{fig:failed-examples}
-\vspace{-.5in}
+\vspace{-.3in}
 \end{figure}
@@
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
-\vspace{-.35in}
+\vspace{-.6in}
 \end{figure}
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
-\vspace{-.5in}
+\vspace{-.25in}
 \caption{{\bf Recognition under varying level of
-random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:} \vspace{1mm}
+random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:} 
+\vspace{3mm}
 {\small
@@
 \label{fig:multipie-occ-rec}
-\vspace{-.5in}
+\vspace{-.25in}
 \end{figure}
@@
 algorithm on each category.
-\begin{table}
-\vspace{-.1in}	
-\centering \caption{Recognition rates on our own database.}
-\vspace{-.2in}
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Test Category & C1 & C2 & C3 & C4  \\
-\hline
-\hline
-Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
-\hline
-\end{tabular}
-\label{tab:UIUC-recognition}
-\vspace{-.4in}
-\end{table}
 Since our focus is on face
@@
 assumptions can be used to improve the recognition performance.
+\begin{table}[h]
+\vspace{-.1in}	
+\centering \caption{Recognition rates on our own database.}
+\vspace{-.2in}
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Test Category & C1 & C2 & C3 & C4  \\
+\hline
+\hline
+Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
+\hline
+\end{tabular}
+\label{tab:UIUC-recognition}
+\vspace{-.4in}
+\end{table}
 

### 2011-03-21 faces_system.tex
Pretty sure this was the revision we submitted in January
Converting to two column...
Working on two column conversion

@@
 
-\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
+%\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
 %\documentclass[9.5pt,journal,final,finalsubmission,twocolumn]{IEEEtran}
+
+\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran} % from the new style file.  supposedly double column.
+
 % *************************************************************************
@@
 \IEEEcompsoctitleabstractindextext{
-\vspace{-15mm}
+\vspace{0mm}
 \begin{abstract}
@@
 using SRC.} \label{fig:promo}
-\vspace{-.25in}
+\vspace{0mm}
 \end{figure}
@@
 \begin{tabular}{cccc}
-\includegraphics[height=1.2in]{figures_cvpr/L1_cropped} &
-\includegraphics[height=1.2in]{figures_cvpr/y_warp_L1} &
-\includegraphics[height=1.2in]{figures_cvpr/y_hat_L1} &
-\includegraphics[height=1.2in]{figures_cvpr/e_L1} \\
-\includegraphics[height=1.2in]{figures_cvpr/L2_cropped} &
-\includegraphics[height=1.2in]{figures_cvpr/y_warp_L2} &
-\includegraphics[height=1.2in]{figures_cvpr/y_hat_L2} &
-\includegraphics[height=1.2in]{figures_cvpr/e_L2} \\
+\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
+\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
+\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
+\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
+\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
+\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
+\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
+\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
 (a) & (b) & (c) & (d)
 \end{tabular}
-\vspace{-.1in}}
+\vspace{0mm}}
 \caption{{\bf Comparing alignment of a subject wearing sunglasses by
@@
 (c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-.4in}
+\vspace{0mm}
 \end{figure}
@@
  algorithm.
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \begin{tabular}{cccc}
-\includegraphics[height=1.1in]{figures_pami/x_tr.png} &
-\includegraphics[height=1.1in]{figures_pami/y_tr.png} &
-\includegraphics[height=1.1in]{figures_pami/theta.png} &
-\includegraphics[height=1.1in]{figures_pami/scale.png}\\
+\includegraphics[height=1in]{figures_pami/x_tr.png} &
+\includegraphics[height=1in]{figures_pami/y_tr.png} &
+\includegraphics[height=1in]{figures_pami/theta.png} &
+\includegraphics[height=1in]{figures_pami/scale.png}\\
 (d)&(e)&(f)& (g)
 \end{tabular}
-\vspace{-.2in}
+\vspace{0mm}
 }
@@
 \label{fig:attraction}
-\vspace{-.4in}
-\end{figure}
+\vspace{0mm}
+\end{figure*}
 
@@
 \centering
-\includegraphics[height=1.8in]{figures_pami/multiscale.png}
-\vspace{-.3in}
+\includegraphics[width=3.5in]{figures_pami/multiscale.png}
+\vspace{0mm}
 \caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
@@
 \label{fig:multiscale}
-\vspace{-.5in}
+\vspace{0mm}
 \end{figure}
@@
 \begin{tabular}{ccccc}
-\includegraphics[height=1in]{figures_cvpr/5} &
-\includegraphics[height=1in]{figures_cvpr/7} &
-\includegraphics[height=1in]{figures_cvpr/09} &
-\includegraphics[height=1in]{figures_cvpr/11} &
-\includegraphics[height=1in]{figures_cvpr/13} \vspace{-.1in} \\ 
+\includegraphics[height=0.55in]{figures_cvpr/5} &
+\includegraphics[height=0.55in]{figures_cvpr/7} &
+\includegraphics[height=0.55in]{figures_cvpr/09} &
+\includegraphics[height=0.55in]{figures_cvpr/11} &
+\includegraphics[height=0.55in]{figures_cvpr/13} \vspace{0mm} \\ 
 (a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=1in]{figures_cvpr/15} &
-\includegraphics[height=1in]{figures_cvpr/17} &
-\includegraphics[height=1in]{figures_cvpr/19} &
-\includegraphics[height=1in]{figures_cvpr/21} &
-\includegraphics[height=1in]{figures_cvpr/3} \vspace{-.1in} \\ 
-(f) & (g) & (h) & (i) & (j) \vspace{-.1in}
+\includegraphics[height=0.55in]{figures_cvpr/15} &
+\includegraphics[height=0.55in]{figures_cvpr/17} &
+\includegraphics[height=0.55in]{figures_cvpr/19} &
+\includegraphics[height=0.55in]{figures_cvpr/21} &
+\includegraphics[height=0.55in]{figures_cvpr/3} \vspace{0mm} \\ 
+(f) & (g) & (h) & (i) & (j) \vspace{0mm}
 \end{tabular}
@@
 to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-.2in}
+}\label{fig:pose-alignment} \vspace{0mm}
 \end{figure}
@@
 \begin{figure}
-\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}\vspace{-.2in}}
+\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}\vspace{0mm}}
 \caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
-\vspace{-.3in}
+\vspace{-2em}
 \label{fig:system}
@@
 \begin{tabular}{cc}
-\includegraphics[height=1.5in]{figures_cvpr/coverage_experiment_asplode.png} &
-\includegraphics[height=1.5in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
+\includegraphics[height=.75in]{figures_cvpr/coverage_experiment_asplode.png} &
+\includegraphics[height=.75in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
 (a) Coverage Experiment & (b) Chosen Illumination Patterns 
-\end{tabular}\vspace{-2mm}
+\end{tabular}\vspace{0mm}
 \caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
@@
 \label{fig:illumination-patterns}\label{fig:illumination-sufficiency}
-\vspace{-.2in}
+\vspace{-.75em}
 \end{figure}
@@
 \begin{tabular}{cc}
-\includegraphics[height=1.5in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.5in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
+\includegraphics[height=1in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
+\includegraphics[height=1in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
 (a) Coverage & (b) Granularity
-\end{tabular}\vspace{-2mm}
+\end{tabular}\vspace{0mm}
 \caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
 \label{fig:illumination-sufficiency}
-\vspace{-.5in}
+\vspace{-4.5em}
 \end{figure}
@@
 \centering
-\includegraphics[width=6.3in]{figures_cvpr/training.png}
-\vspace{-.3in}
+\includegraphics[width=3.5in]{figures_cvpr/training.png}
+\vspace{-2em}
 \end{figure}
@@
 even when there are no impostors, we include results for recognition
-based only on the alignment error residuals(i.e. S=1), shown in row 6.
+based only on the alignment error residuals(i.e. $S=1$), shown in row 6.
 
@@
 \caption{Recognition rates on large-scale Multi-PIE database.}
-\vspace{-.2in}
+\vspace{0mm}
 \centerline{
@@
 \includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{-.1in} \\
-(a) & (b) & (c) & (d) & (e) & (f) \vspace{-.1in}
+\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{0mm} \\
+(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
 \end{tabular}
@@
 \label{fig:failed-examples}
-\vspace{-.3in}
+\vspace{0mm}
 \end{figure}
@@
 \includegraphics[height=1.5in]{figures_pami/example.png} &
-\includegraphics[height=1.5in]{figures_pami/example_new.png} \vspace{-.1in} \\
-(a) & (b) \vspace{-.1in}
+\includegraphics[height=1.5in]{figures_pami/example_new.png} \vspace{0mm} \\
+(a) & (b) \vspace{0mm}
 \end{tabular}
@@
 \label{fig:new-mask}
-\vspace{-.3in}
+\vspace{0mm}
 \end{figure}
@@
 \caption{Recognition rates on single gallery image FERET
-dataset} \vspace{-.1in} \centerline{
+dataset} \vspace{0mm} \centerline{
 \begin{tabular}{|c|c|c|c|c| }
@@
 \end{tabular}
-\label{tab:FERET-recognition} } \vspace{-.4in}
+\label{tab:FERET-recognition} } \vspace{0mm}
 \end{table}
@@
 \includegraphics[width=3in]{figures_pami/pami_roc2.png}
-}\vspace{-.2in}
+}\vspace{0mm}
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
-\vspace{-.6in}
+\vspace{0mm}
 \end{figure}
@@
 example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.\vspace{-.1in}
+recognition rate.\vspace{0mm}
 
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
-\vspace{-.25in}
+\vspace{0mm}
 \caption{{\bf Recognition under varying level of
 random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:} 
-\vspace{3mm}
+\vspace{0mm}
 {\small
@@
 \label{fig:multipie-occ-rec}
-\vspace{-.25in}
+\vspace{0mm}
 \end{figure}
@@
 are insufficient for representing test images under typical indoor and
-outdoor illuminations.\vspace{-0.2in}
+outdoor illuminations.\vspace{0mm}
 
@@
  \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
- \vspace{-.1in}
+ \vspace{0mm}
 \end{figure}
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
-\vspace{-.2in}
+\vspace{0mm}
  \caption{{\bf Representative examples of category C4}.
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.5in}
+\vspace{0mm}
 \end{figure}
@@
 \begin{table}[h]
-\vspace{-.1in}	
+\vspace{0mm}	
 \centering \caption{Recognition rates on our own database.}
-\vspace{-.2in}
+\vspace{0mm}
 \begin{tabular}{|c|c|c|c|c| }
@@
 \label{tab:UIUC-recognition}
-\vspace{-.4in}
+\vspace{0mm}
 \end{table}
@@
 \centerline{
-\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{-0.2in}}
+\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{0mm}}
 \caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.
@@
 \label{fig:occ-block}
-\vspace{-.5in}
+\vspace{0mm}
 \end{figure}
@@
 acknowledges support from a Microsoft Fellowship and
-the Lemelson-Illinois Student Prize. %} %\vspace{-0.2in}
+the Lemelson-Illinois Student Prize. %} %\vspace{0mm}
 

### 2011-03-21 faces_system.tex
Updated table
Rearranged table columns
Added Zihan's updated ROC curve.

--- 2011-01-14 faces_system.tex
@@
 \usepackage{algorithmic}
+
+\usepackage{multirow}
+\usepackage{pict2e}
+\usepackage{slashbox}
 
@@
 even when there are no impostors, we include results for recognition
-based only on the alignment error residuals(i.e. S=1), shown in row 6.
-
+based only on the alignment error residuals(i.e. $S=1$), shown in row 6.
+
+%\begin{table}
+%\caption{Recognition rates on large-scale Multi-PIE database.}
+%\vspace{-.2in}
+%\centerline{
+%\begin{tabular}{|c|c|c|c|c| }
+%\hline
+%Recognition rate & Session 2 & Session 3 & Session 4  \\
+%\hline
+%\hline
+%LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
+%\hline
+%NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
+%\hline
+%NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
+%\hline
+%LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
+%\hline
+%\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
+%\hline
+%{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
+%\hline
+%{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
+%\hline
+%{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+%\hline
+%\end{tabular}
+%\label{tab:MultiPIE-recognition} }
+%\end{table}
 \begin{table}
@@
 \centerline{
-\begin{tabular}{|c|c|c|c|c| }
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
 \hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
+\backslashbox{Rec.}{Align.} 
+& \multicolumn{3}{|c|}{Eye Detector} 
+& \multicolumn{3}{|c|}{Manual} 
+& \multicolumn{3}{|c|}{Iterative} 
+\\
 \hline
+Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
 \hline
-LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
+NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
 \hline
-NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
+NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
 \hline
-NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
+LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
 \hline
-LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
+LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
 \hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+SRC	&			&			&			&			&			&			& 93.9\%	& 93.8\%	& 92.3\% \\
 \hline
@@
 \centerline{
-\includegraphics[width=3in]{figures_pami/pami_roc2.png}
-}\vspace{-.2in}
+	\begin{tabular}{cc}
+	\includegraphics[width=3in]{figures_pami/pami_roc2.png} & 
+	\includegraphics[width=3in]{figures_pami/pami_roc_revision2.png} \\
+	{\bf OLD PLOT, mixed alignment} & {\bf NEW PLOT, all iterative alignment} \\
+	\end{tabular}
+}
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}

### 2011-03-28 faces_system.tex
Wrote explanations for new plots and data

--- 2011-01-14 faces_system.tex
@@
 \section{Tests on Public Databases}\label{sec:multipie}
-In this section and the next section, we conduct comprehensive
-experiments on large-scale face databases to verify the performance
-of our algorithm and system. We
-first test on the largest public face database available that
-is suitable for testing our algorithm, the CMU Multi-PIE.
-One shortcoming of the CMU Multi-PIE database for our purposes
-is that there is no separate set of test images taken under natural illuminations;
-we are left to choose which sets of images to use for testing and training.
-To challenge our algorithm, we choose only a small set of illuminations for the training set,
-yet we include all illuminations in the testing set. In the following section, we will test our
-algorithm on a face dataset that is collected by our own
-system. The goal for that experiment will be to show that with a sufficient set of
-training illuminations for each subject, our algorithm indeed
-works stably and robustly with practical illumination,
-misalignment, pose, and occlusion, as already indicated by our
-experiment shown in Figure \ref{fig:promo}(bottom).
-
-CMU Multi-PIE provides the most extensive test set among public
-datasets. This database contains images of 337 subjects across
-simultaneous variation in pose, expression, and illumination.
-Of these 337 subjects, we use all of the 249 subjects present
-in Session 1 as the training set. The remaining 88 subjects are
-treated as ``impostors'', or invalid images. For each of the
-249 training subjects, we include frontal images of 7 frontal
-illuminations,\footnote{They are illuminations
-$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
-directional illumination, we subtract the ambient-illuminated
-image 0.} taken with neutral expression. As suggested by the
-work of \cite{Georghiades2001-PAMI}, these extreme frontal
-illuminations would be sufficient to linearly represent other
-frontal illuminations, as will also be corroborated by the next
-experiment on our own dataset. For the test set, we use all 20
-illuminations from Sessions 2-4, which were recorded 
-over a period of several months. The dataset is
-challenging due to the large number of subjects, and due to
-natural variation in subject appearance over time. Table
-\ref{tab:MultiPIE-recognition} shows the result of our
-algorithm on each of the 3 testing sessions. Our algorithm
-achieves recognition rates above $90\%$ for all three sessions.
-For the test images, 
-alignment was initialized automatically via the Viola and Jones'
-face detector. 
-To demonstrate that the sparse representation based recognition step is indeed beneficial
-even when there are no impostors, we include results for recognition
-based only on the alignment error residuals(i.e. S=1), shown in row 6.
+In this section and the next section, we conduct comprehensive experiments on
+large-scale face databases to verify the performance of our algorithm and
+system. We first test on the largest public face database available that is
+suitable for testing our algorithm, the CMU Multi-PIE.  One shortcoming of the
+CMU Multi-PIE database for our purposes is that there is no separate set of
+test images taken under natural illuminations; we are left to choose which sets
+of images to use for testing and training.  To challenge our algorithm, we
+choose only a small set of illuminations for the training set, yet we include
+all illuminations in the testing set. In the following section, we will test
+our algorithm on a face dataset that is collected by our own system. The goal
+for that experiment will be to show that with a sufficient set of training
+illuminations for each subject, our algorithm indeed works stably and robustly
+with practical illumination, misalignment, pose, and occlusion, as already
+indicated by our experiment shown in Figure \ref{fig:promo}(bottom).
+
+CMU Multi-PIE provides the most extensive test set among public datasets. This
+database contains images of 337 subjects across simultaneous variation in pose,
+expression, and illumination.  Of these 337 subjects, we use all of the 249
+subjects present in Session 1 as the training set. The remaining 88 subjects
+are treated as ``impostors'', or invalid images. For each of the 249 training
+subjects, we include frontal images of 7 frontal illuminations,\footnote{They
+are illuminations $\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
+directional illumination, we subtract the ambient-illuminated image 0.} taken
+with neutral expression. As suggested by the work of
+\cite{Georghiades2001-PAMI}, these extreme frontal illuminations would be
+sufficient to linearly represent other frontal illuminations, as will also be
+corroborated by the next experiment on our own dataset. For the test set, we
+use all 20 illuminations from Sessions 2-4, which were recorded over a period
+of several months. The dataset is challenging due to the large number of
+subjects, and due to natural variation in subject appearance over time.  Line 5
+of Table \ref{tab:MultiPIE-recognition} shows the result of our recognition
+pipeline on each of the 3 testing sessions. Our algorithm achieves recognition
+rates above $90\%$ for all three sessions.  For the test images, our iterative
+alignment was initialized automatically via the Viola and Jones' face detector.
 
 \begin{table}
-\caption{Recognition rates on large-scale Multi-PIE database.}
+\caption{Recognition rates on the Multi-PIE database for different pairings of alignment
+and recognition stages.}
 \vspace{-.2in}
 \centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
-\hline
-\hline
-LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
-\hline
-NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
-\hline
-NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
-\hline
-LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
-\hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
+\hline
+\backslashbox{Rec.}{Align.} 
+& \multicolumn{3}{|c|}{Face Detector} 
+& \multicolumn{3}{|c|}{Manual} 
+& \multicolumn{3}{|c|}{Iterative} 
+\\
+\hline
+Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
+\hline
+NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
+\hline
+NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
+\hline
+LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
+\hline
+LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
+\hline
+SRC	&			&			&			&			&			&			& 93.9\%	& 93.8\%	& 92.3\% \\
 \hline
@@
 \end{table}
+
 \begin{figure}
@@
 \subsection{Comparison to Existing Work}
+% Classical algorithms
 We compare our result to baseline linear-projection-based
@@
 assume pixel-accurate alignment, they are not expected to work
-well if the test is not well aligned with the training. In
+well if the test image is not well aligned with the training. In
 Table~\ref{tab:MultiPIE-recognition}, we report the results of
-those algorithms with two types of input: 1.\ the output of the
-Viola and Jones' detector, indicated by a subscript ``$d$'';
-2.\ the input face is aligned to the training with manually
-selected outer eye corners, indicated by a subscript ``$m$''.
-Notice that even with careful manual registration, these
-baseline algorithms perform significantly worse than our
-algorithm, which uses input directly from the face detector.
+these classical algorithms with three types of testing image alignment: 
+1.\ alignment from the Viola and Jones' detector,
+2.\ alignment via manually selected outer eye corners, and 
+3.\ the output of our iterative alignment algorithm.
 The performance drop of the LDA algorithm on Multi-PIE reported here
 seems to agree with that reported already in
-\cite{Gross2008-FGR}.
-
-We also compare our result to the very recent work
+\cite{Gross2008-FGR}.  All of the classical algorithms benefit
+greatly from being paired with our iterative alignment algorithm.
+
+% LBP
+Finally, we compare our result to Local Binary Patterns (LBP)
+\cite{Ahonen2006-PAMI}, a local appearance descriptor which is able to capture
+fine details of facial appearance and texture.  Due to its robustness to
+variations in illumination, facial expression, aging and other changes, LBP has
+achieved the state-of-the-art face recognition performance in the scenario when
+only one sample per person is used for training \cite{Tan06facerecognition}. In
+this paper, we follow the same steps as in \cite{Ahonen2006-PAMI} to construct
+an LBP descriptor for each training and testing sample. The $80\times 60$ face
+region is first divided into a regular $10\times 10$ grid of cells, each of
+size $8\times 6$ pixels. Within each cell, the histogram of 59 uniform binary
+patterns is then computed, where the patterns are generated by thresholding 8
+neighboring pixels in a circle of radius 2 using the central pixel value.
+Finally, the local histograms are concatenated to produce the global descriptor
+vector. As suggested in \cite{Ahonen2006-PAMI}, the recognition is performed
+using a nearest neighbor classifier with Chi square distance as the distance
+measure and we report the recognition rates with the same two types of input as
+before.  Recognition results for LBP on Multi-PIE are shown in
+Table~\ref{tab:MultiPIE-recognition}.  We attribute the improvement in
+performance of LBP over SRC in this experiment to its robustness to
+illumination components that are not in the training set.  Some of the test
+images used in this experiment are taken under illuminations not present in the
+training subset.  Note that, like the other classical algorithms, the
+performance of LBP is still sensitive to misalignment. This is notable given
+that LBP is often applied without any special alignment.
+
+\begin{table}
+\caption{Recognition rates on the Multi-PIE database for \cite{Yang2010-CVPR} and variants
+of Algorithm 1}
+\vspace{-.2in}
+\centerline{
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Recognition rate & Session 2 & Session 3 & Session 4  \\
+\hline
+\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
+\hline
+{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
+\hline
+{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
+\hline
+{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+\hline
+\end{tabular}
+\label{tab:MultiPIE-recognition2} }
+\vspace{-.3in}
+\end{table}
+
+To demonstrate that the sparse representation based recognition step is indeed
+beneficial even when there are no impostors, we include results for recognition
+based only on the alignment error residuals(i.e. $S=1$), shown in 
+Table \ref{tab:MultiPIE-recognition2}, row 6.
+
+We also compare our result to the recent work
 \cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
@@
 performance on the CMU Multi-PIE database.
-Table~\ref{tab:MultiPIE-recognition} shows that our algorithm
+Table~\ref{tab:MultiPIE-recognition2} shows that our algorithm
 achieves similar or better performance on different sessions of
 Multi-PIE.
-
-Finally, we compare our result to Local Binary Patterns (LBP)
-\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
-able to capture fine details of facial appearance and texture.
-Due to its robustness to variations in illumination, facial
-expression, aging and other changes, LBP has achieved the
-state-of-the-art face recognition performance in the scenario
-when only one sample per person is used for training
-\cite{Tan06facerecognition}. In this paper, we follow the same
-steps as in \cite{Ahonen2006-PAMI} to construct an LBP
-descriptor for each training and testing sample. The $80\times
-60$ face region is first divided into a regular $10\times 10$
-grid of cells, each of size $8\times 6$ pixels. Within each
-cell, the histogram of 59 uniform binary patterns is then
-computed, where the patterns are generated by thresholding 8
-neighboring pixels in a circle of radius 2 using the central
-pixel value. Finally, the local histograms are concatenated to
-produce the global descriptor vector. As suggested in
-\cite{Ahonen2006-PAMI}, the recognition is performed using a
-nearest neighbor classifier with Chi square distance as the
-distance measure and we report the recognition rates with the
-same two types of input as before. As shown in
-Table~\ref{tab:MultiPIE-recognition}, LBP achieves competitive
-recognition rates given manually aligned training and testing
-samples, which demonstrates its robustness to illumination
-changes. However, its performance is very sensitive to
-misalignment.
 
@@
 
-Figure \ref{fig:roc-multipie} plots the receiver operating
-characteristic (ROC) curves, which are generated by sweeping
-the threshold $t$ through the entire range of possible values
-for each algorithm.\footnote{Rejecting invalid images not in
-the entire database is much more difficult than deciding if two
-face images are the same subject. Figure 8 should not be
-confused with typical ROC curves for face similarity, e.g.,
-\cite{PhillipsP2007}.} We can see that SCI based approach
-significantly outperforms the other three algorithms. Similar
-contrasts between our algorithm and baseline algorithms were
-also observed for SRC in \cite{Wright2009-PAMI}, though on much
-smaller datasets.
-
+Figure \ref{fig:roc-multipie} plots the receiver operating characteristic (ROC)
+curves, which are generated by sweeping the threshold $t$ through the entire
+range of possible values for each algorithm.\footnote{Rejecting invalid images
+not in the entire database is much more difficult than deciding if two face
+images are the same subject. Figure 8 should not be confused with typical ROC
+curves for face similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
+that the SCI based recognition approach significantly outperforms the other
+algorithms, including LBP, even when all algorithms are coupled with our
+proposed iterative alignment.  In the right plot we again see that classical
+algorithms, and even LBP, are very sensitive to alignment.  Similar contrasts
+between our algorithm and baseline algorithms were also observed for SRC in
+\cite{Wright2009-PAMI}, though on much smaller datasets.
 \begin{figure}
 \centerline{
-\includegraphics[width=3in]{figures_pami/pami_roc2.png}
-}\vspace{-.2in}
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
-\vspace{-.6in}
+	\begin{tabular}{cc}
+	\includegraphics[width=3in]{figures_pami/pami_roc_revision2.png} &
+	\includegraphics[width=3in]{figures_pami/pami_roc2.png} \\
+	{\small With iterative alignment} & {\small Classical Algorithms with manual alignment} 
+\\
+	\end{tabular}
+}
+\caption{{\bf ROC curves} for subject validation on Multi-PIE database, (a) with iterative alignment,
+and (b) for the classical algorithms with manual alignment}\label{fig:roc-multipie}
+%\vspace{-.6in}
 \end{figure}

### 2011-03-29 faces_system.tex
Changes from Zihan

@@
 
-CMU Multi-PIE provides the most extensive test set among public datasets. This
-database contains images of 337 subjects across simultaneous variation in pose,
-expression, and illumination.  Of these 337 subjects, we use all of the 249
-subjects present in Session 1 as the training set. The remaining 88 subjects
-are treated as ``impostors'', or invalid images. For each of the 249 training
-subjects, we include frontal images of 7 frontal illuminations,\footnote{They
-are illuminations $\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
-directional illumination, we subtract the ambient-illuminated image 0.} taken
-with neutral expression. As suggested by the work of
-\cite{Georghiades2001-PAMI}, these extreme frontal illuminations would be
-sufficient to linearly represent other frontal illuminations, as will also be
-corroborated by the next experiment on our own dataset. For the test set, we
-use all 20 illuminations from Sessions 2-4, which were recorded over a period
-of several months. The dataset is challenging due to the large number of
-subjects, and due to natural variation in subject appearance over time.  Line 5
-of Table \ref{tab:MultiPIE-recognition} shows the result of our recognition
-pipeline on each of the 3 testing sessions. Our algorithm achieves recognition
-rates above $90\%$ for all three sessions.  For the test images, our iterative
-alignment was initialized automatically via the Viola and Jones' face detector.
+CMU Multi-PIE provides the most extensive test set among public
+datasets. This database contains images of 337 subjects across
+simultaneous variation in pose, expression, and illumination.
+Of these 337 subjects, we use all of the 249 subjects present
+in Session 1 as the training set. The remaining 88 subjects are
+treated as ``impostors'', or invalid images. For each of the
+249 training subjects, we include frontal images of 7 frontal
+illuminations,\footnote{They are illuminations
+$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
+directional illumination, we subtract the ambient-illuminated
+image 0.} taken with neutral expression. As suggested by the
+work of \cite{Georghiades2001-PAMI}, we choose these extreme
+frontal illuminations in the hope that they would linearly
+represent other frontal illuminations well. For the test set,
+we use all 20 illuminations from Sessions 2-4, which were
+recorded over a period of several months. The dataset is
+challenging due to the large number of subjects, and due to
+natural variation in subject appearance over time. Table
+\ref{tab:MultiPIE-recognition2} shows the result of our
+algorithm on each of the 3 testing sessions. Our algorithm
+achieves recognition rates above $90\%$ for all three sessions.
+For the test images, our iterative alignment was initialized
+automatically via the Viola and Jones' face detector. To
+demonstrate that the sparse representation based recognition
+step is indeed beneficial even when there are no impostors, we
+include results for recognition based only on the alignment
+error residuals (i.e. $S=1$), shown in row 1.
 
 \begin{table}
-\caption{Recognition rates on the Multi-PIE database for different pairings of alignment
-and recognition stages.}
-\vspace{-.2in}
+\caption{Recognition rates on the Multi-PIE database for
+Algorithm 1 and \cite{Yang2010-CVPR}} \vspace{-.2in}
 \centerline{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
-\hline
-\backslashbox{Rec.}{Align.} 
-& \multicolumn{3}{|c|}{Face Detector} 
-& \multicolumn{3}{|c|}{Manual} 
-& \multicolumn{3}{|c|}{Iterative} 
-\\
-\hline
-Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
-\hline
-NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
-\hline
-NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
-\hline
-LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
-\hline
-LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
-\hline
-SRC	&			&			&			&			&			&			& 93.9\%	& 93.8\%	& 92.3\% \\
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Recognition rate & Session 2 & Session 3 & Session 4  \\
+\hline
+{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
+\hline
+{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
+\hline
+{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+\hline
+\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
 \hline
 \end{tabular}
-\label{tab:MultiPIE-recognition} }
+\label{tab:MultiPIE-recognition2} }
 \end{table}
@@
 
-Table \ref{tab:MultiPIE-recognition} shows that the
+Table \ref{tab:MultiPIE-recognition2} shows that the
 recognition rates on Multi-PIE indeed increase with this new
@@
 recognition rates of our algorithm are significantly boosted
-with the new window. However, we should mention that the best choice of the window
-is problem-specific and there is
-not a simple guideline to follow. For example, although the new
+with the new window. However, we should mention that the best
+choice of the window is problem-specific and there is not a
+simple guideline to follow. For example, although the new
 window performs better on Multi-PIE, the same window does not
@@
 \subsection{Comparison to Existing Work}
-% Classical algorithms
-We compare our result to baseline linear-projection-based
-algorithms, such as Nearest Neighbor (NN), Nearest Subspace
-(NS) \cite{Lee2005-PAMI}, and Linear Discriminant Analysis
-(LDA) \cite{Belhumeur1997-PAMI}.\footnote{We do not list
-results on PCA \cite{Turk1991-CVPR} as its performance is
-always below that of Nearest Subspace.} Since these algorithms
-assume pixel-accurate alignment, they are not expected to work
-well if the test image is not well aligned with the training. In
-Table~\ref{tab:MultiPIE-recognition}, we report the results of
-these classical algorithms with three types of testing image alignment: 
-1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye corners, and 
-3.\ the output of our iterative alignment algorithm.
-The performance drop of the LDA algorithm on Multi-PIE reported here
-seems to agree with that reported already in
-\cite{Gross2008-FGR}.  All of the classical algorithms benefit
-greatly from being paired with our iterative alignment algorithm.
-
-% LBP
-Finally, we compare our result to Local Binary Patterns (LBP)
-\cite{Ahonen2006-PAMI}, a local appearance descriptor which is able to capture
-fine details of facial appearance and texture.  Due to its robustness to
-variations in illumination, facial expression, aging and other changes, LBP has
-achieved the state-of-the-art face recognition performance in the scenario when
-only one sample per person is used for training \cite{Tan06facerecognition}. In
-this paper, we follow the same steps as in \cite{Ahonen2006-PAMI} to construct
-an LBP descriptor for each training and testing sample. The $80\times 60$ face
-region is first divided into a regular $10\times 10$ grid of cells, each of
-size $8\times 6$ pixels. Within each cell, the histogram of 59 uniform binary
-patterns is then computed, where the patterns are generated by thresholding 8
-neighboring pixels in a circle of radius 2 using the central pixel value.
-Finally, the local histograms are concatenated to produce the global descriptor
-vector. As suggested in \cite{Ahonen2006-PAMI}, the recognition is performed
-using a nearest neighbor classifier with Chi square distance as the distance
-measure and we report the recognition rates with the same two types of input as
-before.  Recognition results for LBP on Multi-PIE are shown in
-Table~\ref{tab:MultiPIE-recognition}.  We attribute the improvement in
-performance of LBP over SRC in this experiment to its robustness to
-illumination components that are not in the training set.  Some of the test
-images used in this experiment are taken under illuminations not present in the
-training subset.  Note that, like the other classical algorithms, the
-performance of LBP is still sensitive to misalignment. This is notable given
-that LBP is often applied without any special alignment.
-
-\begin{table}
-\caption{Recognition rates on the Multi-PIE database for \cite{Yang2010-CVPR} and variants
-of Algorithm 1}
-\vspace{-.2in}
-\centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
-\hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition2} }
-\vspace{-.3in}
-\end{table}
-
-To demonstrate that the sparse representation based recognition step is indeed
-beneficial even when there are no impostors, we include results for recognition
-based only on the alignment error residuals(i.e. $S=1$), shown in 
-Table \ref{tab:MultiPIE-recognition2}, row 6.
-
-We also compare our result to the recent work
+We first compare our result to the recent work
 \cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
-initial registration is again obtained from manually selected
-outer eye corners. Then, a supervised hierarchical sparse
-coding model based on local image descriptors is trained, which
-enjoys certain translation invariant properties. With the same
+initial registration is obtained from manually selected outer
+eye corners. Then, a supervised hierarchical sparse coding
+model based on local image descriptors is trained, which enjoys
+certain translation invariant properties. With the same
 training and testing sets, \cite{Yang2010-CVPR} is able to
-handle any remaining misalignment and achieves state-of-the-art
+handle the remaining misalignment and achieves state-of-the-art
 performance on the CMU Multi-PIE database.
@@
 
-Although our algorithm is not designed for
+% Classical algorithms
+To better examine the effectiveness of our iterative alignment
+algorithm, we next compare our result to baseline
+linear-projection-based algorithms, such as Nearest Neighbor
+(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
+Discriminant Analysis (LDA)
+\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
+PCA \cite{Turk1991-CVPR} as its performance is always below
+that of Nearest Subspace.} Since these algorithms assume
+pixel-accurate alignment, they are not expected to work well if
+the test image is not well aligned with the training. In
+Table~\ref{tab:MultiPIE-recognition}, we report the results of
+these classical algorithms with three types of testing image
+alignment: 1.\ alignment from the Viola and Jones' detector,
+2.\ alignment via manually selected outer eye corners, and 3.\
+the output of our iterative alignment algorithm. The
+performance drop of the LDA algorithm on Multi-PIE reported
+here seems to agree with that reported already in
+\cite{Gross2008-FGR}.  All of the classical algorithms benefit
+greatly from being paired with our iterative alignment
+algorithm.
+
+\begin{table}
+\caption{Recognition rates on the Multi-PIE database for
+different pairings of alignment and recognition stages.}
+\vspace{-.2in} \centerline{
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
+\hline
+\backslashbox{Rec.}{Align.}
+& \multicolumn{3}{|c|}{Face Detector}
+& \multicolumn{3}{|c|}{Manual}
+& \multicolumn{3}{|c|}{Iterative Alignment}
+\\
+\hline
+Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
+\hline
+NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
+\hline
+NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
+\hline
+LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
+\hline
+LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
+\hline
+SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
+\hline
+\end{tabular}
+\label{tab:MultiPIE-recognition} } \vspace{-.4in}
+\end{table}
+
+% LBP
+We also compare our result to Local Binary Patterns (LBP)
+\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
+able to capture fine details of facial appearance and texture.
+Due to its robustness to variations in illumination, facial
+expression, aging and other changes, LBP has achieved the
+state-of-the-art face recognition performance in the scenario
+when only one sample per person is used for training
+\cite{Tan06facerecognition}. In this paper, we follow the same
+steps as in \cite{Ahonen2006-PAMI} to construct an LBP
+descriptor for each training and testing sample. The $80\times
+60$ face region is first divided into a regular $10\times 10$
+grid of cells, each of size $8\times 6$ pixels. Within each
+cell, the histogram of 59 uniform binary patterns is then
+computed, where the patterns are generated by thresholding 8
+neighboring pixels in a circle of radius 2 using the central
+pixel value. Finally, the local histograms are concatenated to
+produce the global descriptor vector. As suggested in
+\cite{Ahonen2006-PAMI}, the recognition is performed using a
+nearest neighbor classifier with Chi square distance as the
+distance measure and we report the recognition rates with the
+same three types of input as before.  As shown in
+Table~\ref{tab:MultiPIE-recognition}, although LBP achieves
+competitive recognition rates given manually aligned training
+and testing samples, demonstrating its robustness to moderate
+misalignment, it still benefits from using the output of our
+iterative alignment algorithm as the input. In addition, like
+the other classical algorithms, the performance of LBP degrades
+dramatically if it is applied directly to the output of a face
+detector. This is notable given that LBP is often applied
+without any special alignment in practice. Finally, we
+attribute the improvement in performance of LBP over SRC in
+this experiment to its robustness to illumination components
+that cannot be linearly interpolated by the training set.
+
+In addition, although our algorithm is not designed for
 recognition when there is only a single gallery image per user,
-we compare its
-performance with LBP within this setting for completeness.
-For this experiment, we use the FERET dataset
+we compare its performance with LBP within this setting for
+completeness. For this experiment, we use the FERET dataset
 \cite{phillips1998feret}, which contains five standard
@@
 
-Figure \ref{fig:roc-multipie} plots the receiver operating characteristic (ROC)
-curves, which are generated by sweeping the threshold $t$ through the entire
-range of possible values for each algorithm.\footnote{Rejecting invalid images
-not in the entire database is much more difficult than deciding if two face
-images are the same subject. Figure 8 should not be confused with typical ROC
-curves for face similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
-that the SCI based recognition approach significantly outperforms the other
-algorithms, including LBP, even when all algorithms are coupled with our
-proposed iterative alignment.  In the right plot we again see that classical
-algorithms, and even LBP, are very sensitive to alignment.  Similar contrasts
-between our algorithm and baseline algorithms were also observed for SRC in
-\cite{Wright2009-PAMI}, though on much smaller datasets.
-\begin{figure}
+\begin{figure}[t]
 \centerline{
@@
 	\includegraphics[width=3in]{figures_pami/pami_roc2.png} \\
-	{\small With iterative alignment} & {\small Classical Algorithms with manual alignment} 
+	{\small All algorithm with iterative alignment} & {\small Classical algorithms with manual alignment}
 \\
@@
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database, (a) with iterative alignment,
-and (b) for the classical algorithms with manual alignment}\label{fig:roc-multipie}
-%\vspace{-.6in}
+and (b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
+\vspace{-.4in}
 \end{figure}
+
+Figure \ref{fig:roc-multipie} plots the receiver operating
+characteristic (ROC) curves, which are generated by sweeping
+the threshold $t$ through the entire range of possible values
+for each algorithm.\footnote{Rejecting invalid images not in
+the entire database is much more difficult than deciding if two
+face images are the same subject. Figure 8 should not be
+confused with typical ROC curves for face similarity, e.g.,
+\cite{PhillipsP2007}.} On the left we can see that the SCI
+based recognition approach significantly outperforms the other
+algorithms, including LBP, even when all algorithms are coupled
+with our proposed iterative alignment.  In the right plot we
+again see that classical algorithms, and even LBP, are very
+sensitive to alignment.  Similar contrasts between our
+algorithm and baseline algorithms were also observed for SRC in
+\cite{Wright2009-PAMI}, though on much smaller datasets.
 
@@
 \label{fig:multipie-occ-rec}
-\vspace{-.25in}
+\vspace{-.4in}
 \end{figure}

### 2011-03-29 rebuttal.tex
fixed typo, added rebuttal

@@
 
-We thank the reviewers for their constructive comments, and in particular
-reviewer 1 for an extra in-depth review.  While we have philosophical
-differences on several points, we deeply appreciate the constructive
-criticisms raised by the reviewers, which have helped us strengthen the paper considerably.
-Below, we respond to comments of the reviewers in order, indicating
-specific changes that we have made to address their concerns.
+We again thank the reviewers and the area chair for their comments.
+%I am mostly satisfied with the responses to the earlier review. The paper is
+%in much better form now and more carefully written.
+Changes have been made to the paper to address all of the changes requested by Reviewer 1,
+including a more complete set of results in Table 2 and Figure 11, and associated 
+discussions.  We give a quick overview of the changes.
 
-\section{Response to Reviewer \#1}
+{\bf
+	Page 1 line 16; "there are many face recognition that fall roughly.."
+	missing the word 'applications'
+}
+The typo on Page 1 is now ammended.
 
-\paragraph{About Presentation:} The reviewer suggests that we can reduce the description
-of the ALM algorithm in Section II-D. We agree that ALM has been extensively studied in the optimization
-literature. We do not claim any novelty here. However, we believe a complete description of the algorithm
-here is important for the TPAMI audience: There are
-many algorithms for solving the associated optimization problem, with drastically different complexity
-and speed. Typical vision researchers might not be so familiar with some of the recent
-developments in these algorithms. In particular some of the more effective improvements to ALM are fairly
-recent and scattered in different literatures. In fact, many of the questions raised by reviewer \#2
-ask for even more detailed explanation about the algorithms. So, we believe that a complete but brief
-description of the ALM-based algorithm is important for the audience of TPAMI to implement and
-test the ideas efficiently and correctly.
+{\bf
+	In Table 1, the baseline results are shown with subscripts 'd' and 'm', i.e.
+	using results of a detector, and manual registration. It seems like an obvious
+	omission not to include results of recognition where input comes from the
+	proposed registration algorithm itself. While the paper states that even
+	manual registration works worse, it is not particularly convincing that manual
+	registration is to be considered better than automatic registration - the
+	human visual sensitivity is pretty bad for a problem like registration.
+}
+We ran the requested experiments.  We agree with the reviewer that automatic registration
+can often out-perform manual registration.  The results back this up, and the text
+is ammended to be more clear on this point.  The new more complete benchmark is shown in
+Table 2, and Figure 11.  All of the classical methods benefit from using our iterative
+alignment procudeure.  This is even true for LBP, which is known to have some robustness
+to moderate misalignment.  The discussions have been updated to cover the new results.
 
-We find the reviewer's other suggestions on presentation useful and have made changes accordingly. 
-In particular, the discussions of gamma and the illumination model have been made more succinct.  As
-requested, the discussion of prior projector-based systems has been expanded.
+{\bf
+	In any case, the extent of manual input is unclear. It is said 'the input face
+	is aligned to the training with manually selected outer eye corners'. If only
+	2 points are manually selected, then what kind of registration can be done
+	with just 2 corresponding points ? Surely not affine alignment. Many more are
+	needed if it is to be done robustly. So it is not surprising that manual
+	alignment is giving worse results than automatic. If my interpretation is
+	wrong, is the entire affine registration being done manually ? If so, as
+	mentioned above the accuracy of the human can be bad.
+}
+Similarity transformations are used for all of the experiments in this section, and
+therefore two manually selected outer eye corners are sufficient.  We have
+made this clear by adding a footnote.
 
-\paragraph{About Positioning and Novelty:} As we have tried to make clear, the
-proposed method does not aim to solve face recognition for all scenarios. It is
-not designed for the case where the training is uncontrolled and limited, and
-hence it may not have any advantages over other face recognition
-techniques/systems specifically designed for such scenarios.  The proposed
-system is an attempt to build a system that is not only accurate and robust
-enough to compete with other biometric measures but also discreet and
-non-intrusive when in use. There are many practical scenarios where the
-proposed setting is very relevant, such as access control, where well
-controlled training images can (and should) be obtained in advance but no well
-controlled test images are needed or available. Particularly, the system should
-be able to handle typical daily changes in facial appearance, such as occlusion
-by glasses, hats, or facial hair and variations in lighting and moderate
-expression. The possibility of achieving fully automated (at recognition time),
-touch-free identification is arguably the main advantage of face recognition
-over other biometric measures such as fingerprint and iris recognition.  To the
-best of our knowledge even for this admittedly restricted scenario, there is no
-practical face recognition system whose performance can meet the standards of
-most access control and security applications. Existing systems have found
-success  mostly in applications that are less demanding in performance, such as
-online image search or photo album organization.
-
-Thanks to the reviewer's comments, we have modified the introduction to make this point more clear, especially clarifying the specific scenario where our method can claim novelty.
-
-\paragraph{About Summary Recommendation:} We agree with the reviewer that there
-is a need for a careful evaluation of our system and comparison to other
-techniques or systems for face recognition. However, because most face
-recognition systems are not designed for the scenario considered by this paper,
-a fair and meaningful comparison is difficult. Nevertheless, to clarify both
-the strengths and limitations of our approach, we have augmented the
-comparisons already performed with comparisons to Local Binary Patterns
-(LBP).  LBP is one of the most popular tools for face verification, and is used
-in many commercial face recognition systems.
-
-We compare both methods on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
-datasets.
-The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is neither controlled nor sufficient.
-
-\paragraph{Detailed Concerns or Comments by Reviewer 1:} \begin{enumerate}
-\item Reviewer 1 is correct that use of our system is largely restricted to
-users who are cooperative (in the sense of non-adversarial). Again, the main
-point of our paper is that we can make identification more reliable and
-convenient, at the relatively small cost of doing a better job of gathering
-data at enrollment time.  Ideally, face recognition will take place
-automatically as a user walks through a door or works in front a desktop.
-
-In addition, if the accuracy and reliability of a face recognition system can
-truly get close to that of fingerprint and iris, it can serve as a
-complementary measure in a multi-modal system and improve the overall security
-for high-stakes applications. To the best of our knowledge, no existing face
-recognition system, with or without cooperative training, comes close to meeting
-the standards of high-stakes access control applications.
-
-In response to the reviewer's question "is occlusion really a problem?", we have found that the use of
-the L1-norm as a robust error function improves performance even when the test
-image does not contain occlusions.  Any boundaries where the intensity changes
-in a misaligned image can result in a large, localized error component.  While
-unoccluded faces are mostly smooth, sharp boundaries still occur where there
-are cast shadows, especially at depth discontinuities such as the nostrils, the
-sides of the nose, and partially open mouths.
-
-\item Reviewer 1's interest in face recognition for terrorist watchlist and law
-enforcement applications is of limited relevance to the evaluation of this
-manuscript.  By claiming that ``the field of face recognition has set itself the
-goal of recognition with as few gallery images as possible,'' the reviewer is
-conflating the constraints of one particular application of face recognition
-(surveillance) with those of face recognition as a whole. This (noble) goal is probably easier to
-set than to meet. Although progress has been made towards this goal, the reviewer
-probably has to agree that after several decades (and several high-profile failures of
-face recognition systems in real-world testing), people are still looking for techniques
-that can achieve the performance required by high-end applications such as homeland
-security. Having such a high, long-term goal should not prevent us from making progress in developing systems that
-can already provide good guarantees of success, even if the conditions are somewhat limited.
-
-A huge part of the message of our paper is that for applications that allow it,
-we can and should be gathering more and higher quality data at enrollment.
-Note that the acquisition of fingerprints has become increasingly complete and
-accurate over years, yet for some reason the same has not happened for face
-data.  Of course, the idea that better training data helps is certainly not new
-or even recent.  What is new is: 1.  The availability of hardware to
-efficiently acquire such data (high intensity, high contrast ratio DLP
-projectors), 2.  Robust and efficient computational tools to use such data in a
-principled fashion, 3. Computer hardware fast enough to even contemplate using
-entire images as features.
-
-\item ``It is conceivable that SRC is applicable to any feature as well - which
-underscores the fact that it is simply a pattern classification toolbox - on
-par with any found in a textbook on pattern recognition. What is it that makes
-it special for face recognition?''
-
-Most general pattern classification algorithms do not take into account the
-physically motivated linear model that is central (though by no means unique)
-to the problem of illumination variation.
-
-\item Our goal is not to achieve the best recognition rates on the NIST family
-of benchmarks (FERET, FRVT, FRGC, MBGC, GBU); rather it is to advance state of
-the art in face recognition given cooperative training images.  The motivating
-application of the NIST data sets is government law enforcement (originally for
-enforcement of drug prohibition, and more recently shifting towards border
-security).  This requires backwards compatibility with old image data, some of
-which was gathered before computers and digital cameras were even in widespread
-use.  Due to these historical reasons, none of these data sets contain multiple
-controlled illuminations.  For example, FRGC contains at most two illuminations
-per person in the gallery.  (Two studio lights were kept on, and a third light
-was switched either on or off.) The techniques used to gather these data sets
-are not a wise choice for applications where users are cooperative, for which
-the constraints on the acquisition are needlessly restrictive.  Furthermore,
-while large public data sets are critical to the rigorous comparison of
-recognition algorithms, their existence should not be used to argue against the
-development of new data acquisition techniques and associated algorithms.
-
-\item We have added comparison with a Local Binary Patterns based recognition system.  Please
-see section IV-B for details.  As can be seen in Tables I and II, and Figure 11, 
-both our system and LBP perform better in the application domains 
-for which they were designed.
-
-\item Although we claim to have a preliminary system that works in a realistic
-access control setting, we have never claimed that our method solves all problems in face
-recognition. We do not know where the reviewer gets that wrong impression.
-In the introduction, and at multiple places later in the manuscript, we emphasize
-the importance of having more restrictive training data for guaranteeing the success
-of our method.
-
-%If it helps, we will try to tone down the rhetoric in XYZ places. UNFINISHED!!
-
-\item Section I-A (Related Work) contains discussion contrasting the
-advantages and disadvantages of our system vs.\ systems incorporating 3D
-models.  Although the proposed techniques have the potential to be extended to
-cases when images of more poses or even a full 3D models are available, this
-paper does not claim to deal with large 3D pose variation.
-
-\item The issues of (mis)alignment and illumination variation are coupled in the
-sense that if one problem is solved, the other problem becomes a lot easier. But solving them
-separately is not sufficient for solving the face recognition problem.
-
-\item If the data fits our model perfectly, at the correct alignment the only
-corruptions are due to occlusions.  Until the images are perfectly aligned,
-there is indeed a component of dense error caused by misalignment.  This dense
-error component is indeed largest where the image gradient is high (or even
-discontinuous), and the derivative of this error component is what drives the
-algorithm to its solution.  We do explicitly  blur and downsample the images to
-mollify registration effects; multiscale alignment is a well-known technique,
-and is not new to our paper.
-
-\item We have removed the reference to scattered light from Section III-A, as there appear to
-be several common uses of the term.  That said, it is true that linear illumination
-models are physically motivated for scenes far more general than for convex
-Lambertian objects.  Maxwell's equations in free space are linear, and materials
-that create non-linear boundary conditions are the exception
-rather than the rule.  Without proportionality between radiance and irradiance
-the concept of a Bidirectional Reflectance Distribution Function and its
-generalizations are not even meaningful.
-
-\item Rigorous justification of the choice of illuminations is a subject of
-continuing research.  Our current choice is motivated empirically by the
-experiments described in the paper.
-
-\item To our knowledge, we are the first to use projectors for illumination in
-face recognition applications.  Section I-A of the previous submission already
-cites several other applications that use projectors as illumination sources.
-The requested references propose a technique for improving the SNR of projector-based
-illumination systems for certain classes of objects.  We have added
-discussion of this to the paper in Section III-B.
-
-\item  It is true that one or both stages of the algorithm could be substituted
-with other techniques.  That said, the registration and recognition stages are
-coherent in the sense that they are motivated by same modeling assumptions
-(illumination model, sparse occlusion, etc.), they use the same robust penalty
-on image error (the L1-norm), and they use the same numerical optimization
-techniques (ALM).  While conceptual simplicity is a secondary concern in
-algorithm design, starting with a minimalist foundation benefits both
-practitioners as well as researchers.  Other researchers have already published
-more complex extensions to our work.
-
-\item  There is no extensive computation involved in
-    Step 13 of Algorithm 1. In fact, for each candidate
-    subject $i\in {k_1, \ldots, k_s}$ for recognition, Step
-    13 simply uses the transformation $\tau_i$ obtained
-    from the iterative alignment stage (Step 2-9) to
-    re-sample each original training
-    image of that subject, so that this new set of training
-    samples is aligned with the test sample. Nevertheless,
-    it worth mentioning that, since we perform alignment
-    separately for each class in the gallery, the
-    computation for the alignment stage (Step 2-9) is
-    linear in the size of the gallery. However, thanks to
-    its separability, the alignment stage can be easily
-    parallelized.
-
-    The variation in alignment across classes is indeed
-    small, but still changes the objective enough to affect
-    recognition results. Aligning to a generic face (or a
-    gallery based on ``aligned'' images of many people) is
-    a good idea, and one we have considered as well.  If
-    nothing else, it could be used to improve alignment
-    initialization instead of using raw face detector
-    output.  It may also be useful to use the roughly
-    aligned test image to quickly eliminate some of the
-    classes from further consideration.  Again, we have
-    deliberately tried to keep the algorithm as minimalistic
-    as possible for this publication.
-
-\item The adaptation of the Lucas-Kanade algorithm to both the L1 norm cost
-function and a linear illumination model is novel to our knowledge.  As
-discussed earlier, we disagree with Reviewer 1's contention that data
-acquisition is not a critical part of the face recognition problem -- careful (training)
-data acquisition has been critical for almost all other biometric identification
-systems (e.g.\ fingerprint and iris)! Why should face recognition be different?
-
-We believe we have demonstrated that by using a well-thought-out combination of existing ideas
-(iterative image alignment, L1-error function, SRC, using projectors for
-illumination), one can build a recognition system that is conceptually simple, well
-motivated, and competitive with state-of-the-art recognition systems for access
-control scenarios.
-
-\end{enumerate}
-
-\section{Response to Reviewer \#2}
-Reviewer 2 raises five questions, which we will now address.  Since the
-recognition optimization is convex, and convergence is therefore not an issue, we assume
-that (1)-(3) are asking about the iterative alignment scheme.
-
-\begin{enumerate}
-\item (i) ``How often an infinite loop is encountered, and how it is handled."
-
-We place a hard limit on the maximum number of alignment iterations, so the
-algorithm is guaranteed to terminate in finite time.  In practice, the
-alignment algorithm tends to converge within 4-10 iterations for each
-resolution of the multi-scale implementation. This low number of iterations is expected, since the iterative alignment scheme is locally quadratically convergent. We also place hard bounds on how
-large the deformation can be, and terminate alignment early if these bounds are
-reached.  In the cases where the alignment ``slips off'' the face, this is how
-alignment terminates.
-
-\item (ii) ``How often the starting point happens to be located in a wrong basin of
-attraction, i.e. a wrong face. (If there is any way to know this.)''
-
-Since we align the test image to the training users individually, we avoid the
-potential problem of the test image converging to a local minimum corresponding
-to a wrong face.
-
-\item (iii) ``How often the solution gets drifted away into a wrong basin of
-attraction. (If there is any way to know this.)''
-
-In Section II-E, we have conducted careful experiments to measure the region of attraction
-for the proposed alignment method. We have studied the region of attraction of the proposed
-method for both 2D and 3D pose variation, as well as how it can be improved with a proper
-multi-scale implementation.
-
-\item (iv) ``What is the criteria or threshold for convergence? If the threshold is
-stringent, the algorithms are likely to get stuck in infinite loops, and if the
-threshold is liberal, premature convergence would occur. How are such
-thresholds picked and how universal are they? Does every new dataset need its
-own threshold?''
-
-For the $\ell_1$ minimizations we terminate the ALM
-algorithm when the relative changes of primal variables in
-two consecutive iterations become small, as suggested by
-[27]. That is,
-\begin{displaymath}
-\frac{\|\x_{k+1}-\x_k\|_2}{\|\x_k\|_2} < \epsilon \quad
-\textup{and} \quad \frac{\|\e_{k+1}-\e_k\|_2}{\|\e_k\|_2} <
-\epsilon,
-\end{displaymath}
-where $\epsilon>0$ is the tolerance. With the suggested
-criteria, the $\ell_1$ minimizations are guaranteed to converge
-eventually, although convergence may be very slow if the constants
-are poorly tuned. For the alignment
-iteration, we terminate when the change of objective value
-(\ie $\|\e\|_1$) in two consecutive iterations becomes
-small. We use the same set of convergence thresholds for
-all the data sets we ran against.
-
-\item (v) ``The problem, however, is that if we minimize a function one variable at
-a time, we may actually end up increasing the function $(L_{\mu}(x,e,\lambda_{k}))$ value.''
-
-This may be true for general nonlinear optimization problem. However for convex problems
-we are solving here, the alternating direction methods have been proven to converge, see
-[27] and related work by the same authors.
-
-\end{enumerate}
-
-\section{Response to Reviewer \#3}
-Reviewer 3 makes the following requests:
-
-\begin{enumerate}
-\item ``I would have liked to see the experiment for the case S=1, without
-impostors in the dataset, to judge the effect of the ``Sparse
-Concentration Index''.
-
-We have added experiment results to Table I demonstrating the utility of the
-sparse representation(SR) based recognition stage for classification.  In
-general, we see a few percentage points of improvement in recognition rate when
-using the representation residual obtained from the SR-based recognition stage
-(with $S=10$) over the recognition based purely on alignment error (with
-$S$=1).  While this effect is desirable, as discussed in the paper, it is
-worthy emphasizing that the SR-based recognition also enables us to compute
-SCI, which has demonstrated its power for impostor rejection.
-
-\item ``I would also retitle the paper to include ``robust measure''
-instead of ``sparse representation", as the L1 Cost is just a robust
-cost, and the system would probably work as well or better with other
-(even non-convex) robust cost functions.''
-
-We hope reviewer will find our above justification for the (Sparse
-Representation based) recognition convincing.  If so, the reviewer may
-therefore also agree that the current title is indeed appropriate.
-
-\item ``It might be possible to streamline the introduction a bit more, such that it
-clearly describes the algorithm before going into the details.''
-
-We have attempted to make the overall structure of our system clear earlier in the
-paper, by clarifying early on that SRC forms the recognition stage of our system
-in the introduction, as well as by making it clear in the caption of the motivating example
-is not the final pipeline.
-
-\item ``Using and requiring several input images of a person asks for a
-comment on the advantage or difference to a system that uses
-a 3D representation of a face. Please comment on this!"
-
-The primary advantage of systems that internally use a 3D model of human faces
-is that they have the potential to work over a wider range of pose variation in
-the test image.  In general, systems using 2D models are simpler and easier to
-engineer.  As far as data acquisition is concerned, 2D cameras still have a far
-higher resolution and lower noise levels than 3D cameras.  Based on the principles
-of photometric stereo, our set of 2D images may very well encode more useful shape
-information about a user's face than a single low-resolution depth image, especially if
-recognition is to be performed on 2D input images.
-
-In addition, the sparse representation techniques are not limited to frontal 2D images.
-They can be easily applied to 2D images from multiple poses or to establish alignment
-between a 2D image and a renderings from a 3D face model (if available). These are all good directions
-for future research.
-
-We have modified Section I-A and the Conclusion to address these issues.
-
-\end{enumerate}
+{\bf
+	Some comments on this would be useful to better understand the results shown in table 1 and 2.
+}
+As already discussed, the discussion of these results has been updated.
 
 \end{document}
- a user's face than a single low-resolution depth image, especially if
-recognition is to be performed on 2D input images.
-
-In addition, the sparse representation techniques are not limited to frontal 2D images.
-They can be

### 2011-03-29 faces_system.tex
Just a couple small changes.
fixed typo, added rebuttal

@@
 field of face recognition research.  However, there are many face recognition
-that fall roughly between these extremes, where very high recognition
+applications that fall roughly between these extremes, where very high recognition
 performance is desired, but the users in the gallery are still allies of the
@@
 challenging due to the large number of subjects, and due to
-natural variation in subject appearance over time. Table
-\ref{tab:MultiPIE-recognition2} shows the result of our
+natural variation in subject appearance over time. 
+Table \ref{tab:MultiPIE-recognition2} shows the result of our
 algorithm on each of the 3 testing sessions. Our algorithm
@@
 alignment: 1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye corners, and 3.\
+2.\ alignment via manually selected outer eye corners,
+\footnote{Two manually clicked points are sufficient to define
+a similarity transformation.  All of the experiments in this section are carried out with
+similarity transformations.}
+ and 3.\
 the output of our iterative alignment algorithm. The
@@
 when only one sample per person is used for training
-\cite{Tan06facerecognition}. In this paper, we follow the same
+\cite{Tan06facerecognition}. 
+In this paper, we follow the same
 steps as in \cite{Ahonen2006-PAMI} to construct an LBP
@@
 misalignment, it still benefits from using the output of our
-iterative alignment algorithm as the input. In addition, like
+iterative alignment algorithm as the input. 
+In addition, like
 the other classical algorithms, the performance of LBP degrades
@@
 the entire database is much more difficult than deciding if two
-face images are the same subject. Figure 8 should not be
+face images are the same subject. Figure \ref{fig:roc-multipie} should not be
 confused with typical ROC curves for face similarity, e.g.,
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
-\vspace{-.2in}
+\vspace{-.0in}
  \caption{{\bf Representative examples of category C4}.
@@
 Using a well-though-out combination of existing ideas
-(iterative image alignment, l1-error function, SRC, using projectors for
+(iterative image alignment, $\ell_1$-error function, SRC, using projectors for
 illumination), we have proposed a system for recognizing human faces

### 2011-03-31 rebuttal.tex
More numbers from Zihan

@@
 Changes have been made to the paper to address all of the changes requested by Reviewer 1,
-including a more complete set of results in Table 2 and Figure 11, and associated 
+including a more complete set of results in Table 2, Table 3 and Figure 11, and associated
 discussions.  We give a quick overview of the changes.
@@
 }
-The typo on Page 1 is now ammended.
+
+The typo on Page 1 is now amended.
 
@@
 }
+
 We ran the requested experiments.  We agree with the reviewer that automatic registration
 can often out-perform manual registration.  The results back this up, and the text
-is ammended to be more clear on this point.  The new more complete benchmark is shown in
+is amended to be more clear on this point.  The new more complete benchmark is shown in
 Table 2, and Figure 11.  All of the classical methods benefit from using our iterative
-alignment procudeure.  This is even true for LBP, which is known to have some robustness
+alignment procedure.  This is even true for LBP, which is known to have some robustness
 to moderate misalignment.  The discussions have been updated to cover the new results.
@@
 }
+
 Similarity transformations are used for all of the experiments in this section, and
@@
 }
-As already discussed, the discussion of these results has been updated.
+
+The discussion of these results (now Table 2 and 3 in the new
+draft) has been updated. For the single gallery image FERET
+experiment (Table 3), since our algorithm is not designed for
+such scenario, it fails when the illumination changes and other
+variations seriously violate its assumption. We found that LBP
+with manual alignment achieves the best performance in this
+experiment.
 

### 2011-03-31 faces_system.tex
More numbers from Zihan

@@
 alignment: 1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye corners,
-\footnote{Two manually clicked points are sufficient to define
-a similarity transformation.  All of the experiments in this section are carried out with
-similarity transformations.}
- and 3.\
-the output of our iterative alignment algorithm. The
+2.\ alignment via manually selected outer eye
+corners,\footnote{Two manually clicked points are sufficient to
+define a similarity transformation. All of the experiments in
+this section are carried out with similarity transformations.}
+and 3.\ the output of our iterative alignment algorithm. The
 performance drop of the LDA algorithm on Multi-PIE reported
@@
 \end{table}
+
+%\begin{table}
+%\caption{Recognition rates on the Multi-PIE database for
+%different pairings of alignment and recognition stages.}
+%\vspace{-.2in} \centerline{
+%\begin{tabular}{|c|c|c|c|c| } \hline
+%Recognition rate & Session 2 & Session 3 & Session 4  \\
+%\hline
+%LDA($d/m/i$) & 5.1 / 49.4 / 91.0\% & 5.9 / 44.3 / 89.9\% & 4.3 / 47.9 / 88.1\% \\
+%\hline
+%NN($d/m/i$) & 26.4 / 67.3 / 73.5\% & 24.7 / 66.2 / 69.6\% & 21.9 / 62.8 / 69.3\% \\
+%\hline
+%NS($d/m/i$) & 30.8 / 77.6 / 84.5\% & 29.4 / 74.3 / 82.3\% & 24.6 / 73.4 / 81.4\% \\
+%\hline
+%LBP($d/m/i$) & 39.9 / 93.3 / {\bf 95.2}\% & 38.1 / 91.2 / {\bf 94.7}\% & 33.9 / 92.9 / {\bf 93.5}\% \\
+%\hline
+%Algorithm 1 & 93.9\% & 93.8\% & 92.3\% \\
+%\hline
+%\end{tabular}
+%\label{tab:MultiPIE-recognition} }\vspace{-.4in}
+%\end{table}
 
@@
 our algorithm on the four test sets, with input directly
-obtained from the Viola and Jones' detector, and the
-performance of LBP with input obtained from both face detector
-and manual alignment using the eye coordinates.
+obtained from the Viola and Jones' detector. We also report the
+performance of LBP with the same three types of input as before
+in Table~\ref{tab:FERET-recognition}, where we use letters
+``$d$'', ``$m$'' and ``$i$'' to indicate face detector, manual
+alignment and our iterative alignment algorithm, respectively.
 
@@
 \hline
-LBP$_d$ (LBP$_m$) & 54.8 (96.6)\% & 10.3 (58.8)\% & 29.8 (71.6)\% & 19.8 (61.5)\% \\
+LBP($d/m/i$) & 54.8 / {\bf 96.6} / 94.5 \% & 10.3 / {\bf 58.8} / 42.8\% & 29.8 / {\bf 71.6} / 46.5\% & 19.8 / {\bf 61.5} / 21.1\% \\
 \hline
@@
 `fb', in which the illumination is similar to the training and
-the mere variation in facial expression is handled well by the sparse error
-model. For the other three test sets, our algorithm fails because the illumination
-changes and other variations seriously violate the assumptions of our method.
-On the other hand, while LBP achieves the best recognition rates given
+the mere variation in facial expression is handled well by the
+sparse error model. For the other three test sets, our
+algorithm fails because the illumination changes and other
+variations seriously violate the assumptions of our method.
+This also explains why LBP performs worse with our iterative
+alignment algorithm, compared to manual alignment. On the other
+hand, while LBP achieves the best recognition rates given
 manually aligned training and testing samples, its performance

### 2011-03-31 faces_system.tex
Minor change to make a sentence read better.

@@
 
-%\begin{table}
-%\caption{Recognition rates on the Multi-PIE database for
-%different pairings of alignment and recognition stages.}
-%\vspace{-.2in} \centerline{
-%\begin{tabular}{|c|c|c|c|c| } \hline
-%Recognition rate & Session 2 & Session 3 & Session 4  \\
-%\hline
-%LDA($d/m/i$) & 5.1 / 49.4 / 91.0\% & 5.9 / 44.3 / 89.9\% & 4.3 / 47.9 / 88.1\% \\
-%\hline
-%NN($d/m/i$) & 26.4 / 67.3 / 73.5\% & 24.7 / 66.2 / 69.6\% & 21.9 / 62.8 / 69.3\% \\
-%\hline
-%NS($d/m/i$) & 30.8 / 77.6 / 84.5\% & 29.4 / 74.3 / 82.3\% & 24.6 / 73.4 / 81.4\% \\
-%\hline
-%LBP($d/m/i$) & 39.9 / 93.3 / {\bf 95.2}\% & 38.1 / 91.2 / {\bf 94.7}\% & 33.9 / 92.9 / {\bf 93.5}\% \\
-%\hline
-%Algorithm 1 & 93.9\% & 93.8\% & 92.3\% \\
-%\hline
-%\end{tabular}
-%\label{tab:MultiPIE-recognition} }\vspace{-.4in}
-%\end{table}
-
 % LBP
@@
 obtained from the Viola and Jones' detector. We also report the
-performance of LBP with the same three types of input as before
-in Table~\ref{tab:FERET-recognition}, where we use letters
-``$d$'', ``$m$'' and ``$i$'' to indicate face detector, manual
-alignment and our iterative alignment algorithm, respectively.
+performance of LBP with the same three types of input as before:
+we use letters
+``$d$'', ``$m$'', and ``$i$'' to indicate face detector, manual
+alignment, and our iterative alignment algorithm, respectively.
 

### 2011-03-31 rebuttal.tex
Fixed a sentence in the rebuttal

@@
 Changes have been made to the paper to address all of the changes requested by Reviewer 1,
-including a more complete set of results in Table 2, Table 3 and Figure 11, and associated
+including a more complete set of results in Table 2, Table 3, Figure 11, and associated
 discussions.  We give a quick overview of the changes.
@@
 
-The discussion of these results (now Table 2 and 3 in the new
+The discussion of these results (seen in Table 2 and 3 in the new
 draft) has been updated. For the single gallery image FERET
-experiment (Table 3), since our algorithm is not designed for
-such scenario, it fails when the illumination changes and other
-variations seriously violate its assumption. We found that LBP
+experiment (Table 3), our algorithm is not designed for
+this scenario, and it fails when the illumination changes and other
+variations seriously violate its assumptions (this was true in the previous
+revision as well). We found that LBP
 with manual alignment achieves the best performance in this

### 2011-03-31 faces_system.tex
Fixed spacing

@@
 
+\vspace{-.2in}
 \subsection{Recognition with Pose and Expression} We now run tests of
@@
 
+\vspace{-.1in}
 \section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
@@
 
+\vspace{-.1in}
 \subsection{Necessity of Rear Illuminations} To see how
@@
 are insufficient for representing test images under typical indoor and
-outdoor illuminations.\vspace{-0.2in}
+outdoor illuminations.
+\vspace{-0.1in}
 
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.5in}
+\vspace{-.4in}
 \end{figure}
@@
 \label{fig:occ-block}
-\vspace{-.5in}
+\vspace{-.3in}
 \end{figure}

### 2011-03-31 rebuttal.tex
Fixed spacing

@@
 
+\vspace{-.2in}
 We again thank the reviewers and the area chair for their comments.
@@
 	2 points are manually selected, then what kind of registration can be done
-	with just 2 corresponding points ? Surely not affine alignment. Many more are
+	with just 2 corresponding points? Surely not affine alignment. Many more are
 	needed if it is to be done robustly. So it is not surprising that manual
 	alignment is giving worse results than automatic. If my interpretation is
-	wrong, is the entire affine registration being done manually ? If so, as
+	wrong, is the entire affine registration being done manually? If so, as
 	mentioned above the accuracy of the human can be bad.

### 2011-03-31 faces_system.tex

@@
 \begin{tabular}{ccccc}
-\includegraphics[height=1in]{figures_cvpr/5} &
-\includegraphics[height=1in]{figures_cvpr/7} &
-\includegraphics[height=1in]{figures_cvpr/09} &
-\includegraphics[height=1in]{figures_cvpr/11} &
-\includegraphics[height=1in]{figures_cvpr/13} \vspace{-.1in} \\
+\includegraphics[height=0.55in]{figures_cvpr/5} &
+\includegraphics[height=0.55in]{figures_cvpr/7} &
+\includegraphics[height=0.55in]{figures_cvpr/09} &
+\includegraphics[height=0.55in]{figures_cvpr/11} &
+\includegraphics[height=0.55in]{figures_cvpr/13} \vspace{0mm} \\ 
 (a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=1in]{figures_cvpr/15} &
-\includegraphics[height=1in]{figures_cvpr/17} &
-\includegraphics[height=1in]{figures_cvpr/19} &
-\includegraphics[height=1in]{figures_cvpr/21} &
-\includegraphics[height=1in]{figures_cvpr/3} \vspace{-.1in} \\
-(f) & (g) & (h) & (i) & (j) \vspace{-.1in}
+\includegraphics[height=0.55in]{figures_cvpr/15} &
+\includegraphics[height=0.55in]{figures_cvpr/17} &
+\includegraphics[height=0.55in]{figures_cvpr/19} &
+\includegraphics[height=0.55in]{figures_cvpr/21} &
+\includegraphics[height=0.55in]{figures_cvpr/3} \vspace{0mm} \\ 
+(f) & (g) & (h) & (i) & (j) \vspace{0mm}
 \end{tabular}
@@
 \caption{Recognition rates on the Multi-PIE database for
-Algorithm 1 and \cite{Yang2010-CVPR}} \vspace{-.2in}
+Algorithm 1 and \cite{Yang2010-CVPR}}
 \centerline{
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
-\vspace{-.25in}
+\vspace{0mm}
 \caption{{\bf Recognition under varying level of
@@
 \label{fig:multipie-occ-rec}
-\vspace{-.4in}
+\vspace{0mm}
 \end{figure}
@@
 \includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
-\vspace{-.0in}
+\vspace{0mm}
  \caption{{\bf Representative examples of category C4}.
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.4in}
+\vspace{0mm}
 \end{figure}
@@
 \label{fig:occ-block}
-\vspace{-.3in}
+\vspace{0mm}
 \end{figure}

### 2011-04-01 faces_system.tex
Removed dead code

@@
 in Section \ref{sec:multipie}.
-%The strength of the
-%filter depends on how much the warping downsamples the image.
-%Although the scale component of the warping may vary with each
-%step of the iterative alignment, it is safe in practice to
-%choose a fixed conservative smoothing filter (erring slightly
-%on the side of more smoothing). Whenever a transformation
-%operator is not shown for a vector in our notation (i.e. for
-%the columns of $A$ or $J$), it means that a suitable
-%transformation has already been applied to a larger image
-%(otherwise, the dimensions of the arrays would not agree).
 
@@
 14 for global sparse representation.
-%The most classical
-%solution to these $\ell^1$-minimization problems is the so
-%called \emph{primal-dual interior-point} method
-%\cite{BoydS2004}. Conceptually, it operates in an iterative
-%fashion, and a linear system is solved at each iteration.
-%Therefore, its complexity per iteration is at least $O(n^3)$,
-%where $n$ is the total number of training images for the face
-%recognition problem. Meanwhile, it is also observed in
-%\cite{YangA2010-pp} that its computational cost with respect to
-%the dimension of each sample, $m$, grows much fast than other
-%alternative methods. Hence for real-world applications, as the
-%image resolution and number of classes grow, this method will
-%soon become too slow to be useful.
 Fortunately, many fast algorithms for solving these problems have been proposed over the
 past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
-%We refer the interested readers to
-%\cite{YangA2010-pp} for a comprehensive survey in the
-%literature and a detailed study of the performance of five
-%representative approaches on solving the two $\ell^1$-minimization
-%problems in Step 6 and 14 of
-%Algorithm~\ref{alg:deformable-src}.
 That work suggests that \emph{Augmented Lagrange Multiplier}
@@
 higher accuracy with a fixed computational budget.
-%method, is always
-%among the most accurate and fastest algorithms and achieves
-%better performance in both effectiveness and efficiency
-%compared to other first-order methods.
-%Most importantly, as first-order methods, they require only
-%lightweight vector operations and matrix-vector
-%multiplications, contrast to many other methods, such as
-%primal-dual interior-point method and the homotopy method
-%\cite{DonohoD2006-pp}, which all involve expensive operations
-%such as solving a linear system. For the rest of this section,
-%we briefly describe the ALM method.
 
@@
 
-%
-%Note
-%that, however, Algorithm~\ref{alg:alm} differs from
-%\cite{YangJ2009-pp} mainly in the way that the sub-problem
-%associated with $\x$ is solved. In \cite{YangJ2009-pp}, only an
-%approximation to the solution is obtained by just computing one
-%iteration of the Conjugate Gradient algorithm. As pointed out
-%by \cite{YangA2010-pp}, although this further reduces the
-%complexity of the algorithm, it converges very slowly when the
-%condition number of the matrix $A^TA$ is large, which is indeed
-%the case for the face recognition problem.
-
 We have selected this algorithm because it strikes the best
@@
 \noindent{2) \em Multiscale Implementation.}
-        %The
-    %gain here is a reduction in the computational cost. We
-    %briefly explain how this works here. In our system, a
-    %truncated Gaussian kernel is used for blurring. In
-    %order to ensure accurate operation, we change the
-    %window size of the Gaussian kernel in proportion to its
-    %standard deviation. Consequently, achieving a large
-    %amount of blur requires a large convolution
-    %mask\footnote{It is possible to apply a low-bandwidth
-    %Gaussian kernel multiple times to achieve the same
-    %effect of a single high-bandwidth kernel. However, the
-    %time required will be at least as great as performing
-    %the single large blur.} and hence is computationally
-    %expensive.
-    %To avoid using a large window, we observe that, by
-    %down-sampling the image, a smaller blur kernel can
-    %achieve similar blurring effect. However, working at a
-    %lower resolution degrades the alignment precision.
-    %Therefore, we use a multi-scale framework to achieve
-    %computational efficiency and high precision at the same
-    %time. It starts from the smallest scale with largest
-%blur. The strong blur can handle large misalignment, but
-%the low resolution results in poor precision. We then use
-%the estimated alignment to initialize a similar process
-%that works at a higher resolution and with smaller blur.
-%This process repeats and thus the solution is computed
-%progressively.
 Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost.
 Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
-%To see the relationship between number of scales and the
-%standard deviation $\sigma$ of the Gaussian kernel applied
-%between scales,
 We use the same 7 illuminations in
@@
     initialize our alignment algorithm.
-%whenever it succeeds. For extreme poses where the detector fails (usually for poses $> 45^\circ$), we use manual initialization by putting a reasonable square window around the face.
 Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
@@
  pose variation is beyond what we intend to handle with our formulation.
-%While we don't intend to perform 2D recognition with this much pose,
-%the alignment is at least stable.
 \begin{figure}
@@
 
-% % META: Most of this is about understanding the Basi Result.  For the Paper, don't try to derive the result; just state it.
-% % Derivation of linearity using strong assumptions
-% First, let us assume that
-% the illumination if the subject's face is distant.  This model
-% will be a good approximation as long as distance to the nearest
-% light source is much larger than the person's face.  If we
-% further assume that the object is convex so that there is no
-% self-shadowing, the illumination incident on a surface patch
-% will depend only on the surface normal, and not on the location on the surface.
-% Furthermore, if the object is Lambertian, the image intensity will not depend on the angle between patch surface normal and the sensor, or on the depth of the patch.
-Since each pixel in the image corresponds to
-% some patch on the object, the vector of image intensities is a
-% linear function of the radiance of the corresponding portions
-% of the object.
-% % Motivation of low dimensionality from spherical harmonics, low-pass interpretation
-% Consider the vector space of all illuminations
-% of the object.  These are just positive functions (or more
-% generally, distributions) defined on the sphere, where each
-% point on the sphere corresponds to a direction of incoming
-% light.  Consider the vector space formed by the light exiting
-% the object.  This too is a positive function defined on the
-% sphere, where now each point on the sphere corresponds to the
-% normal of a patch on the object (or any other patch with the
-% same normal).  Under the Lambertian assumption, the radiance of
-% a patch depends on the cosine of the incident angle of incoming
-% light, and is integrated over the half-sphere of illumination
-% that the patch can "see". Thus, the surface of the object acts
-% on the sphere in a manner very similar to a half-cosine
-% approximation to a low-pass filter for a one-dimensional
-% signal.  Thus the energy leaving the object is
-% disproportionately concentrated in the subspace corresponding
-% to low spatial frequencies.  Since the image is a linear
-% function of this, the space of images of an object will also
-% tend to fall on (the positive portion of) a subspace.   In
-% \cite{Basri2003-PAMI}, Basri showed using spherical harmonic
-% basis functions that nine basis illuminations (corresponding to
-% the lowest frequency spherical harmonics) result in training
-% images that do a good job of linearly interpolating all other
-% training images.  It should be noted, however, that these basis
-% illuminations are not strictly positive, and thus neither are
-% the training images they generate (when rendered on a computer
-% that can handle negative illumination).  \footnote{Note that
-% while the spherical harmonic basis functions have regions where
-% they are negative, this does no necessary preclude this
-% theoretical result from being put to use in practice.  If the
-% geometry of the training illumination system were well
-% calibrated, it would be possible to generate approximations of
-% the positive and (rectified) negative components of the basis
-% illuminations separately, taking the difference of the
-% resulting two images, and storing it using a signed datatype.
-% Combinations of these images may not be positive, but they may
-% still be good enough to be useful.}
-
 The strongest theoretical results so far regarding the relationship
@@
 The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
-%Under those assumptions, the incident and reflected light are represented using distributions on a sphere and can thus be studied in a spherical harmonic basis.
-%The Lambertian nature of the object acts as a low-pass filter between the incident and reflected light distributions,
-%and as a result, the set of images of the object end up lying very
-%close to the subspace corresponding to low frequency spherical harmonics.\footnote{This is highly condensed; please refer to \cite{Basri2003-PAMI} for a  complete description of this result.}
 all images of the object can be well approximated by linear combinations of
@@
 systems is misguided for several reasons.
-% Non-physicality
-%Neither the basis illuminations, nor the basis images used
-%in the Basri analysis are physical.  Like Fourier bases, they have negative components,
-%and are thus not, at least directly, physically realizable.
-% This alone is not
-% a showstopper, however: one could, for example, decompose the basis images into
-% their positive and negative parts and use them separately.  If the images
-% are being rendered from a 3D model, ``negative'' light can be permitted.
-%There are ways to engineer around this difficulty\footnote{such as allowing negative light while rendering from a 3D model, decomposing mixed-sign illuminations and images into positive negative parts, and enforcing positivity only in the output image}, but the photometric and geometric assumptions behind this model limit its applicability even more severely.
-%Studio photographers and movie-makers
-%go to great lengths to control their effects through makeup and lighting.
-% Not only does specularity allow some of the high angular frequency components of the
-% incident illumination to affect the outgoing illumination, it even prevents
-% the relationship between incoming and outgoing light from even being expressable as a convolution.
-% Indeed, in the extreme case of an image of a mirror-surface sphere, the image is basically a warped
-% version of the sphere's incident illumination, and there is no reason to believe the set of
-% images is any lower dimensional than the set of illuminations.
 Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
 and they all do so in a way that violates the modeling assumptions of the Basri analysis.
-%Specifically, self-shadowing and inter-reflection break the simple relationship between the outgoing light
-%and the resulting images, and specularity even invalidates the representation of outgoing light that is used.
-% itself,
-%since it is no longer independent of viewing angle.
-%%If the surface has a specular component, the relationship between incoming and outgoing light
-%is no longer expressible as a convolution, and if there is self-shadowing or inter-reflection the illumination
-%of each patch will depend not only on its location on the surface, but on the illumination of other pixels in the image.
 
@@
 
-%Assuming that no pixels are shared by multiple illuminations is important;
-%for example, any other non-negative basis illuminations formed from positive linear
-%combinations of our basis illuminations would span a cone of lower
-%volume, and thus the resulting training images might not be
-%able to represent as wide a variety of testing images without using
-%negative coefficients in $\x$.
-%\footnote{This is true if $\x>0$ is enforced;
-%the effect of negative coefficients in $\x$ is still under study.}
-
 Assuming that the sequentially illuminated regions do not overlap results in a
@@
     representation error becomes, with diminishing returns.
-%\end{itemize}
-%\begin{itemize}
 \item {\em Granularity Experiment.} In the second
@@
 \end{figure}
-
-%\subsection{Should Positivity in the Representation Coefficients
-%be Enforced?} One critical issue in linear illumination models
-%is whether to enforce nonnegativity in the coefficients $\x$,
-%i.e. whether to model illumination using a cone or a subspace.
-%Some authors, i.e. \cite{Basri2003-PAMI}, have chosen to allow
-%negative components in the coefficient vector $\x$ when
-%representing one image as a sum of other images, and only
-%enforce that the linear combination $A\x$ be positive
-%everywhere.  Unfortunately, for non-convex objects such as
-%faces, enforcing non-negativity of the representation is not
-%sufficient to guarantee that the resulting image of the object
-%is physical.  For example, consider a Lambertian planar object
-%with a cylindrical hole drilled in it that is aligned with the
-%optical axis of an orthographic camera.  Consider two images:
-%one under illumination from a distant point source aligned with
-%the hole, and one off of the hole axis.  Positively weighting
-%the on-axis image and negatively weighting the on-axis image
-%can result in an image where the bottom of the hole is more
-%strongly illuminated than the surrounding plane.  The image is
-%positive, but it is clearly non-physical.  It is worth
-%emphasizing that this phenomenon is a result of some points on
-%an object seeing a restricted subset of the illumination sphere
-%(a violation of convexity),  and not a result of the
-%directionality of the light chosen in the thought experiment.
-%While the number of coefficients that actually end up negative
-%after optimization is usually very small, we have observed
-%pathological cases where the shadows next to the user's nose in
-%the reconstructed image invert in value.
-
-%Instead of using a smaller set of training images, allowing the
-%coefficients to go negative to try to represent more test
-%illuminations, and risking over-fitting, we decided to enforce
-%non-negativity of the coefficients.  Nonnegative combinations
-%of images are guaranteed to correspond to physically plausible
-%images.  all physical illuminations unless the training images
-%actually span the boundary of the illumination cone. Because we
-%have a flexible acquisition system, we can directly generate a
-%set of illuminations that span most of the illumination cone,
-%without resorting to negative coefficients and risking
-%overfitting.  Thus, in Algorithm 1, we have enforced that the
-%coefficients $\x$ be non-negative.
-
-%One of the main factors that complicates face recognition, and computer vision in general, is that two images of the same person's face can be very different, even if the pose is carefully controlled and all points on the person's face are visible in both images.
-The most safe, reliable, and general assumption that can be made about the imaging process is that it is linear with respect to the intensity of different light sources.
-The set of images of a given object is therefore a cone in the non-negative orthant of the pixel basis.
-All recognition algorithms that rely on vision for training rely on images sampled from this cone.
 
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-% Use this command to get the appendices' numbers in "A", "B" instead of the
-% default capitalized Roman numerals ("I", "II", etc.).
-% However, the capital letter form may result in awkward subsection numbers
-% (such as "A-A"). Capitalized Roman numerals are the default.
-%\useRomanappendicesfalse
-%
-%\appendices
-%\section{Proof of the First Zonklar Equation}
-%Appendix one text goes here.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-%\section{}
-%Appendix two text goes here.
-
-% use section* for acknowledgement
 \section*{Acknowledgment}
-% optional entry into table of contents (if used)
-%\addcontentsline{toc}{section}{Acknowledgment}
 This work was supported by grants NSF IIS 08-49292, NSF ECCS
@@
 
-% references section
-% NOTE: BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% standard IEEE bibliography style from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
 \bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,faces}
 \bibliography{faces}
-
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
 

### 2011-04-01 faces_system.tex
Replaced shrink with \mathcal{T}
Figure 9 fits better.

@@
 \begin{equation}
-\textup{shrink}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
+\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
 \end{equation}
@@
 \begin{equation}
-\e_{k+1} = \textup{shrink}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
+\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
 \end{equation}
@@
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
+\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
 \frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
@@
 \WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
+\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
 \frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
@@
 
+\newcommand{\tempwidth}[0]{0.6in}
 \begin{figure}
-\centering
 {\small
-\begin{tabular}{cccccc}
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{0mm} \\
+\begin{tabular}{c@{}c@{}c@{}c@{}c@{}c}
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
+\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
 (a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
@@
 \label{fig:failed-examples}
-\vspace{0mm}
+\vspace{-15mm}
 \end{figure}

### 2011-04-01 faces_system.tex
Got more figures to fit.

@@
 
-\newcommand{\tempwidth}[0]{0.6in}
+\newcommand{\tempwidth}[0]{0.58in}
 \begin{figure}
+\centering
 {\small
-\begin{tabular}{c@{}c@{}c@{}c@{}c@{}c}
+\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
 \includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
@@
 
-\begin{table}
+\begin{table*}
+\centering
 \caption{Recognition rates on the Multi-PIE database for
 different pairings of alignment and recognition stages.}
-\vspace{-.2in} \centerline{
+\centerline{
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
@@
 \end{tabular}
-\label{tab:MultiPIE-recognition} } \vspace{-.4in}
-\end{table}
+\label{tab:MultiPIE-recognition} } 
+\vspace{-.3in}
+\end{table*}
 
@@
 
+\renewcommand{\tempwidth}[0]{1.85in}
 \begin{figure}[t]
+\vspace{.2in}
 \centerline{
-	\begin{tabular}{cc}
-	\includegraphics[width=3in]{figures_pami/pami_roc_revision2.png} &
-	\includegraphics[width=3in]{figures_pami/pami_roc2.png} \\
-	{\small All algorithm with iterative alignment} & {\small Classical algorithms with manual alignment}
-\\
+	\begin{tabular}{@{}c@{}c@{}}
+	\includegraphics[trim=.9in 0in 1.2in 0in, clip, width=\tempwidth]{figures_pami/pami_roc_revision2} &
+	\includegraphics[trim=.9in 0in 1.2in 0in, clip, width=\tempwidth]{figures_pami/pami_roc2} \\
+	{\small (a)} & {\small (b)} \\
 	\end{tabular}
 }
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database, (a) with iterative alignment,
-and (b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
+\caption{{\bf ROC curves} for subject validation on Multi-PIE database, 
+(a) for all algorithms with iterative alignment, and 
+(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
 \vspace{-.4in}
@@
 
+\newcommand{\tempscale}{0.45}
 \begin{figure}
 \centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
 \vspace{2mm}
 \centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
 \vspace{2mm}
 \centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG}
 \vspace{-.2in}

### 2011-04-01 cvpr09_faces.tex
Addded our old cvpr file for reference

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{pdfsync}
\usepackage{helvet}
\usepackage{url}

% Include other packages here, before hyperref.
\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\subj}{\mathrm{subj}}

\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
%\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

 \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{1291} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Towards a Practical Face Recognition System: \\ Robust Registration and
Illumination by Sparse Representation\vspace{-4mm}}

\author{Andrew Wagner, John Wright, Arvind Ganesh, Zihan Zhou, Yi Ma\\
University of Illinois at Urbana-Champaign, 1308 W. Main st. Urbana, IL 61801\\
{\tt\small \{awagner, jnwright, abalasu2, zzhou7, yima\}@illinois.edu}}

\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}\vspace{-2mm}
Most contemporary face recognition algorithms work well under
laboratory conditions but degrade when tested in less-controlled environments. This is mostly due to the difficulty of simultaneously 
handling variations in illumination, alignment, pose, and occlusion.
In this paper, we propose a simple and practical  face recognition system that achieves a high degree of  
robustness and stability to all these variations. We demonstrate how to use tools from sparse representation 
to align a test face image with a set of frontal training 
images in the presence of significant registration error and occlusion. We thoroughly characterize the region 
of attraction for our alignment algorithm on public face datasets such as Multi-PIE. 
We further study how to obtain a sufficient set of training illuminations for linearly  
interpolating practical lighting conditions. We have implemented a complete face recognition
system, including a projector-based training acquisition system, in order to evaluate how our algorithms work under practical testing conditions. We show  
that our system can efficiently and effectively recognize faces under  
a variety of realistic conditions, using only 
frontal images under the proposed illuminations as training.\vspace{-10mm}
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}\vspace{-2mm}
Automatic face recognition remains one of the most active areas in computer vision. While classical algorithms \cite{Turk1991-CVPR,Belhumeur1997-PAMI} remain popular for their speed and simplicity, they tend to fail on large-scale, practical tests, falling short of the ultimate goal of truly automating face recognition for real-world applications such as access control for facilities, computer systems and automatic teller machines. These applications are interesting both for their potential sociological impact and also because they allow the possibility of carefully controlling the acquisition of the training data, allowing more tractable and reliable solutions.\footnote{Face recognition with less-controlled training samples taken under uncontrolled scenarios remains an active research area as well \cite{LFW}.} In this setting, one promising recent direction, set forth in \cite{Wright2008-PAMI}, casts the recognition problem as one of finding a sparse representation of the test image in terms of the training set as a whole, up to some sparse error due to occlusion.

\begin{figure}[t]
\centering
\begin{tabular}{cc}
\includegraphics[height=0.9in]{Figures/promo/case1/detector.png}& \hspace{3mm}
\includegraphics[height=0.9in,width=1.7in]{Figures/promo/case1/sci_with_axis_face_case1.png} \\
\includegraphics[height=0.9in]{Figures/promo/alignment_and_detector.png}& \hspace{3mm}
\includegraphics[height=0.9in,width=1.7in]{Figures/promo/case2/sci_with_axis_face_case2.png} \\
\includegraphics[height=0.9in]{Figures/promo/case3/alignment.png} & \hspace{3mm}
\includegraphics[height=0.9in,width=1.7in]{Figures/promo/case3/sci_with_axis_face_case3.png}
\end{tabular}
\caption{{\bf Compound effect of registration and illumination}. The task is to identify the girl among 20 subjects, by computing the sparse representation of her input face with respect to the entire training set. The absolute sum of the coefficients associated with each subject is plotted on the right. We also show the faces reconstructed with each subject's training images weighted by the associated sparse coefficients. The red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The input face is from Viola and Jones' face detector (the black box) and all 38 illuminations specified in Section \ref{sec:illumination} are used in the training.  {\bf Middle:} The input face is well-aligned (the white box) with the training by our algorithm specified in Section \ref{sec:registration} but only 24 frontal illuminations are used in the training for recognition (see Section \ref{sec:illumination}). {\bf Bottom:} Informative representation obtained by using both well-aligned input face and sufficient (all 38) illuminations in the training. \vspace{-9mm}} 
\label{fig:promo}
\end{figure}

While that work achieves impressive results on public datasets taken under controlled laboratory conditions such as Extended Yale B \cite{Georghiades2001-PAMI}, it fails to address two critical aspects of real world face recognition: significant variations in both the {\em image domain} and in the {\em image value}. We illustrate this with an example in  Figure \ref{fig:promo}. The task is to identify the girl among 20 subjects. If the test face image, say obtained from an off-the-shelf face detector, has even a small amount of registration error against the training images (caused by mild pose, scale, or misalignment), the representation is no longer informative, even if sufficient illuminations are present in the training as shown in Figure \ref{fig:promo} top. In addition, in order to sufficiently interpolate the illumination of a typical indoor (or outdoor) environment, illuminations from behind the subject are also needed in the training. Otherwise, even for perfectly aligned test images, the representation will not necessarily be sparse or informative, as shown by the example in Figure \ref{fig:promo} middle. Unfortunately, most public face databases lack images with a significant component of rear (more than 90 degrees from frontal) illumination, either for training or testing.\vspace{-4mm}

\paragraph{Contributions.} In this paper, we show how the two {\em strongly coupled} issues of registration and illumination can be naturally addressed within the sparse representation framework. We show that face registration, a challenging nonlinear problem, can be solved by a series of linear programs that iteratively minimize the sparsity of the registration error. This leads to an efficient and effective alignment algorithm for face images that works for a large range of variation in translation, rotation, scale, and pose, even when the face is only partially visible due to eyeglasses, hats, closed eyes and open mouth, sensor saturation, etc.  We also propose a sufficient, if not the smallest, set of training illuminations that is capable of interpolating typical indoor and outdoor lighting, along with a practical hardware system for capturing them.  Finally, we demonstrate the effectiveness of the proposed new methods with a complete face recognition system that is {\em simple, stable, and scalable}. 
The proposed algorithm performs robust automatic recognition of subjects from loosely controlled images taken both indoors and outdoors, using labeled frontal views of the subjects' faces under the proposed illuminations for training and an off-the-shelf face detector\footnote{In this paper, we use the OpenCV implementation of the Viola and Jones' face detector \cite{Viola2004-IJCV}.} to detect faces in images. \vspace{-3mm}

\section{Handling Practical Registration Error}\label{sec:registration}\vspace{-2mm}
As demonstrated in Figure \ref{fig:promo} top, the main limitation of the {\em sparse representation and classification} (SRC) algorithm of \cite{Wright2008-PAMI} is the assumption of pixel-accurate alignment between the test image and the training set. This leads to brittleness under pose and misalignment, making it inappropriate for deployment outside a laboratory setting. In this section, we show how this weakness can be rectified while still preserving the conceptual simplicity and good recognition performance of SRC.

SRC assumes access to a database of multiple registered training images per subject, taken under varying illuminations. The images of subject $i$, stacked as vectors, form a matrix $A_i \in \Re^{m \times n_i}$. Taken together, all of the images form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ] \in \Re^{m \times n}$. As argued in \cite{Wright2008-PAMI}, a well-aligned test image $\y_0$ can be represented as a sparse linear combination $A \x_0$ of all of the images in the database,\footnote{We assume the illuminations in the training set are sufficient. We will address how to ensure illumination sufficiency in the next section.} plus a sparse error $\e_0$ due to occlusion. The sparse representation can be recovered by minimizing the sum or the 1-norm\footnote{The 1-norm of a vector $\x$ is the sum of absolute values of the entries.} of $\x$ and $\e$:\vspace{-1mm}
\begin{equation}
\min \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
\label{eqn:robust-l1}\vspace{-1mm}
\end{equation}
Now suppose that $\y_0$ is subject to some pose or misalignment, so that instead of observing $\y_0$, we observe the warped image $\y = \y_0 \circ \tau^{-1}$, for some transformation $\tau \in T$ where $T$ is a finite-dimensional group of transformations acting on the image domain.  The transformed image $\y$ no longer has a sparse representation of the form $\y = A \x_0 + \e_0$, and naively applying the algorithm of \cite{Wright2008-PAMI} is no longer appropriate, as seen in Figure \ref{fig:promo} top. \vspace{-3mm}

\paragraph{Batch and individual alignment.} Notice that if the true deformation $\tau^{-1}$ can be found, then we can apply its inverse $\tau$ to the test image and it again becomes possible to find a sparse representation of the resulting image, as $\y \circ \tau = A \x_0 + \e_0$.  This sparsity provides a strong cue for finding the correct deformation $\tau$: conceptually, one would like to seek a transformation $\tau$ that allows the sparsest representation, by solving\vspace{-1mm}
\begin{equation} \label{eqn:L1-L1-conceptual}
\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e. \vspace{-1mm}
\end{equation}
For fixed $\tau$, this problem is jointly convex in $\x$ and $\e$. However, as a simultaneous optimization over the coefficients $\x$, error representation $\e$, and transformation $\tau$, it is a difficult, nonconvex optimization problem. One source of difficulty is the presence of multiple faces in the matrix $A$: \eqref{eqn:L1-L1-conceptual} has many local minima that correspond to aligning $\y$ to different subjects. In this sense, the misaligned recognition problem differs from the well-aligned version studied in \cite{Wright2008-PAMI}. For the well-aligned case, it is possible to directly solve for a global representation, with no concern for local minima. With possible misalignment, it is more appropriate to seek the best alignment of the test face with each subject $i$:\vspace{-1mm}
\begin{equation} \label{eqn:per-subject-L1}
\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.\vspace{-1mm}
\end{equation}
We no longer penalize $\| \x \|_1$, since $A_i$ consists of only images of subject $i$ and so $\x$ is no longer expected to be sparse.\vspace{-3mm}

\paragraph{Alignment via iterative $\ell^1$-minimization.} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:\vspace{-1mm}
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.\vspace{-1mm}
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the Jacobian of $\y \circ \tau$ with respect to the transformation parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The above equation is underdetermined if we allow the registration error $\e$ to be arbitrary. Near the correct alignment we expect the aligned testing image to differ from $A_i \x$ only for the minority of the pixels corrupted by occlusions. Thus, we seek a deformation step $\Delta \tau$ that best sparsifies of the registration error $\e$, in terms of its $\ell^1$-norm: \vspace{-1mm}
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y + J \Delta \tau = A_i \x + \e.\vspace{-1mm}
\label{eqn:L1-align}
\end{equation}
Notice that this is different from the popular choice that minimizes the 2-norm of the registration error:\vspace{-1mm}
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y + J \Delta \tau = A_i \x + \e,\vspace{-1mm}
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step $\Delta  \tau$ by solving the least-square problem: $\min \|\y + J\Delta \tau - A_i \x \|_2$. Empirically, we find that if there is only small noise between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and $\eqref{eqn:L2-align}$ have similar performance.  However, if there are occlusions in $\y_0$, iterative $\ell^1$-minimization \eqref{eqn:L1-align} is significantly better than iterative $\ell^2$-minimization \eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an example.
\begin{figure}[t]
\centering
\begin{tabular}{cccc}
\includegraphics[scale=0.214,clip=true]{figures/L1_cropped} &
\includegraphics[scale=0.75,clip=true]{figures/y_warp_L1} &
\includegraphics[scale=0.75,clip=true]{figures/y_hat_L1} &
\includegraphics[scale=0.75,clip=true]{figures/e_L1} \\
\includegraphics[scale=0.214,clip=true]{figures/L2_cropped} &
\includegraphics[scale=0.75,clip=true]{figures/y_warp_L2} &
\includegraphics[scale=0.75,clip=true]{figures/y_hat_L2} &
\includegraphics[scale=0.75,clip=true]{figures/e_L2} \\
(a) & (b) & (c) & (d)
\end{tabular}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by 
$\ell^1$ and $\ell^2$ minimization.} 
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:} 
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$; 
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. \vspace{-3mm}}\label{fig:L1-L2-align}
\vspace{-5mm}\end{figure}

In addition to normalizing the training images (which is done once), it is important to normalize the warped testing image $\y \circ \tau$ as the algorithm runs.  Without normalization, the algorithm may fall into a degenerate global minimum corresponding to expanding a single black pixel in the test image.  Normalization is done by replacing the linearization of $\y \circ \tau$ with a linearization of the normalized version $\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.  The proposed alignment algorithm can be easily extended to work in a {\em multiscale} fashion, with benefits both in convergence behavior and computational cost.  The alignment algorithm is simply run to completion on progressively less down-sampled versions of the training and testing images, using the result of one level to initialize the next.  \vspace{-3mm}

\paragraph{Robust recognition by sparse representation.}
Once the best transformation $\tau_i$ has been computed for each subject $i$, the training sets $A_i$ can be aligned to $\y$, and a global sparse representation problem of the form \eqref{eqn:robust-l1} can be solved to obtain a discriminative representation in terms of the entire training set. Moreover, the per-subject alignment residuals $\| \e \|_1$ can be used to prune unpromising candidates from the global optimization, leaving a much smaller and more efficiently solvable problem. The complete optimization procedure is summarized as Algorithm \ref{alg:deformable-src}. \vspace{-2mm}
\begin{algorithm}[thb]
\begin{small}
\caption{\bf \small  (Deformable Sparse Recovery and Classification for Face Recognition).} \label{alg:deformable-src}
\begin{algorithmic}[1]
\STATE {\bf Input:} Frontal training images $A_1, A_2, \ldots, A_K \in \Re^{m\times n_i}$ for $K$ subjects,  a test image $\y\in\Re^m$ and a deformation group $T$ considered.
\STATE {\bf for} each subject $k$, 
\STATE \hspace{3mm} $\tau^{(0)} \leftarrow I$.
\STATE \hspace{3mm} {\bf do}
\STATE \hspace{6mm} $\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\y(\tau)  \bigr|_{\tau^{(i)}} $;\vspace{1mm}
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \; \mathrm{s.t.} \; \tilde \y + J \Delta \tau = A_k \x + \e, \; \x \ge \0.$
\STATE \hspace{6mm} $\tau^{(i+1)} \leftarrow \tau^{(i)} + \Delta \tau$; 
\STATE \hspace{3mm} {\bf while} $\| \tau^{(i+1)} - \tau^{(i)} \| \ge \varepsilon$.
\STATE {\bf end}
\STATE Keep the top $S$ candidates $k_1, \ldots, k_S$ with the smallest residuals $\|\e\|_1$. 
\STATE Set $A \leftarrow \big[ A_{k_1} \circ \tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots \mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. 
\STATE Solve the $\ell^1$-minimization problem: \vspace{-2mm}
$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \text{subj} \;\; \y = A \x + \e, \;\x \ge 0.\vspace{-2mm}$$
\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \hat{\x}_i \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
\end{algorithmic}
\end{small}
\end{algorithm}\vspace{-2mm}

The most important free parameter in Algorithm \ref{alg:deformable-src} is the class of deformations $T$. In our experiments, we typically use 2D similarity transformations, $T = \mathbb{SE}(2)\times \Re_+$, for compensating error incurred by face detector, or 2D projective transformations, $T = \mathbb{GL}(3)$, for handling some pose variation. The parameter $S$ decides how many top candidates get considered together to provide a sparse representation for the test image. If $S = 1$, the algorithm reduces to classification by registration error; but considering the test image might be an invalid subject, we typically choose $S = 10$. Since valid images have a sparse representation in terms of this larger set, we can reject invalid test images using the {\em sparsity concentration index} proposed in \cite{Wright2008-PAMI}. We have implemented a fast linear program for our algorithm in C. Running on a 2.8GHz Mac Pro, alignment takes 0.65 second per subject for our database. \vspace{-3mm}

\paragraph{Simulations and experiments on region of attraction.} We now perform simulations and experiments demonstrating the effectiveness of the individual alignment procedure outlined in the previous section, and clarifying its operating range. We delay large-scale recognition experiments to Section \ref{sec:experiment}, after we have discussed the issue of illumination in the next section.\vspace{-2mm}
\begin{enumerate}	
\item{\em 2D Deformation.}  We first verify the effectiveness of our alignment algorithm with images from the CMU Multi-PIE Database \cite{Gross2008-FGR}. We select 120 subjects in Session 2, use 11 illuminations per person from Session 2 for training, and test on one new illumination from Session 3.\footnote{The training are illuminations $\{0, 1,3,5,7,9,11,13,14,16,18\}$ of \cite{Gross2008-FGR}, and the testing is the illumination 10. } We manually select eye corners in both training and testing as the ground truth for registration. We down-sample the images to $80\times 60$ pixels\footnote{Unless otherwise stated, this will be the default resolution at which we prepare all our training and testing datasets and run all our experiments.} and the distance between the two outer eye corners are normalized to be 50 pixels for each person. We introduce artificial deformation to the testing image with a combination of translation or rotation. We consider a registration successful if the difference between the final registration error is within 1\% of the error by manual registration.  Figure \ref{fig:attraction} shows the percentage of successful registrations for the 120 subjects for each artificial deformation. The results suggest that our algorithm works extremely well with translation up to 20\% of the eye distance (or 10 pixels) in all directions and up to $30^\circ$ in-plane rotation. We have also tested our alignment algorithm with scale variation and it can handle up to 25\% change in scale.

We have gathered the statistics of the Viola and Jones' face detector on the Multi-PIE datasets. For 4,600 frontal images of 230 subjects under 20 different illuminations, using manual registration as the ground truth, the average misalignment error of the detected faces is about 6 pixels and the variation in scale is 17\%. This falls safely inside the range of attraction for our alignment algorithm.
\begin{figure}[t]
\begin{minipage}{1.5in}
\includegraphics[height=1.36in]{figures/translation_fig3.png}
%\centerline{\small a) $x$ \& $y$ translation}
\end{minipage}\hspace{3mm}
\begin{minipage}{1.5in}
\includegraphics[height=1.42in, width = 1.9in]{figures/translation_rotation_fig1.png}
%\centerline{\small b) $y$-translation \& rotation}
\end{minipage}
\vspace{.1in}
\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.  The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees. {\bf Left:} Simultaneous translation in $x$ and $y$ directions. More than $90\%$ of the subjects were correctly aligned for any combination of $x$ and $y$ translations, each upto $0.2$. {\bf Right:} Simultaneous translation in $y$ direction and in-plane rotation $\theta$. More than 90\% of the subjects were correctly aligned for any combination of $y$ translation upto $0.2$ and $\theta$ upto $25^\circ$.}
\label{fig:attraction}
\end{figure}

\vspace{-2mm}
\item{\em 3D Pose Variation.} As densely sampled pose and illumination face images are not available in any of the public databases, including Multi-PIE, we have collected our own dataset using our own system (to be introduced in the next section.) We use frontal face images of a subject under the 38 illuminations proposed in the next section as training. For testing, we collect the image of the subject under a typical indoor lighting condition at pose ranging from $-90^\circ$ to $+90^\circ$ with step size 5.625$^\circ$, a total of 33 poses. We use Viola and Jones' face detector to initialize our alignment algorithm. 
%whenever it succeeds. For extreme poses where the detector fails (usually for poses $> 45^\circ$), we use manual initialization by putting a reasonable square window around the face. 
Figure \ref{fig:pose-alignment} shows the typical alignment results of our algorithm, working surprisingly well with poses up to $\pm 45^\circ$. \vspace{-4mm}
\begin{figure}[t]
\begin{tabular}{ccccc}
\includegraphics[scale=0.085,clip=true]{figures/5} &
\includegraphics[scale=0.085,clip=true]{figures/7} &
\includegraphics[scale=0.085,clip=true]{figures/09} &
\includegraphics[scale=0.085,clip=true]{figures/11} &
\includegraphics[scale=0.085,clip=true]{figures/13} \\
(a) & (b) & (c) & (d) & (e)\\
\includegraphics[scale=0.085,clip=true]{figures/15} &
\includegraphics[scale=0.085,clip=true]{figures/17} &
\includegraphics[scale=0.085,clip=true]{figures/19} &
\includegraphics[scale=0.085,clip=true]{figures/21} &
\includegraphics[scale=0.085,clip=true]{figures/3} \\
(f) & (g) & (h) & (i) & (j)
\end{tabular}
\caption{{\bf Aligning different poses to frontal training images.} {\bf (a) to (i):}  good alignment for pose from $-45^{\circ}$
to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).\vspace{-4mm}
}\label{fig:pose-alignment}
\end{figure}
\end{enumerate}

\paragraph{Relationship to existing work.} 
Our modification to SRC roots solidly in the tradition of adding deformation-robustness to face recognition algorithms \cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}. However, the only previous work to investigate face alignment in the context of sparse signal representation and SRC is the work of \cite{Huang2008-CVPR}. They consider the case where the training images themselves are misaligned and allow one deformation per training image. They linearize the training rather than the test, which is computationally more costly as it effectively triples the size of the training set. In addition, as they align the test image to all subjects simultaneously, it potentially is more prone to local minima when the number of subjects increases, as we will see in the following experimental comparisons. \vspace{-1mm}
\begin{enumerate}
\item {\em Extended Yale B.} In this experiment, we have used the exact experimental settings in \cite{Huang2008-CVPR}. 
20 subjects are selected and each has 32 frontal images (selected at random) as training and another 32 for testing. An artificial translation
of 10 pixels (in both $x$ and $y$ directions) is introduced to the test. For our algorithm we down-sample all the images
to $88\times 80$ for memory reasons, whereas the work of \cite{Huang2008-CVPR} uses random projections. Our algorithm achieves the recognition rate 88.59\% which is on par with the result reported in \cite{Huang2008-CVPR}.
However, this special setting is disadvantageous to our algorithm: The use of cropped test images introduces boundary effects, and the presence of very extreme illuminations makes enforcing nonnegativity of $\x$ (as in Algorithm 1 less appropriate. We further discuss the justification for nonnegativity in the next section.\vspace{-2mm}
\item {\em CMU Multi-PIE.} In this experiment, we choose 160 subjects from the CMU Multi-PIE, 11 training images from Session 2
and 1 test image from Session 3 per person. The setting is exactly the same as the previous experiment on 2D deformation, 
except that we have more subjects. We again 
work with downsampled images of size $80\times 60$. An artificial translation of 5 
pixels (in both $x$ and $y$ directions) was induced in the test image. The algorithm of \cite{Huang2008-CVPR} achieves
a recognition rate of 73.75\%,\footnote{That algorithm has two free parameters - $l$ and $d$. For this experiment we chose 
$l = 1$ and $d = 514$ (higher values may get a better recognition rate at the expense of higher running time).
} while ours does 90.625\%. \end{enumerate}\vspace{-4mm}

\section{Handling Practical Illumination Variation}\label{sec:illumination}\vspace{-2mm}
In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly interpolated by a finite number of training illuminations. It has been shown that for a convex Lambertian surface, one only needs about nine basis illuminations to linearly interpolate all other illuminations \cite{Basri2003-PAMI}. Although a human face is neither perfectly Lambertian nor convex, it has been observed in various empirical studies that one can often get away using a similar number of frontal illuminations to interpolate a wide range of new frontal illuminations that taken under the same laboratory conditions \cite{Georghiades2001-PAMI}. This is the case for many public face datasets, including AR, ORL, PIE, and Multi-PIE.

Unfortunately, we have found that in practice, a training database consisting purely of frontal illuminations is not sufficient to linearly interpolate images of a faces taken under typical indoor or outdoor conditions (see the experiment conducted in Section \ref{sec:own-data}). The representation computed is not always sparse or informative, as shown by the example in Figure \ref{fig:promo}. Subsequently, the recognition could become inaccurate. 
%Additional experimental results in the next section will further confirm this point. 
Thus, to ensure our algorithm works in practice, we need to find a set of training illuminations that are indeed {\em sufficient} to linearly interpolate variety of practical indoor and outdoor illuminations.\vspace{-3mm}

\paragraph{Capturing a sufficient set of training illuminations.}   To this end, we have designed a system that can illuminate the subject from all directions above horizontal, while acquiring the subject's frontal images. A sketch of the system is shown in Figure \ref{fig:system}: The illumination system consists of four projectors that display various bright patterns onto the three white walls in the corner of a dark room.  The light reflects off of the walls and illuminates the user's head indirectly.  After taking the frontal illuminations we rotate the chair by 180 degrees and take pictures from the opposite direction.  Having two cameras speeds the process since only the chair needs to be moved in between frontal and rear illuminations.  
Our projector-based system has several advantages over flash-based illumination systems:\vspace{-2mm}
\begin{itemize}
\item The illuminations can be defined in software.\vspace{-2mm}
\item It is easy to capture many different illuminations.\vspace{-2mm}
\item There is no need to mount cameras on walls or construct a large dome.\vspace{-2mm}
\item No custom hardware is needed for a basic system.\vspace{-2mm}
\end{itemize}
\begin{figure}
\centerline{\hspace{-0.1in}\includegraphics[width=2.3in]{figures/camera_rig.pdf}}
\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.\vspace{-8mm}}
\label{fig:system}
\end{figure}
With our projector system, our choice of illuminations is constrained only by the need to achieve a good SNR for representing typical test images and a reasonably short total acquisition time.\footnote{Better SNR can be achieved with more illuminations but that will increase the capture time for each subject.}  We ran two experiments to guide our choice of illuminations for our large-scale experiments:\vspace{-2mm}  
\begin{itemize}
\item {\em Coverage Experiment.} In the first experiment we attempt to determine what coverage of the sphere is required to achieve good interpolation for test images.  The subject was illuminated by 100 (50 front, 50 back) illuminations arranged in concentric rings centered at the front camera.  Subsets of the training images were chosen, starting at the front camera and adding a ring at a time.  Each time a ring was added to the training illumination set, the average $\ell^1$ registration error (residual) for a set of test images taken under sunlight was computed and plotted in Figure \ref{fig:illumination-patterns} (a).  The more rings of training illuminations are added, the lower the representation error becomes, with diminishing returns.\vspace{-2mm}
\item {\em Granularity Experiment.} In the second experiment we attempt to determine how finely divided the illumination sphere should be.  At the first granularity level, the projectors  illuminate the covered area uniformly.  At each subsequent granularity level each illuminated cell is divided in two along its longer side but intensity doubled.  For each granularity level the average $\ell^1$ registration error is computed as in the coverage experiment and shown in Figure \ref{fig:illumination-sufficiency} (b).  Again, diminishing returns are observed as more illuminations are added.\vspace{-3mm}  
\end{itemize}
\begin{figure}[h]
\begin{minipage}{1.5in}~\hspace{-0mm}
\includegraphics[height=1.1in]{figures/coverage_experiment_asplode.png}
\centerline{\small (a) Coverage Experiment}
\end{minipage}\hspace{5mm}
\begin{minipage}{1.5in}
\includegraphics[height=1.1in]{figures/final_cvpr_illuminations_asplode.png}
\centerline{\small (b) Chosen Illumination Patterns}
\end{minipage}
\vspace{.1in}
\caption{{\bf Illumination patterns.}   The cells are illuminated in sequence.  For rear illuminations the sequence is reversed.  In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations. The four rectangular regions correspond to the four projectors.\vspace{-3mm}}
\label{fig:illumination-patterns}
\end{figure}
\begin{figure}[h]
\begin{minipage}{1.5in}
\includegraphics[height=1.3in]{figures/illum_results/coverage_sunset.pdf}
\centerline{\small (a) Coverage }
\end{minipage}\hspace{3mm}
\begin{minipage}{1.5in}
\includegraphics[height=1.3in]{figures/illum_results/granularity_sunset.pdf}
\centerline{\small (b) Granularity}
\end{minipage}
\vspace{.1in}
\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. \vspace{-0.5cm}}
\label{fig:illumination-sufficiency}
\end{figure}

\vspace{-3mm}\paragraph{Chosen illumination patterns.} In the plot for the coverage experiment, Figure \ref{fig:illumination-sufficiency} (a), we clearly see two plateau regions: one is after 4 rings and one is after 10 rings. The first four rings represent the typical frontal illuminations, which are present in most public face datasets; however, we see that the residual stabilizes after 10 rings which include some illuminations from the back of the subject. This suggests that although the frontal illuminations can span majority of illumination on the face, some illuminations from the back are needed in the training to emulate the effect of ambient illumination from all directions. In the plot for the granularity experiment, Figure \ref{fig:illumination-sufficiency} (b), we observe that the residual reaches a plateau after four divisions, corresponding to a total of 32 illuminations. Based on the results from both experiments, we decide to partition the area covered by the first 10 rings into a total of 38 cells, whose layout is explained in Figure \ref{fig:illumination-patterns} (b). For our large-scale experiments, we have collected those illuminations for all our subjects.\footnote{It is very likely that with more careful experiments, we can further reduce the number of illuminations needed. Especially some of the frontal illuminations might be redundant. But we keep those in our training anyway as the additional images do not add too much cost to our alignment and recognition algorithm.}

See below for the 38 images for one subject:
\begin{figure}[h]
\includegraphics[width=3.3in]{figures/training.png}
\end{figure}
\vspace{-6mm}
\paragraph{The role of nonnegativity.} 
One critical issue in linear illumination models is whether to enforce nonnegativity in the coefficients $\x$: whether to model illumination using a cone or a subspace. Nonnegative combinations are guaranteed to correspond to physically plausible illuminations, but will not be sufficient to represent all physical illuminations unless the training images actually span the boundary of the illumination cone. Because we have a flexible acquisition system, we can directly generate a set of illuminations that span most of the illumination cone, without resorting to negative coefficients and risking overfitting. Thus, in Algorithm 1, we have enforced $\x$ to be non-negative.

%%

%Once we have sufficient number of illuminations, we may expect that any new illumination falls into the convex hull spanned by them, known as the illumination cone. In this case, when we try to find the sparse representation of a test image from equation \eqref{eqn:robust-l1}, we also expect the coefficients $\x$ to be non-negative. This ensures the representation to be physically meaningful -- an image that has negative values. This point is reflected in the linear program used in Algorithm 1.

%In addition, the linear illumination model requires that the camera has captured the illuminance faithfully. But that is not the case for most cameras: older analog video cameras often use NTSC defined gamma compression, while most modern digital consumer and SLR cameras record in sRGB defined gamma compression.\footnote{For more info, see the website \url{http://en.wikipedia.org/wiki/Gamma_correction}}  Since not all algorithms depend on a linear model for the illumination, there are public face recognition databases that do not specify the gamma compression of their images.  The performance of the algorithm presented in this paper benefits greatly if proper gamma decompression has been applied to the training and testing images.

%{\bf Yi: show how sparse representation improves for subject X with the new training illuminations (Andrew).}

%\jw{I have not touched this section much.}

%One of the main factors that complicates face recognition, and computer vision in general, is that two images of the same person's face can be very different, even if the pose is carefully controlled and all points on the person's face are visible in both images.  The most safe, reliable, and general assumption that can be made about the imaging process is that it is linear with respect to the intensity of different light sources.  The set of images of a given object is therefore a cone in the non-negative orthant of the pixel basis.  All recognition algorithms that rely on vision for training rely on images sampled from this cone.

%A lot of algorithms make much stronger assumptions than linearity in the interest of simplifying the problem.  For example, if the object under study is convex, the lighting is distant, and the surface reflectance is Lambertian, it has been shown that most of the variation of the reflected light with respect to the surface normal will be concentrated in the low frequencies with respect to the direction of the surface normal.  This provides a good justification for relying on a reduced number of illuminations for your training set, provided that they are well chosen.  Indeed, when most of the illumination in the training and testing images is frontal (as it is in most public face recognition databases), the importance of non-convexity is significantly decreased.

%\jw{The below thoughts on Gamma compression should go somewhere in this section.}
%\paragraph{Compensating for Gamma Compression.}
%Any algorithm based on representing a testing image as a weighted sum of training images must make sure that the image intensity has been coded linearly with image intensity.  Unfortunately, this cannot be taken for granted.  While most modern machine vision cameras capture linearly in intensity, older analog video cameras often used NTSC defined gamma compression, while most modern digital consumer and SLR cameras record in sRGB defined gamma compression.  Since not all algorithms depend on a linear image coding, there are public face recognition databases that do not specify the gamma compression of their images.  The performance of the algorithm presented in this paper benefits greatly from proper gamma decompression of the training and testing images.

%
%%Improved robustness to illumination variation can be achieved by integrating the design of the algorithm with the design of the training image capture system.

%\subsection{Should Positivity in the Representation Coefficients be Enforced?}
%For non-convex objects, enforcing non-negativity of the representation is not sufficient to guarantee that the resulting image of the object is physical.  For example, consider a Lambertian planar object with a cylindrical hole drilled in it that is aligned with the optical axis of an orthographic camera.  Consider two images: one under illumination from a distant point source aligned with the hole, and one off of the hole axis.  Positively weighting the on-axis image and negatively weighting the on-axis image can result in an image where the bottom of the hole is more strongly illuminated than the surrounding plane.  The image is positive, but it is clearly non-physical.  It is worth emphasizing that this phenomenon is a result of some points on an object seeing a restricted subset of the illumination sphere,  and not a result of the directionality of the light chosen in the thought experiment.

%		Why doesn't Spherical Harmonics solve the problem of what your training images should be?
%			-Illuminations must be non-negative.  Could take positive and negative parts of the image separately, but we're not allowing negative coefficients.

%	\subsection{Motivation}
%Unfortunately, {\em human faces are non-convex and non-lambertian enough to cause big problems for face recognition systems!}  When algorithm \ref{eqn:final_algorithm} is run on images of people taken under lighting conditions, the error image spends a lot of its energy on areas in the image where there is either a specularity (a violation of the Lambertian assumption), or self-shadowing (a violation of the assumption that the object is convex).
%	\subsection{Experiments}

%		\begin{itemize}
%		\item Show that a multii-pie-like subset of illuminations fails to represent lobby and outdoor images; show a single example of per-user alignment borking.
%		\item Show that a multii-pie-like subset of illuminations fails to represent lobby and outdoor images, compute (low) success rate of per-user alignment over a bunch of test images.
%		\item Show that the complete set of training illuminations succeeds on the same images; show a single example of per-user alignment working.
%		\item Show that the complete set of training illuminations succeeds on the same images, compute (high) success rate of per-user alignment over a bunch of test images. \end{itemize}

%		\subsection{Experimentally Nail down what training images are actually necessary}
%			\begin{itemize}
%				\item Increase coverage, expanding outwards from frontal.
%				\item Gradually increase granularity, show where performance jumps.  Have to express granularity in terms of percentage of the illumination sphere.
%				\item Test if partitions of the illumination sphere perform better than point sources at the centers of the regions.  Regions may be better due to lowpass effect on cast shadow boundaries.		
%			\end{itemize}

\section{Overall System Evaluation}\label{sec:experiment}\vspace{-2mm}
In this section, to verify the performance of our algorithm and system, we conduct comprehensive experiments on large-scale face databases. We first test on the largest public face database available that is suitable for testing our algorithm, the CMU Multi-PIE. The goal is to show that our algorithm can indeed be used to achieve good performance on such a dataset with test images obtained from an off-the-shelf face detector, even though we can only use a small number of, not necessarily sufficient, training illuminations. We then test our algorithm on a face dataset that is collected by our own system. The goal is to show that with a sufficient set of training illuminations for each subject, our algorithm indeed works stably and robustly with practical illumination, misalignment, pose, and occlusion, as already indicated by our experiment shown in Figure \ref{fig:promo} bottom.

\subsection{Tests on public databases}\vspace{-2mm}
CMU Multi-PIE provides the most extensive test of our algorithm among public datasets. This database contains images of 337 subjects across simultaneous variation in pose, expression, and illumination. Of these 337 subjects, we use all the 249 subjects present in Session 1 as a training set. The remaining 88 subjects are considered ``outliers'' or invalid images. For each of the 249 training subjects, we include frontal images of 7 frontal illuminations\footnote{They are illuminations $\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}.  For each directional illumination, we subtract the ambient-illuminated image 0.}, taken with neutral expression. As suggested by the work of \cite{Georghiades2001-PAMI}, these extreme frontal illuminations would be sufficient to interpolate other frontal illuminations, as will also be corroborated by the next experiment on our own dataset. For the test set, we use all 20 illuminations from Sessions 2-4, which were recorded at distinct times over a period of several months. The dataset is challenging due to the large number of subjects, and due to natural variation in subject appearance over time. Table \ref{tab:multipie} shows the result of our algorithm on each of the 3 testing sessions. Our algorithm achieves recognition rates above $90\%$ for all three sessions, with input {\em directly} obtained from the Viola and Jones' face detector -- no manual intervention. We compare our result to baseline linear-projection-based algorithms, such as Nearest Neighbor (NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear Discriminant Analysis (LDA) \cite{Belhumeur1997-PAMI}.\footnote{We do not list results on PCA \cite{Turk1991-CVPR} as its performance is always below that of Nearest Subspace.} Since these algorithms assume pixel-accurate alignment, they are not expected to work well if the test is not well-aligned with the training. In the table of Figure \ref{tab:multipie}, we report the results of those algorithms with two types of input: 1.\ the output of the Viola and Jones' detector, indicated by a subscript ``$d$''; 2.\ the input face is aligned to the training with manually selected outer eye corners, indicated by a subscript ``$m$''. Notice that, despite careful manual registration, these baseline algorithms perform significantly worse than our algorithm, which uses input directly from the face detector. The performance of the LDA algorithm on Multi-PIE reported here seems to agree with that reported already in \cite{Gross2008-FGR}.\vspace{-2mm}

\paragraph{Subject validation.} We test the algorithms' ability to reject invalid images of the 88 subjects not appearing in the training database. Figure 8 (bottom) plots the receiver operating characteristic (ROC) curve for each algorithm.\footnote{Rejecting invalid images not in the entire database is much more difficult than deciding if two face images are the same subject. Figure 8 should not be confused with typical ROC curves for face similarity, e.g., \cite{PhillipsP2007}.} Similar contrasts between our algorithm and baseline algorithms were also observed for SRC in \cite{Wright2008-PAMI}, though on much smaller datasets.

\vspace{-3mm}
\begin{figure}[t]
\centerline{
\begin{small}
\begin{tabular}{|l|c|c|c|c| }
\hline
Rec. Rates & Session 2 & Session 3 & Session 4  \\
\hline
\hline
LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
\hline
NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
\hline
NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
\hline
{Algorithm 1} & {\bf 91.4} \% & {\bf 90.3} \% & {\bf 90.2} \% \\
\hline
\end{tabular}
\end{small}
\label{tab:MultiPIE-recognition}
}
\centerline{
\includegraphics[width=2.8in]{figures/roc3.png} 
}
\caption{{\bf Large-scale experiments on Multi PIE.} {\bf Top:} Recognition rates; {\bf Bottom:} ROC curves for our algorithm (labeled as ``$\ell_1$''), compared with those for NN$_m$, NS$_m$, and LDA$_m$.} \label{tab:multipie}
\end{figure}

\paragraph{Cause of errors.} Our algorithm's errors are mostly caused by a few subjects who significantly change their appearances between sessions (such as hair, facial hair, and eyeglasses). Some representative examples are shown in Figure \ref{fig:failed-examples}. In fact, for those subjects, alignment and recognition fail on almost all test illuminations.\vspace{-3mm} %This may be due the limited number of illuminations present in the training. \vspace{-4mm}
\begin{figure}[t]
\centering
\includegraphics[scale=0.2320,clip=true]{figures/079_01_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/130_01_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/163_01_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/175_01_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/118_01_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/223_01_01_051_08.png} \\
\includegraphics[scale=0.2320,clip=true]{figures/079_02_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/130_02_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/163_02_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/175_02_01_051_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/118_02_01_140_08.png} 
\includegraphics[scale=0.2320,clip=true]{figures/223_02_01_140_08.png} 
\caption{{\bf Representative examples of failed Multi-PIE subjects}. {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2 -- first four are frontal and the last two with $15^\circ$ pose. Notice the change of hair style and facial hair, which makes alignment fail on those subjects, actually regardless of test image illuminations.\vspace{-3mm}}\label{fig:failed-examples}
\end{figure}

\paragraph{Pose and expression.} We also run limited tests of our algorithm on images with pose and expression in Multi-PIE.  Using the same training as above, we test our algorithm on images in Session 2 with $15^\circ$ pose, for all 20 illuminations. The recognition rate is 77.5\%. We also test our algorithm on images in Session 3 with smile. For illumination 0 (ambient), the rate is 58.5\%, for illumination 10, the rate is 68.6\%.\vspace{-2mm}

\subsection{Tests on our own datasets}\label{sec:own-data}\vspace{-2mm}
Using the training acquisition system that we have described in the previous section, Figure \ref{fig:system}, we have collected the frontal view of 74 subjects {\em without eyeglasses} under 38 illuminations shown in Figure \ref{fig:illumination-patterns}. For testing our algorithm, we have also taken 593 images of these subjects with a different camera under a variety of practical conditions.\vspace{-4mm}

\paragraph{Limitation of frontal illuminations.}
To see how training illuminations affect the performance of our algorithm in practice, we now compare how well a few frontal illuminations can interpolate: 1. other frontal illuminations taken under the same laboratory conditions, and 2. typical indoor and outdoor illuminations. To this end, we select 20 subjects from the face database acquired by our system and use 7 illuminations per subject as training. The illuminations are chosen to be similar to the 7 illuminations used in the previous experiment on Multi-PIE.\footnote{We use the illumination set $\{6, 9, 12, 13, 18, 21, 22\}$ shown in Figure \ref{fig:illumination-patterns}(b) to mimic the illumination set $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test our algorithm on the remaining $24 - 7 = 17$ frontal illuminations for all the 20 subjects. The recognition rate is $99.7\%$, nearly perfect. We also test our algorithm on $173$ frontal images of these subjects taken under a variety of indoor and outdoor conditions (in category 1 specified below), similar to the one shown in Figure \ref{fig:promo}, and the recognition drops down to $93.6\%$. One would expect the rate to drop even further when the number of subjects increases.\vspace{-2mm}

\paragraph{Large-scale test with sufficient training illuminations.}
Now we use all 74 subjects and 38 illuminations in the training and test on 593 images taken under a variety of conditions. Based on the main variability in the test images, we have partitioned them into five main categories:%\vspace{-2mm}
\begin{small}
\begin{description}
\item[C1:] 242 images of 47 subjects without eyeglasses, generally frontal view, under a variety of practical illuminations (indoor and outdoor) (Fig. 10, row 1). %\vspace{-2mm}
\item[C2:] 109 images of 23 subjects with eyeglasses (Fig. 10, row 2).%\vspace{-2mm}
\item[C3:] 19 images of 14 subjects with sunglasses (Fig. 10, row 3).%\vspace{-2mm}
\item[C4:] 100 images of 40 subjects with noticeable expressions, poses, mild blur, and sometimes occlusion (Fig. 11, both rows).%\vspace{-2mm}
\item[C5:] 123 images of 17 subjects with little control (out of focus, motion blur, significant pose, large occlusion, funny faces, extreme expressions) (Fig. 12, both rows).%\vspace{-2mm}
\end{description}
\end{small}
We apply Viola and Jones' face detector on these images and directly use the detected faces as the input to our algorithm. The table below reports the performance of our algorithm on each category. The errors include failures of the face detector on some of the more challenging images.\vspace{-1mm}
\begin{table}[h]	
\centering
\begin{tabular}{|l|c|c|c|c|c| }
\hline
Test Categories & C1 & C2 & C3 & C4 & C5  \\
\hline
\hline
Rec. Rates (\%) &  95.9 & 91.5 & 63.2 & 73.7 & 53.5 \\
\hline
\end{tabular} \vspace{-1mm}
\end{table}

\begin{figure}
\centering
\includegraphics[scale=0.25,clip=true]{figures/examples/1/DSC_1319.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/1/DSC_1531.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/1/DSC_1574.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/1/DSC_1622.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/1/DSC_1786.jpg} \\
\includegraphics[scale=0.25,clip=true]{figures/examples/2/DSC_1448.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/2/DSC_1585.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/2/DSC_1588.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/2/DSC_1666.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/2/DSC_1874.jpg} \\
\includegraphics[scale=0.25,clip=true]{figures/examples/3/DSC_1572.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/3/DSC_1587.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/3/DSC_1623.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/3/DSC_1652.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/3/DSC_1675.jpg} 
 \caption{{\bf Representative examples of categories 1-3}. One row for each category.\vspace{1.0mm}}\label{fig:examples1-3}
%\end{figure}
%\begin{figure}[h]
\centering
\includegraphics[scale=0.25,clip=true]{figures/examples/4/success/DSC_1579.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/success/DSC_1839.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/success/DSC_1883.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/success/DSC_1922.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/success/DSC_2000.jpg} \\
\includegraphics[scale=0.25,clip=true]{figures/examples/4/failure/DSC_1568.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/failure/DSC_1578.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/failure/DSC_1500.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/failure/DSC_1930.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/4/failure/DSC_1985.jpg} 
 \caption{{\bf Representative examples of category 4}. Top row: successful examples. Bottom row: failures.}\label{fig:examples4}
\vspace{1mm}
\centering
\includegraphics[scale=0.25,clip=true]{figures/examples/5/success/DSC_1664.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/success/DSC_1777.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/success/DSC_1871.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/success/DSC_1913.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/success/DSC_2005.jpg} \\
\includegraphics[scale=0.25,clip=true]{figures/examples/5/failure/DSC_1669.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/failure/DSC_1800.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/failure/DSC_1870.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/failure/DSC_1917.jpg} 
\includegraphics[scale=0.25,clip=true]{figures/examples/5/failure/DSC_2008.jpg} 
 \caption{{\bf Representative examples of category 5}. Top row: successful examples. Bottom row: failures.\vspace{-4mm}}\label{fig:examples5}
\end{figure}

\vspace{-5mm}
\section{Conclusion}\vspace{-2mm}
We have proposed a new algorithm and system for recognizing human faces from images taken under practical conditions. The proposed system is very {\em simple} and hence the results are easy to reproduce. The proposed algorithm is {\em scalable} both in terms of computational complexity and recognition performance. The system is directly compatible with off-the-shelf face detectors and achieves extremely {\em stable} performance under a wide range of variations in illumination, misalignment, pose, and occlusion. We achieve very good recognition performance on large-scale tests with public datasets and our practical face images, using only frontal 2D images in the training without any explicit 3D face model. Our implementation still has plenty of room for further engineering improvements.\vspace{-2mm}

%
{\small
\bibliographystyle{ieee}
\bibliography{cvpr09_faces}
}

\end{document}

### 2011-04-02 faces_system.tex
Fixed figures.  Need Bios.  0.7 pages over

@@
 }
-\label{fig:illumination-patterns}\label{fig:illumination-sufficiency}
+\label{fig:illumination-patterns}
 \vspace{-.75em}
@@
 \centering
-\begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
+\begin{tabular}{@{}c@{}c@{}}
+\includegraphics[height=1.35in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
+\includegraphics[height=1.35in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
 (a) Coverage & (b) Granularity
@@
 \begin{figure}
+\vspace{-1em}
 \centering
 {\small
-\begin{tabular}{cc}
-\includegraphics[height=1.5in]{figures_pami/example.png} &
-\includegraphics[height=1.5in]{figures_pami/example_new.png} \vspace{0mm} \\
-(a) & (b) \vspace{0mm}
+\begin{tabular}{@{}cc@{}}
+\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example.png} &
+\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example_new.png} \vspace{0mm} \\
+Default window. & Proposed window. \vspace{0mm}
 \end{tabular}
 }
-\caption{{\bf Choosing different sampling windows.} (a) The default window. (b) The newly proposed window.}
+\caption{{\bf Choosing different sampling windows.}}
 \label{fig:new-mask}
-\vspace{0mm}
+\vspace{-1em}
 \end{figure}
@@
 \label{tab:MultiPIE-recognition} } 
-\vspace{-.3in}
+\vspace{-.1in}
 \end{table*}
@@
 \begin{table}
-\caption{Recognition rates on single gallery image FERET
-dataset} \vspace{0mm} \centerline{
+\caption{Performance on single gallery image FERET dataset} 
+\vspace{-3mm} \centerline{
 \begin{tabular}{|c|c|c|c|c| }
 \hline
-Recognition rate & fb & fc & dup1 & dup2 \\
-\hline
-\hline
-LBP($d/m/i$) & 54.8 / {\bf 96.6} / 94.5 \% & 10.3 / {\bf 58.8} / 42.8\% & 29.8 / {\bf 71.6} / 46.5\% & 19.8 / {\bf 61.5} / 21.1\% \\
-\hline
-{Algorithm 1} & 95.2\% & 28.4\% & 46.1\% & 20.3\% \\
+Recognition rate \% & fb & fc & dup1 & dup2 \\
+\hline
+$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
+\hline
+$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
+\hline
+$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
+\hline
+{Algorithm 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
 \hline
 \end{tabular}
-\label{tab:FERET-recognition} } \vspace{0mm}
+\label{tab:FERET-recognition} } 
+\vspace{-12mm}
 \end{table}
@@
 \begin{figure}[t]
-\vspace{.2in}
+\vspace{-.15in}
 \centerline{
@@
 example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.\vspace{0mm}
-
-\begin{figure}[t]
+recognition rate.
+\vspace{0mm}
+
+\newcommand{\tempscale}{0.40}
+\begin{figure}
 \centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ10.png}
-\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ20.png}
-\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ30.png}
-\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ40.png}
-\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
-\vspace{0mm}
+\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ10.png} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ20.png} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ30.png} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ40.png} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ50.png}  \\
+\end{tabular}
+\vspace{-1mm}
 \caption{{\bf Recognition under varying level of
 random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:}
-\vspace{3mm}
+\vspace{2mm}
 {\small
@@
 Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
-\hline
 \hline
@@
 \label{fig:multipie-occ-rec}
-\vspace{0mm}
+\vspace{-.5in}
 \end{figure}
@@
 
-\newcommand{\tempscale}{0.45}
+\renewcommand{\tempscale}{0.45}
 \begin{figure}
 \centering
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG}
+\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG} &
 \includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
-\vspace{2mm}
-\centering
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG} &
 \includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
-\vspace{2mm}
-\centering
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG}
-\vspace{-.2in}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} \\
+\end{tabular}
 \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
- \vspace{0mm}
+\vspace{-.1in}
 \end{figure}
@@
 \centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
-\vspace{2mm}
-\centering
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG}
-\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
-\vspace{0mm}
+\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG} &
+\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG} \\
+\end{tabular}
  \caption{{\bf Representative examples of category C4}.
@@
 Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{0mm}
+\vspace{-.55in}
 \end{figure}
@@
 \centerline{
-\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{0mm}}
+\includegraphics[width=3in]{figures_pami/occ_block.png}\vspace{0mm}}
 \caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.

### 2011-04-02 faces_system.tex
Added 6 Arvinds

@@
 % simpler here.
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell.eps}}]{Michael Shell}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Andrew Wagner}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{John Wright}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Hossein Mobahi}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Yi Ma}
+Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
+is currently a PhD candidate in the Electrical \& Computer Engineering
+Department at the University of Illinois, Urbana-Champaign. His research
+interests include compressed sensing, computer vision, and machine learning.
+His recent work focuses on low-rank matrix recovery techniques for
+batch image alignment and texture rectification.  \end{IEEEbiography}
+
 % or if you just want to reserve a space for a photo:

### 2011-04-04 faces_system.tex
Added my biography, switched to subfig

@@
 \usepackage{amssymb}
-\usepackage{subfigure}
+
+%\usepackage{subfigure}
+
+%\ifCLASSOPTIONcompsoc
+ %\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+%\else
+ %\usepackage[caption=false,font=footnotesize]{subfig}
+%\fi
+\usepackage[caption=false,font=footnotesize]{subfig}
+
 \usepackage{pdfsync}
@@
 
+% insert where needed to balance the two columns on the last page
+%\newpage
+
 % trigger a \newpage just before the given reference
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Andrew Wagner}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
+Andrew Wagner received his Bachelor's degree in General Engineering,
+and his Master's degree in Electrical Engineering, from the University
+of Illinois at Urbana-Champaign, where he is currently a Ph.D. 
+candidate in Electrical \& Computer Engineering.
+His research interests include robotics, computer vision, and optimal control.
+His recent work focuses on parallel algorithms for sparsity
+based face recognition.\end{IEEEbiography}
 
@@
 
-% or if you just want to reserve a space for a photo:
-
-%
-%\begin{biography}{Andrew Wagner}
-%Biography text here.
-%\end{biography}
-%
-%\begin{biography}{John Wright}
-%Biography text here.
-%\end{biography}
-%
-%\begin{biography}{Arvind Ganesh}
-%Biography text here.
-%\end{biography}
-%
-%\begin{biography}{Zihan Zhou}
-%Biography text here.
-%\end{biography}
-%
-%\begin{biography}{Yi Ma}
-%Biography text here.
-%\end{biography}
-%
-
 % if you will not have a photo
@@
 %\end{biographynophoto}
-
-% insert where needed to balance the two columns on the last page
-%\newpage
 
@@
 
-% that's all folks
 \end{document}

### 2011-04-04 faces_system.tex
Vspacing from zihan

@@
 %\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
-%\documentclass[9.5pt,journal,final,finalsubmission,twocolumn]{IEEEtran}
+%\documentclass[9.5pt,journal,final,compsoc,twocolumn]{IEEEtran}
 
@@
 \usepackage{amssymb}
-
-%\usepackage{subfigure}
-
-%\ifCLASSOPTIONcompsoc
- %\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
- %\usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-\usepackage[caption=false,font=footnotesize]{subfig}
-
+\usepackage{subfigure}
 \usepackage{pdfsync}
@@
 \usepackage{slashbox}
+
+% \usepackage[small]{caption}
 
@@
 \begin{abstract}
-Many classic and contemporary face recognition algorithms work well on public data sets, but degrade sharply when they are  used in a real recognition system.
-This is mostly due to the difficulty of simultaneously handling variations in illumination, image misalignment, and occlusion in the test image.
-We consider a scenario where the training images are well controlled, and test images are only loosely controlled.
-We propose a conceptually simple face recognition system that achieves a high degree of robustness and stability to illumination variation, image misalignment, and partial occlusion.
-The system uses tools from sparse representation to align a test face image to a set of frontal training images.
-The region of attraction of our alignment algorithm is computed empirically for public face datasets such as Multi-PIE.
-We demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under uncontrolled illumination.
-In order to evaluate how our algorithms work under practical testing conditions, we have implemented a complete face recognition system, including a projector-based training acquisition system.
-Our system can efficiently and effectively recognize faces under a variety of realistic conditions, using only frontal images under the proposed illuminations as training.\vspace{0mm}
+Many classic and contemporary face recognition algorithms work
+well on public data sets, but degrade sharply when they are
+used in a real recognition system. This is mostly due to the
+difficulty of simultaneously handling variations in
+illumination, image misalignment, and occlusion in the test
+image. We consider a scenario where the training images are
+well controlled, and test images are only loosely controlled.
+We propose a conceptually simple face recognition system that
+achieves a high degree of robustness and stability to
+illumination variation, image misalignment, and partial
+occlusion. The system uses tools from sparse representation to
+align a test face image to a set of frontal training images.
+The region of attraction of our alignment algorithm is computed
+empirically for public face datasets such as Multi-PIE. We
+demonstrate how to capture a set of training images with enough
+illumination variation that they span test images taken under
+uncontrolled illumination. In order to evaluate how our
+algorithms work under practical testing conditions, we have
+implemented a complete face recognition system, including a
+projector-based training acquisition system. Our system can
+efficiently and effectively recognize faces under a variety of
+realistic conditions, using only frontal images under the
+proposed illuminations as training.\vspace{0mm}
 \end{abstract}
@@
 
-\PARstart{F}{ace} recognition applications to date have fallen into roughly two
-categories.  Face recognition has recently seen a lot of success in a family of
-less-demanding applications such as online image search and family photo album
-organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto).
-At the other end of the tractability spectrum there are the terrorist watchlist
-and mass surveillance applications that have for the most part dominated the
-field of face recognition research.  However, there are many face recognition
-applications that fall roughly between these extremes, where very high recognition
-performance is desired, but the users in the gallery are still allies of the
-system rather than adversaries.  These applications include access control for
-secure facilities (e.g., prisons, office buildings), computer systems,
-automobiles, or automatic teller machines, where controlled gallery images can
-be obtained in advance.  These applications are very interesting due to their
-potential sociological impact.  Since the gallery subjects are allies, rather
-than opponents, of the recognition system, this creates the possibility of
-carefully controlling the acquisition of the training data. While the same can
-be said for other biometrics such as fingerprints and iris recognition, face
-recognition has the potential of working with test data that is much less
-controlled, allowing the access control system to be made less intrusive to the
-users of the system. To some extent, the goal of this paper is to show how a
-reliable face recognition system can be built for this restricted,
-but still important, scenario.
+\PARstart{F}{ace} recognition applications to date have fallen
+into roughly two categories.  Face recognition has recently
+seen a lot of success in a family of less-demanding
+applications such as online image search and family photo album
+organization (e.g.\ Google Picassa, Microsoft Photo Gallery,
+and Apple iPhoto). At the other end of the tractability
+spectrum there are the terrorist watchlist and mass
+surveillance applications that have for the most part dominated
+the field of face recognition research.  However, there are
+many face recognition applications that fall roughly between
+these extremes, where very high recognition performance is
+desired, but the users in the gallery are still allies of the
+system rather than adversaries.  These applications include
+access control for secure facilities (e.g., prisons, office
+buildings), computer systems, automobiles, or automatic teller
+machines, where controlled gallery images can be obtained in
+advance.  These applications are very interesting due to their
+potential sociological impact.  Since the gallery subjects are
+allies, rather than opponents, of the recognition system, this
+creates the possibility of carefully controlling the
+acquisition of the training data. While the same can be said
+for other biometrics such as fingerprints and iris recognition,
+face recognition has the potential of working with test data
+that is much less controlled, allowing the access control
+system to be made less intrusive to the users of the system. To
+some extent, the goal of this paper is to show how a reliable
+face recognition system can be built for this restricted, but
+still important, scenario.
 
@@
 
-\begin{figure} \centering \begin{tabular}{cc}
+\begin{figure}
+\centering \begin{tabular}{cc}
 \includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
@@
 illuminations are used in the training. Both are necessary for correct recognition
-using SRC.} \label{fig:promo}
-\vspace{0mm}
+using SRC.}\label{fig:promo}
+\vspace{-4mm}
 \end{figure}
@@
 \subsection{Related Work}
-
-We briefly review existing techniques for recognition, image registration, and
-handling of illumination variation.  Our system is based purely on 2D
-techniques.  This fact immediately distinguishes our approach from systems
-that either require a 3D data as an input, or attempt to estimate a 3D model
-from 2D input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
-techniques can achieve better robustness to pose variation given
-a sufficiently accurate 3D model, for access control applications where only
-moderate pose variation is present, the proposed method will be more
-than sufficient.   Note that 2D images of faces
-under varying illuminations already contain 3D shape-related information,\footnote{In principle,
-one can recover the 3D shape of the face from multiple illuminations using photometric stereo.}
-and this information can be leveraged by 2D algorithms for alignment and recognition
-even if shape is not reconstructed explicitly.
+We briefly review existing techniques for recognition, image
+registration, and handling of illumination variation.  Our
+system is based purely on 2D techniques.  This fact immediately
+distinguishes our approach from systems that either require a
+3D data as an input, or attempt to estimate a 3D model from 2D
+input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While
+these techniques can achieve better robustness to pose
+variation given a sufficiently accurate 3D model, for access
+control applications where only moderate pose variation is
+present, the proposed method will be more than sufficient.
+Note that 2D images of faces under varying illuminations
+already contain 3D shape-related information,\footnote{In
+principle, one can recover the 3D shape of the face from
+multiple illuminations using photometric stereo.} and this
+information can be leveraged by 2D algorithms for alignment and
+recognition even if shape is not reconstructed explicitly.
 %In addition, the proposed sparsity-based robust
@@
 the purpose of face recognition.
-
 \subsection{Contributions} In this paper, we show how
@@
 \begin{tabular}{ccc}
-\includegraphics[height=1.4in]{figures_pami/x_y_roa.png} &
-\includegraphics[height=1.4in]{figures_pami/y_theta_roa.png} &
-\includegraphics[height=1.4in]{figures_pami/y_s_roa.png}\\
+\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
+\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
+\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
 (a)&(b)&(c)
@@
 \centering
-\includegraphics[width=3.5in]{figures_pami/multiscale.png}
-\vspace{0mm}
+\includegraphics[width=3in]{figures_pami/multiscale.png}
+\vspace{-2mm}
 \caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
@@
 \label{fig:multiscale}
-\vspace{0mm}
+\vspace{-4mm}
 \end{figure}
 
-\noindent{3) \em 3D Pose Variation.} As densely sampled pose and
+\noindent{3) \em 3D Pose Variation.} As densely sampled pose
+and
     illumination face images are not available in any of
@@
 \begin{tabular}{ccccc}
-\includegraphics[height=0.55in]{figures_cvpr/5} &
-\includegraphics[height=0.55in]{figures_cvpr/7} &
-\includegraphics[height=0.55in]{figures_cvpr/09} &
-\includegraphics[height=0.55in]{figures_cvpr/11} &
-\includegraphics[height=0.55in]{figures_cvpr/13} \vspace{0mm} \\ 
+\includegraphics[height=0.5in]{figures_cvpr/5} &
+\includegraphics[height=0.5in]{figures_cvpr/7} &
+\includegraphics[height=0.5in]{figures_cvpr/09} &
+\includegraphics[height=0.5in]{figures_cvpr/11} &
+\includegraphics[height=0.5in]{figures_cvpr/13} \vspace{0mm} \\
 (a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=0.55in]{figures_cvpr/15} &
-\includegraphics[height=0.55in]{figures_cvpr/17} &
-\includegraphics[height=0.55in]{figures_cvpr/19} &
-\includegraphics[height=0.55in]{figures_cvpr/21} &
-\includegraphics[height=0.55in]{figures_cvpr/3} \vspace{0mm} \\ 
+\includegraphics[height=0.5in]{figures_cvpr/15} &
+\includegraphics[height=0.5in]{figures_cvpr/17} &
+\includegraphics[height=0.5in]{figures_cvpr/19} &
+\includegraphics[height=0.5in]{figures_cvpr/21} &
+\includegraphics[height=0.5in]{figures_cvpr/3} \vspace{0mm} \\
 (f) & (g) & (h) & (i) & (j) \vspace{0mm}
@@
 to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{0mm}
+}\label{fig:pose-alignment} \vspace{-4mm}
 \end{figure}
-
 \subsection{Comparison with Related Work}
@@
 \begin{figure}
-\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}\vspace{0mm}}
+\centering
+\includegraphics[height=2in]{figures_cvpr/camera_rig.pdf}
 \caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
@@
 \begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height=1.35in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.35in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
+\includegraphics[height=1.2in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
+\includegraphics[height=1.2in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
 (a) Coverage & (b) Granularity
@@
 
- In the plot for the
-coverage experiment, Figure \ref{fig:illumination-sufficiency}(a),
+In the plot for the coverage experiment, Figure
+\ref{fig:illumination-sufficiency}(a),
  we clearly see two plateau regions: one is after 4 rings
@@
 \centering
-\includegraphics[width=3.5in]{figures_cvpr/training.png}
+\includegraphics[width=3.2in]{figures_cvpr/training.png}
 \vspace{-2em}
@@
 \hline
-{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
+{Alg. 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
+\hline
+{Alg. 1} & 93.9\% & 93.8\% & 92.3\% \\
+\hline
+{Alg. 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
 \hline
@@
 \begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
+\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
 \includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
@@
 \includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
+\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
 \includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
@@
 \includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
+\hspace{-2mm}(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
 \end{tabular}
@@
 \label{fig:failed-examples}
-\vspace{-15mm}
+\vspace{-4mm}
 \end{figure}
@@
 \hline
-{Algorithm 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
+{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
 \hline
@@
 thresholding the error residuals of NN, NS, LDA, and LBP.
-
-\renewcommand{\tempwidth}[0]{1.85in}
 \begin{figure}[t]
-\vspace{-.15in}
-\centerline{
+{\small
+\centering
 \begin{tabular}{@{}c@{}c@{}}
-	\includegraphics[trim=.9in 0in 1.2in 0in, clip, width=\tempwidth]{figures_pami/pami_roc_revision2} &
-	\includegraphics[trim=.9in 0in 1.2in 0in, clip, width=\tempwidth]{figures_pami/pami_roc2} \\
-	{\small (a)} & {\small (b)} \\
+\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
+\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
+(a) & \hspace{2mm}(b) \\
 \end{tabular}
 }
+\vspace{-2mm}
 \caption{{\bf ROC curves} for subject validation on Multi-PIE database,
@@
 (b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-.4in}
+\vspace{-2mm}
 \end{figure}
@@
 the entire database is much more difficult than deciding if two
-face images are the same subject. Figure \ref{fig:roc-multipie} should not be
-confused with typical ROC curves for face similarity, e.g.,
-\cite{PhillipsP2007}.} On the left we can see that the SCI
-based recognition approach significantly outperforms the other
-algorithms, including LBP, even when all algorithms are coupled
-with our proposed iterative alignment.  In the right plot we
-again see that classical algorithms, and even LBP, are very
-sensitive to alignment.  Similar contrasts between our
-algorithm and baseline algorithms were also observed for SRC in
-\cite{Wright2009-PAMI}, though on much smaller datasets.
+face images are the same subject. Figure \ref{fig:roc-multipie}
+should not be confused with typical ROC curves for face
+similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
+that the SCI based recognition approach significantly
+outperforms the other algorithms, including LBP, even when all
+algorithms are coupled with our proposed iterative alignment.
+In the right plot we again see that classical algorithms, and
+even LBP, are very sensitive to alignment.  Similar contrasts
+between our algorithm and baseline algorithms were also
+observed for SRC in \cite{Wright2009-PAMI}, though on much
+smaller datasets.
 
@@
 recognition rate.
-\vspace{0mm}
 
@@
 
-\vspace{-.2in}
 \subsection{Recognition with Pose and Expression} We now run tests of
@@
 \vspace{0mm}
-\centering \caption{Recognition rates on our own database.}
-\vspace{0mm}
+\centering \caption{Recognition rates on our own
+database.} \vspace{0mm}
 \begin{tabular}{|c|c|c|c|c| }
@@
 \begin{figure}
-\centerline{
-\includegraphics[width=3in]{figures_pami/occ_block.png}\vspace{0mm}}
+\centering
+\includegraphics[width=2.6in]{figures_pami/occ_block.png}
 \caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.

### 2011-04-05 faces_system.tex
Hacking to get spacing acceptable

@@
 
-% make the title area
-
 \begin{IEEEkeywords}
@@
 \IEEEdisplaynotcompsoctitleabstractindextext
+\vspace{-9mm} % HACK
 
@@
 using SRC.}\label{fig:promo}
-\vspace{-4mm}
+\vspace{-6.9mm}
 \end{figure}
-%\subsection{Compound Effect of Registration and Illumination}
 
@@
 in the gallery images for good performance.
-%Furthermore, the experiments were carried out on public datasets taken under
-%controlled laboratory conditions such as Extended Yale B
-%\cite{Georghiades2001-PAMI}, which do not contain realistic test images.
-%two critical aspects of real world face recognition must be addressed:  significant variations in both the {\em image domain} and in
-%the {\em image value}.
 We illustrate the compounded effect of both alignment and illumination
@@
 recognition method proposed by \cite{Wright2009-PAMI}.
-%Unfortunately, most public face databases lack images with a
-%significant component of rear (more than 90 degrees from
-%frontal) illumination, either for training or testing.
 In this paper, we examine how to handle alignment and illumination simultaneously
@@
 
+\vspace{-0mm}
 \subsection{Related Work}
@@
 recognition even if shape is not reconstructed explicitly.
-%In addition, the proposed sparsity-based robust
-%alignment and recognition techniques are not limited to 2D data and can be easily extended
-%to 3D data if necessary.
 
@@
 (c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{0mm}
+\vspace{-8.0mm}
 \end{figure}
@@
 \end{algorithm}
+%\vspace{-4mm}
 
@@
 \label{alg:alm}
+%\vspace{-4mm}
 \end{algorithm}
@@
 
-\subsection{Simulations and Experiments on Region of
-Attraction} We will now present three experimental results
-demonstrating the effectiveness of the individual alignment
-procedure outlined in the previous section. They show the sufficiency of the
-region of attraction, verify effectiveness of the multiscale extension,
-and show stability to small pose variations.  We delay large-scale recognition experiments
-to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
+\subsection{Experiments on Region of Attraction} We will now present three
+experimental results demonstrating the effectiveness of the individual
+alignment procedure outlined in the previous section. They show the sufficiency
+of the region of attraction, verify effectiveness of the multiscale extension,
+and show stability to small pose variations.  We delay large-scale recognition
+experiments to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
 have discussed the issue of illumination in the next section.
@@
 
-\section{Handling Practical Illumination Variation}\label{sec:illumination}
+\section{Handling Illumination Variation}\label{sec:illumination}
 In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.
@@
 
-\subsection{Improving the Performance by Choosing Different Sampling Windows}
+\subsection{Improving the Sampling Window}
 Our algorithm's errors are mostly caused by a few subjects who
@@
 
-\vspace{-.1in}
+%\vspace{-.1in}
 \section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
@@
 a different camera under a variety of practical conditions.
-
 \vspace{-.1in}
+
 \subsection{Necessity of Rear Illuminations} To see how

### 2011-04-05 faces_system.tex
Added john's bio

@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{John Wright}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/john}}]{John Wright}
+John Wright received his PhD in Electrical Engineering from the University of Illinois at Urbana-Champaign in October 2009.
+He is currently a researcher in the Visual Computing group at Microsoft Research Asia.
+His research focuses on developing provably correct and efficient tools for recovering low-dimensional structure in high-dimensional datasets, even when data are missing or grossly corrupted.
+These techniques address critical estimation problems in imaging and vision applications such as automatic face recognition, video stabilization and tracking, image and data segmentation, and more.
+They also find application outside of vision, for example in web data analysis and bioinformatics.
+His work has received a number of awards and honors, including the 2009 Lemelson-Illinois Prize for Innovation for his work on robust face recognition, the 2009 UIUC Martin Award for Excellence in Graduate Research, a 2008-2010 Microsoft Research Fellowship, a Carver fellowship, and a UIUC Bronze Tablet award.
+\end{IEEEbiography}
 

### 2011-04-05 faces_system.tex
Working on bios

--- 2011-04-05 faces_system.tex
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Andrew Wagner}
-Andrew Wagner received his Bachelor's degree in General Engineering,
-and his Master's degree in Electrical Engineering, from the University
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
+received his Bachelor's degree in General Engineering in 2003,
+and his Master's degree in Electrical Engineering in 2006, from the University
 of Illinois at Urbana-Champaign, where he is currently a Ph.D.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{John Wright}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/john}}]{John
+Wright} received his PhD in Electrical Engineering from the
+University of Illinois at Urbana-Champaign in October 2009. He is currently a
+researcher in the Visual Computing group at Microsoft Research Asia.  His
+research focuses on developing provably correct and efficient tools for
+recovering low-dimensional structure in corrupted high-dimensional datasets.
+His work has received a number of awards, including the 2009 Lemelson-Illinois
+Prize for Innovation, the 2009 UIUC Martin Award for Excellence in Graduate
+Research, a 2008-2010 Microsoft Research Fellowship, a Carver fellowship, and a
+UIUC Bronze Tablet award.  \end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
+received his Bachelor's and Master's degrees, both in Electrical
 Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
-
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
+received... {\bf PLACEHOLDER}
+\end{IEEEbiography}
 
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Hossein Mobahi}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Yi Ma}
-Arvind Ganesh received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
+received... {\bf PLACEHOLDER}
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
+Ma} received two Bachelors degree in Automation and Applied Mathematics
+from  Tsinghua University, Beijing, China, in 1995. He received an Master
+degree in Electrical Engineering and Computer Sciences (EECS) in 1997, a second
+Master degree in Mathematics in 2000, and the Ph.D. degree in EECS in 2000, all
+from the University of California at Berkeley. He is currently an associate
+professor (with tenure) at the Department of Electrical and Computer
+Engineering, University of Illinois at Urbana-Champaign, and since January 2009
+has also served as research manager for the Visual Computing Group at Microsoft
+Research Asia, Beijing, China.  \end{IEEEbiography}
+
+%He is an associate
+%editor of the  IEEE Transactions on Pattern Analysis and Machine Intelligence
+%since 2007 and the International Journal of Computer Vision since 2010. He has
+%also served as the chief guest editor for special issues for the Proceedings of
+%IEEE and the IEEE Signal Processing Magazine. He serves as Area Chair for NIPS
+%2011, Program Board Memember for ICCV 2011, and Program Chair for ICCV 2013.
 
 % if you will not have a photo
-%\begin{biographynophoto}{John Doe}
-%Biography text here.
+%\begin{biographynophoto}{John Doe} Biography text here.
 %\end{biographynophoto}

### 2011-04-05 faces_system.tex
Working on biography alignment
Added hossein's bio

@@
 
-% insert where needed to balance the two columns on the last page
-%\newpage
-
 % trigger a \newpage just before the given reference
@@
 % the document is modified later
-%\IEEEtriggeratref{8}
+%\IEEEtriggeratref{40}
 % The "triggered" command can be changed if desired:
@@
 
+\newcommand{\biospace}{\vspace{-2em}}
+
 % biography section
-%
-% If you had an eps/pdf photo file (graphicx package needed)
-% the extra braces prevent the LaTeX parser from getting confused
-% when it sees the complicated \includegraphics command within an
-% optional argument. You can create your own macro to make things
-% simpler here.
-
+\biospace
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
@@
 
+\biospace
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/john}}]{John
@@
 
+\biospace
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
@@
 
+\biospace
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Hossein Mobahi}
-received... {\bf PLACEHOLDER}
-\end{IEEEbiography}
-
+\biospace
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]
+{bios/hossein}}]{Hossein Mobahi} received his Bachelor's and Master's degrees
+in Iran, both in Computer Engineering from Azad University (Tehran-South) in
+2003  and University of Tehran in 2005 respectively. He is currently a PhD
+candidate in the Computer Science Department at the University of Illinois,
+Urbana-Champaign. His research interests include pattern classification,
+clustering and optimization.  His recent research focuses on iterative
+smoothing for image alignment.  \end{IEEEbiography}
+
+\biospace
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
@@
 Research Asia, Beijing, China.  \end{IEEEbiography}
-
-%He is an associate
-%editor of the  IEEE Transactions on Pattern Analysis and Machine Intelligence
-%since 2007 and the International Journal of Computer Vision since 2010. He has
-%also served as the chief guest editor for special issues for the Proceedings of
-%IEEE and the IEEE Signal Processing Magazine. He serves as Area Chair for NIPS
-%2011, Program Board Memember for ICCV 2011, and Program Chair for ICCV 2013.
-
-% if you will not have a photo
-%\begin{biographynophoto}{John Doe} Biography text here.
-%\end{biographynophoto}
-
-%\begin{biographynophoto}{Jane Doe}
-%Biography text here.
-%$\end{biographynophoto}
+\vfill
 
@@
 
-%\vfill
-
 % Can be used to pull up biographies so that the bottom of the last one

### 2011-04-05 bios/bios.tex
Moved bios to separate file

--- 2011-04-05 faces_system.tex
@@
-%% bare_jrnl.tex
-%% V1.1
-%% 2002/08/13
-%% by Michael Shell
-%% mshell@ece.gatech.edu
-%%
-%% NOTE: This text file uses MS Windows line feed conventions. When (human)
-%% reading this file on other platforms, you may have to use a text
-%% editor that can handle lines terminated by the MS Windows line feed
-%% characters (0x0D 0x0A).
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.6 or later) with an IEEE journal paper.
-%%
-%% Support sites:
-%% http://www.ieee.org
-%% and/or
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
-%%
-%% This code is offered as-is - no warranty - user assumes all risk.
-%% Free to use, distribute and modify.
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% Testflow can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/testflow
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print. Authors are encouraged to use U.S. letter paper when
-% submitting to IEEE. Use the testflow package mentioned above to verify
-% correct handling of both paper sizes by the author's LaTeX system.
-%
-% Also note that the "draftcls", not "draft", option should be used if
-% it is desired that the figures are to be displayed in draft mode.
-
-% *************************************************************************
-% You can now use this template both for submitting to peer review and,
-% if your paper is accepted, for sending in final publication materials.
-% To change from peer review format (single column, double-spaced) to
-% final publication format (double column, single-spaced), just move the
-% comment-line sign (%) from one \documentclass line to the other.
-% The first is for peer review format(single column), the second is for final publication(double column).
-
-%\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
-%\documentclass[9.5pt,journal,final,compsoc,twocolumn]{IEEEtran}
-
-\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran} % from the new style file.  supposedly double column.
-
-% *************************************************************************
-
-\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amscd}
-\usepackage{amssymb}
-\usepackage{subfigure}
-\usepackage{pdfsync}
-\usepackage{helvet}
-\usepackage{url}
-\usepackage{color}
-
-% Include other packages here, before hyperref.
-\usepackage{mydefs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-\usepackage{multirow}
-\usepackage{pict2e}
-\usepackage{slashbox}
-
-% \usepackage[small]{caption}
-
-\newcommand{\subj}{\textup{subj to}}
-
-\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
-%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
-%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}
-
-\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}
-
-\begin{document}
-%
-% paper title
-\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
-Illumination by Sparse Representation}
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
-        John~Wright,~\IEEEmembership{Member,~IEEE,}
-        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
-        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
-        Hossein~Mobahi,
-        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{A.
-Wagner, A.
-Ganesh, Z.
-Zhou, and Y.
-Ma are with the Dept.
-of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign.
-H.
-Mobahi is with the Computer Science Dept.
-at the University of Illinois at Urbana-Champaign.
-J.
-Wright and Y.
-Ma are with Microsoft Research Asia.
-Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W.
-Main st.
-Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-
-% If you want to put a publisher's ID mark on the page
-% (can leave text blank if you just want to see how the
-% text height on the first page will be reduced by IEEE)
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
-
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%% ABSTRACT
-\IEEEcompsoctitleabstractindextext{
-\vspace{0mm}
-\begin{abstract}
-Many classic and contemporary face recognition algorithms work
-well on public data sets, but degrade sharply when they are
-used in a real recognition system. This is mostly due to the
-difficulty of simultaneously handling variations in
-illumination, image misalignment, and occlusion in the test
-image. We consider a scenario where the training images are
-well controlled, and test images are only loosely controlled.
-We propose a conceptually simple face recognition system that
-achieves a high degree of robustness and stability to
-illumination variation, image misalignment, and partial
-occlusion. The system uses tools from sparse representation to
-align a test face image to a set of frontal training images.
-The region of attraction of our alignment algorithm is computed
-empirically for public face datasets such as Multi-PIE. We
-demonstrate how to capture a set of training images with enough
-illumination variation that they span test images taken under
-uncontrolled illumination. In order to evaluate how our
-algorithms work under practical testing conditions, we have
-implemented a complete face recognition system, including a
-projector-based training acquisition system. Our system can
-efficiently and effectively recognize faces under a variety of
-realistic conditions, using only frontal images under the
-proposed illuminations as training.\vspace{0mm}
-\end{abstract}
-
-\begin{IEEEkeywords}
-Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
-Sparse Representation,  Error Correction, Validation and Outlier Rejection.
-\end{IEEEkeywords}}
-
-\maketitle
-
-\IEEEdisplaynotcompsoctitleabstractindextext
-\vspace{-9mm} % HACK
-
-\IEEEpeerreviewmaketitle
-
-%%%%%%%%% BODY TEXT
-\section{Introduction}
-%Automatic face recognition remains one of the most active areas in computer vision.
-
-\PARstart{F}{ace} recognition applications to date have fallen
-into roughly two categories.  Face recognition has recently
-seen a lot of success in a family of less-demanding
-applications such as online image search and family photo album
-organization (e.g.\ Google Picassa, Microsoft Photo Gallery,
-and Apple iPhoto). At the other end of the tractability
-spectrum there are the terrorist watchlist and mass
-surveillance applications that have for the most part dominated
-the field of face recognition research.  However, there are
-many face recognition applications that fall roughly between
-these extremes, where very high recognition performance is
-desired, but the users in the gallery are still allies of the
-system rather than adversaries.  These applications include
-access control for secure facilities (e.g., prisons, office
-buildings), computer systems, automobiles, or automatic teller
-machines, where controlled gallery images can be obtained in
-advance.  These applications are very interesting due to their
-potential sociological impact.  Since the gallery subjects are
-allies, rather than opponents, of the recognition system, this
-creates the possibility of carefully controlling the
-acquisition of the training data. While the same can be said
-for other biometrics such as fingerprints and iris recognition,
-face recognition has the potential of working with test data
-that is much less controlled, allowing the access control
-system to be made less intrusive to the users of the system. To
-some extent, the goal of this paper is to show how a reliable
-face recognition system can be built for this restricted, but
-still important, scenario.
-
-Very few recognition systems specifically target applications where many well-
-controlled training images are available.  Of these, the classical holistic
-subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
-are well known for their speed and simplicity, as well as for their natural
-extension to linear illumination models.  However, their performance has been
-shown to be extremely brittle not only to alignment variation, but to even
-minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
-is slightly open. One promising recent direction, set forth in
-\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
-sparse representation of the test image in terms of the training set as a
-whole, up to some sparse error due to occlusion. A \emph{sparse
-representation-based classification} (SRC) method is then proposed for
-recognition. The main idea is that the sparse nonzero coefficients should
-concentrate on the training samples with the same class label as the test
-sample. SRC has demonstrated striking recognition performance despite severe
-occlusion or corruption by solving a simple convex program.
-%For this reason, the final recognition stage of our proposed system will
-%consist of SRC, and the alignment stage that precedes it will be based on
-%similar techniques.
-
-\begin{figure}
-\centering \begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
-\end{tabular} \caption{{\bf Effects of registration and illumination on
-Recognition}. In this example we identify the girl among 20 subjects, by
-computing the sparse representation of her input face with respect to the
-entire training set. The absolute sum of the coefficients associated with each
-subject is plotted on the right. We also show the faces reconstructed with each
-subject's training images weighted by the associated sparse coefficients. The
-red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
-input face is from Viola and Jones' face detector (the black box) and all 38
-illuminations specified in Section \ref{sec:illumination} are used in the
-training.  {\bf Middle:} The input face is well-aligned (the white box) with
-the training by our algorithm specified in Section \ref{sec:registration} but
-only 24 frontal illuminations are used in the training for recognition (see
-Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and
-a sufficient set (all 38) of
-illuminations are used in the training. Both are necessary for correct recognition
-using SRC.}\label{fig:promo}
-\vspace{-6.9mm}
-\end{figure}
-
-Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
-between the test and training images, and it requires a rich set of illuminations
-in the gallery images for good performance.
-We illustrate the compounded effect of both alignment and illumination
-with an example in Figure \ref{fig:promo}.
-The task is to identify the girl among 20 subjects. If the test
-face image, say obtained from an off-the-shelf face detector,
-has even a small amount of registration error against the
-training images (caused by mild pose, scale, or misalignment),
-the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative,
-even if sufficient illuminations are present in the training, as shown in Figure
-\ref{fig:promo}(top). Additionally, in order to span
-the illuminations of a typical indoor (or outdoor) environment,
-illuminations from behind the subject are needed in the
-training set. Otherwise, even for perfectly aligned test images,
-the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
-informative, as shown by the example in Figure
-\ref{fig:promo}(middle). Clearly, both good alignment, as well
-as sufficient training images are needed to ensure success of the sparsity-based
-recognition method proposed by \cite{Wright2009-PAMI}.
-In this paper, we examine how to handle alignment and illumination simultaneously
-in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
-closer to practical use.
-
-\vspace{-0mm}
-\subsection{Related Work}
-We briefly review existing techniques for recognition, image
-registration, and handling of illumination variation.  Our
-system is based purely on 2D techniques.  This fact immediately
-distinguishes our approach from systems that either require a
-3D data as an input, or attempt to estimate a 3D model from 2D
-input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While
-these techniques can achieve better robustness to pose
-variation given a sufficiently accurate 3D model, for access
-control applications where only moderate pose variation is
-present, the proposed method will be more than sufficient.
-Note that 2D images of faces under varying illuminations
-already contain 3D shape-related information,\footnote{In
-principle, one can recover the 3D shape of the face from
-multiple illuminations using photometric stereo.} and this
-information can be leveraged by 2D algorithms for alignment and
-recognition even if shape is not reconstructed explicitly.
-
-% Alignment
-In holistic recognition algorithms, correspondence between points in the test
-image and in the training must be achieved.  A long line of research exists on
-using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
-Active Shape Models \cite{cootes1992active} to register images against a
-relatively high-dimensional model of plausible face appearances, often
-leveraging face-specific contours.  While these model-based techniques have
-advantages in dealing with variations in expression and pose, they may add
-unnecessary complexity to applications where subjects normally present a neutral face
-or only have moderate expression. We prefer to focus on deformations with far fewer
-degrees of freedom, i.e. similarity transformations, and to use the training images themselves
-as the appearance model. Iterative registration in this spirit dates at least back to the
-Lucas-Kanade algorithm \cite{lucas1981iterative}.
-
-Whereas much of the early work on image registration is aimed at the
-problem of registering nearly identical images, say by minimizing a sum of
-squared distances or maximizing normalized correlation, here we must confront several physical factors
-simultaneously: misalignment, illumination variations, and corrupted pixels.
-As we discuss further below, illumination variation can be dealt by
-expressing the test image as a linear combination of an appropriate set of
-training images. Similar representations have been exploited in
-illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
-For robustness to gross errors, the $\ell^1$-norm of the residual is a more
-appropriate objective function than the classical $\ell^2$-norm. Its use here
-is loosely motivated by theoretical results due to Cand\`{e}s and Tao
-\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
-lead us to pose the registration problem as the search for a set of
-transformations and illumination coefficients that minimize the $\ell^1$-norm
-of the representation error. We solve this problem using a generalized
-Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
-minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
-these problems can also be solved efficiently using recently developed
-first-order techniques for $\ell^1$-minimization, which are reviewed in
-\cite{YangA2010-pp}.
-
-% Illumination
-Researchers have tried various techniques to deal with illumination variation.
-In almost all recognition algorithms where only a single gallery image is
-available per individual, illumination effects are regarded as a nuisance that
-must be removed before the algorithm can continue.  This is typically done by
-making statistical assumptions about how illumination affects the image, and
-using those assumptions to extract a new representation that is claimed to be
-illumination invariant.  Recent examples include \cite{chen2006total} and
-\cite{zhou2007appearance}.  However, despite these efforts, truly
-illumination-invariant features are difficult to obtain from a single input
-image.  We argue that if one has the luxury of designing the acquisition system
-and the application demands a high recognition rate,
-it is then unwise to limit the gallery to a
-single image per person.  We therefore take the strategy of sampling many
-gallery images of each individual under varying illuminations.  These images
-are used as the basis for either a convex cone model
-\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
-\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
-projector based light stage.  While similar systems have been used for
-other applications, to our knowledge, we
-are the first to use projectors to indirectly illuminate a subject's face for
-the purpose of face recognition.
-\subsection{Contributions} In this paper, we show how
-registration and illumination
-can be simultaneously addressed within a robust sparse representation
-framework. We show that face registration, a challenging
-nonlinear problem, can be solved by a series of linear programs
-that iteratively minimize the sparsity of the registration
-error. This leads to an efficient and effective alignment
-algorithm for face images that works for a large range of
-variation in translation, rotation, and scale, even when
-the face is only partially visible due to eyeglasses, closed
-eyes and open mouth, sensor saturation, etc.  We also propose a
-sufficient set of training illuminations
-that is capable of linearly representing typical indoor and outdoor
-lighting, along with a practical hardware system for capturing
-them.
-
-% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.
-While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.
-
-% We use tools from sparse representation
-% to align a test face image with a set of frontal training
-% images in the presence of registration error, illumination variation, and partial occlusion. The region
-% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
-% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
-% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions.
-We show
-% that our system can efficiently and effectively recognize faces under
-% a variety of realistic conditions, using only
-% frontal images under the proposed illuminations as training.
-
-%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.
-We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.
-Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.
-We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.
-We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.
-
-We then demonstrate the effectiveness of the proposed new
-methods with a complete face recognition system that is {\em
-simple, stable, and scalable}. The proposed system performs
-robust automatic recognition of subjects from loosely
-controlled probe images taken both indoors and outdoors,
-using a gallery of
-frontal views of the subjects' faces under the proposed
-illuminations. An off-the-shelf face
-detector\footnote{We use the OpenCV
-implementation of the Viola and Jones' face detector
-\cite{Viola2004-IJCV}.} is used to detect faces in the test images.
-
-We conduct extensive experiments on the proposed system with
-both public databases and a face database that is collected by
-our own acquisition system. Our experimental results on
-large-scale public face databases show that our algorithm
-indeed achieves very good performance on these databases,
-exceeding or competing with the state-of-the-art algorithms.
-Additionally, our experimental results on our own database
-clearly demonstrate that our system not only works well with
-images taken under controlled laboratory conditions, but is
-capable of handling practical indoor and outdoor illuminations as well.
-
-\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
-we derive our robust registration and recognition algorithm within the sparse
-representation framework. We elaborate on algorithmic implementation issues,
-conduct region of attraction experiments with respect to both 2D in-plane
-deformation and 3D pose variation, and discuss its relationship to existing
-work. Section \ref{sec:illumination} is dedicated to our training acquisition
-system. Using this system, we investigate empirically how many training
-illuminations are required to handle practical illumination variations, and
-suggest a sufficient set of 38 training illuminations. Extensive experiments on
-a large-scale public database and on our own database are conducted in Section
-\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
-proposed system. Section \ref{sec:conclusion} concludes our work with
-discussion of promising future directions.
-
-\section{Robust Alignment}\label{sec:registration}
-As demonstrated in Figure \ref{fig:promo}(top), the main
-limitation of the {\em sparse representation and
-classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
-the assumption of pixel-accurate alignment between the test
-image and the training set. This leads to brittleness under
-pose and misalignment, making it inappropriate for deployment
-outside a laboratory setting. In this section, we show how this
-weakness can be rectified while still preserving the conceptual
-simplicity and good recognition performance of SRC.
-
-SRC assumes access to a database of multiple registered
-training images per subject, taken under varying illuminations.
-The images of subject $i$, stacked as vectors, form a matrix
-$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
-form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
-\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
-well-aligned test image $\y_0$ can be represented as a sparse
-linear combination $A \x_0$ of all of the images in the
-database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
-sufficiency in the next section.} plus a sparse error $\e_0$
-due to corrupted pixels. The sparse representation can be recovered by
-minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
-vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
-$\x$ and $\e$:
-\begin{equation}
-\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
-\label{eqn:robust-l1}
-\end{equation}
-Now suppose that $\y_0$ is subject to some pose or
-misalignment, so that instead of observing $\y_0$, we observe
-the warped image $\y = \y_0 \circ \tau^{-1}$, for some
-transformation $\tau \in T$ where $T$ is a finite-dimensional
-group of transformations acting on the image domain.  The
-transformed image $\y$ no longer has a sparse representation of
-the form $\y = A \x_0 + \e_0$, and naively applying the
-algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
-as seen in Figure \ref{fig:promo}(top).
-
-\subsection{Batch and Individual Alignment} If the
-true deformation $\tau^{-1}$ can be found, then we can apply
-its inverse $\tau$ to the test image and it again becomes
-possible to find a sparse representation of the resulting
-image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
-  This sparsity
-provides a strong cue for finding the correct deformation
-$\tau$: conceptually, one would like to seek a transformation
-$\tau$ that allows the sparsest representation, via
-\begin{equation} \label{eqn:L1-L1-conceptual}
-\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
-\end{equation}
-For fixed $\tau$, this problem is jointly convex in $\x$ and
-$\e$. However, as a simultaneous optimization over the
-coefficients $\x$, error representation $\e$, and
-transformation $\tau$, it is a difficult, nonconvex
-optimization problem. One source of difficulty is the presence
-of multiple faces in the matrix $A$:
-\eqref{eqn:L1-L1-conceptual} has many local minima that
-correspond to aligning $\y$ to different subjects. In this
-sense, the misaligned recognition problem differs from the
-well-aligned version studied in \cite{Wright2009-PAMI}. For the
-well-aligned case, it is possible to directly solve for a
-global representation, with no concern for local minima. With
-possible misalignment, it is more appropriate to seek the best
-alignment of the test face with each subject $i$:
-\begin{equation} \label{eqn:per-subject-L1}
-\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
-\end{equation}
-We no longer penalize $\| \x \|_1$, since $A_i$ consists of
-only images of subject $i$ and so $\x$ is no longer expected to
-be sparse.
-
-\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector.
-We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
-\begin{equation}
-\y\circ \tau + J \Delta \tau = A_i \x + \e.
-\end{equation}
-Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
-Jacobian of $\y \circ \tau$ with respect to the transformation
-parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
-above equation is underdetermined if we allow the registration
-error $\e$ to be arbitrary. At the correct alignment we
-expect the test image to differ from $A_i \x$ only
-for the minority of the pixels corrupted by occlusions. Thus,
-we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
-\begin{equation}
-\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
-\label{eqn:L1-align}
-\end{equation}
-This is different from the popular choice that
-minimizes the $\ell^2$-norm of the registration error:
-\begin{equation}
-\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
-\label{eqn:L2-align}
-\end{equation}
-which is also equivalent to finding the deformation step
-$\Delta  \tau$ by solving the least-square problem:
-$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
-\|_2$. Empirically, we find that if there is only small noise
-between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
-$\eqref{eqn:L2-align}$ have similar performance.  However, if
-there are occlusions in $\y_0$, sequential
-$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
-better than sequential $\ell^2$-minimization
-\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
-example.
-
-The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images.
-Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}.
-While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm.
-In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution.
-We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.
-
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{cccc}
-\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
-\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
-\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
-\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
-(a) & (b) & (c) & (d)
-\end{tabular}
-\vspace{0mm}}
-\caption{{\bf Comparing alignment of a subject wearing sunglasses by
-$\ell^1$ and $\ell^2$ minimization.}
-{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
-result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
-given by the face detector, {\em Red (solid):} Alignment result shown on the same
-face; (b) warped testing image using the estimated transformation $\y_0$;
-(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-8.0mm}
-\end{figure}
-
-In addition to normalizing the training images (which is done
-once), it is important to normalize the warped testing image
-$\y \circ \tau$ as the algorithm runs.  Without normalization,
-the algorithm may fall into a degenerate global minimum
-corresponding to zooming in on a dark region of the test
-image.  Normalization is done by replacing the linearization of
-$\y \circ \tau$ with a linearization of the normalized version
-$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
-The proposed alignment algorithm can be easily extended to work
-in a {\em multiscale} fashion, with benefits both in
-convergence behavior and computational cost.  The alignment
-algorithm is simply run to completion on progressively less
-downsampled versions of the training and testing images, using
-the result of one level to initialize the next.
-
-\subsection{Robust Recognition by Sparse Representation} Once
-the best transformation $\tau_i$ has been computed for each
-subject $i$, the training sets $A_i$ can be aligned to $\y$,
-and a global sparse representation problem of the form
-\eqref{eqn:robust-l1} can be solved to obtain a discriminative
-representation in terms of the entire training set. Moreover,
-the per-subject alignment residuals $\| \e \|_1$ can be used to
-prune unpromising candidates from the global optimization,
-leaving a much smaller and more efficiently solvable problem.
-The complete optimization procedure is summarized as Algorithm
-\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
-considered together to provide a sparse representation for the
-test image. If $S = 1$, the algorithm reduces to classification
-by registration error; but considering the test image might be
-an invalid subject, we typically choose $S = 10$. Since valid
-images have a sparse representation in terms of this larger
-set, we can reject invalid test images using the {\em sparsity
-concentration index} proposed in \cite{Wright2009-PAMI}.
-The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
-selects coefficients from the vector $\x$ corresponding to subject $i$.
-
-Another important free parameter in Algorithm
-\ref{alg:deformable-src} is the class of deformations $T$. In
-our experiments, we typically use 2D similarity
-transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
-Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
-detector, or 2D projective transformations, $T =
-\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
-This class of transformations is able to represent distortion
-in a perspective image of a planar object.}$, for handling some
-pose variation.
-
-In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
-performance of our system, based on the observation that the face detector output may be poorly centered on
-the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
-instead of aligning the training sets to the
-original $\y$ directly obtained from the face detector, we
-compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
-\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
-is believed to be better centered, and update $\y$ according to
-$\bar{\tau}$. For the 2D similarity transformations, which are
-used in our system when initialized by the face detector, a
-transformation $\tau$ can be parameterized as $\tau = (\tau^1,
-\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
-the translations in $x$- and $y$-axis, $\tau^3$ represents the
-rotation angle and $\tau^4$ represents the scale. Then the
-average transformation is simply obtained by taking the
-component-wise mean:
-\begin{displaymath}
-\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
-\tau_{k_S}^i) / S, i = 1,2,3,4.
-\end{displaymath}
-Finally, the training sets are aligned to the new $\y$.
-
-\begin{algorithm}[thb]
-\caption{\bf (Deformable Sparse Recovery and Classification for
-Face Recognition)} \label{alg:deformable-src}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
-$\y\in\Re^m$ and a deformation group $T$.
-\STATE
-{\bf for} each subject $i$,
-\STATE \hspace{3mm} $\tau^{(0)}
-\leftarrow I$.
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
-\STATE \hspace{6mm}
-$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
-\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
-\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
-\STATE
-\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$;
-\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-residuals $\|\e\|_1$. \STATE Compute an average transformation
-$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
-\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem: \hspace{2mm}
-$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
-\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
-\end{small}
-\end{algorithmic}
-\end{algorithm}
-%\vspace{-4mm}
-
-The transformation $\tau$ defines a mapping between the
-coordinates of pixels in the large original image and a smaller
-(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
-artifacts in the downsampled image, one should apply a
-smoothing filter to the original image. For a simple implementation, a rectangular window with
-regular sampling can used, but in general, the small image need
-not be regularly sampled in pixel coordinates.  For example,
-the sample locations could be arbitrarily selected from within a
-``face shaped'' area. We will discuss how choosing
-different windows can affect the performance of our algorithm
-in Section \ref{sec:multipie}.
-
-\subsection{System Implementation}
-In this section, we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}. It is
-not hard to see that its computational complexity is dominated
-by the two steps where the $\ell^1$-norm minimization problems
-are solved; namely Step 6 for iterative registration, and Step
-14 for global sparse representation.
-Fortunately, many fast algorithms for solving these problems have been proposed over the
-past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
-That work suggests that \emph{Augmented Lagrange Multiplier}
-(ALM) algorithms \cite{Bertsekas1982} strike a good balance
-between scalability and accuracy: as first order methods, they
-require only lightweight vector operations and matrix-vector
-multiplications at each iteration, making them preferable to
-more classical solutions such as interior point methods.
-However, compared to other first-order methods, they achieve
-higher accuracy with a fixed computational budget.
-
-We use Step 14 as an example to illustrate the ALM
-method, since solving Step 6 is very similar. Recall that in
-Step 14 the problem we are interested in is:
-\begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
-A \x + \e.
-\end{equation}
-Its corresponding augmented Lagrangian function is
-\begin{equation}
-L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
-\end{equation}
-where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
-penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
-(\x,\e,\blamda)$ by alternating between optimizing with respect
-to the primal variables $\x, \e$ and updating the dual variable
-$\blamda$, with the other fixed, as follows:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:alm}
-\end{equation}
-Although updating $\blamda$ is trivial,
-minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
-$\x$ and $\e$ could still be costly. To further reduce the
-complexity of the problem, we adopt an approach used in
-\cite{YangJ2009-pp}, called \emph{alternating direction
-method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
-over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
-\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:adm}
-\end{equation}
-As the objective function is convex and alternation is between two
-terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).
-
-In order to discuss the solution to the above subproblems, we
-need to define the following soft-thresholding operator for a
-vector $\x$ and a scalar $\alpha \geq 0$:
-\begin{equation}
-\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
-\end{equation}
-where all the operations are performed component-wise. It is
-easy to show that the subproblem with respect to $\e$ has a
-closed-form solution given by the soft-thresholding operator:
-\begin{equation}
-\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
-\end{equation}
-To solve the subproblem associated with $\x$, we
-apply a first-order $\ell^1$-minimization method,
-called \emph{fast iterative shrinkage-threshold algorithm}
-(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
-iteratively minimize a quadratic approximation $Q(\x, \z)$ to
-$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
-carefully chosen in order to achieve a good convergence
-rate. We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
-\begin{algorithm}[t]
-\caption{\bf (Augmented Lagrange Multiplier Method for Global
-Recognition)}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
-\mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
-\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
-\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
-\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
-\ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{small}
-\end{algorithmic}
-\label{alg:alm}
-%\vspace{-4mm}
-\end{algorithm}
-
-We have selected this algorithm because it strikes the best
-balance between speed, accuracy, and scalability for our problem out of
-many algorithms that we have tested. We refer the interested reader to
-\cite{YangA2010-pp} for a more detailed discussion of competing
-approaches.  On a Mac Pro with
-Dual-Core 2.66GHz Xeon processors and 4GB memory,
-running on our database containing images size $80\times 60$
-pixels from 109 subjects under 38 illuminations,
-our C implementation of Algorithm~\ref{alg:deformable-src} takes
-about 0.60 seconds per subject for alignment and about 2.0
-seconds for global recognition. Compared to the highly
-customized interior point method used in the conference version
-of this paper \cite{Wagner2009-CVPR}, this new algorithm is
-only slightly faster for per subject alignment. However, it is
-much simpler to implement and it achieves a
-\emph{speedup of more than a factor of 10} for global
-recognition!
-
-\subsection{Experiments on Region of Attraction} We will now present three
-experimental results demonstrating the effectiveness of the individual
-alignment procedure outlined in the previous section. They show the sufficiency
-of the region of attraction, verify effectiveness of the multiscale extension,
-and show stability to small pose variations.  We delay large-scale recognition
-experiments to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
-have discussed the issue of illumination in the next section.
-
-\noindent {1) \em 2D Deformation.}  We first verify the
-    effectiveness of our alignment algorithm with images
-    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
-    We select all the subjects in Session 1, use 7
-    illuminations per person from Session 1 for training,
-    and test on one new illumination from Session
-    2.\footnote{The training are illuminations $\{0, 1, 7,
-    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
-    testing is the illumination 10. } We manually select
-    eye corners in both training and testing as the ground
-    truth for registration. We downsample the images to
-    $80\times 60$ pixels\footnote{Unless otherwise stated,
-    this will be the default resolution at which we prepare
-    all our training and testing datasets.} and the distance between the two outer
-    eye corners is normalized to be 50 pixels for each
-    person. We introduce artificial deformation to the
-    testing image with a combination of translation,
-    rotation and scaling. We further use the alignment
-    error $\|\e\|_1$ as an indicator of success. Let $r_0$
-    be the alignment error obtained by aligning a test
-    image to the training images without any artificial
-    perturbation. When the test image is artificially
-    perturbed and aligned, resulting in an alignment error
-    $r$, we consider the alignment successful if $|r - r_0 | \leq
-    0.01r_0$. Figure \ref{fig:attraction} shows the
-    percentage of successful registrations for all subjects
-    for each artificial deformation. The results suggest
-    that our algorithm works extremely well with
-    translation up to 20\% of the eye distance (or 10
-    pixels) in all directions and up to $30^\circ$ in-plane
-    rotation. We have also tested our alignment algorithm
-    with scale variation and it can handle up to 15\%
-    change in scale.
-
-We have gathered the statistics of the Viola and Jones'
- face detector on the Multi-PIE dataset. For 4,600 frontal
- images of 230 subjects under 20 different illuminations,
- using manual registration as the ground truth, the average
- misalignment error of the detected faces is about 6 pixels
- and the average variation in scale is 8\%. This falls
- safely inside the region of attraction for our alignment
- algorithm.
-\begin{figure*}
-\centering
-{\small
-\begin{tabular}{ccc}
-\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
-(a)&(b)&(c)
-\end{tabular}
-\begin{tabular}{cccc}
-\includegraphics[height=1in]{figures_pami/x_tr.png} &
-\includegraphics[height=1in]{figures_pami/y_tr.png} &
-\includegraphics[height=1in]{figures_pami/theta.png} &
-\includegraphics[height=1in]{figures_pami/scale.png}\\
-(d)&(e)&(f)& (g)
-\end{tabular}
-\vspace{0mm}
-}
-\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.
-The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees.
-{\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions.
-(b) Simultaneous translation in $y$ direction and in-plane rotation.
-(c) Simultaneous translation in $y$ direction and scale variation.
-{\bf Bottom row:} (d) Translation in $x$ direction only.
-(e) Translation in $y$ direction only.
-(f) In-plane rotation only.
-(g) Scale variation only.}
-\label{fig:attraction}
-\vspace{0mm}
-\end{figure*}
-
-\noindent{2) \em Multiscale Implementation.}
-Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost.
-Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
-We use the same 7 illuminations in
-Session 1 as training, and all 20 illuminations in the same
-session as testing. We introduce artificial deformation in
-both $x$ and $y$ directions up to 16 pixels in the
-$80\times 60$ frame, with a step size of 4 pixels, i.e.,
-$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
-\{-16,-12,\ldots,12,16\}$. We consider an alignment
-successful if the estimated coordinates of the eye-corners
-are within 1 pixel from the ground truth in the original
-image.  In Figure \ref{fig:multiscale}, we report the
-alignment success rate, averaged over the artificially
-perturbed initial deformations, as a function of the
-standard deviation of the Gaussian kernel $\sigma$, for
-three choices of the number of scales. As one can see,
-using multiscale indeed improves the performance, and when
-3 scales are used, a smaller convolution kernel can achieve
-a similar performance compared to a much larger kernel when
-only 2 scales are used.
-\begin{figure}
-\centering
-\includegraphics[width=3in]{figures_pami/multiscale.png}
-\vspace{-2mm}
-\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
-A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
-\label{fig:multiscale}
-\vspace{-4mm}
-\end{figure}
-
-\noindent{3) \em 3D Pose Variation.} As densely sampled pose
-and
-    illumination face images are not available in any of
-    the public databases, including Multi-PIE, we have
-    collected our own dataset using our own system (to be
-    introduced in the next section). We use frontal face
-    images of a subject under the 38 illuminations proposed
-    in the next section as training. For testing, we
-    collect images of the subject under a typical indoor
-    lighting condition at pose ranging from $-90^\circ$ to
-    $+90^\circ$ with step size 5.625$^\circ$, a total of 33
-    poses. We use Viola and Jones' face detector to
-    initialize our alignment algorithm.
-Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
-to $\pm 45^\circ$.
-Note that this level of out-of-plane
- pose variation is beyond what we intend to handle with our formulation.
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{ccccc}
-\includegraphics[height=0.5in]{figures_cvpr/5} &
-\includegraphics[height=0.5in]{figures_cvpr/7} &
-\includegraphics[height=0.5in]{figures_cvpr/09} &
-\includegraphics[height=0.5in]{figures_cvpr/11} &
-\includegraphics[height=0.5in]{figures_cvpr/13} \vspace{0mm} \\
-(a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=0.5in]{figures_cvpr/15} &
-\includegraphics[height=0.5in]{figures_cvpr/17} &
-\includegraphics[height=0.5in]{figures_cvpr/19} &
-\includegraphics[height=0.5in]{figures_cvpr/21} &
-\includegraphics[height=0.5in]{figures_cvpr/3} \vspace{0mm} \\
-(f) & (g) & (h) & (i) & (j) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
-to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-4mm}
-\end{figure}
-\subsection{Comparison with Related Work}
-Our modification to SRC roots solidly in the tradition of
-adding deformation-robustness to face recognition algorithms
-\cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}.
-However, the only previous work to investigate face alignment
-in the context of sparse signal representation and SRC is the
-work of \cite{Huang2008-CVPR}. They consider the case where the
-training images themselves are misaligned and allow one
-deformation per training image. They linearize the training
-rather than the test, which is computationally more costly as
-it effectively triples the size of the training set. In
-addition, as they align the test image to all subjects
-simultaneously, it potentially is more prone to local minima
-when the number of subjects increases, as we will see in the
-following experimental comparisons.
-\begin{enumerate}
-\item {\em Extended Yale B.} In this experiment, we have
-    used the same experimental settings as in
-    \cite{Huang2008-CVPR}. 20 subjects are selected and
-    each has 32 frontal images (selected at random) as
-    training and another 32 for testing. An artificial
-    translation of 10 pixels (in both $x$ and $y$
-    directions) is introduced to the test image. For our
-    algorithm we downsample all the images to $88\times
-    80$ for memory reasons, whereas the work of
-    \cite{Huang2008-CVPR} uses random projections.
-Note that the use of cropped images in this experiment introduces image boundary effects.
-    Our
-    algorithm achieves the recognition rate 93.7\%,
-    compared to 89.1\% recognition rate reported in
-    \cite{Huang2008-CVPR}.
-\item {\em CMU Multi-PIE.} In this experiment, we choose
-    all subjects from the CMU Multi-PIE database, 7
-    training images from Session 1 and 1 test image from
-    Session 2 per person. The setting is exactly the same
-    as the previous experiment on 2D deformation. We again
-    work with downsampled images of size $80\times 60$ pixels. An
-    artificial translation of 5 pixels (in both $x$ and $y$
-    directions) was induced in the test image. The
-    algorithm of \cite{Huang2008-CVPR} achieves a
-    recognition rate of 67.5\%,\footnote{That algorithm has
-    two free parameters - $l$ and $d$, which govern the tradeoff between
-	accuracy and run-time. For this experiment
-    we chose $l = 1$ and $d = 514$.} while ours achieves 92.2\%.
-    \end{enumerate}
-
-\section{Handling Illumination Variation}\label{sec:illumination}
-In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.
-Under what conditions is this a reasonable assumption to make?
-What can we say from first principles about how the training images should be chosen?
-
-\subsection{The Illumination Model}
-
-The strongest theoretical results so far regarding the relationship
-between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
-The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
-all images of the object can be well approximated by linear combinations of
-nine (properly chosen) basis images.  The basis images have mixed sign, and
-their illuminations consist of the lowest frequency spherical harmonics.
-While this is a very important result for understanding the image
-formation process, the direct application of this result in most practical
-systems is misguided for several reasons.
-Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
-and they all do so in a way that violates the modeling assumptions of the Basri analysis.
-
-Fortunately, even with these effects, for most materials the relationship between
-illumination and image is still linear,\footnote{Materials that break
-this assumption include fluorescent materials and the photochromic (``Transition'') lenses
-in some eyeglasses.  Most materials emit light in proportion to their
-incident light.} provided the sensor has a linear response curve.\footnote{Proper handling of gamma encoding is an important consideration for
-practitioners.  Most cameras apply a non-linear and often undocumented response
-curve to captured images.  A slight degradation of performance will occur if
-gamma compressed images are treated as if they were linear.  We recommend the use
-of cameras with well documented response curves that can be inverted when the
-image file is loaded.}
-For a more in-depth study
-of the relationship between illumination and images, we refer the reader to
-\cite{belhumeur1998set}.
-While the relationship between illuminations and images is linear,
-only positive weights are allowed; the space of all images of an object with
-fixed pose and varying illumination is a convex cone lying in the positive
-orthant. The question becomes, how many images does it take to do a good job
-of representing images sampled from this cone?
-
-It has been observed in various empirical studies that
-one can get away with using a small number of frontal
-illuminations to linearly represent a wide range of new frontal
-illuminations, when they are all taken under the same laboratory conditions
-\cite{Georghiades2001-PAMI}. This is the case for many public
-face datasets, including AR, ORL, PIE, and Multi-PIE.
-Unfortunately, we have found that in practice, a training
-database consisting purely of frontal illuminations is not
-sufficient to linearly represent images of a faces taken
-under typical indoor or outdoor conditions (see the experiment
-conducted in Section \ref{sec:own-data}). As illustrated by the
-example in Figure \ref{fig:promo}, an insufficient number of
-training illuminations can result in recognition failure. To
-ensure our algorithm works in practice, we need to find a set
-of training illuminations that are indeed {\em sufficient} to
-linearly represent a wide variety of practical indoor and
-outdoor illuminations.
-
-\subsection{Projector-based Illumination System}
-
-We have designed a system that can acquire frontal images of a subject while
-simultaneously illuminating the subject from all directions above horizontal. A sketch of the
-system is shown in Figure \ref{fig:system}: The illumination
-system consists of four projectors that display various bright
-patterns onto the three white walls in the corner of a dark
-room.  The light reflects off of the walls and illuminates the
-user's head indirectly.  After taking the frontal illuminations
-we rotate the chair by 180 degrees and take pictures from the
-opposite direction.  Having two cameras speeds the process
-since only the chair needs to be moved in between frontal and
-rear illuminations. Our projector-based system has several
-advantages over flash-based illumination systems for face recognition:
-\begin{itemize}
-\item The illuminations can be modified in software, rather than hardware.
-\item It is easy to capture many different illuminations quickly.
-\item Good coverage and distant illumination can be achieved simultaneously.
-\item There is no need to mount anything on the walls or construct a large dome.
-\item The system can be assembled from off-the-shelf hardware.
-\end{itemize}
-\begin{figure}
-\centering
-\includegraphics[height=2in]{figures_cvpr/camera_rig.pdf}
-\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
-\vspace{-2em}
-\label{fig:system}
-\end{figure}
-With our projector system, our choice of illuminations is
-constrained only by the need to achieve a good
-SNR\footnote{Since illuminations with more pixels illuminated
-will have a better SNR (provided they don't saturate), there is
-an engineering tradeoff between the SNR and the number of
-training images.}, avoid saturation, and achieve
-a reasonably short acquisition time.  Two simplifying
-assumptions that we make are that every pixel is either turned
-fully on or off in every illumination, and that the illuminated
-regions do not overlap.
-
-Assuming that each pixel is fully on or off enables us to guarantee
-that each illumination image has the same overall intensity, merely
-by guaranteeing that we illuminate the same number of pixels in each image.\footnote{Since DLP projectors may have dramatically different response
-curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
-Since our algorithm depends only on  the
-linearity between the illuminations and the images, and not on the
-relative intensities of the illuminations, the designer has the freedom to choose the overall intensity of the illuminations
-to prevent saturation or low SNR, in a sort of offline exposure control.
-
-Assuming that the sequentially illuminated regions do not overlap results in a
-set of training images that span a larger cone than a similar number of
-overlapping regions.  This results in training images that require fewer
-negative coefficients in $\x$ to represent test images under natural
-illuminations.  The effect of negative coefficients in $\x$ appears to depend
-partly on how the test images are taken and is still under study.
-
-{\em Relationship to existing work:} Most light stages used for face recognition have
-been constructed for the purpose of creating public data sets to study
-illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
-light stages have been used for computer graphics purposes
-\cite{debevec2000acquiring, jones2005performance}.
-The light source can be
-moved around manually \cite{masselus2002free}, but this may result in poor
-consistency of illuminations between users.  Structured light applications use projectors to
-directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
-reconstruction, but this is very disturbing to the user.
-Y.\ Schechner \cite{schechner2007multiplexing}
-studies techniques for multiplexing illumination that can dramatically reduce the noise
-of the demultiplexed images for certain classes of objects and cameras.
-While these techniques have not been incorporated into the current
-system, they fit elegantly into our framework and will likely be used
-in future implementations.  We stress that use of this multiplexing technique
-is independent from the choice of original (directional) illuminations.
-
-\subsection{Choice of Illumination Patterns}
-
-We ran two experiments to guide our choice of illuminations for
-our large-scale experiments:
-\begin{enumerate}
-\item {\em Coverage Experiment.} In the first experiment we
-    attempt to determine what coverage of the sphere is
-    required to achieve good interpolation for test images.
-    The subject was illuminated by 100 (50 front, 50 back)
-    illuminations arranged in concentric rings centered at
-    the front camera.  Subsets of the training images were
-    chosen, starting at the front camera and adding a ring
-    at a time.  Each time a ring was added to the training
-    illumination set, the average $\ell^1$ registration
-    error (residual) for a set of test images taken under
-    sunlight was computed and plotted in Figure
-    \ref{fig:illumination-patterns}(a).  The more rings of
-    training illuminations are added, the lower the
-    representation error becomes, with diminishing returns.
-\item {\em Granularity Experiment.} In the second
-    experiment we attempt to determine how finely divided
-    the illumination sphere should be.  At the first
-    granularity level, the projectors  illuminate the
-    covered area uniformly.  At each subsequent granularity
-    level each illuminated cell is divided in two along its
-    longer side but intensity doubled.  For each
-    granularity level the average $\ell^1$ registration
-    error is computed as in the coverage experiment and
-    shown in Figure \ref{fig:illumination-sufficiency}(b).
-    Again, diminishing returns are observed as more
-    illuminations are added.
-\end{enumerate}
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-\includegraphics[height=.75in]{figures_cvpr/coverage_experiment_asplode.png} &
-\includegraphics[height=.75in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
-(a) Coverage Experiment & (b) Chosen Illumination Patterns
-\end{tabular}\vspace{0mm}
-\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
-For rear illuminations the sequence is reversed.
-In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations.
-The four rectangular regions correspond to the four projectors.
-}
-\label{fig:illumination-patterns}
-\vspace{-.75em}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
-(a) Coverage & (b) Granularity
-\end{tabular}\vspace{0mm}
-\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
-\label{fig:illumination-sufficiency}
-\vspace{-4.5em}
-\end{figure}
-
-In the plot for the coverage experiment, Figure
-\ref{fig:illumination-sufficiency}(a),
- we clearly see two plateau regions: one is after 4 rings
-and one is after 10 rings. The first four rings represent the
-typical frontal illuminations, which are present in most public
-face datasets; however, we see that the residual stabilizes
-after 10 rings which include some illuminations from the back
-of the subject. This suggests that although the frontal
-illuminations account for most of the illumination on the face,
-some illuminations from the back are needed in the training set to
-represent images with illumination coming from all directions.
-In the plot for the granularity experiment, Figure
-\ref{fig:illumination-sufficiency}(b), we observe that the
-residual reaches a plateau after four divisions, corresponding
-to a total of 32 illuminations. Based on the results from both
-experiments, we decide to partition the area covered by the
-first 10 rings into a total of 38 cells, whose layout is
-explained in Figure \ref{fig:illumination-patterns}(b). For
-our large-scale experiments, we have collected those
-illuminations for all our subjects.\footnote{It is possible
-that with further experimentation a reduced set of illuminations
-can be found that performs as well or better.}
-
-See below for the 38 training images of one subject:
-\begin{figure}[h]
-\centering
-\includegraphics[width=3.2in]{figures_cvpr/training.png}
-\vspace{-2em}
-\end{figure}
-
-\section{Tests on Public Databases}\label{sec:multipie}
-In this section and the next section, we conduct comprehensive experiments on
-large-scale face databases to verify the performance of our algorithm and
-system. We first test on the largest public face database available that is
-suitable for testing our algorithm, the CMU Multi-PIE.  One shortcoming of the
-CMU Multi-PIE database for our purposes is that there is no separate set of
-test images taken under natural illuminations; we are left to choose which sets
-of images to use for testing and training.  To challenge our algorithm, we
-choose only a small set of illuminations for the training set, yet we include
-all illuminations in the testing set. In the following section, we will test
-our algorithm on a face dataset that is collected by our own system. The goal
-for that experiment will be to show that with a sufficient set of training
-illuminations for each subject, our algorithm indeed works stably and robustly
-with practical illumination, misalignment, pose, and occlusion, as already
-indicated by our experiment shown in Figure \ref{fig:promo}(bottom).
-
-CMU Multi-PIE provides the most extensive test set among public
-datasets. This database contains images of 337 subjects across
-simultaneous variation in pose, expression, and illumination.
-Of these 337 subjects, we use all of the 249 subjects present
-in Session 1 as the training set. The remaining 88 subjects are
-treated as ``impostors'', or invalid images. For each of the
-249 training subjects, we include frontal images of 7 frontal
-illuminations,\footnote{They are illuminations
-$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
-directional illumination, we subtract the ambient-illuminated
-image 0.} taken with neutral expression. As suggested by the
-work of \cite{Georghiades2001-PAMI}, we choose these extreme
-frontal illuminations in the hope that they would linearly
-represent other frontal illuminations well. For the test set,
-we use all 20 illuminations from Sessions 2-4, which were
-recorded over a period of several months. The dataset is
-challenging due to the large number of subjects, and due to
-natural variation in subject appearance over time.
-Table \ref{tab:MultiPIE-recognition2} shows the result of our
-algorithm on each of the 3 testing sessions. Our algorithm
-achieves recognition rates above $90\%$ for all three sessions.
-For the test images, our iterative alignment was initialized
-automatically via the Viola and Jones' face detector. To
-demonstrate that the sparse representation based recognition
-step is indeed beneficial even when there are no impostors, we
-include results for recognition based only on the alignment
-error residuals (i.e. $S=1$), shown in row 1.
-
-\begin{table}
-\caption{Recognition rates on the Multi-PIE database for
-Algorithm 1 and \cite{Yang2010-CVPR}}
-\centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
-\hline
-{Alg. 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Alg. 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Alg. 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
-\hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition2} }
-\end{table}
-
-\newcommand{\tempwidth}[0]{0.58in}
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-\hspace{-2mm}(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Representative failures from Multi-PIE}.
-{\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2.
-Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
-\label{fig:failed-examples}
-\vspace{-4mm}
-\end{figure}
-
-\subsection{Improving the Sampling Window}
-Our algorithm's errors are mostly caused by a few subjects who
-significantly change their appearances between sessions (such
-as hair, facial hair, and eyeglasses). Some representative
-examples are shown in Figure \ref{fig:failed-examples}. For those subjects, alignment and recognition fail on
-almost all test illuminations.
-
-Meanwhile, this observation also suggests that we might be able
-to improve the performance of our method by carefully choosing
-a face region which is less affected by the above factors for
-recognition. In particular, since the forehead region is likely
-to be affected by the change of hair style, we try replacing
-the previous $80 \times 60$ canonical frame with a new
-window that better excludes the forehead. We adjust the
-resolution of the window to keep $m$ approximately constant. In addition,
-we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
-the observation that in many cases the corners
-actually contain background. An example of the new window
-is shown in Figure~\ref{fig:new-mask}.
-
-\begin{figure}
-\vspace{-1em}
-\centering
-{\small
-\begin{tabular}{@{}cc@{}}
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example.png} &
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example_new.png} \vspace{0mm} \\
-Default window. & Proposed window. \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Choosing different sampling windows.}}
-\label{fig:new-mask}
-\vspace{-1em}
-\end{figure}
-
-Table \ref{tab:MultiPIE-recognition2} shows that the
-recognition rates on Multi-PIE indeed increase with this new
-window. In addition, Figure \ref{fig:failed-examples}(a), (b),
-and (c) illustrate three representative subjects for which the
-recognition rates of our algorithm are significantly boosted
-with the new window. However, we should mention that the best
-choice of the window is problem-specific and there is not a
-simple guideline to follow. For example, although the new
-window performs better on Multi-PIE, the same window does not
-help at all on our own database, which will be introduced in
-the next section. This is because most of the training and
-testing images in our database are taken on the same day so the
-variation in hair style is very small. Hence, excluding the
-forehead part may actually result in loss of useful
-discriminative information.
-
-\subsection{Comparison to Existing Work}
-We first compare our result to the recent work
-\cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
-initial registration is obtained from manually selected outer
-eye corners. Then, a supervised hierarchical sparse coding
-model based on local image descriptors is trained, which enjoys
-certain translation invariant properties. With the same
-training and testing sets, \cite{Yang2010-CVPR} is able to
-handle the remaining misalignment and achieves state-of-the-art
-performance on the CMU Multi-PIE database.
-Table~\ref{tab:MultiPIE-recognition2} shows that our algorithm
-achieves similar or better performance on different sessions of
-Multi-PIE.
-
-% Classical algorithms
-To better examine the effectiveness of our iterative alignment
-algorithm, we next compare our result to baseline
-linear-projection-based algorithms, such as Nearest Neighbor
-(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
-Discriminant Analysis (LDA)
-\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
-PCA \cite{Turk1991-CVPR} as its performance is always below
-that of Nearest Subspace.} Since these algorithms assume
-pixel-accurate alignment, they are not expected to work well if
-the test image is not well aligned with the training. In
-Table~\ref{tab:MultiPIE-recognition}, we report the results of
-these classical algorithms with three types of testing image
-alignment: 1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye
-corners,\footnote{Two manually clicked points are sufficient to
-define a similarity transformation. All of the experiments in
-this section are carried out with similarity transformations.}
-and 3.\ the output of our iterative alignment algorithm. The
-performance drop of the LDA algorithm on Multi-PIE reported
-here seems to agree with that reported already in
-\cite{Gross2008-FGR}.  All of the classical algorithms benefit
-greatly from being paired with our iterative alignment
-algorithm.
-
-\begin{table*}
-\centering
-\caption{Recognition rates on the Multi-PIE database for
-different pairings of alignment and recognition stages.}
-\centerline{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
-\hline
-\backslashbox{Rec.}{Align.}
-& \multicolumn{3}{|c|}{Face Detector}
-& \multicolumn{3}{|c|}{Manual}
-& \multicolumn{3}{|c|}{Iterative Alignment}
-\\
-\hline
-Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
-\hline
-NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
-\hline
-NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
-\hline
-LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
-\hline
-LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
-\hline
-SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition} }
-\vspace{-.1in}
-\end{table*}
-
-% LBP
-We also compare our result to Local Binary Patterns (LBP)
-\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
-able to capture fine details of facial appearance and texture.
-Due to its robustness to variations in illumination, facial
-expression, aging and other changes, LBP has achieved the
-state-of-the-art face recognition performance in the scenario
-when only one sample per person is used for training
-\cite{Tan06facerecognition}.
-In this paper, we follow the same
-steps as in \cite{Ahonen2006-PAMI} to construct an LBP
-descriptor for each training and testing sample. The $80\times
-60$ face region is first divided into a regular $10\times 10$
-grid of cells, each of size $8\times 6$ pixels. Within each
-cell, the histogram of 59 uniform binary patterns is then
-computed, where the patterns are generated by thresholding 8
-neighboring pixels in a circle of radius 2 using the central
-pixel value. Finally, the local histograms are concatenated to
-produce the global descriptor vector. As suggested in
-\cite{Ahonen2006-PAMI}, the recognition is performed using a
-nearest neighbor classifier with Chi square distance as the
-distance measure and we report the recognition rates with the
-same three types of input as before.  As shown in
-Table~\ref{tab:MultiPIE-recognition}, although LBP achieves
-competitive recognition rates given manually aligned training
-and testing samples, demonstrating its robustness to moderate
-misalignment, it still benefits from using the output of our
-iterative alignment algorithm as the input.
-In addition, like
-the other classical algorithms, the performance of LBP degrades
-dramatically if it is applied directly to the output of a face
-detector. This is notable given that LBP is often applied
-without any special alignment in practice. Finally, we
-attribute the improvement in performance of LBP over SRC in
-this experiment to its robustness to illumination components
-that cannot be linearly interpolated by the training set.
-
-In addition, although our algorithm is not designed for
-recognition when there is only a single gallery image per user,
-we compare its performance with LBP within this setting for
-completeness. For this experiment, we use the FERET dataset
-\cite{phillips1998feret}, which contains five standard
-partitions: `fa' is the gallery containing 1196 frontal images
-of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
-sets of probe images. The testing sets differ from the training
-in facial expression (`fb'), illumination (`fc'), aging (`dup1'
-) and long aging (`dup2'). In fact, except for `fb', we notice
-significant changes of illumination in all the other three test
-sets. For the training, we again crop and normalize the face
-region from each original image to an $80\times 60$ window
-using manually marked eye coordinates \cite{Deng2010-PR}. In
-Table~\ref{tab:FERET-recognition}, we report the performance of
-our algorithm on the four test sets, with input directly
-obtained from the Viola and Jones' detector. We also report the
-performance of LBP with the same three types of input as before:
-we use letters
-``$d$'', ``$m$'', and ``$i$'' to indicate face detector, manual
-alignment, and our iterative alignment algorithm, respectively.
-
-\begin{table}
-\caption{Performance on single gallery image FERET dataset}
-\vspace{-3mm} \centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate \% & fb & fc & dup1 & dup2 \\
-\hline
-$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
-\hline
-$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
-\hline
-$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
-\hline
-{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
-\hline
-\end{tabular}
-\label{tab:FERET-recognition} }
-\vspace{-12mm}
-\end{table}
-
-As expected, our algorithm does not perform well except for
-`fb', in which the illumination is similar to the training and
-the mere variation in facial expression is handled well by the
-sparse error model. For the other three test sets, our
-algorithm fails because the illumination changes and other
-variations seriously violate the assumptions of our method.
-This also explains why LBP performs worse with our iterative
-alignment algorithm, compared to manual alignment. On the other
-hand, while LBP achieves the best recognition rates given
-manually aligned training and testing samples, its performance
-degrades drastically when the input is obtained directly from
-the face detector. It is also worth noting that similar poor
-performance of LBP, as well as other descriptors, has been
-observed on the Labeled Face in the Wild (LFW) database, where
-the training is uncontrolled and limited and the input is
-directly obtained from the face detector \cite{Wolf2008-ECCV}.
-
-All of these experimental results confirm that
-both illumination and alignment need to be simultaneously handled
-well in order to achieve accurate face recognition, even when there is
-no obvious occlusion or corruption in the test.
-
-\subsection{Subject Validation}
-
-We test the algorithms' ability to reject invalid images of the
-88 subjects not appearing in the training database. As
-mentioned before, the \emph{sparsity concentration index} (SCI)
-is used as the outlier rejection rule. Given the sparse
-representation $\x$ of a test image with respect to $K$
-training classes, the SCI measures how concentrated the
-coefficients are on a single class in the dataset and is
-defined as in \cite{Wright2009-PAMI}:
-\begin{displaymath}
-\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
-\|\x\|_1 - 1}{K - 1} \in [0,1] .
-\end{displaymath}
-It is easy to see that if $\textup{SCI}(\x) = 1$, the test
-image is represented using images from one single subject
-class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
-evenly over all classes. Thus, we can choose a threshold $t \in
-[0,1]$ for the proposed method and accept a test image as valid
-if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
-invalid. We compare this classifier to classifiers based on
-thresholding the error residuals of NN, NS, LDA, and LBP.
-\begin{figure}[t]
-{\small
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
-(a) & \hspace{2mm}(b) \\
-\end{tabular}
-}
-\vspace{-2mm}
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database,
-(a) for all algorithms with iterative alignment, and
-(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-2mm}
-\end{figure}
-
-Figure \ref{fig:roc-multipie} plots the receiver operating
-characteristic (ROC) curves, which are generated by sweeping
-the threshold $t$ through the entire range of possible values
-for each algorithm.\footnote{Rejecting invalid images not in
-the entire database is much more difficult than deciding if two
-face images are the same subject. Figure \ref{fig:roc-multipie}
-should not be confused with typical ROC curves for face
-similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
-that the SCI based recognition approach significantly
-outperforms the other algorithms, including LBP, even when all
-algorithms are coupled with our proposed iterative alignment.
-In the right plot we again see that classical algorithms, and
-even LBP, are very sensitive to alignment.  Similar contrasts
-between our algorithm and baseline algorithms were also
-observed for SRC in \cite{Wright2009-PAMI}, though on much
-smaller datasets.
-
-\subsection{Recognition with Synthetic Random Block Occlusion}
-
-We further test the robustness of our $\ell^1$-norm based
-algorithm to synthetic occlusion. We simulate various levels of
-occlusion from 10\% to 50\% by replacing a randomly located
-block of the face image with an image of a baboon, as shown in
-Figure~\ref{fig:multipie-occ-rec}. In this experiment, to avoid
-any other factors that may contribute to extra occlusion of the
-face (such as the change of hair style), we choose illumination
-10 from Session 1\footnote{This is the same session as the
-training set.} as testing. The rest of the experimental setting
-remains unchanged. The table in
-Figure~\ref{fig:multipie-occ-rec} shows that our algorithm is
-indeed capable of handling a moderate amount of occlusion. For
-example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.
-
-\newcommand{\tempscale}{0.40}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ10.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ20.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ30.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ40.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ50.png}  \\
-\end{tabular}
-\vspace{-1mm}
-\caption{{\bf Recognition under varying level of
-random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:}
-\vspace{2mm}
-{\small
-\begin{tabular}{|c|c|c|c|c|c| }
-\hline
-Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
-\hline
-Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
-\hline
-\end{tabular}
-}
-\label{fig:multipie-occ-rec}
-\vspace{-.5in}
-\end{figure}
-
-\subsection{Recognition with Pose and Expression} We now run tests of
-our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
-Using the same training set as above, we test our algorithm on
-images in Session 2 with $15^\circ$ pose, for all 20
-illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
-algorithm on images in Session 3 with smile, again for all 20
-illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
-the performance of our method will be significantly improved if pose and expression data
-are available in the training.
-
-%\vspace{-.1in}
-\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
-system we described in Section \ref{sec:illumination}, and shown in Figure
-\ref{fig:system}, we have collected the frontal view of 109
-subjects {\em without eyeglasses} under 38 illuminations shown
-in Figure \ref{fig:illumination-patterns}. For testing our
-algorithm, we have also taken 935 images of these subjects with
-a different camera under a variety of practical conditions.
-\vspace{-.1in}
-
-\subsection{Necessity of Rear Illuminations} To see how
-training illuminations affect the performance of our algorithm
-in practice, we now compare how well a few frontal
-illuminations can linearly represent: 1. other frontal illuminations
-taken under the same laboratory conditions, and 2. typical
-indoor and outdoor illuminations. To this end, we use the face
-database acquired by our system and use 7 illuminations per
-subject as training. The illuminations are chosen to be similar
-to the 7 illuminations used in the previous experiment on
-Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
-13, 18, 21, 22\}$ shown in Figure
-\ref{fig:illumination-patterns}(b) to mimic the illuminations
- $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
-our algorithm on the remaining $24 - 7 = 17$ frontal
-illuminations for all the subjects. The recognition rate is
-$99.8\%$, nearly perfect. We also test our algorithm on 310
-indoor images and 168 outdoor images of these subjects taken
-under a variety of lighting conditions (category 1 and 2
-specified below), similar to the one shown in Figure
-\ref{fig:promo}, and the recognition rates for indoor and
-outdoor images drop down to $94.2\%$ and $89.2\%$,
-respectively. This is a strong indication that
-frontal illuminations taken under laboratory conditions
-are insufficient for representing test images under typical indoor and
-outdoor illuminations.
-\vspace{-0.1in}
-
-\renewcommand{\tempscale}{0.45}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} \\
-\end{tabular}
-\caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
-\vspace{-.1in}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG} \\
-\end{tabular}
- \caption{{\bf Representative examples of category C4}.
-Top row: successful examples with our method using overlapping blocks.
-Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.55in}
-\end{figure}
-
-\subsection{Large-Scale Test with Sufficient Training
-Illuminations} Now we use all 109 subjects and 38 illuminations
-in the training and test on 935 images taken under a variety of
-practical illuminations and conditions. We have manually partitioned the test images into four main
-categories:
-\begin{description}
-\item[C1:] 310 \emph{indoor} images of 72 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 1).
-\item[C2:] 168 \emph{outdoor} images of 48 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 2).
-\item[C3:] 211 images of 32 subjects with \emph{eyeglasses}
-    (Fig.~\ref{fig:examples1-3}, row 3).
-\item[C4:] 246 images of 56 subjects with \emph{sunglasses}
-    (Fig.~\ref{fig:examples4}).
-\end{description}
-We apply Viola and Jones' face detector on these images and
-directly use the detected faces as the input to our algorithm.
-Table~\ref{tab:UIUC-recognition} reports the performance of our
-algorithm on each category.
-Since our focus is on face
-recognition, the errors do not include failures of the face
-detector on some of the more challenging images.
-As one can see, our algorithm achieves higher than 95\%
-recognition rates on categories 1-3. Furthermore, using the
-full set of 38 illuminations indeed improves the performance of
-our system under practical illumination conditions compared to
-only using a small subset of 7 illuminations. However, the
-performance dramatically drops when the faces are occluded by
-various types of sunglasses, which could cover up to 40\% of
-the entire face. Given the previous experimental results on
-synthetic random block occlusions, and given that the
-illuminations are more challenging, the result is not
-surprising. In the next subsection, we will show how additional
-assumptions can be used to improve the recognition performance.
-\begin{table}[h]
-\vspace{0mm}
-\centering \caption{Recognition rates on our own
-database.} \vspace{0mm}
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Test Category & C1 & C2 & C3 & C4  \\
-\hline
-\hline
-Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
-\hline
-\end{tabular}
-\label{tab:UIUC-recognition}
-\vspace{0mm}
-\end{table}
-
-\subsection{Improving the Performance with Occlusion using Overlapping Blocks}
-A traditional approach to improve the performance of face
-recognition under severe occlusion is to use subregions instead
-the entire face as a whole. This idea has been explored in many
-earlier works; see \cite{Pentland1994-CVPR, Wright2009-PAMI}
-for examples. Since in most real world cases the occlusion is contiguous, it is reasonable to argue that a minority of the
-subregions are likely to be affected by the occlusion. In this
-paper, we adopt the same idea and partition the face into four
-overlapping blocks to better handle sunglasses. This
-scheme is illustrated in Figure~\ref{fig:occ-block}. Notice
-that in this example three out of the four blocks are partially
-or almost completely occluded. In our experiment, each block is
-of size $90\times 48$ and covers about two-fifths of the entire
-face. The testing and training sets are partitioned in the same
-way. We then independently apply
-Algorithm~\ref{alg:deformable-src} and compute a sparse
-representation after registration for each block independently
-with respect to the training set. The recognition
-results for individual blocks are then aggregated by voting.
-
-\begin{figure}
-\centering
-\includegraphics[width=2.6in]{figures_pami/occ_block.png}
-\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.
-(b) The four overlapping blocks.
-(c) The sparse representation is calculated after alignment for each block independently.
-The red lines correspond to his true identity.
-(d) The true identity is successfully recovered by voting based on the SCI scores.}
-\label{fig:occ-block}
-\vspace{0mm}
-\end{figure}
-
-In this experiment, we found that the using the \emph{sparsity
-concentration index} (SCI) scores for voting achieves higher
-recognition rate than the residual measure used in Algorithm~\ref{alg:deformable-src}, on
-category 4 (sunglasses) of our database. The recognition rate
-is increased to 78.3\%, compared to 40.9\% obtained without
-this partition scheme. This is another evidence of the superior
-ability of SCI on subject validation, since a heavily occluded
-block can be regarded as an outlier for recognition and should
-be rejected while voting.
-
-However, we should point out that a major problem with this
-approach is that occlusion cannot always be expected to fall within
- any fixed partition of the face image. Therefore, the
-proposed scheme should only be viewed as an example which shows
-that the performance under occlusion can be boosted by
-leveraging local information of a face as well as global information. We
-leave the investigation of more general models (e.g., MRF \cite{ZhouZ2009}) for face
-recognition with both misalignment and occlusion as an
-interesting future work.
-
-\section{Conclusion}\label{sec:conclusion}
-Using a well-though-out combination of existing ideas
-(iterative image alignment, $\ell_1$-error function, SRC, using projectors for
-illumination), we have proposed a system for recognizing human faces
-from images taken under practical conditions that is conceptually simple, well
-motivated, and competitive with state-of-the-art recognition systems for access
-control scenarios.
-
-The system achieves extremely stable performance under
-a wide range of variations in illumination, misalignment, and even under small amounts of
-pose and occlusion. We achieve very good recognition performance on
-large-scale tests with public datasets as well as our practical face
-images, while using only frontal 2D images in the gallery and no
-explicit 3D face model.
-Our system could potentially be extended to better handle large pose
-and expression, either by incorporating training images with different poses or
-expressions or by explicitly modeling and compensating the associated deformations
-in the alignment stage.
-
-Another important direction for future
-investigation is to extend the alignment algorithm to better
-tackle contiguous occlusion. We have demonstrated that misalignment can be naturally handled within the
-sparse representation framework. More complicated models for
-spatial continuity, such as Markov random fields, have also
-been successfully integrated into the computation of a sparse
-representation of well-aligned test images
-\cite{Cevher2008-NIPS, ZhouZ2009}. A unified approach
-for face alignment and recognition in the presence of
-contiguous occlusion remains an open problem.
-
-\section*{Acknowledgment}
-This work was supported by grants NSF IIS 08-49292, NSF ECCS
-07-01676, and ONR N00014-09-1-0230. JW thanks Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
-Mathematics for discussions related to this work, and
-acknowledges support from a Microsoft Fellowship and
-the Lemelson-Illinois Student Prize. %} %\vspace{0mm}
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{40}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-\bibliographystyle{IEEEtran}
-\bibliography{faces}
-
+% Biography section
 \newcommand{\biospace}{\vspace{-2em}}
-
-% biography section
 \biospace
@@
 %\enlargethispage{-5in}
-
-\end{document}

### 2011-04-05 faces_system.tex
Moved bios to separate file

@@
 % the document is modified later
-%\IEEEtriggeratref{40}
+%\IEEEtriggeratref{42}
 % The "triggered" command can be changed if desired:
@@
 
-\newcommand{\biospace}{\vspace{-2em}}
-
-% biography section
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
-received his Bachelor's degree in General Engineering in 2003,
-and his Master's degree in Electrical Engineering in 2006, from the University
-of Illinois at Urbana-Champaign, where he is currently a Ph.D.
-candidate in Electrical \& Computer Engineering.
-His research interests include robotics, computer vision, and optimal control.
-His recent work focuses on parallel algorithms for sparsity
-based face recognition.\end{IEEEbiography}
-
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/john}}]{John
-Wright} received his PhD in Electrical Engineering from the
-University of Illinois at Urbana-Champaign in October 2009. He is currently a
-researcher in the Visual Computing group at Microsoft Research Asia.  His
-research focuses on developing provably correct and efficient tools for
-recovering low-dimensional structure in corrupted high-dimensional datasets.
-His work has received a number of awards, including the 2009 Lemelson-Illinois
-Prize for Innovation, the 2009 UIUC Martin Award for Excellence in Graduate
-Research, a 2008-2010 Microsoft Research Fellowship, a Carver fellowship, and a
-UIUC Bronze Tablet award.  \end{IEEEbiography}
-
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
-received his Bachelor's and Master's degrees, both in Electrical
-Engineering, from the Indian Institute of Technology, Madras, India in 2006. He
-is currently a PhD candidate in the Electrical \& Computer Engineering
-Department at the University of Illinois, Urbana-Champaign. His research
-interests include compressed sensing, computer vision, and machine learning.
-His recent work focuses on low-rank matrix recovery techniques for
-batch image alignment and texture rectification.  \end{IEEEbiography}
-
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
-received... {\bf PLACEHOLDER}
-\end{IEEEbiography}
-
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]
-{bios/hossein}}]{Hossein Mobahi} received his Bachelor's and Master's degrees
-in Iran, both in Computer Engineering from Azad University (Tehran-South) in
-2003  and University of Tehran in 2005 respectively. He is currently a PhD
-candidate in the Computer Science Department at the University of Illinois,
-Urbana-Champaign. His research interests include pattern classification,
-clustering and optimization.  His recent research focuses on iterative
-smoothing for image alignment.  \end{IEEEbiography}
-
-\biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
-Ma} received two Bachelors degree in Automation and Applied Mathematics
-from  Tsinghua University, Beijing, China, in 1995. He received an Master
-degree in Electrical Engineering and Computer Sciences (EECS) in 1997, a second
-Master degree in Mathematics in 2000, and the Ph.D. degree in EECS in 2000, all
-from the University of California at Berkeley. He is currently an associate
-professor (with tenure) at the Department of Electrical and Computer
-Engineering, University of Illinois at Urbana-Champaign, and since January 2009
-has also served as research manager for the Visual Computing Group at Microsoft
-Research Asia, Beijing, China.  \end{IEEEbiography}
-\vfill
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
+\input{bios/bios}
 

### 2011-04-06 bios/bios.tex
Added Zihan's bio, working on supp. material

@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
-received... {\bf PLACEHOLDER}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou}
+Zihan Zhou received the bachelor's degree in automation from Tsinghua
+University, China, in 2007. Since then, he has been with University of Illinois
+at Urbana-Champaign, where he received the MS degree in electrical and computer
+engineering and where he is now working toward the PhD degree. During the
+summer of 2009, he was a research intern with Microsoft Research Asia, Beijing.
+His research interests include computer vision, signal processing and machine
+learning. He is a student member of the IEEE.
 \end{IEEEbiography}

### 2011-04-06 supplementary_material.tex
Added Zihan's bio, working on supp. material

--- 2011-04-05 faces_system.tex
@@
-%% bare_jrnl.tex
-%% V1.1
-%% 2002/08/13
-%% by Michael Shell
-%% mshell@ece.gatech.edu
-%%
-%% NOTE: This text file uses MS Windows line feed conventions. When (human)
-%% reading this file on other platforms, you may have to use a text
-%% editor that can handle lines terminated by the MS Windows line feed
-%% characters (0x0D 0x0A).
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.6 or later) with an IEEE journal paper.
-%%
-%% Support sites:
-%% http://www.ieee.org
-%% and/or
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
-%%
-%% This code is offered as-is - no warranty - user assumes all risk.
-%% Free to use, distribute and modify.
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% Testflow can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/testflow
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print. Authors are encouraged to use U.S. letter paper when
-% submitting to IEEE. Use the testflow package mentioned above to verify
-% correct handling of both paper sizes by the author's LaTeX system.
-%
-% Also note that the "draftcls", not "draft", option should be used if
-% it is desired that the figures are to be displayed in draft mode.
-
-% *************************************************************************
-% You can now use this template both for submitting to peer review and,
-% if your paper is accepted, for sending in final publication materials.
-% To change from peer review format (single column, double-spaced) to
-% final publication format (double column, single-spaced), just move the
-% comment-line sign (%) from one \documentclass line to the other.
-% The first is for peer review format(single column), the second is for final publication(double column).
-
-%\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
-%\documentclass[9.5pt,journal,final,compsoc,twocolumn]{IEEEtran}
-
-\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran} % from the new style file.  supposedly double column.
-
-% *************************************************************************
+\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
 
@@
 \title{Towards a Practical Face Recognition System: \\ Robust Alignment and
-Illumination by Sparse Representation}
+Illumination by Sparse Representation\\
+(Supplementary Materials)}
 %
@@
 % after the title page when using the twoside option.
-
-% If you want to put a publisher's ID mark on the page
-% (can leave text blank if you just want to see how the
-% text height on the first page will be reduced by IEEE)
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
-
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%% ABSTRACT
-\IEEEcompsoctitleabstractindextext{
-\vspace{0mm}
-\begin{abstract}
-Many classic and contemporary face recognition algorithms work
-well on public data sets, but degrade sharply when they are
-used in a real recognition system. This is mostly due to the
-difficulty of simultaneously handling variations in
-illumination, image misalignment, and occlusion in the test
-image. We consider a scenario where the training images are
-well controlled, and test images are only loosely controlled.
-We propose a conceptually simple face recognition system that
-achieves a high degree of robustness and stability to
-illumination variation, image misalignment, and partial
-occlusion. The system uses tools from sparse representation to
-align a test face image to a set of frontal training images.
-The region of attraction of our alignment algorithm is computed
-empirically for public face datasets such as Multi-PIE. We
-demonstrate how to capture a set of training images with enough
-illumination variation that they span test images taken under
-uncontrolled illumination. In order to evaluate how our
-algorithms work under practical testing conditions, we have
-implemented a complete face recognition system, including a
-projector-based training acquisition system. Our system can
-efficiently and effectively recognize faces under a variety of
-realistic conditions, using only frontal images under the
-proposed illuminations as training.\vspace{0mm}
-\end{abstract}
 

### 2011-04-06 faces_system.tex
Typesets properly without appendix, spills with appendix.

--- 2011-04-05 faces_system.tex
@@
 \usepackage{color}
+
+\usepackage{verbatim}
 
@@
 
-\begin{algorithm}[thb]
-\caption{\bf (Deformable Sparse Recovery and Classification for
-Face Recognition)} \label{alg:deformable-src}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
-$\y\in\Re^m$ and a deformation group $T$.
-\STATE
-{\bf for} each subject $i$,
-\STATE \hspace{3mm} $\tau^{(0)}
-\leftarrow I$.
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
-\STATE \hspace{6mm}
-$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
-\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
-\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
-\STATE
-\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$;
-\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-residuals $\|\e\|_1$. \STATE Compute an average transformation
-$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
-\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem: \hspace{2mm}
-$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
-\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
-\end{small}
-\end{algorithmic}
-\end{algorithm}
-%\vspace{-4mm}
+\input{deformable-src.tex}
 
@@
 \subsection{System Implementation}
-In this section, we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}. It is
-not hard to see that its computational complexity is dominated
-by the two steps where the $\ell^1$-norm minimization problems
-are solved; namely Step 6 for iterative registration, and Step
-14 for global sparse representation.
-Fortunately, many fast algorithms for solving these problems have been proposed over the
-past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
-That work suggests that \emph{Augmented Lagrange Multiplier}
-(ALM) algorithms \cite{Bertsekas1982} strike a good balance
-between scalability and accuracy: as first order methods, they
-require only lightweight vector operations and matrix-vector
-multiplications at each iteration, making them preferable to
-more classical solutions such as interior point methods.
-However, compared to other first-order methods, they achieve
-higher accuracy with a fixed computational budget.
-
-We use Step 14 as an example to illustrate the ALM
-method, since solving Step 6 is very similar. Recall that in
-Step 14 the problem we are interested in is:
-\begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
-A \x + \e.
-\end{equation}
-Its corresponding augmented Lagrangian function is
-\begin{equation}
-L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
-\end{equation}
-where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
-penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
-(\x,\e,\blamda)$ by alternating between optimizing with respect
-to the primal variables $\x, \e$ and updating the dual variable
-$\blamda$, with the other fixed, as follows:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:alm}
-\end{equation}
-Although updating $\blamda$ is trivial,
-minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
-$\x$ and $\e$ could still be costly. To further reduce the
-complexity of the problem, we adopt an approach used in
-\cite{YangJ2009-pp}, called \emph{alternating direction
-method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
-over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
-\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:adm}
-\end{equation}
-As the objective function is convex and alternation is between two
-terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).
-
-In order to discuss the solution to the above subproblems, we
-need to define the following soft-thresholding operator for a
-vector $\x$ and a scalar $\alpha \geq 0$:
-\begin{equation}
-\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
-\end{equation}
-where all the operations are performed component-wise. It is
-easy to show that the subproblem with respect to $\e$ has a
-closed-form solution given by the soft-thresholding operator:
-\begin{equation}
-\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
-\end{equation}
-To solve the subproblem associated with $\x$, we
-apply a first-order $\ell^1$-minimization method,
-called \emph{fast iterative shrinkage-threshold algorithm}
-(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
-iteratively minimize a quadratic approximation $Q(\x, \z)$ to
-$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
-carefully chosen in order to achieve a good convergence
-rate. We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
-\begin{algorithm}[t]
-\caption{\bf (Augmented Lagrange Multiplier Method for Global
-Recognition)}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
-\mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
-\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
-\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
-\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
-\ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{small}
-\end{algorithmic}
-\label{alg:alm}
-%\vspace{-4mm}
-\end{algorithm}
-
+The runtime of Algorithm~\ref{alg:deformable-src} is dominated
+by the time spent solving two qualitatively similar $\ell_1$ minimization problems.
+We have developed custom solvers for this purpose based on 
+\emph{Augmented Lagrange Multiplier} (ALM) algorithm.  
 We have selected this algorithm because it strikes the best
 balance between speed, accuracy, and scalability for our problem out of
-many algorithms that we have tested. We refer the interested reader to
-\cite{YangA2010-pp} for a more detailed discussion of competing
-approaches.  On a Mac Pro with
+many algorithms that we have tested. We refer the reader to our
+supplementary materials for a more in-depth discussion of our
+solvers.  
+For a more detailed discussion of competing
+approaches, we refer the interested reader to \cite{YangA2010-pp} 
+On a Mac Pro with
 Dual-Core 2.66GHz Xeon processors and 4GB memory,
@@
 (b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-2mm}
+\vspace{-2.5em}
 \end{figure}
@@
 \label{fig:occ-block}
-\vspace{0mm}
+\vspace{-3em}
 \end{figure}
@@
 
-\newcommand{\biospace}{\vspace{-2em}}
-
-% biography section
+% Biography section
+\newcommand{\biospace}{\vspace{-4em}}
 \biospace
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
-received... {\bf PLACEHOLDER}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou}
+Zihan Zhou received the bachelor's degree in automation from Tsinghua
+University, China, in 2007. Since then, he has been with University of Illinois
+at Urbana-Champaign, where he received the MS degree in electrical and computer
+engineering and where he is now working toward the PhD degree. During the
+summer of 2009, he was a research intern with Microsoft Research Asia, Beijing.
+His research interests include computer vision, signal processing and machine
+learning. 
 \end{IEEEbiography}
@@
 Research Asia, Beijing, China.  \end{IEEEbiography}
-\vfill
 
@@
 
+\begin{comment}
+\appendix[L1 Minimization via Augmented Lagrange Multiplier]
+In this Appendix we discuss the computational issues related to
+the implementation of Algorithm~\ref{alg:deformable-src}, which is
+repeated here for convenience. It is
+not hard to see that its computational complexity is dominated
+by the two steps where the $\ell^1$-norm minimization problems
+are solved; namely Step 6 for iterative registration, and Step
+14 for global sparse representation.
+Fortunately, many fast algorithms for solving these problems have been proposed over the
+past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
+That work suggests that \emph{Augmented Lagrange Multiplier}
+(ALM) algorithms \cite{Bertsekas1982} strike a good balance
+between scalability and accuracy: as first order methods, they
+require only lightweight vector operations and matrix-vector
+multiplications at each iteration, making them preferable to
+more classical solutions such as interior point methods.
+However, compared to other first-order methods, they achieve
+higher accuracy with a fixed computational budget.
+
+We use Step 14 as an example to illustrate the ALM
+method, since solving Step 6 is very similar. Recall that in
+Step 14 the problem we are interested in is:
+\begin{equation}
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
+A \x + \e.
+\end{equation}
+Its corresponding augmented Lagrangian function is
+\begin{equation}
+L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
+\end{equation}
+where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
+penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
+(\x,\e,\blamda)$ by alternating between optimizing with respect
+to the primal variables $\x, \e$ and updating the dual variable
+$\blamda$, with the other fixed, as follows:
+\begin{equation}
+\left \{
+\begin{array}{lll}
+(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
+\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
+\end{array}
+\right .
+\label{eq:alm}
+\end{equation}
+Although updating $\blamda$ is trivial,
+minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
+$\x$ and $\e$ could still be costly. To further reduce the
+complexity of the problem, we adopt an approach used in
+\cite{YangJ2009-pp}, called \emph{alternating direction
+method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
+over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
+\begin{equation}
+\left \{
+\begin{array}{lll}
+\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
+\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
+\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
+\end{array}
+\right .
+\label{eq:adm}
+\end{equation}
+As the objective function is convex and alternation is between two
+terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).
+
+In order to discuss the solution to the above subproblems, we
+need to define the following soft-thresholding operator for a
+vector $\x$ and a scalar $\alpha \geq 0$:
+\begin{equation}
+\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
+\end{equation}
+where all the operations are performed component-wise. It is
+easy to show that the subproblem with respect to $\e$ has a
+closed-form solution given by the soft-thresholding operator:
+\begin{equation}
+\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
+\end{equation}
+To solve the subproblem associated with $\x$, we
+apply a first-order $\ell^1$-minimization method,
+called \emph{fast iterative shrinkage-threshold algorithm}
+(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
+iteratively minimize a quadratic approximation $Q(\x, \z)$ to
+$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
+carefully chosen in order to achieve a good convergence
+rate. We summarize the entire ALM
+algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
+in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
+\begin{algorithm}[t]
+\caption{\bf (Augmented Lagrange Multiplier Method for Global
+Recognition)}
+\begin{algorithmic}[1]
+\begin{small}
+\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
+\mathbf{0}$.
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
+\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
+\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
+\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+\sqrt{1+4t_l^2}\right)$;
+\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+\ENDWHILE
+\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\ENDWHILE \STATE
+{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{small}
+\end{algorithmic}
+\label{alg:alm}
+%\vspace{-4mm}
+\end{algorithm}
+
+We have selected this algorithm because it strikes the best
+balance between speed, accuracy, and scalability for our problem out of
+many algorithms that we have tested. We refer the interested reader to
+\cite{YangA2010-pp} for a more detailed discussion of competing
+approaches.  On a Mac Pro with
+Dual-Core 2.66GHz Xeon processors and 4GB memory,
+running on our database containing images size $80\times 60$
+pixels from 109 subjects under 38 illuminations,
+our C implementation of Algorithm~\ref{alg:deformable-src} takes
+about 0.60 seconds per subject for alignment and about 2.0
+seconds for global recognition. Compared to the highly
+customized interior point method used in the conference version
+of this paper \cite{Wagner2009-CVPR}, this new algorithm is
+only slightly faster for per subject alignment. However, it is
+much simpler to implement and it achieves a
+\emph{speedup of more than a factor of 10} for global
+recognition!
+\end{comment}
+
 \end{document}

### 2011-04-06 supplementary_material.tex
Typesets properly without appendix, spills with appendix.

@@
 
-\begin{IEEEkeywords}
-Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
-Sparse Representation,  Error Correction, Validation and Outlier Rejection.
-\end{IEEEkeywords}}
-
 \maketitle
@@
 %%%%%%%%% BODY TEXT
-\section{Introduction}
-%Automatic face recognition remains one of the most active areas in computer vision.
+\section{L1 Minimization via Augmented Lagrange Multiplier}
 
-\PARstart{F}{ace} recognition applications to date have fallen
-into roughly two categories.  Face recognition has recently
-seen a lot of success in a family of less-demanding
-applications such as online image search and family photo album
-organization (e.g.\ Google Picassa, Microsoft Photo Gallery,
-and Apple iPhoto). At the other end of the tractability
-spectrum there are the terrorist watchlist and mass
-surveillance applications that have for the most part dominated
-the field of face recognition research.  However, there are
-many face recognition applications that fall roughly between
-these extremes, where very high recognition performance is
-desired, but the users in the gallery are still allies of the
-system rather than adversaries.  These applications include
-access control for secure facilities (e.g., prisons, office
-buildings), computer systems, automobiles, or automatic teller
-machines, where controlled gallery images can be obtained in
-advance.  These applications are very interesting due to their
-potential sociological impact.  Since the gallery subjects are
-allies, rather than opponents, of the recognition system, this
-creates the possibility of carefully controlling the
-acquisition of the training data. While the same can be said
-for other biometrics such as fingerprints and iris recognition,
-face recognition has the potential of working with test data
-that is much less controlled, allowing the access control
-system to be made less intrusive to the users of the system. To
-some extent, the goal of this paper is to show how a reliable
-face recognition system can be built for this restricted, but
-still important, scenario.
-
-Very few recognition systems specifically target applications where many well-
-controlled training images are available.  Of these, the classical holistic
-subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
-are well known for their speed and simplicity, as well as for their natural
-extension to linear illumination models.  However, their performance has been
-shown to be extremely brittle not only to alignment variation, but to even
-minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
-is slightly open. One promising recent direction, set forth in
-\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
-sparse representation of the test image in terms of the training set as a
-whole, up to some sparse error due to occlusion. A \emph{sparse
-representation-based classification} (SRC) method is then proposed for
-recognition. The main idea is that the sparse nonzero coefficients should
-concentrate on the training samples with the same class label as the test
-sample. SRC has demonstrated striking recognition performance despite severe
-occlusion or corruption by solving a simple convex program.
-%For this reason, the final recognition stage of our proposed system will
-%consist of SRC, and the alignment stage that precedes it will be based on
-%similar techniques.
-
-\begin{figure}
-\centering \begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
-\end{tabular} \caption{{\bf Effects of registration and illumination on
-Recognition}. In this example we identify the girl among 20 subjects, by
-computing the sparse representation of her input face with respect to the
-entire training set. The absolute sum of the coefficients associated with each
-subject is plotted on the right. We also show the faces reconstructed with each
-subject's training images weighted by the associated sparse coefficients. The
-red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
-input face is from Viola and Jones' face detector (the black box) and all 38
-illuminations specified in Section \ref{sec:illumination} are used in the
-training.  {\bf Middle:} The input face is well-aligned (the white box) with
-the training by our algorithm specified in Section \ref{sec:registration} but
-only 24 frontal illuminations are used in the training for recognition (see
-Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and
-a sufficient set (all 38) of
-illuminations are used in the training. Both are necessary for correct recognition
-using SRC.}\label{fig:promo}
-\vspace{-6.9mm}
-\end{figure}
-
-Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
-between the test and training images, and it requires a rich set of illuminations
-in the gallery images for good performance.
-We illustrate the compounded effect of both alignment and illumination
-with an example in Figure \ref{fig:promo}.
-The task is to identify the girl among 20 subjects. If the test
-face image, say obtained from an off-the-shelf face detector,
-has even a small amount of registration error against the
-training images (caused by mild pose, scale, or misalignment),
-the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative,
-even if sufficient illuminations are present in the training, as shown in Figure
-\ref{fig:promo}(top). Additionally, in order to span
-the illuminations of a typical indoor (or outdoor) environment,
-illuminations from behind the subject are needed in the
-training set. Otherwise, even for perfectly aligned test images,
-the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
-informative, as shown by the example in Figure
-\ref{fig:promo}(middle). Clearly, both good alignment, as well
-as sufficient training images are needed to ensure success of the sparsity-based
-recognition method proposed by \cite{Wright2009-PAMI}.
-In this paper, we examine how to handle alignment and illumination simultaneously
-in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
-closer to practical use.
-
-\vspace{-0mm}
-\subsection{Related Work}
-We briefly review existing techniques for recognition, image
-registration, and handling of illumination variation.  Our
-system is based purely on 2D techniques.  This fact immediately
-distinguishes our approach from systems that either require a
-3D data as an input, or attempt to estimate a 3D model from 2D
-input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While
-these techniques can achieve better robustness to pose
-variation given a sufficiently accurate 3D model, for access
-control applications where only moderate pose variation is
-present, the proposed method will be more than sufficient.
-Note that 2D images of faces under varying illuminations
-already contain 3D shape-related information,\footnote{In
-principle, one can recover the 3D shape of the face from
-multiple illuminations using photometric stereo.} and this
-information can be leveraged by 2D algorithms for alignment and
-recognition even if shape is not reconstructed explicitly.
-
-% Alignment
-In holistic recognition algorithms, correspondence between points in the test
-image and in the training must be achieved.  A long line of research exists on
-using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
-Active Shape Models \cite{cootes1992active} to register images against a
-relatively high-dimensional model of plausible face appearances, often
-leveraging face-specific contours.  While these model-based techniques have
-advantages in dealing with variations in expression and pose, they may add
-unnecessary complexity to applications where subjects normally present a neutral face
-or only have moderate expression. We prefer to focus on deformations with far fewer
-degrees of freedom, i.e. similarity transformations, and to use the training images themselves
-as the appearance model. Iterative registration in this spirit dates at least back to the
-Lucas-Kanade algorithm \cite{lucas1981iterative}.
-
-Whereas much of the early work on image registration is aimed at the
-problem of registering nearly identical images, say by minimizing a sum of
-squared distances or maximizing normalized correlation, here we must confront several physical factors
-simultaneously: misalignment, illumination variations, and corrupted pixels.
-As we discuss further below, illumination variation can be dealt by
-expressing the test image as a linear combination of an appropriate set of
-training images. Similar representations have been exploited in
-illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
-For robustness to gross errors, the $\ell^1$-norm of the residual is a more
-appropriate objective function than the classical $\ell^2$-norm. Its use here
-is loosely motivated by theoretical results due to Cand\`{e}s and Tao
-\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
-lead us to pose the registration problem as the search for a set of
-transformations and illumination coefficients that minimize the $\ell^1$-norm
-of the representation error. We solve this problem using a generalized
-Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
-minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
-these problems can also be solved efficiently using recently developed
-first-order techniques for $\ell^1$-minimization, which are reviewed in
-\cite{YangA2010-pp}.
-
-% Illumination
-Researchers have tried various techniques to deal with illumination variation.
-In almost all recognition algorithms where only a single gallery image is
-available per individual, illumination effects are regarded as a nuisance that
-must be removed before the algorithm can continue.  This is typically done by
-making statistical assumptions about how illumination affects the image, and
-using those assumptions to extract a new representation that is claimed to be
-illumination invariant.  Recent examples include \cite{chen2006total} and
-\cite{zhou2007appearance}.  However, despite these efforts, truly
-illumination-invariant features are difficult to obtain from a single input
-image.  We argue that if one has the luxury of designing the acquisition system
-and the application demands a high recognition rate,
-it is then unwise to limit the gallery to a
-single image per person.  We therefore take the strategy of sampling many
-gallery images of each individual under varying illuminations.  These images
-are used as the basis for either a convex cone model
-\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
-\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
-projector based light stage.  While similar systems have been used for
-other applications, to our knowledge, we
-are the first to use projectors to indirectly illuminate a subject's face for
-the purpose of face recognition.
-\subsection{Contributions} In this paper, we show how
-registration and illumination
-can be simultaneously addressed within a robust sparse representation
-framework. We show that face registration, a challenging
-nonlinear problem, can be solved by a series of linear programs
-that iteratively minimize the sparsity of the registration
-error. This leads to an efficient and effective alignment
-algorithm for face images that works for a large range of
-variation in translation, rotation, and scale, even when
-the face is only partially visible due to eyeglasses, closed
-eyes and open mouth, sensor saturation, etc.  We also propose a
-sufficient set of training illuminations
-that is capable of linearly representing typical indoor and outdoor
-lighting, along with a practical hardware system for capturing
-them.
-
-% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.
-While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.
-
-% We use tools from sparse representation
-% to align a test face image with a set of frontal training
-% images in the presence of registration error, illumination variation, and partial occlusion. The region
-% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
-% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
-% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions.
-We show
-% that our system can efficiently and effectively recognize faces under
-% a variety of realistic conditions, using only
-% frontal images under the proposed illuminations as training.
-
-%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.
-We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.
-Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.
-We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.
-We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.
-
-We then demonstrate the effectiveness of the proposed new
-methods with a complete face recognition system that is {\em
-simple, stable, and scalable}. The proposed system performs
-robust automatic recognition of subjects from loosely
-controlled probe images taken both indoors and outdoors,
-using a gallery of
-frontal views of the subjects' faces under the proposed
-illuminations. An off-the-shelf face
-detector\footnote{We use the OpenCV
-implementation of the Viola and Jones' face detector
-\cite{Viola2004-IJCV}.} is used to detect faces in the test images.
-
-We conduct extensive experiments on the proposed system with
-both public databases and a face database that is collected by
-our own acquisition system. Our experimental results on
-large-scale public face databases show that our algorithm
-indeed achieves very good performance on these databases,
-exceeding or competing with the state-of-the-art algorithms.
-Additionally, our experimental results on our own database
-clearly demonstrate that our system not only works well with
-images taken under controlled laboratory conditions, but is
-capable of handling practical indoor and outdoor illuminations as well.
-
-\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
-we derive our robust registration and recognition algorithm within the sparse
-representation framework. We elaborate on algorithmic implementation issues,
-conduct region of attraction experiments with respect to both 2D in-plane
-deformation and 3D pose variation, and discuss its relationship to existing
-work. Section \ref{sec:illumination} is dedicated to our training acquisition
-system. Using this system, we investigate empirically how many training
-illuminations are required to handle practical illumination variations, and
-suggest a sufficient set of 38 training illuminations. Extensive experiments on
-a large-scale public database and on our own database are conducted in Section
-\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
-proposed system. Section \ref{sec:conclusion} concludes our work with
-discussion of promising future directions.
-
-\section{Robust Alignment}\label{sec:registration}
-As demonstrated in Figure \ref{fig:promo}(top), the main
-limitation of the {\em sparse representation and
-classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
-the assumption of pixel-accurate alignment between the test
-image and the training set. This leads to brittleness under
-pose and misalignment, making it inappropriate for deployment
-outside a laboratory setting. In this section, we show how this
-weakness can be rectified while still preserving the conceptual
-simplicity and good recognition performance of SRC.
-
-SRC assumes access to a database of multiple registered
-training images per subject, taken under varying illuminations.
-The images of subject $i$, stacked as vectors, form a matrix
-$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
-form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
-\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
-well-aligned test image $\y_0$ can be represented as a sparse
-linear combination $A \x_0$ of all of the images in the
-database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
-sufficiency in the next section.} plus a sparse error $\e_0$
-due to corrupted pixels. The sparse representation can be recovered by
-minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
-vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
-$\x$ and $\e$:
-\begin{equation}
-\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
-\label{eqn:robust-l1}
-\end{equation}
-Now suppose that $\y_0$ is subject to some pose or
-misalignment, so that instead of observing $\y_0$, we observe
-the warped image $\y = \y_0 \circ \tau^{-1}$, for some
-transformation $\tau \in T$ where $T$ is a finite-dimensional
-group of transformations acting on the image domain.  The
-transformed image $\y$ no longer has a sparse representation of
-the form $\y = A \x_0 + \e_0$, and naively applying the
-algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
-as seen in Figure \ref{fig:promo}(top).
-
-\subsection{Batch and Individual Alignment} If the
-true deformation $\tau^{-1}$ can be found, then we can apply
-its inverse $\tau$ to the test image and it again becomes
-possible to find a sparse representation of the resulting
-image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
-  This sparsity
-provides a strong cue for finding the correct deformation
-$\tau$: conceptually, one would like to seek a transformation
-$\tau$ that allows the sparsest representation, via
-\begin{equation} \label{eqn:L1-L1-conceptual}
-\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
-\end{equation}
-For fixed $\tau$, this problem is jointly convex in $\x$ and
-$\e$. However, as a simultaneous optimization over the
-coefficients $\x$, error representation $\e$, and
-transformation $\tau$, it is a difficult, nonconvex
-optimization problem. One source of difficulty is the presence
-of multiple faces in the matrix $A$:
-\eqref{eqn:L1-L1-conceptual} has many local minima that
-correspond to aligning $\y$ to different subjects. In this
-sense, the misaligned recognition problem differs from the
-well-aligned version studied in \cite{Wright2009-PAMI}. For the
-well-aligned case, it is possible to directly solve for a
-global representation, with no concern for local minima. With
-possible misalignment, it is more appropriate to seek the best
-alignment of the test face with each subject $i$:
-\begin{equation} \label{eqn:per-subject-L1}
-\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
-\end{equation}
-We no longer penalize $\| \x \|_1$, since $A_i$ consists of
-only images of subject $i$ and so $\x$ is no longer expected to
-be sparse.
-
-\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector.
-We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
-\begin{equation}
-\y\circ \tau + J \Delta \tau = A_i \x + \e.
-\end{equation}
-Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
-Jacobian of $\y \circ \tau$ with respect to the transformation
-parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
-above equation is underdetermined if we allow the registration
-error $\e$ to be arbitrary. At the correct alignment we
-expect the test image to differ from $A_i \x$ only
-for the minority of the pixels corrupted by occlusions. Thus,
-we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
-\begin{equation}
-\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
-\label{eqn:L1-align}
-\end{equation}
-This is different from the popular choice that
-minimizes the $\ell^2$-norm of the registration error:
-\begin{equation}
-\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
-\label{eqn:L2-align}
-\end{equation}
-which is also equivalent to finding the deformation step
-$\Delta  \tau$ by solving the least-square problem:
-$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
-\|_2$. Empirically, we find that if there is only small noise
-between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
-$\eqref{eqn:L2-align}$ have similar performance.  However, if
-there are occlusions in $\y_0$, sequential
-$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
-better than sequential $\ell^2$-minimization
-\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
-example.
-
-The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images.
-Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}.
-While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm.
-In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution.
-We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.
-
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{cccc}
-\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
-\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
-\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
-\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
-(a) & (b) & (c) & (d)
-\end{tabular}
-\vspace{0mm}}
-\caption{{\bf Comparing alignment of a subject wearing sunglasses by
-$\ell^1$ and $\ell^2$ minimization.}
-{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
-result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
-given by the face detector, {\em Red (solid):} Alignment result shown on the same
-face; (b) warped testing image using the estimated transformation $\y_0$;
-(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-8.0mm}
-\end{figure}
-
-In addition to normalizing the training images (which is done
-once), it is important to normalize the warped testing image
-$\y \circ \tau$ as the algorithm runs.  Without normalization,
-the algorithm may fall into a degenerate global minimum
-corresponding to zooming in on a dark region of the test
-image.  Normalization is done by replacing the linearization of
-$\y \circ \tau$ with a linearization of the normalized version
-$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
-The proposed alignment algorithm can be easily extended to work
-in a {\em multiscale} fashion, with benefits both in
-convergence behavior and computational cost.  The alignment
-algorithm is simply run to completion on progressively less
-downsampled versions of the training and testing images, using
-the result of one level to initialize the next.
-
-\subsection{Robust Recognition by Sparse Representation} Once
-the best transformation $\tau_i$ has been computed for each
-subject $i$, the training sets $A_i$ can be aligned to $\y$,
-and a global sparse representation problem of the form
-\eqref{eqn:robust-l1} can be solved to obtain a discriminative
-representation in terms of the entire training set. Moreover,
-the per-subject alignment residuals $\| \e \|_1$ can be used to
-prune unpromising candidates from the global optimization,
-leaving a much smaller and more efficiently solvable problem.
-The complete optimization procedure is summarized as Algorithm
-\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
-considered together to provide a sparse representation for the
-test image. If $S = 1$, the algorithm reduces to classification
-by registration error; but considering the test image might be
-an invalid subject, we typically choose $S = 10$. Since valid
-images have a sparse representation in terms of this larger
-set, we can reject invalid test images using the {\em sparsity
-concentration index} proposed in \cite{Wright2009-PAMI}.
-The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
-selects coefficients from the vector $\x$ corresponding to subject $i$.
-
-Another important free parameter in Algorithm
-\ref{alg:deformable-src} is the class of deformations $T$. In
-our experiments, we typically use 2D similarity
-transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
-Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
-detector, or 2D projective transformations, $T =
-\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
-This class of transformations is able to represent distortion
-in a perspective image of a planar object.}$, for handling some
-pose variation.
-
-In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
-performance of our system, based on the observation that the face detector output may be poorly centered on
-the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
-instead of aligning the training sets to the
-original $\y$ directly obtained from the face detector, we
-compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
-\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
-is believed to be better centered, and update $\y$ according to
-$\bar{\tau}$. For the 2D similarity transformations, which are
-used in our system when initialized by the face detector, a
-transformation $\tau$ can be parameterized as $\tau = (\tau^1,
-\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
-the translations in $x$- and $y$-axis, $\tau^3$ represents the
-rotation angle and $\tau^4$ represents the scale. Then the
-average transformation is simply obtained by taking the
-component-wise mean:
-\begin{displaymath}
-\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
-\tau_{k_S}^i) / S, i = 1,2,3,4.
-\end{displaymath}
-Finally, the training sets are aligned to the new $\y$.
-
-\begin{algorithm}[thb]
-\caption{\bf (Deformable Sparse Recovery and Classification for
-Face Recognition)} \label{alg:deformable-src}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
-$\y\in\Re^m$ and a deformation group $T$.
-\STATE
-{\bf for} each subject $i$,
-\STATE \hspace{3mm} $\tau^{(0)}
-\leftarrow I$.
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
-\STATE \hspace{6mm}
-$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
-\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
-\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
-\STATE
-\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$;
-\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-residuals $\|\e\|_1$. \STATE Compute an average transformation
-$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
-\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem: \hspace{2mm}
-$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
-\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
-\end{small}
-\end{algorithmic}
-\end{algorithm}
-%\vspace{-4mm}
-
-The transformation $\tau$ defines a mapping between the
-coordinates of pixels in the large original image and a smaller
-(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
-artifacts in the downsampled image, one should apply a
-smoothing filter to the original image. For a simple implementation, a rectangular window with
-regular sampling can used, but in general, the small image need
-not be regularly sampled in pixel coordinates.  For example,
-the sample locations could be arbitrarily selected from within a
-``face shaped'' area. We will discuss how choosing
-different windows can affect the performance of our algorithm
-in Section \ref{sec:multipie}.
-
-\subsection{System Implementation}
-In this section, we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}. It is
+\PARstart{I}{n} these supplementary materials we discuss the computational issues related to
+the implementation of Algorithm~\ref{alg:deformable-src}, which is
+repeated here for convenience. It is
 not hard to see that its computational complexity is dominated
@@
 higher accuracy with a fixed computational budget.
+\input{deformable-src.tex}
 
@@
 
-\subsection{Experiments on Region of Attraction} We will now present three
-experimental results demonstrating the effectiveness of the individual
-alignment procedure outlined in the previous section. They show the sufficiency
-of the region of attraction, verify effectiveness of the multiscale extension,
-and show stability to small pose variations.  We delay large-scale recognition
-experiments to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
-have discussed the issue of illumination in the next section.
-
-\noindent {1) \em 2D Deformation.}  We first verify the
-    effectiveness of our alignment algorithm with images
-    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
-    We select all the subjects in Session 1, use 7
-    illuminations per person from Session 1 for training,
-    and test on one new illumination from Session
-    2.\footnote{The training are illuminations $\{0, 1, 7,
-    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
-    testing is the illumination 10. } We manually select
-    eye corners in both training and testing as the ground
-    truth for registration. We downsample the images to
-    $80\times 60$ pixels\footnote{Unless otherwise stated,
-    this will be the default resolution at which we prepare
-    all our training and testing datasets.} and the distance between the two outer
-    eye corners is normalized to be 50 pixels for each
-    person. We introduce artificial deformation to the
-    testing image with a combination of translation,
-    rotation and scaling. We further use the alignment
-    error $\|\e\|_1$ as an indicator of success. Let $r_0$
-    be the alignment error obtained by aligning a test
-    image to the training images without any artificial
-    perturbation. When the test image is artificially
-    perturbed and aligned, resulting in an alignment error
-    $r$, we consider the alignment successful if $|r - r_0 | \leq
-    0.01r_0$. Figure \ref{fig:attraction} shows the
-    percentage of successful registrations for all subjects
-    for each artificial deformation. The results suggest
-    that our algorithm works extremely well with
-    translation up to 20\% of the eye distance (or 10
-    pixels) in all directions and up to $30^\circ$ in-plane
-    rotation. We have also tested our alignment algorithm
-    with scale variation and it can handle up to 15\%
-    change in scale.
-
-We have gathered the statistics of the Viola and Jones'
- face detector on the Multi-PIE dataset. For 4,600 frontal
- images of 230 subjects under 20 different illuminations,
- using manual registration as the ground truth, the average
- misalignment error of the detected faces is about 6 pixels
- and the average variation in scale is 8\%. This falls
- safely inside the region of attraction for our alignment
- algorithm.
-\begin{figure*}
-\centering
-{\small
-\begin{tabular}{ccc}
-\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
-(a)&(b)&(c)
-\end{tabular}
-\begin{tabular}{cccc}
-\includegraphics[height=1in]{figures_pami/x_tr.png} &
-\includegraphics[height=1in]{figures_pami/y_tr.png} &
-\includegraphics[height=1in]{figures_pami/theta.png} &
-\includegraphics[height=1in]{figures_pami/scale.png}\\
-(d)&(e)&(f)& (g)
-\end{tabular}
-\vspace{0mm}
-}
-\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.
-The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees.
-{\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions.
-(b) Simultaneous translation in $y$ direction and in-plane rotation.
-(c) Simultaneous translation in $y$ direction and scale variation.
-{\bf Bottom row:} (d) Translation in $x$ direction only.
-(e) Translation in $y$ direction only.
-(f) In-plane rotation only.
-(g) Scale variation only.}
-\label{fig:attraction}
-\vspace{0mm}
-\end{figure*}
-
-\noindent{2) \em Multiscale Implementation.}
-Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost.
-Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
-We use the same 7 illuminations in
-Session 1 as training, and all 20 illuminations in the same
-session as testing. We introduce artificial deformation in
-both $x$ and $y$ directions up to 16 pixels in the
-$80\times 60$ frame, with a step size of 4 pixels, i.e.,
-$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
-\{-16,-12,\ldots,12,16\}$. We consider an alignment
-successful if the estimated coordinates of the eye-corners
-are within 1 pixel from the ground truth in the original
-image.  In Figure \ref{fig:multiscale}, we report the
-alignment success rate, averaged over the artificially
-perturbed initial deformations, as a function of the
-standard deviation of the Gaussian kernel $\sigma$, for
-three choices of the number of scales. As one can see,
-using multiscale indeed improves the performance, and when
-3 scales are used, a smaller convolution kernel can achieve
-a similar performance compared to a much larger kernel when
-only 2 scales are used.
-\begin{figure}
-\centering
-\includegraphics[width=3in]{figures_pami/multiscale.png}
-\vspace{-2mm}
-\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
-A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
-\label{fig:multiscale}
-\vspace{-4mm}
-\end{figure}
-
-\noindent{3) \em 3D Pose Variation.} As densely sampled pose
-and
-    illumination face images are not available in any of
-    the public databases, including Multi-PIE, we have
-    collected our own dataset using our own system (to be
-    introduced in the next section). We use frontal face
-    images of a subject under the 38 illuminations proposed
-    in the next section as training. For testing, we
-    collect images of the subject under a typical indoor
-    lighting condition at pose ranging from $-90^\circ$ to
-    $+90^\circ$ with step size 5.625$^\circ$, a total of 33
-    poses. We use Viola and Jones' face detector to
-    initialize our alignment algorithm.
-Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
-to $\pm 45^\circ$.
-Note that this level of out-of-plane
- pose variation is beyond what we intend to handle with our formulation.
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{ccccc}
-\includegraphics[height=0.5in]{figures_cvpr/5} &
-\includegraphics[height=0.5in]{figures_cvpr/7} &
-\includegraphics[height=0.5in]{figures_cvpr/09} &
-\includegraphics[height=0.5in]{figures_cvpr/11} &
-\includegraphics[height=0.5in]{figures_cvpr/13} \vspace{0mm} \\
-(a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=0.5in]{figures_cvpr/15} &
-\includegraphics[height=0.5in]{figures_cvpr/17} &
-\includegraphics[height=0.5in]{figures_cvpr/19} &
-\includegraphics[height=0.5in]{figures_cvpr/21} &
-\includegraphics[height=0.5in]{figures_cvpr/3} \vspace{0mm} \\
-(f) & (g) & (h) & (i) & (j) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
-to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-4mm}
-\end{figure}
-\subsection{Comparison with Related Work}
-Our modification to SRC roots solidly in the tradition of
-adding deformation-robustness to face recognition algorithms
-\cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}.
-However, the only previous work to investigate face alignment
-in the context of sparse signal representation and SRC is the
-work of \cite{Huang2008-CVPR}. They consider the case where the
-training images themselves are misaligned and allow one
-deformation per training image. They linearize the training
-rather than the test, which is computationally more costly as
-it effectively triples the size of the training set. In
-addition, as they align the test image to all subjects
-simultaneously, it potentially is more prone to local minima
-when the number of subjects increases, as we will see in the
-following experimental comparisons.
-\begin{enumerate}
-\item {\em Extended Yale B.} In this experiment, we have
-    used the same experimental settings as in
-    \cite{Huang2008-CVPR}. 20 subjects are selected and
-    each has 32 frontal images (selected at random) as
-    training and another 32 for testing. An artificial
-    translation of 10 pixels (in both $x$ and $y$
-    directions) is introduced to the test image. For our
-    algorithm we downsample all the images to $88\times
-    80$ for memory reasons, whereas the work of
-    \cite{Huang2008-CVPR} uses random projections.
-Note that the use of cropped images in this experiment introduces image boundary effects.
-    Our
-    algorithm achieves the recognition rate 93.7\%,
-    compared to 89.1\% recognition rate reported in
-    \cite{Huang2008-CVPR}.
-\item {\em CMU Multi-PIE.} In this experiment, we choose
-    all subjects from the CMU Multi-PIE database, 7
-    training images from Session 1 and 1 test image from
-    Session 2 per person. The setting is exactly the same
-    as the previous experiment on 2D deformation. We again
-    work with downsampled images of size $80\times 60$ pixels. An
-    artificial translation of 5 pixels (in both $x$ and $y$
-    directions) was induced in the test image. The
-    algorithm of \cite{Huang2008-CVPR} achieves a
-    recognition rate of 67.5\%,\footnote{That algorithm has
-    two free parameters - $l$ and $d$, which govern the tradeoff between
-	accuracy and run-time. For this experiment
-    we chose $l = 1$ and $d = 514$.} while ours achieves 92.2\%.
-    \end{enumerate}
-
-\section{Handling Illumination Variation}\label{sec:illumination}
-In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.
-Under what conditions is this a reasonable assumption to make?
-What can we say from first principles about how the training images should be chosen?
-
-\subsection{The Illumination Model}
-
-The strongest theoretical results so far regarding the relationship
-between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
-The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
-all images of the object can be well approximated by linear combinations of
-nine (properly chosen) basis images.  The basis images have mixed sign, and
-their illuminations consist of the lowest frequency spherical harmonics.
-While this is a very important result for understanding the image
-formation process, the direct application of this result in most practical
-systems is misguided for several reasons.
-Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
-and they all do so in a way that violates the modeling assumptions of the Basri analysis.
-
-Fortunately, even with these effects, for most materials the relationship between
-illumination and image is still linear,\footnote{Materials that break
-this assumption include fluorescent materials and the photochromic (``Transition'') lenses
-in some eyeglasses.  Most materials emit light in proportion to their
-incident light.} provided the sensor has a linear response curve.\footnote{Proper handling of gamma encoding is an important consideration for
-practitioners.  Most cameras apply a non-linear and often undocumented response
-curve to captured images.  A slight degradation of performance will occur if
-gamma compressed images are treated as if they were linear.  We recommend the use
-of cameras with well documented response curves that can be inverted when the
-image file is loaded.}
-For a more in-depth study
-of the relationship between illumination and images, we refer the reader to
-\cite{belhumeur1998set}.
-While the relationship between illuminations and images is linear,
-only positive weights are allowed; the space of all images of an object with
-fixed pose and varying illumination is a convex cone lying in the positive
-orthant. The question becomes, how many images does it take to do a good job
-of representing images sampled from this cone?
-
-It has been observed in various empirical studies that
-one can get away with using a small number of frontal
-illuminations to linearly represent a wide range of new frontal
-illuminations, when they are all taken under the same laboratory conditions
-\cite{Georghiades2001-PAMI}. This is the case for many public
-face datasets, including AR, ORL, PIE, and Multi-PIE.
-Unfortunately, we have found that in practice, a training
-database consisting purely of frontal illuminations is not
-sufficient to linearly represent images of a faces taken
-under typical indoor or outdoor conditions (see the experiment
-conducted in Section \ref{sec:own-data}). As illustrated by the
-example in Figure \ref{fig:promo}, an insufficient number of
-training illuminations can result in recognition failure. To
-ensure our algorithm works in practice, we need to find a set
-of training illuminations that are indeed {\em sufficient} to
-linearly represent a wide variety of practical indoor and
-outdoor illuminations.
-
-\subsection{Projector-based Illumination System}
-
-We have designed a system that can acquire frontal images of a subject while
-simultaneously illuminating the subject from all directions above horizontal. A sketch of the
-system is shown in Figure \ref{fig:system}: The illumination
-system consists of four projectors that display various bright
-patterns onto the three white walls in the corner of a dark
-room.  The light reflects off of the walls and illuminates the
-user's head indirectly.  After taking the frontal illuminations
-we rotate the chair by 180 degrees and take pictures from the
-opposite direction.  Having two cameras speeds the process
-since only the chair needs to be moved in between frontal and
-rear illuminations. Our projector-based system has several
-advantages over flash-based illumination systems for face recognition:
-\begin{itemize}
-\item The illuminations can be modified in software, rather than hardware.
-\item It is easy to capture many different illuminations quickly.
-\item Good coverage and distant illumination can be achieved simultaneously.
-\item There is no need to mount anything on the walls or construct a large dome.
-\item The system can be assembled from off-the-shelf hardware.
-\end{itemize}
-\begin{figure}
-\centering
-\includegraphics[height=2in]{figures_cvpr/camera_rig.pdf}
-\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
-\vspace{-2em}
-\label{fig:system}
-\end{figure}
-With our projector system, our choice of illuminations is
-constrained only by the need to achieve a good
-SNR\footnote{Since illuminations with more pixels illuminated
-will have a better SNR (provided they don't saturate), there is
-an engineering tradeoff between the SNR and the number of
-training images.}, avoid saturation, and achieve
-a reasonably short acquisition time.  Two simplifying
-assumptions that we make are that every pixel is either turned
-fully on or off in every illumination, and that the illuminated
-regions do not overlap.
-
-Assuming that each pixel is fully on or off enables us to guarantee
-that each illumination image has the same overall intensity, merely
-by guaranteeing that we illuminate the same number of pixels in each image.\footnote{Since DLP projectors may have dramatically different response
-curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
-Since our algorithm depends only on  the
-linearity between the illuminations and the images, and not on the
-relative intensities of the illuminations, the designer has the freedom to choose the overall intensity of the illuminations
-to prevent saturation or low SNR, in a sort of offline exposure control.
-
-Assuming that the sequentially illuminated regions do not overlap results in a
-set of training images that span a larger cone than a similar number of
-overlapping regions.  This results in training images that require fewer
-negative coefficients in $\x$ to represent test images under natural
-illuminations.  The effect of negative coefficients in $\x$ appears to depend
-partly on how the test images are taken and is still under study.
-
-{\em Relationship to existing work:} Most light stages used for face recognition have
-been constructed for the purpose of creating public data sets to study
-illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
-light stages have been used for computer graphics purposes
-\cite{debevec2000acquiring, jones2005performance}.
-The light source can be
-moved around manually \cite{masselus2002free}, but this may result in poor
-consistency of illuminations between users.  Structured light applications use projectors to
-directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
-reconstruction, but this is very disturbing to the user.
-Y.\ Schechner \cite{schechner2007multiplexing}
-studies techniques for multiplexing illumination that can dramatically reduce the noise
-of the demultiplexed images for certain classes of objects and cameras.
-While these techniques have not been incorporated into the current
-system, they fit elegantly into our framework and will likely be used
-in future implementations.  We stress that use of this multiplexing technique
-is independent from the choice of original (directional) illuminations.
-
-\subsection{Choice of Illumination Patterns}
-
-We ran two experiments to guide our choice of illuminations for
-our large-scale experiments:
-\begin{enumerate}
-\item {\em Coverage Experiment.} In the first experiment we
-    attempt to determine what coverage of the sphere is
-    required to achieve good interpolation for test images.
-    The subject was illuminated by 100 (50 front, 50 back)
-    illuminations arranged in concentric rings centered at
-    the front camera.  Subsets of the training images were
-    chosen, starting at the front camera and adding a ring
-    at a time.  Each time a ring was added to the training
-    illumination set, the average $\ell^1$ registration
-    error (residual) for a set of test images taken under
-    sunlight was computed and plotted in Figure
-    \ref{fig:illumination-patterns}(a).  The more rings of
-    training illuminations are added, the lower the
-    representation error becomes, with diminishing returns.
-\item {\em Granularity Experiment.} In the second
-    experiment we attempt to determine how finely divided
-    the illumination sphere should be.  At the first
-    granularity level, the projectors  illuminate the
-    covered area uniformly.  At each subsequent granularity
-    level each illuminated cell is divided in two along its
-    longer side but intensity doubled.  For each
-    granularity level the average $\ell^1$ registration
-    error is computed as in the coverage experiment and
-    shown in Figure \ref{fig:illumination-sufficiency}(b).
-    Again, diminishing returns are observed as more
-    illuminations are added.
-\end{enumerate}
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-\includegraphics[height=.75in]{figures_cvpr/coverage_experiment_asplode.png} &
-\includegraphics[height=.75in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
-(a) Coverage Experiment & (b) Chosen Illumination Patterns
-\end{tabular}\vspace{0mm}
-\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
-For rear illuminations the sequence is reversed.
-In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations.
-The four rectangular regions correspond to the four projectors.
-}
-\label{fig:illumination-patterns}
-\vspace{-.75em}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
-(a) Coverage & (b) Granularity
-\end{tabular}\vspace{0mm}
-\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
-\label{fig:illumination-sufficiency}
-\vspace{-4.5em}
-\end{figure}
-
-In the plot for the coverage experiment, Figure
-\ref{fig:illumination-sufficiency}(a),
- we clearly see two plateau regions: one is after 4 rings
-and one is after 10 rings. The first four rings represent the
-typical frontal illuminations, which are present in most public
-face datasets; however, we see that the residual stabilizes
-after 10 rings which include some illuminations from the back
-of the subject. This suggests that although the frontal
-illuminations account for most of the illumination on the face,
-some illuminations from the back are needed in the training set to
-represent images with illumination coming from all directions.
-In the plot for the granularity experiment, Figure
-\ref{fig:illumination-sufficiency}(b), we observe that the
-residual reaches a plateau after four divisions, corresponding
-to a total of 32 illuminations. Based on the results from both
-experiments, we decide to partition the area covered by the
-first 10 rings into a total of 38 cells, whose layout is
-explained in Figure \ref{fig:illumination-patterns}(b). For
-our large-scale experiments, we have collected those
-illuminations for all our subjects.\footnote{It is possible
-that with further experimentation a reduced set of illuminations
-can be found that performs as well or better.}
-
-See below for the 38 training images of one subject:
-\begin{figure}[h]
-\centering
-\includegraphics[width=3.2in]{figures_cvpr/training.png}
-\vspace{-2em}
-\end{figure}
-
-\section{Tests on Public Databases}\label{sec:multipie}
-In this section and the next section, we conduct comprehensive experiments on
-large-scale face databases to verify the performance of our algorithm and
-system. We first test on the largest public face database available that is
-suitable for testing our algorithm, the CMU Multi-PIE.  One shortcoming of the
-CMU Multi-PIE database for our purposes is that there is no separate set of
-test images taken under natural illuminations; we are left to choose which sets
-of images to use for testing and training.  To challenge our algorithm, we
-choose only a small set of illuminations for the training set, yet we include
-all illuminations in the testing set. In the following section, we will test
-our algorithm on a face dataset that is collected by our own system. The goal
-for that experiment will be to show that with a sufficient set of training
-illuminations for each subject, our algorithm indeed works stably and robustly
-with practical illumination, misalignment, pose, and occlusion, as already
-indicated by our experiment shown in Figure \ref{fig:promo}(bottom).
-
-CMU Multi-PIE provides the most extensive test set among public
-datasets. This database contains images of 337 subjects across
-simultaneous variation in pose, expression, and illumination.
-Of these 337 subjects, we use all of the 249 subjects present
-in Session 1 as the training set. The remaining 88 subjects are
-treated as ``impostors'', or invalid images. For each of the
-249 training subjects, we include frontal images of 7 frontal
-illuminations,\footnote{They are illuminations
-$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
-directional illumination, we subtract the ambient-illuminated
-image 0.} taken with neutral expression. As suggested by the
-work of \cite{Georghiades2001-PAMI}, we choose these extreme
-frontal illuminations in the hope that they would linearly
-represent other frontal illuminations well. For the test set,
-we use all 20 illuminations from Sessions 2-4, which were
-recorded over a period of several months. The dataset is
-challenging due to the large number of subjects, and due to
-natural variation in subject appearance over time.
-Table \ref{tab:MultiPIE-recognition2} shows the result of our
-algorithm on each of the 3 testing sessions. Our algorithm
-achieves recognition rates above $90\%$ for all three sessions.
-For the test images, our iterative alignment was initialized
-automatically via the Viola and Jones' face detector. To
-demonstrate that the sparse representation based recognition
-step is indeed beneficial even when there are no impostors, we
-include results for recognition based only on the alignment
-error residuals (i.e. $S=1$), shown in row 1.
-
-\begin{table}
-\caption{Recognition rates on the Multi-PIE database for
-Algorithm 1 and \cite{Yang2010-CVPR}}
-\centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
-\hline
-{Alg. 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Alg. 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Alg. 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
-\hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition2} }
-\end{table}
-
-\newcommand{\tempwidth}[0]{0.58in}
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-\hspace{-2mm}(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Representative failures from Multi-PIE}.
-{\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2.
-Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
-\label{fig:failed-examples}
-\vspace{-4mm}
-\end{figure}
-
-\subsection{Improving the Sampling Window}
-Our algorithm's errors are mostly caused by a few subjects who
-significantly change their appearances between sessions (such
-as hair, facial hair, and eyeglasses). Some representative
-examples are shown in Figure \ref{fig:failed-examples}. For those subjects, alignment and recognition fail on
-almost all test illuminations.
-
-Meanwhile, this observation also suggests that we might be able
-to improve the performance of our method by carefully choosing
-a face region which is less affected by the above factors for
-recognition. In particular, since the forehead region is likely
-to be affected by the change of hair style, we try replacing
-the previous $80 \times 60$ canonical frame with a new
-window that better excludes the forehead. We adjust the
-resolution of the window to keep $m$ approximately constant. In addition,
-we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
-the observation that in many cases the corners
-actually contain background. An example of the new window
-is shown in Figure~\ref{fig:new-mask}.
-
-\begin{figure}
-\vspace{-1em}
-\centering
-{\small
-\begin{tabular}{@{}cc@{}}
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example.png} &
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example_new.png} \vspace{0mm} \\
-Default window. & Proposed window. \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Choosing different sampling windows.}}
-\label{fig:new-mask}
-\vspace{-1em}
-\end{figure}
-
-Table \ref{tab:MultiPIE-recognition2} shows that the
-recognition rates on Multi-PIE indeed increase with this new
-window. In addition, Figure \ref{fig:failed-examples}(a), (b),
-and (c) illustrate three representative subjects for which the
-recognition rates of our algorithm are significantly boosted
-with the new window. However, we should mention that the best
-choice of the window is problem-specific and there is not a
-simple guideline to follow. For example, although the new
-window performs better on Multi-PIE, the same window does not
-help at all on our own database, which will be introduced in
-the next section. This is because most of the training and
-testing images in our database are taken on the same day so the
-variation in hair style is very small. Hence, excluding the
-forehead part may actually result in loss of useful
-discriminative information.
-
-\subsection{Comparison to Existing Work}
-We first compare our result to the recent work
-\cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
-initial registration is obtained from manually selected outer
-eye corners. Then, a supervised hierarchical sparse coding
-model based on local image descriptors is trained, which enjoys
-certain translation invariant properties. With the same
-training and testing sets, \cite{Yang2010-CVPR} is able to
-handle the remaining misalignment and achieves state-of-the-art
-performance on the CMU Multi-PIE database.
-Table~\ref{tab:MultiPIE-recognition2} shows that our algorithm
-achieves similar or better performance on different sessions of
-Multi-PIE.
-
-% Classical algorithms
-To better examine the effectiveness of our iterative alignment
-algorithm, we next compare our result to baseline
-linear-projection-based algorithms, such as Nearest Neighbor
-(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
-Discriminant Analysis (LDA)
-\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
-PCA \cite{Turk1991-CVPR} as its performance is always below
-that of Nearest Subspace.} Since these algorithms assume
-pixel-accurate alignment, they are not expected to work well if
-the test image is not well aligned with the training. In
-Table~\ref{tab:MultiPIE-recognition}, we report the results of
-these classical algorithms with three types of testing image
-alignment: 1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye
-corners,\footnote{Two manually clicked points are sufficient to
-define a similarity transformation. All of the experiments in
-this section are carried out with similarity transformations.}
-and 3.\ the output of our iterative alignment algorithm. The
-performance drop of the LDA algorithm on Multi-PIE reported
-here seems to agree with that reported already in
-\cite{Gross2008-FGR}.  All of the classical algorithms benefit
-greatly from being paired with our iterative alignment
-algorithm.
-
-\begin{table*}
-\centering
-\caption{Recognition rates on the Multi-PIE database for
-different pairings of alignment and recognition stages.}
-\centerline{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
-\hline
-\backslashbox{Rec.}{Align.}
-& \multicolumn{3}{|c|}{Face Detector}
-& \multicolumn{3}{|c|}{Manual}
-& \multicolumn{3}{|c|}{Iterative Alignment}
-\\
-\hline
-Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
-\hline
-NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
-\hline
-NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
-\hline
-LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
-\hline
-LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
-\hline
-SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition} }
-\vspace{-.1in}
-\end{table*}
-
-% LBP
-We also compare our result to Local Binary Patterns (LBP)
-\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
-able to capture fine details of facial appearance and texture.
-Due to its robustness to variations in illumination, facial
-expression, aging and other changes, LBP has achieved the
-state-of-the-art face recognition performance in the scenario
-when only one sample per person is used for training
-\cite{Tan06facerecognition}.
-In this paper, we follow the same
-steps as in \cite{Ahonen2006-PAMI} to construct an LBP
-descriptor for each training and testing sample. The $80\times
-60$ face region is first divided into a regular $10\times 10$
-grid of cells, each of size $8\times 6$ pixels. Within each
-cell, the histogram of 59 uniform binary patterns is then
-computed, where the patterns are generated by thresholding 8
-neighboring pixels in a circle of radius 2 using the central
-pixel value. Finally, the local histograms are concatenated to
-produce the global descriptor vector. As suggested in
-\cite{Ahonen2006-PAMI}, the recognition is performed using a
-nearest neighbor classifier with Chi square distance as the
-distance measure and we report the recognition rates with the
-same three types of input as before.  As shown in
-Table~\ref{tab:MultiPIE-recognition}, although LBP achieves
-competitive recognition rates given manually aligned training
-and testing samples, demonstrating its robustness to moderate
-misalignment, it still benefits from using the output of our
-iterative alignment algorithm as the input.
-In addition, like
-the other classical algorithms, the performance of LBP degrades
-dramatically if it is applied directly to the output of a face
-detector. This is notable given that LBP is often applied
-without any special alignment in practice. Finally, we
-attribute the improvement in performance of LBP over SRC in
-this experiment to its robustness to illumination components
-that cannot be linearly interpolated by the training set.
-
-In addition, although our algorithm is not designed for
-recognition when there is only a single gallery image per user,
-we compare its performance with LBP within this setting for
-completeness. For this experiment, we use the FERET dataset
-\cite{phillips1998feret}, which contains five standard
-partitions: `fa' is the gallery containing 1196 frontal images
-of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
-sets of probe images. The testing sets differ from the training
-in facial expression (`fb'), illumination (`fc'), aging (`dup1'
-) and long aging (`dup2'). In fact, except for `fb', we notice
-significant changes of illumination in all the other three test
-sets. For the training, we again crop and normalize the face
-region from each original image to an $80\times 60$ window
-using manually marked eye coordinates \cite{Deng2010-PR}. In
-Table~\ref{tab:FERET-recognition}, we report the performance of
-our algorithm on the four test sets, with input directly
-obtained from the Viola and Jones' detector. We also report the
-performance of LBP with the same three types of input as before:
-we use letters
-``$d$'', ``$m$'', and ``$i$'' to indicate face detector, manual
-alignment, and our iterative alignment algorithm, respectively.
-
-\begin{table}
-\caption{Performance on single gallery image FERET dataset}
-\vspace{-3mm} \centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate \% & fb & fc & dup1 & dup2 \\
-\hline
-$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
-\hline
-$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
-\hline
-$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
-\hline
-{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
-\hline
-\end{tabular}
-\label{tab:FERET-recognition} }
-\vspace{-12mm}
-\end{table}
-
-As expected, our algorithm does not perform well except for
-`fb', in which the illumination is similar to the training and
-the mere variation in facial expression is handled well by the
-sparse error model. For the other three test sets, our
-algorithm fails because the illumination changes and other
-variations seriously violate the assumptions of our method.
-This also explains why LBP performs worse with our iterative
-alignment algorithm, compared to manual alignment. On the other
-hand, while LBP achieves the best recognition rates given
-manually aligned training and testing samples, its performance
-degrades drastically when the input is obtained directly from
-the face detector. It is also worth noting that similar poor
-performance of LBP, as well as other descriptors, has been
-observed on the Labeled Face in the Wild (LFW) database, where
-the training is uncontrolled and limited and the input is
-directly obtained from the face detector \cite{Wolf2008-ECCV}.
-
-All of these experimental results confirm that
-both illumination and alignment need to be simultaneously handled
-well in order to achieve accurate face recognition, even when there is
-no obvious occlusion or corruption in the test.
-
-\subsection{Subject Validation}
-
-We test the algorithms' ability to reject invalid images of the
-88 subjects not appearing in the training database. As
-mentioned before, the \emph{sparsity concentration index} (SCI)
-is used as the outlier rejection rule. Given the sparse
-representation $\x$ of a test image with respect to $K$
-training classes, the SCI measures how concentrated the
-coefficients are on a single class in the dataset and is
-defined as in \cite{Wright2009-PAMI}:
-\begin{displaymath}
-\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
-\|\x\|_1 - 1}{K - 1} \in [0,1] .
-\end{displaymath}
-It is easy to see that if $\textup{SCI}(\x) = 1$, the test
-image is represented using images from one single subject
-class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
-evenly over all classes. Thus, we can choose a threshold $t \in
-[0,1]$ for the proposed method and accept a test image as valid
-if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
-invalid. We compare this classifier to classifiers based on
-thresholding the error residuals of NN, NS, LDA, and LBP.
-\begin{figure}[t]
-{\small
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
-(a) & \hspace{2mm}(b) \\
-\end{tabular}
-}
-\vspace{-2mm}
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database,
-(a) for all algorithms with iterative alignment, and
-(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-2mm}
-\end{figure}
-
-Figure \ref{fig:roc-multipie} plots the receiver operating
-characteristic (ROC) curves, which are generated by sweeping
-the threshold $t$ through the entire range of possible values
-for each algorithm.\footnote{Rejecting invalid images not in
-the entire database is much more difficult than deciding if two
-face images are the same subject. Figure \ref{fig:roc-multipie}
-should not be confused with typical ROC curves for face
-similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
-that the SCI based recognition approach significantly
-outperforms the other algorithms, including LBP, even when all
-algorithms are coupled with our proposed iterative alignment.
-In the right plot we again see that classical algorithms, and
-even LBP, are very sensitive to alignment.  Similar contrasts
-between our algorithm and baseline algorithms were also
-observed for SRC in \cite{Wright2009-PAMI}, though on much
-smaller datasets.
-
-\subsection{Recognition with Synthetic Random Block Occlusion}
-
-We further test the robustness of our $\ell^1$-norm based
-algorithm to synthetic occlusion. We simulate various levels of
-occlusion from 10\% to 50\% by replacing a randomly located
-block of the face image with an image of a baboon, as shown in
-Figure~\ref{fig:multipie-occ-rec}. In this experiment, to avoid
-any other factors that may contribute to extra occlusion of the
-face (such as the change of hair style), we choose illumination
-10 from Session 1\footnote{This is the same session as the
-training set.} as testing. The rest of the experimental setting
-remains unchanged. The table in
-Figure~\ref{fig:multipie-occ-rec} shows that our algorithm is
-indeed capable of handling a moderate amount of occlusion. For
-example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.
-
-\newcommand{\tempscale}{0.40}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ10.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ20.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ30.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ40.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ50.png}  \\
-\end{tabular}
-\vspace{-1mm}
-\caption{{\bf Recognition under varying level of
-random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:}
-\vspace{2mm}
-{\small
-\begin{tabular}{|c|c|c|c|c|c| }
-\hline
-Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
-\hline
-Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
-\hline
-\end{tabular}
-}
-\label{fig:multipie-occ-rec}
-\vspace{-.5in}
-\end{figure}
-
-\subsection{Recognition with Pose and Expression} We now run tests of
-our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
-Using the same training set as above, we test our algorithm on
-images in Session 2 with $15^\circ$ pose, for all 20
-illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
-algorithm on images in Session 3 with smile, again for all 20
-illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
-the performance of our method will be significantly improved if pose and expression data
-are available in the training.
-
-%\vspace{-.1in}
-\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
-system we described in Section \ref{sec:illumination}, and shown in Figure
-\ref{fig:system}, we have collected the frontal view of 109
-subjects {\em without eyeglasses} under 38 illuminations shown
-in Figure \ref{fig:illumination-patterns}. For testing our
-algorithm, we have also taken 935 images of these subjects with
-a different camera under a variety of practical conditions.
-\vspace{-.1in}
-
-\subsection{Necessity of Rear Illuminations} To see how
-training illuminations affect the performance of our algorithm
-in practice, we now compare how well a few frontal
-illuminations can linearly represent: 1. other frontal illuminations
-taken under the same laboratory conditions, and 2. typical
-indoor and outdoor illuminations. To this end, we use the face
-database acquired by our system and use 7 illuminations per
-subject as training. The illuminations are chosen to be similar
-to the 7 illuminations used in the previous experiment on
-Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
-13, 18, 21, 22\}$ shown in Figure
-\ref{fig:illumination-patterns}(b) to mimic the illuminations
- $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
-our algorithm on the remaining $24 - 7 = 17$ frontal
-illuminations for all the subjects. The recognition rate is
-$99.8\%$, nearly perfect. We also test our algorithm on 310
-indoor images and 168 outdoor images of these subjects taken
-under a variety of lighting conditions (category 1 and 2
-specified below), similar to the one shown in Figure
-\ref{fig:promo}, and the recognition rates for indoor and
-outdoor images drop down to $94.2\%$ and $89.2\%$,
-respectively. This is a strong indication that
-frontal illuminations taken under laboratory conditions
-are insufficient for representing test images under typical indoor and
-outdoor illuminations.
-\vspace{-0.1in}
-
-\renewcommand{\tempscale}{0.45}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} \\
-\end{tabular}
-\caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
-\vspace{-.1in}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG} \\
-\end{tabular}
- \caption{{\bf Representative examples of category C4}.
-Top row: successful examples with our method using overlapping blocks.
-Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.55in}
-\end{figure}
-
-\subsection{Large-Scale Test with Sufficient Training
-Illuminations} Now we use all 109 subjects and 38 illuminations
-in the training and test on 935 images taken under a variety of
-practical illuminations and conditions. We have manually partitioned the test images into four main
-categories:
-\begin{description}
-\item[C1:] 310 \emph{indoor} images of 72 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 1).
-\item[C2:] 168 \emph{outdoor} images of 48 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 2).
-\item[C3:] 211 images of 32 subjects with \emph{eyeglasses}
-    (Fig.~\ref{fig:examples1-3}, row 3).
-\item[C4:] 246 images of 56 subjects with \emph{sunglasses}
-    (Fig.~\ref{fig:examples4}).
-\end{description}
-We apply Viola and Jones' face detector on these images and
-directly use the detected faces as the input to our algorithm.
-Table~\ref{tab:UIUC-recognition} reports the performance of our
-algorithm on each category.
-Since our focus is on face
-recognition, the errors do not include failures of the face
-detector on some of the more challenging images.
-As one can see, our algorithm achieves higher than 95\%
-recognition rates on categories 1-3. Furthermore, using the
-full set of 38 illuminations indeed improves the performance of
-our system under practical illumination conditions compared to
-only using a small subset of 7 illuminations. However, the
-performance dramatically drops when the faces are occluded by
-various types of sunglasses, which could cover up to 40\% of
-the entire face. Given the previous experimental results on
-synthetic random block occlusions, and given that the
-illuminations are more challenging, the result is not
-surprising. In the next subsection, we will show how additional
-assumptions can be used to improve the recognition performance.
-\begin{table}[h]
-\vspace{0mm}
-\centering \caption{Recognition rates on our own
-database.} \vspace{0mm}
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Test Category & C1 & C2 & C3 & C4  \\
-\hline
-\hline
-Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
-\hline
-\end{tabular}
-\label{tab:UIUC-recognition}
-\vspace{0mm}
-\end{table}
-
-\subsection{Improving the Performance with Occlusion using Overlapping Blocks}
-A traditional approach to improve the performance of face
-recognition under severe occlusion is to use subregions instead
-the entire face as a whole. This idea has been explored in many
-earlier works; see \cite{Pentland1994-CVPR, Wright2009-PAMI}
-for examples. Since in most real world cases the occlusion is contiguous, it is reasonable to argue that a minority of the
-subregions are likely to be affected by the occlusion. In this
-paper, we adopt the same idea and partition the face into four
-overlapping blocks to better handle sunglasses. This
-scheme is illustrated in Figure~\ref{fig:occ-block}. Notice
-that in this example three out of the four blocks are partially
-or almost completely occluded. In our experiment, each block is
-of size $90\times 48$ and covers about two-fifths of the entire
-face. The testing and training sets are partitioned in the same
-way. We then independently apply
-Algorithm~\ref{alg:deformable-src} and compute a sparse
-representation after registration for each block independently
-with respect to the training set. The recognition
-results for individual blocks are then aggregated by voting.
-
-\begin{figure}
-\centering
-\includegraphics[width=2.6in]{figures_pami/occ_block.png}
-\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.
-(b) The four overlapping blocks.
-(c) The sparse representation is calculated after alignment for each block independently.
-The red lines correspond to his true identity.
-(d) The true identity is successfully recovered by voting based on the SCI scores.}
-\label{fig:occ-block}
-\vspace{0mm}
-\end{figure}
-
-In this experiment, we found that the using the \emph{sparsity
-concentration index} (SCI) scores for voting achieves higher
-recognition rate than the residual measure used in Algorithm~\ref{alg:deformable-src}, on
-category 4 (sunglasses) of our database. The recognition rate
-is increased to 78.3\%, compared to 40.9\% obtained without
-this partition scheme. This is another evidence of the superior
-ability of SCI on subject validation, since a heavily occluded
-block can be regarded as an outlier for recognition and should
-be rejected while voting.
-
-However, we should point out that a major problem with this
-approach is that occlusion cannot always be expected to fall within
- any fixed partition of the face image. Therefore, the
-proposed scheme should only be viewed as an example which shows
-that the performance under occlusion can be boosted by
-leveraging local information of a face as well as global information. We
-leave the investigation of more general models (e.g., MRF \cite{ZhouZ2009}) for face
-recognition with both misalignment and occlusion as an
-interesting future work.
-
-\section{Conclusion}\label{sec:conclusion}
-Using a well-though-out combination of existing ideas
-(iterative image alignment, $\ell_1$-error function, SRC, using projectors for
-illumination), we have proposed a system for recognizing human faces
-from images taken under practical conditions that is conceptually simple, well
-motivated, and competitive with state-of-the-art recognition systems for access
-control scenarios.
-
-The system achieves extremely stable performance under
-a wide range of variations in illumination, misalignment, and even under small amounts of
-pose and occlusion. We achieve very good recognition performance on
-large-scale tests with public datasets as well as our practical face
-images, while using only frontal 2D images in the gallery and no
-explicit 3D face model.
-Our system could potentially be extended to better handle large pose
-and expression, either by incorporating training images with different poses or
-expressions or by explicitly modeling and compensating the associated deformations
-in the alignment stage.
-
-Another important direction for future
-investigation is to extend the alignment algorithm to better
-tackle contiguous occlusion. We have demonstrated that misalignment can be naturally handled within the
-sparse representation framework. More complicated models for
-spatial continuity, such as Markov random fields, have also
-been successfully integrated into the computation of a sparse
-representation of well-aligned test images
-\cite{Cevher2008-NIPS, ZhouZ2009}. A unified approach
-for face alignment and recognition in the presence of
-contiguous occlusion remains an open problem.
-
-\section*{Acknowledgment}
-This work was supported by grants NSF IIS 08-49292, NSF ECCS
-07-01676, and ONR N00014-09-1-0230. JW thanks Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
-Mathematics for discussions related to this work, and
-acknowledges support from a Microsoft Fellowship and
-the Lemelson-Illinois Student Prize. %} %\vspace{0mm}
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{42}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
 \bibliographystyle{IEEEtran}

### 2011-04-06 appendix_real.tex
Automating pdf creation...

--- 2011-04-05 faces_system.tex
@@
-%% bare_jrnl.tex
-%% V1.1
-%% 2002/08/13
-%% by Michael Shell
-%% mshell@ece.gatech.edu
-%%
-%% NOTE: This text file uses MS Windows line feed conventions. When (human)
-%% reading this file on other platforms, you may have to use a text
-%% editor that can handle lines terminated by the MS Windows line feed
-%% characters (0x0D 0x0A).
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.6 or later) with an IEEE journal paper.
-%%
-%% Support sites:
-%% http://www.ieee.org
-%% and/or
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
-%%
-%% This code is offered as-is - no warranty - user assumes all risk.
-%% Free to use, distribute and modify.
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% Testflow can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/testflow
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print. Authors are encouraged to use U.S. letter paper when
-% submitting to IEEE. Use the testflow package mentioned above to verify
-% correct handling of both paper sizes by the author's LaTeX system.
-%
-% Also note that the "draftcls", not "draft", option should be used if
-% it is desired that the figures are to be displayed in draft mode.
-
-% *************************************************************************
-% You can now use this template both for submitting to peer review and,
-% if your paper is accepted, for sending in final publication materials.
-% To change from peer review format (single column, double-spaced) to
-% final publication format (double column, single-spaced), just move the
-% comment-line sign (%) from one \documentclass line to the other.
-% The first is for peer review format(single column), the second is for final publication(double column).
-
-%\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
-%\documentclass[9.5pt,journal,final,compsoc,twocolumn]{IEEEtran}
-
-\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran} % from the new style file.  supposedly double column.
-
-% *************************************************************************
-
-\usepackage{times}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amscd}
-\usepackage{amssymb}
-\usepackage{subfigure}
-\usepackage{pdfsync}
-\usepackage{helvet}
-\usepackage{url}
-\usepackage{color}
-
-% Include other packages here, before hyperref.
-\usepackage{mydefs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-\usepackage{multirow}
-\usepackage{pict2e}
-\usepackage{slashbox}
-
-% \usepackage[small]{caption}
-
-\newcommand{\subj}{\textup{subj to}}
-
-\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
-%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
-%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}
-
-\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}
-
-\begin{document}
-%
-% paper title
-\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
-Illumination by Sparse Representation}
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
-        John~Wright,~\IEEEmembership{Member,~IEEE,}
-        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
-        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
-        Hossein~Mobahi,
-        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{A.
-Wagner, A.
-Ganesh, Z.
-Zhou, and Y.
-Ma are with the Dept.
-of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign.
-H.
-Mobahi is with the Computer Science Dept.
-at the University of Illinois at Urbana-Champaign.
-J.
-Wright and Y.
-Ma are with Microsoft Research Asia.
-Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W.
-Main st.
-Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-
-% If you want to put a publisher's ID mark on the page
-% (can leave text blank if you just want to see how the
-% text height on the first page will be reduced by IEEE)
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
-
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%% ABSTRACT
-\IEEEcompsoctitleabstractindextext{
-\vspace{0mm}
-\begin{abstract}
-Many classic and contemporary face recognition algorithms work
-well on public data sets, but degrade sharply when they are
-used in a real recognition system. This is mostly due to the
-difficulty of simultaneously handling variations in
-illumination, image misalignment, and occlusion in the test
-image. We consider a scenario where the training images are
-well controlled, and test images are only loosely controlled.
-We propose a conceptually simple face recognition system that
-achieves a high degree of robustness and stability to
-illumination variation, image misalignment, and partial
-occlusion. The system uses tools from sparse representation to
-align a test face image to a set of frontal training images.
-The region of attraction of our alignment algorithm is computed
-empirically for public face datasets such as Multi-PIE. We
-demonstrate how to capture a set of training images with enough
-illumination variation that they span test images taken under
-uncontrolled illumination. In order to evaluate how our
-algorithms work under practical testing conditions, we have
-implemented a complete face recognition system, including a
-projector-based training acquisition system. Our system can
-efficiently and effectively recognize faces under a variety of
-realistic conditions, using only frontal images under the
-proposed illuminations as training.\vspace{0mm}
-\end{abstract}
-
-\begin{IEEEkeywords}
-Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
-Sparse Representation,  Error Correction, Validation and Outlier Rejection.
-\end{IEEEkeywords}}
-
-\maketitle
-
-\IEEEdisplaynotcompsoctitleabstractindextext
-\vspace{-9mm} % HACK
-
-\IEEEpeerreviewmaketitle
-
-%%%%%%%%% BODY TEXT
-\section{Introduction}
-%Automatic face recognition remains one of the most active areas in computer vision.
-
-\PARstart{F}{ace} recognition applications to date have fallen
-into roughly two categories.  Face recognition has recently
-seen a lot of success in a family of less-demanding
-applications such as online image search and family photo album
-organization (e.g.\ Google Picassa, Microsoft Photo Gallery,
-and Apple iPhoto). At the other end of the tractability
-spectrum there are the terrorist watchlist and mass
-surveillance applications that have for the most part dominated
-the field of face recognition research.  However, there are
-many face recognition applications that fall roughly between
-these extremes, where very high recognition performance is
-desired, but the users in the gallery are still allies of the
-system rather than adversaries.  These applications include
-access control for secure facilities (e.g., prisons, office
-buildings), computer systems, automobiles, or automatic teller
-machines, where controlled gallery images can be obtained in
-advance.  These applications are very interesting due to their
-potential sociological impact.  Since the gallery subjects are
-allies, rather than opponents, of the recognition system, this
-creates the possibility of carefully controlling the
-acquisition of the training data. While the same can be said
-for other biometrics such as fingerprints and iris recognition,
-face recognition has the potential of working with test data
-that is much less controlled, allowing the access control
-system to be made less intrusive to the users of the system. To
-some extent, the goal of this paper is to show how a reliable
-face recognition system can be built for this restricted, but
-still important, scenario.
-
-Very few recognition systems specifically target applications where many well-
-controlled training images are available.  Of these, the classical holistic
-subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
-are well known for their speed and simplicity, as well as for their natural
-extension to linear illumination models.  However, their performance has been
-shown to be extremely brittle not only to alignment variation, but to even
-minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
-is slightly open. One promising recent direction, set forth in
-\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
-sparse representation of the test image in terms of the training set as a
-whole, up to some sparse error due to occlusion. A \emph{sparse
-representation-based classification} (SRC) method is then proposed for
-recognition. The main idea is that the sparse nonzero coefficients should
-concentrate on the training samples with the same class label as the test
-sample. SRC has demonstrated striking recognition performance despite severe
-occlusion or corruption by solving a simple convex program.
-%For this reason, the final recognition stage of our proposed system will
-%consist of SRC, and the alignment stage that precedes it will be based on
-%similar techniques.
-
-\begin{figure}
-\centering \begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
-\end{tabular} \caption{{\bf Effects of registration and illumination on
-Recognition}. In this example we identify the girl among 20 subjects, by
-computing the sparse representation of her input face with respect to the
-entire training set. The absolute sum of the coefficients associated with each
-subject is plotted on the right. We also show the faces reconstructed with each
-subject's training images weighted by the associated sparse coefficients. The
-red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
-input face is from Viola and Jones' face detector (the black box) and all 38
-illuminations specified in Section \ref{sec:illumination} are used in the
-training.  {\bf Middle:} The input face is well-aligned (the white box) with
-the training by our algorithm specified in Section \ref{sec:registration} but
-only 24 frontal illuminations are used in the training for recognition (see
-Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and
-a sufficient set (all 38) of
-illuminations are used in the training. Both are necessary for correct recognition
-using SRC.}\label{fig:promo}
-\vspace{-6.9mm}
-\end{figure}
-
-Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
-between the test and training images, and it requires a rich set of illuminations
-in the gallery images for good performance.
-We illustrate the compounded effect of both alignment and illumination
-with an example in Figure \ref{fig:promo}.
-The task is to identify the girl among 20 subjects. If the test
-face image, say obtained from an off-the-shelf face detector,
-has even a small amount of registration error against the
-training images (caused by mild pose, scale, or misalignment),
-the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative,
-even if sufficient illuminations are present in the training, as shown in Figure
-\ref{fig:promo}(top). Additionally, in order to span
-the illuminations of a typical indoor (or outdoor) environment,
-illuminations from behind the subject are needed in the
-training set. Otherwise, even for perfectly aligned test images,
-the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
-informative, as shown by the example in Figure
-\ref{fig:promo}(middle). Clearly, both good alignment, as well
-as sufficient training images are needed to ensure success of the sparsity-based
-recognition method proposed by \cite{Wright2009-PAMI}.
-In this paper, we examine how to handle alignment and illumination simultaneously
-in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
-closer to practical use.
-
-\vspace{-0mm}
-\subsection{Related Work}
-We briefly review existing techniques for recognition, image
-registration, and handling of illumination variation.  Our
-system is based purely on 2D techniques.  This fact immediately
-distinguishes our approach from systems that either require a
-3D data as an input, or attempt to estimate a 3D model from 2D
-input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While
-these techniques can achieve better robustness to pose
-variation given a sufficiently accurate 3D model, for access
-control applications where only moderate pose variation is
-present, the proposed method will be more than sufficient.
-Note that 2D images of faces under varying illuminations
-already contain 3D shape-related information,\footnote{In
-principle, one can recover the 3D shape of the face from
-multiple illuminations using photometric stereo.} and this
-information can be leveraged by 2D algorithms for alignment and
-recognition even if shape is not reconstructed explicitly.
-
-% Alignment
-In holistic recognition algorithms, correspondence between points in the test
-image and in the training must be achieved.  A long line of research exists on
-using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
-Active Shape Models \cite{cootes1992active} to register images against a
-relatively high-dimensional model of plausible face appearances, often
-leveraging face-specific contours.  While these model-based techniques have
-advantages in dealing with variations in expression and pose, they may add
-unnecessary complexity to applications where subjects normally present a neutral face
-or only have moderate expression. We prefer to focus on deformations with far fewer
-degrees of freedom, i.e. similarity transformations, and to use the training images themselves
-as the appearance model. Iterative registration in this spirit dates at least back to the
-Lucas-Kanade algorithm \cite{lucas1981iterative}.
-
-Whereas much of the early work on image registration is aimed at the
-problem of registering nearly identical images, say by minimizing a sum of
-squared distances or maximizing normalized correlation, here we must confront several physical factors
-simultaneously: misalignment, illumination variations, and corrupted pixels.
-As we discuss further below, illumination variation can be dealt by
-expressing the test image as a linear combination of an appropriate set of
-training images. Similar representations have been exploited in
-illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
-For robustness to gross errors, the $\ell^1$-norm of the residual is a more
-appropriate objective function than the classical $\ell^2$-norm. Its use here
-is loosely motivated by theoretical results due to Cand\`{e}s and Tao
-\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
-lead us to pose the registration problem as the search for a set of
-transformations and illumination coefficients that minimize the $\ell^1$-norm
-of the representation error. We solve this problem using a generalized
-Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
-minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
-these problems can also be solved efficiently using recently developed
-first-order techniques for $\ell^1$-minimization, which are reviewed in
-\cite{YangA2010-pp}.
-
-% Illumination
-Researchers have tried various techniques to deal with illumination variation.
-In almost all recognition algorithms where only a single gallery image is
-available per individual, illumination effects are regarded as a nuisance that
-must be removed before the algorithm can continue.  This is typically done by
-making statistical assumptions about how illumination affects the image, and
-using those assumptions to extract a new representation that is claimed to be
-illumination invariant.  Recent examples include \cite{chen2006total} and
-\cite{zhou2007appearance}.  However, despite these efforts, truly
-illumination-invariant features are difficult to obtain from a single input
-image.  We argue that if one has the luxury of designing the acquisition system
-and the application demands a high recognition rate,
-it is then unwise to limit the gallery to a
-single image per person.  We therefore take the strategy of sampling many
-gallery images of each individual under varying illuminations.  These images
-are used as the basis for either a convex cone model
-\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
-\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
-projector based light stage.  While similar systems have been used for
-other applications, to our knowledge, we
-are the first to use projectors to indirectly illuminate a subject's face for
-the purpose of face recognition.
-\subsection{Contributions} In this paper, we show how
-registration and illumination
-can be simultaneously addressed within a robust sparse representation
-framework. We show that face registration, a challenging
-nonlinear problem, can be solved by a series of linear programs
-that iteratively minimize the sparsity of the registration
-error. This leads to an efficient and effective alignment
-algorithm for face images that works for a large range of
-variation in translation, rotation, and scale, even when
-the face is only partially visible due to eyeglasses, closed
-eyes and open mouth, sensor saturation, etc.  We also propose a
-sufficient set of training illuminations
-that is capable of linearly representing typical indoor and outdoor
-lighting, along with a practical hardware system for capturing
-them.
-
-% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.
-While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.
-
-% We use tools from sparse representation
-% to align a test face image with a set of frontal training
-% images in the presence of registration error, illumination variation, and partial occlusion. The region
-% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
-% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
-% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions.
-We show
-% that our system can efficiently and effectively recognize faces under
-% a variety of realistic conditions, using only
-% frontal images under the proposed illuminations as training.
-
-%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.
-We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.
-Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.
-We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.
-We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.
-
-We then demonstrate the effectiveness of the proposed new
-methods with a complete face recognition system that is {\em
-simple, stable, and scalable}. The proposed system performs
-robust automatic recognition of subjects from loosely
-controlled probe images taken both indoors and outdoors,
-using a gallery of
-frontal views of the subjects' faces under the proposed
-illuminations. An off-the-shelf face
-detector\footnote{We use the OpenCV
-implementation of the Viola and Jones' face detector
-\cite{Viola2004-IJCV}.} is used to detect faces in the test images.
-
-We conduct extensive experiments on the proposed system with
-both public databases and a face database that is collected by
-our own acquisition system. Our experimental results on
-large-scale public face databases show that our algorithm
-indeed achieves very good performance on these databases,
-exceeding or competing with the state-of-the-art algorithms.
-Additionally, our experimental results on our own database
-clearly demonstrate that our system not only works well with
-images taken under controlled laboratory conditions, but is
-capable of handling practical indoor and outdoor illuminations as well.
-
-\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
-we derive our robust registration and recognition algorithm within the sparse
-representation framework. We elaborate on algorithmic implementation issues,
-conduct region of attraction experiments with respect to both 2D in-plane
-deformation and 3D pose variation, and discuss its relationship to existing
-work. Section \ref{sec:illumination} is dedicated to our training acquisition
-system. Using this system, we investigate empirically how many training
-illuminations are required to handle practical illumination variations, and
-suggest a sufficient set of 38 training illuminations. Extensive experiments on
-a large-scale public database and on our own database are conducted in Section
-\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
-proposed system. Section \ref{sec:conclusion} concludes our work with
-discussion of promising future directions.
-
-\section{Robust Alignment}\label{sec:registration}
-As demonstrated in Figure \ref{fig:promo}(top), the main
-limitation of the {\em sparse representation and
-classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
-the assumption of pixel-accurate alignment between the test
-image and the training set. This leads to brittleness under
-pose and misalignment, making it inappropriate for deployment
-outside a laboratory setting. In this section, we show how this
-weakness can be rectified while still preserving the conceptual
-simplicity and good recognition performance of SRC.
-
-SRC assumes access to a database of multiple registered
-training images per subject, taken under varying illuminations.
-The images of subject $i$, stacked as vectors, form a matrix
-$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
-form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
-\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
-well-aligned test image $\y_0$ can be represented as a sparse
-linear combination $A \x_0$ of all of the images in the
-database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
-sufficiency in the next section.} plus a sparse error $\e_0$
-due to corrupted pixels. The sparse representation can be recovered by
-minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
-vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
-$\x$ and $\e$:
-\begin{equation}
-\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
-\label{eqn:robust-l1}
-\end{equation}
-Now suppose that $\y_0$ is subject to some pose or
-misalignment, so that instead of observing $\y_0$, we observe
-the warped image $\y = \y_0 \circ \tau^{-1}$, for some
-transformation $\tau \in T$ where $T$ is a finite-dimensional
-group of transformations acting on the image domain.  The
-transformed image $\y$ no longer has a sparse representation of
-the form $\y = A \x_0 + \e_0$, and naively applying the
-algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
-as seen in Figure \ref{fig:promo}(top).
-
-\subsection{Batch and Individual Alignment} If the
-true deformation $\tau^{-1}$ can be found, then we can apply
-its inverse $\tau$ to the test image and it again becomes
-possible to find a sparse representation of the resulting
-image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
-  This sparsity
-provides a strong cue for finding the correct deformation
-$\tau$: conceptually, one would like to seek a transformation
-$\tau$ that allows the sparsest representation, via
-\begin{equation} \label{eqn:L1-L1-conceptual}
-\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
-\end{equation}
-For fixed $\tau$, this problem is jointly convex in $\x$ and
-$\e$. However, as a simultaneous optimization over the
-coefficients $\x$, error representation $\e$, and
-transformation $\tau$, it is a difficult, nonconvex
-optimization problem. One source of difficulty is the presence
-of multiple faces in the matrix $A$:
-\eqref{eqn:L1-L1-conceptual} has many local minima that
-correspond to aligning $\y$ to different subjects. In this
-sense, the misaligned recognition problem differs from the
-well-aligned version studied in \cite{Wright2009-PAMI}. For the
-well-aligned case, it is possible to directly solve for a
-global representation, with no concern for local minima. With
-possible misalignment, it is more appropriate to seek the best
-alignment of the test face with each subject $i$:
-\begin{equation} \label{eqn:per-subject-L1}
-\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
-\end{equation}
-We no longer penalize $\| \x \|_1$, since $A_i$ consists of
-only images of subject $i$ and so $\x$ is no longer expected to
-be sparse.
-
-\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector.
-We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
-\begin{equation}
-\y\circ \tau + J \Delta \tau = A_i \x + \e.
-\end{equation}
-Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
-Jacobian of $\y \circ \tau$ with respect to the transformation
-parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
-above equation is underdetermined if we allow the registration
-error $\e$ to be arbitrary. At the correct alignment we
-expect the test image to differ from $A_i \x$ only
-for the minority of the pixels corrupted by occlusions. Thus,
-we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
-\begin{equation}
-\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
-\label{eqn:L1-align}
-\end{equation}
-This is different from the popular choice that
-minimizes the $\ell^2$-norm of the registration error:
-\begin{equation}
-\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
-\label{eqn:L2-align}
-\end{equation}
-which is also equivalent to finding the deformation step
-$\Delta  \tau$ by solving the least-square problem:
-$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
-\|_2$. Empirically, we find that if there is only small noise
-between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
-$\eqref{eqn:L2-align}$ have similar performance.  However, if
-there are occlusions in $\y_0$, sequential
-$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
-better than sequential $\ell^2$-minimization
-\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
-example.
-
-The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images.
-Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}.
-While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm.
-In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution.
-We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.
-
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{cccc}
-\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
-\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
-\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
-\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
-(a) & (b) & (c) & (d)
-\end{tabular}
-\vspace{0mm}}
-\caption{{\bf Comparing alignment of a subject wearing sunglasses by
-$\ell^1$ and $\ell^2$ minimization.}
-{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
-result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
-given by the face detector, {\em Red (solid):} Alignment result shown on the same
-face; (b) warped testing image using the estimated transformation $\y_0$;
-(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-8.0mm}
-\end{figure}
-
-In addition to normalizing the training images (which is done
-once), it is important to normalize the warped testing image
-$\y \circ \tau$ as the algorithm runs.  Without normalization,
-the algorithm may fall into a degenerate global minimum
-corresponding to zooming in on a dark region of the test
-image.  Normalization is done by replacing the linearization of
-$\y \circ \tau$ with a linearization of the normalized version
-$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
-The proposed alignment algorithm can be easily extended to work
-in a {\em multiscale} fashion, with benefits both in
-convergence behavior and computational cost.  The alignment
-algorithm is simply run to completion on progressively less
-downsampled versions of the training and testing images, using
-the result of one level to initialize the next.
-
-\subsection{Robust Recognition by Sparse Representation} Once
-the best transformation $\tau_i$ has been computed for each
-subject $i$, the training sets $A_i$ can be aligned to $\y$,
-and a global sparse representation problem of the form
-\eqref{eqn:robust-l1} can be solved to obtain a discriminative
-representation in terms of the entire training set. Moreover,
-the per-subject alignment residuals $\| \e \|_1$ can be used to
-prune unpromising candidates from the global optimization,
-leaving a much smaller and more efficiently solvable problem.
-The complete optimization procedure is summarized as Algorithm
-\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
-considered together to provide a sparse representation for the
-test image. If $S = 1$, the algorithm reduces to classification
-by registration error; but considering the test image might be
-an invalid subject, we typically choose $S = 10$. Since valid
-images have a sparse representation in terms of this larger
-set, we can reject invalid test images using the {\em sparsity
-concentration index} proposed in \cite{Wright2009-PAMI}.
-The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
-selects coefficients from the vector $\x$ corresponding to subject $i$.
-
-Another important free parameter in Algorithm
-\ref{alg:deformable-src} is the class of deformations $T$. In
-our experiments, we typically use 2D similarity
-transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
-Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
-detector, or 2D projective transformations, $T =
-\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
-This class of transformations is able to represent distortion
-in a perspective image of a planar object.}$, for handling some
-pose variation.
-
-In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
-performance of our system, based on the observation that the face detector output may be poorly centered on
-the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
-instead of aligning the training sets to the
-original $\y$ directly obtained from the face detector, we
-compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
-\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
-is believed to be better centered, and update $\y$ according to
-$\bar{\tau}$. For the 2D similarity transformations, which are
-used in our system when initialized by the face detector, a
-transformation $\tau$ can be parameterized as $\tau = (\tau^1,
-\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
-the translations in $x$- and $y$-axis, $\tau^3$ represents the
-rotation angle and $\tau^4$ represents the scale. Then the
-average transformation is simply obtained by taking the
-component-wise mean:
-\begin{displaymath}
-\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
-\tau_{k_S}^i) / S, i = 1,2,3,4.
-\end{displaymath}
-Finally, the training sets are aligned to the new $\y$.
-
-\begin{algorithm}[thb]
-\caption{\bf (Deformable Sparse Recovery and Classification for
-Face Recognition)} \label{alg:deformable-src}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
-$\y\in\Re^m$ and a deformation group $T$.
-\STATE
-{\bf for} each subject $i$,
-\STATE \hspace{3mm} $\tau^{(0)}
-\leftarrow I$.
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
-\STATE \hspace{6mm}
-$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
-\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
-\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
-\STATE
-\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$;
-\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-residuals $\|\e\|_1$. \STATE Compute an average transformation
-$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
-\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem: \hspace{2mm}
-$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
-\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
-\end{small}
-\end{algorithmic}
-\end{algorithm}
-%\vspace{-4mm}
-
-The transformation $\tau$ defines a mapping between the
-coordinates of pixels in the large original image and a smaller
-(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
-artifacts in the downsampled image, one should apply a
-smoothing filter to the original image. For a simple implementation, a rectangular window with
-regular sampling can used, but in general, the small image need
-not be regularly sampled in pixel coordinates.  For example,
-the sample locations could be arbitrarily selected from within a
-``face shaped'' area. We will discuss how choosing
-different windows can affect the performance of our algorithm
-in Section \ref{sec:multipie}.
-
-\subsection{System Implementation}
-In this section, we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}. It is
+\appendix[L1 Minimization via Augmented Lagrange Multiplier]
+In this Appendix we discuss the computational issues related to
+the implementation of Algorithm~\ref{alg:deformable-src}, which is
+repeated here for convenience. It is
 not hard to see that its computational complexity is dominated
@@
 recognition!
-
-\subsection{Experiments on Region of Attraction} We will now present three
-experimental results demonstrating the effectiveness of the individual
-alignment procedure outlined in the previous section. They show the sufficiency
-of the region of attraction, verify effectiveness of the multiscale extension,
-and show stability to small pose variations.  We delay large-scale recognition
-experiments to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
-have discussed the issue of illumination in the next section.
-
-\noindent {1) \em 2D Deformation.}  We first verify the
-    effectiveness of our alignment algorithm with images
-    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
-    We select all the subjects in Session 1, use 7
-    illuminations per person from Session 1 for training,
-    and test on one new illumination from Session
-    2.\footnote{The training are illuminations $\{0, 1, 7,
-    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
-    testing is the illumination 10. } We manually select
-    eye corners in both training and testing as the ground
-    truth for registration. We downsample the images to
-    $80\times 60$ pixels\footnote{Unless otherwise stated,
-    this will be the default resolution at which we prepare
-    all our training and testing datasets.} and the distance between the two outer
-    eye corners is normalized to be 50 pixels for each
-    person. We introduce artificial deformation to the
-    testing image with a combination of translation,
-    rotation and scaling. We further use the alignment
-    error $\|\e\|_1$ as an indicator of success. Let $r_0$
-    be the alignment error obtained by aligning a test
-    image to the training images without any artificial
-    perturbation. When the test image is artificially
-    perturbed and aligned, resulting in an alignment error
-    $r$, we consider the alignment successful if $|r - r_0 | \leq
-    0.01r_0$. Figure \ref{fig:attraction} shows the
-    percentage of successful registrations for all subjects
-    for each artificial deformation. The results suggest
-    that our algorithm works extremely well with
-    translation up to 20\% of the eye distance (or 10
-    pixels) in all directions and up to $30^\circ$ in-plane
-    rotation. We have also tested our alignment algorithm
-    with scale variation and it can handle up to 15\%
-    change in scale.
-
-We have gathered the statistics of the Viola and Jones'
- face detector on the Multi-PIE dataset. For 4,600 frontal
- images of 230 subjects under 20 different illuminations,
- using manual registration as the ground truth, the average
- misalignment error of the detected faces is about 6 pixels
- and the average variation in scale is 8\%. This falls
- safely inside the region of attraction for our alignment
- algorithm.
-\begin{figure*}
-\centering
-{\small
-\begin{tabular}{ccc}
-\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
-(a)&(b)&(c)
-\end{tabular}
-\begin{tabular}{cccc}
-\includegraphics[height=1in]{figures_pami/x_tr.png} &
-\includegraphics[height=1in]{figures_pami/y_tr.png} &
-\includegraphics[height=1in]{figures_pami/theta.png} &
-\includegraphics[height=1in]{figures_pami/scale.png}\\
-(d)&(e)&(f)& (g)
-\end{tabular}
-\vspace{0mm}
-}
-\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.
-The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees.
-{\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions.
-(b) Simultaneous translation in $y$ direction and in-plane rotation.
-(c) Simultaneous translation in $y$ direction and scale variation.
-{\bf Bottom row:} (d) Translation in $x$ direction only.
-(e) Translation in $y$ direction only.
-(f) In-plane rotation only.
-(g) Scale variation only.}
-\label{fig:attraction}
-\vspace{0mm}
-\end{figure*}
-
-\noindent{2) \em Multiscale Implementation.}
-Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost.
-Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
-We use the same 7 illuminations in
-Session 1 as training, and all 20 illuminations in the same
-session as testing. We introduce artificial deformation in
-both $x$ and $y$ directions up to 16 pixels in the
-$80\times 60$ frame, with a step size of 4 pixels, i.e.,
-$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
-\{-16,-12,\ldots,12,16\}$. We consider an alignment
-successful if the estimated coordinates of the eye-corners
-are within 1 pixel from the ground truth in the original
-image.  In Figure \ref{fig:multiscale}, we report the
-alignment success rate, averaged over the artificially
-perturbed initial deformations, as a function of the
-standard deviation of the Gaussian kernel $\sigma$, for
-three choices of the number of scales. As one can see,
-using multiscale indeed improves the performance, and when
-3 scales are used, a smaller convolution kernel can achieve
-a similar performance compared to a much larger kernel when
-only 2 scales are used.
-\begin{figure}
-\centering
-\includegraphics[width=3in]{figures_pami/multiscale.png}
-\vspace{-2mm}
-\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations.
-A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
-\label{fig:multiscale}
-\vspace{-4mm}
-\end{figure}
-
-\noindent{3) \em 3D Pose Variation.} As densely sampled pose
-and
-    illumination face images are not available in any of
-    the public databases, including Multi-PIE, we have
-    collected our own dataset using our own system (to be
-    introduced in the next section). We use frontal face
-    images of a subject under the 38 illuminations proposed
-    in the next section as training. For testing, we
-    collect images of the subject under a typical indoor
-    lighting condition at pose ranging from $-90^\circ$ to
-    $+90^\circ$ with step size 5.625$^\circ$, a total of 33
-    poses. We use Viola and Jones' face detector to
-    initialize our alignment algorithm.
-Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
-to $\pm 45^\circ$.
-Note that this level of out-of-plane
- pose variation is beyond what we intend to handle with our formulation.
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{ccccc}
-\includegraphics[height=0.5in]{figures_cvpr/5} &
-\includegraphics[height=0.5in]{figures_cvpr/7} &
-\includegraphics[height=0.5in]{figures_cvpr/09} &
-\includegraphics[height=0.5in]{figures_cvpr/11} &
-\includegraphics[height=0.5in]{figures_cvpr/13} \vspace{0mm} \\
-(a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=0.5in]{figures_cvpr/15} &
-\includegraphics[height=0.5in]{figures_cvpr/17} &
-\includegraphics[height=0.5in]{figures_cvpr/19} &
-\includegraphics[height=0.5in]{figures_cvpr/21} &
-\includegraphics[height=0.5in]{figures_cvpr/3} \vspace{0mm} \\
-(f) & (g) & (h) & (i) & (j) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
-to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-4mm}
-\end{figure}
-\subsection{Comparison with Related Work}
-Our modification to SRC roots solidly in the tradition of
-adding deformation-robustness to face recognition algorithms
-\cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}.
-However, the only previous work to investigate face alignment
-in the context of sparse signal representation and SRC is the
-work of \cite{Huang2008-CVPR}. They consider the case where the
-training images themselves are misaligned and allow one
-deformation per training image. They linearize the training
-rather than the test, which is computationally more costly as
-it effectively triples the size of the training set. In
-addition, as they align the test image to all subjects
-simultaneously, it potentially is more prone to local minima
-when the number of subjects increases, as we will see in the
-following experimental comparisons.
-\begin{enumerate}
-\item {\em Extended Yale B.} In this experiment, we have
-    used the same experimental settings as in
-    \cite{Huang2008-CVPR}. 20 subjects are selected and
-    each has 32 frontal images (selected at random) as
-    training and another 32 for testing. An artificial
-    translation of 10 pixels (in both $x$ and $y$
-    directions) is introduced to the test image. For our
-    algorithm we downsample all the images to $88\times
-    80$ for memory reasons, whereas the work of
-    \cite{Huang2008-CVPR} uses random projections.
-Note that the use of cropped images in this experiment introduces image boundary effects.
-    Our
-    algorithm achieves the recognition rate 93.7\%,
-    compared to 89.1\% recognition rate reported in
-    \cite{Huang2008-CVPR}.
-\item {\em CMU Multi-PIE.} In this experiment, we choose
-    all subjects from the CMU Multi-PIE database, 7
-    training images from Session 1 and 1 test image from
-    Session 2 per person. The setting is exactly the same
-    as the previous experiment on 2D deformation. We again
-    work with downsampled images of size $80\times 60$ pixels. An
-    artificial translation of 5 pixels (in both $x$ and $y$
-    directions) was induced in the test image. The
-    algorithm of \cite{Huang2008-CVPR} achieves a
-    recognition rate of 67.5\%,\footnote{That algorithm has
-    two free parameters - $l$ and $d$, which govern the tradeoff between
-	accuracy and run-time. For this experiment
-    we chose $l = 1$ and $d = 514$.} while ours achieves 92.2\%.
-    \end{enumerate}
-
-\section{Handling Illumination Variation}\label{sec:illumination}
-In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.
-Under what conditions is this a reasonable assumption to make?
-What can we say from first principles about how the training images should be chosen?
-
-\subsection{The Illumination Model}
-
-The strongest theoretical results so far regarding the relationship
-between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
-The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
-all images of the object can be well approximated by linear combinations of
-nine (properly chosen) basis images.  The basis images have mixed sign, and
-their illuminations consist of the lowest frequency spherical harmonics.
-While this is a very important result for understanding the image
-formation process, the direct application of this result in most practical
-systems is misguided for several reasons.
-Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
-and they all do so in a way that violates the modeling assumptions of the Basri analysis.
-
-Fortunately, even with these effects, for most materials the relationship between
-illumination and image is still linear,\footnote{Materials that break
-this assumption include fluorescent materials and the photochromic (``Transition'') lenses
-in some eyeglasses.  Most materials emit light in proportion to their
-incident light.} provided the sensor has a linear response curve.\footnote{Proper handling of gamma encoding is an important consideration for
-practitioners.  Most cameras apply a non-linear and often undocumented response
-curve to captured images.  A slight degradation of performance will occur if
-gamma compressed images are treated as if they were linear.  We recommend the use
-of cameras with well documented response curves that can be inverted when the
-image file is loaded.}
-For a more in-depth study
-of the relationship between illumination and images, we refer the reader to
-\cite{belhumeur1998set}.
-While the relationship between illuminations and images is linear,
-only positive weights are allowed; the space of all images of an object with
-fixed pose and varying illumination is a convex cone lying in the positive
-orthant. The question becomes, how many images does it take to do a good job
-of representing images sampled from this cone?
-
-It has been observed in various empirical studies that
-one can get away with using a small number of frontal
-illuminations to linearly represent a wide range of new frontal
-illuminations, when they are all taken under the same laboratory conditions
-\cite{Georghiades2001-PAMI}. This is the case for many public
-face datasets, including AR, ORL, PIE, and Multi-PIE.
-Unfortunately, we have found that in practice, a training
-database consisting purely of frontal illuminations is not
-sufficient to linearly represent images of a faces taken
-under typical indoor or outdoor conditions (see the experiment
-conducted in Section \ref{sec:own-data}). As illustrated by the
-example in Figure \ref{fig:promo}, an insufficient number of
-training illuminations can result in recognition failure. To
-ensure our algorithm works in practice, we need to find a set
-of training illuminations that are indeed {\em sufficient} to
-linearly represent a wide variety of practical indoor and
-outdoor illuminations.
-
-\subsection{Projector-based Illumination System}
-
-We have designed a system that can acquire frontal images of a subject while
-simultaneously illuminating the subject from all directions above horizontal. A sketch of the
-system is shown in Figure \ref{fig:system}: The illumination
-system consists of four projectors that display various bright
-patterns onto the three white walls in the corner of a dark
-room.  The light reflects off of the walls and illuminates the
-user's head indirectly.  After taking the frontal illuminations
-we rotate the chair by 180 degrees and take pictures from the
-opposite direction.  Having two cameras speeds the process
-since only the chair needs to be moved in between frontal and
-rear illuminations. Our projector-based system has several
-advantages over flash-based illumination systems for face recognition:
-\begin{itemize}
-\item The illuminations can be modified in software, rather than hardware.
-\item It is easy to capture many different illuminations quickly.
-\item Good coverage and distant illumination can be achieved simultaneously.
-\item There is no need to mount anything on the walls or construct a large dome.
-\item The system can be assembled from off-the-shelf hardware.
-\end{itemize}
-\begin{figure}
-\centering
-\includegraphics[height=2in]{figures_cvpr/camera_rig.pdf}
-\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
-\vspace{-2em}
-\label{fig:system}
-\end{figure}
-With our projector system, our choice of illuminations is
-constrained only by the need to achieve a good
-SNR\footnote{Since illuminations with more pixels illuminated
-will have a better SNR (provided they don't saturate), there is
-an engineering tradeoff between the SNR and the number of
-training images.}, avoid saturation, and achieve
-a reasonably short acquisition time.  Two simplifying
-assumptions that we make are that every pixel is either turned
-fully on or off in every illumination, and that the illuminated
-regions do not overlap.
-
-Assuming that each pixel is fully on or off enables us to guarantee
-that each illumination image has the same overall intensity, merely
-by guaranteeing that we illuminate the same number of pixels in each image.\footnote{Since DLP projectors may have dramatically different response
-curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
-Since our algorithm depends only on  the
-linearity between the illuminations and the images, and not on the
-relative intensities of the illuminations, the designer has the freedom to choose the overall intensity of the illuminations
-to prevent saturation or low SNR, in a sort of offline exposure control.
-
-Assuming that the sequentially illuminated regions do not overlap results in a
-set of training images that span a larger cone than a similar number of
-overlapping regions.  This results in training images that require fewer
-negative coefficients in $\x$ to represent test images under natural
-illuminations.  The effect of negative coefficients in $\x$ appears to depend
-partly on how the test images are taken and is still under study.
-
-{\em Relationship to existing work:} Most light stages used for face recognition have
-been constructed for the purpose of creating public data sets to study
-illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
-light stages have been used for computer graphics purposes
-\cite{debevec2000acquiring, jones2005performance}.
-The light source can be
-moved around manually \cite{masselus2002free}, but this may result in poor
-consistency of illuminations between users.  Structured light applications use projectors to
-directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
-reconstruction, but this is very disturbing to the user.
-Y.\ Schechner \cite{schechner2007multiplexing}
-studies techniques for multiplexing illumination that can dramatically reduce the noise
-of the demultiplexed images for certain classes of objects and cameras.
-While these techniques have not been incorporated into the current
-system, they fit elegantly into our framework and will likely be used
-in future implementations.  We stress that use of this multiplexing technique
-is independent from the choice of original (directional) illuminations.
-
-\subsection{Choice of Illumination Patterns}
-
-We ran two experiments to guide our choice of illuminations for
-our large-scale experiments:
-\begin{enumerate}
-\item {\em Coverage Experiment.} In the first experiment we
-    attempt to determine what coverage of the sphere is
-    required to achieve good interpolation for test images.
-    The subject was illuminated by 100 (50 front, 50 back)
-    illuminations arranged in concentric rings centered at
-    the front camera.  Subsets of the training images were
-    chosen, starting at the front camera and adding a ring
-    at a time.  Each time a ring was added to the training
-    illumination set, the average $\ell^1$ registration
-    error (residual) for a set of test images taken under
-    sunlight was computed and plotted in Figure
-    \ref{fig:illumination-patterns}(a).  The more rings of
-    training illuminations are added, the lower the
-    representation error becomes, with diminishing returns.
-\item {\em Granularity Experiment.} In the second
-    experiment we attempt to determine how finely divided
-    the illumination sphere should be.  At the first
-    granularity level, the projectors  illuminate the
-    covered area uniformly.  At each subsequent granularity
-    level each illuminated cell is divided in two along its
-    longer side but intensity doubled.  For each
-    granularity level the average $\ell^1$ registration
-    error is computed as in the coverage experiment and
-    shown in Figure \ref{fig:illumination-sufficiency}(b).
-    Again, diminishing returns are observed as more
-    illuminations are added.
-\end{enumerate}
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-\includegraphics[height=.75in]{figures_cvpr/coverage_experiment_asplode.png} &
-\includegraphics[height=.75in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
-(a) Coverage Experiment & (b) Chosen Illumination Patterns
-\end{tabular}\vspace{0mm}
-\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
-For rear illuminations the sequence is reversed.
-In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations.
-The four rectangular regions correspond to the four projectors.
-}
-\label{fig:illumination-patterns}
-\vspace{-.75em}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
-(a) Coverage & (b) Granularity
-\end{tabular}\vspace{0mm}
-\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
-\label{fig:illumination-sufficiency}
-\vspace{-4.5em}
-\end{figure}
-
-In the plot for the coverage experiment, Figure
-\ref{fig:illumination-sufficiency}(a),
- we clearly see two plateau regions: one is after 4 rings
-and one is after 10 rings. The first four rings represent the
-typical frontal illuminations, which are present in most public
-face datasets; however, we see that the residual stabilizes
-after 10 rings which include some illuminations from the back
-of the subject. This suggests that although the frontal
-illuminations account for most of the illumination on the face,
-some illuminations from the back are needed in the training set to
-represent images with illumination coming from all directions.
-In the plot for the granularity experiment, Figure
-\ref{fig:illumination-sufficiency}(b), we observe that the
-residual reaches a plateau after four divisions, corresponding
-to a total of 32 illuminations. Based on the results from both
-experiments, we decide to partition the area covered by the
-first 10 rings into a total of 38 cells, whose layout is
-explained in Figure \ref{fig:illumination-patterns}(b). For
-our large-scale experiments, we have collected those
-illuminations for all our subjects.\footnote{It is possible
-that with further experimentation a reduced set of illuminations
-can be found that performs as well or better.}
-
-See below for the 38 training images of one subject:
-\begin{figure}[h]
-\centering
-\includegraphics[width=3.2in]{figures_cvpr/training.png}
-\vspace{-2em}
-\end{figure}
-
-\section{Tests on Public Databases}\label{sec:multipie}
-In this section and the next section, we conduct comprehensive experiments on
-large-scale face databases to verify the performance of our algorithm and
-system. We first test on the largest public face database available that is
-suitable for testing our algorithm, the CMU Multi-PIE.  One shortcoming of the
-CMU Multi-PIE database for our purposes is that there is no separate set of
-test images taken under natural illuminations; we are left to choose which sets
-of images to use for testing and training.  To challenge our algorithm, we
-choose only a small set of illuminations for the training set, yet we include
-all illuminations in the testing set. In the following section, we will test
-our algorithm on a face dataset that is collected by our own system. The goal
-for that experiment will be to show that with a sufficient set of training
-illuminations for each subject, our algorithm indeed works stably and robustly
-with practical illumination, misalignment, pose, and occlusion, as already
-indicated by our experiment shown in Figure \ref{fig:promo}(bottom).
-
-CMU Multi-PIE provides the most extensive test set among public
-datasets. This database contains images of 337 subjects across
-simultaneous variation in pose, expression, and illumination.
-Of these 337 subjects, we use all of the 249 subjects present
-in Session 1 as the training set. The remaining 88 subjects are
-treated as ``impostors'', or invalid images. For each of the
-249 training subjects, we include frontal images of 7 frontal
-illuminations,\footnote{They are illuminations
-$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
-directional illumination, we subtract the ambient-illuminated
-image 0.} taken with neutral expression. As suggested by the
-work of \cite{Georghiades2001-PAMI}, we choose these extreme
-frontal illuminations in the hope that they would linearly
-represent other frontal illuminations well. For the test set,
-we use all 20 illuminations from Sessions 2-4, which were
-recorded over a period of several months. The dataset is
-challenging due to the large number of subjects, and due to
-natural variation in subject appearance over time.
-Table \ref{tab:MultiPIE-recognition2} shows the result of our
-algorithm on each of the 3 testing sessions. Our algorithm
-achieves recognition rates above $90\%$ for all three sessions.
-For the test images, our iterative alignment was initialized
-automatically via the Viola and Jones' face detector. To
-demonstrate that the sparse representation based recognition
-step is indeed beneficial even when there are no impostors, we
-include results for recognition based only on the alignment
-error residuals (i.e. $S=1$), shown in row 1.
-
-\begin{table}
-\caption{Recognition rates on the Multi-PIE database for
-Algorithm 1 and \cite{Yang2010-CVPR}}
-\centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate & Session 2 & Session 3 & Session 4  \\
-\hline
-{Alg. 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
-\hline
-{Alg. 1} & 93.9\% & 93.8\% & 92.3\% \\
-\hline
-{Alg. 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
-\hline
-\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition2} }
-\end{table}
-
-\newcommand{\tempwidth}[0]{0.58in}
-\begin{figure}
-\centering
-{\small
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-\hspace{-2mm}(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Representative failures from Multi-PIE}.
-{\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2.
-Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
-\label{fig:failed-examples}
-\vspace{-4mm}
-\end{figure}
-
-\subsection{Improving the Sampling Window}
-Our algorithm's errors are mostly caused by a few subjects who
-significantly change their appearances between sessions (such
-as hair, facial hair, and eyeglasses). Some representative
-examples are shown in Figure \ref{fig:failed-examples}. For those subjects, alignment and recognition fail on
-almost all test illuminations.
-
-Meanwhile, this observation also suggests that we might be able
-to improve the performance of our method by carefully choosing
-a face region which is less affected by the above factors for
-recognition. In particular, since the forehead region is likely
-to be affected by the change of hair style, we try replacing
-the previous $80 \times 60$ canonical frame with a new
-window that better excludes the forehead. We adjust the
-resolution of the window to keep $m$ approximately constant. In addition,
-we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
-the observation that in many cases the corners
-actually contain background. An example of the new window
-is shown in Figure~\ref{fig:new-mask}.
-
-\begin{figure}
-\vspace{-1em}
-\centering
-{\small
-\begin{tabular}{@{}cc@{}}
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example.png} &
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example_new.png} \vspace{0mm} \\
-Default window. & Proposed window. \vspace{0mm}
-\end{tabular}
-}
-\caption{{\bf Choosing different sampling windows.}}
-\label{fig:new-mask}
-\vspace{-1em}
-\end{figure}
-
-Table \ref{tab:MultiPIE-recognition2} shows that the
-recognition rates on Multi-PIE indeed increase with this new
-window. In addition, Figure \ref{fig:failed-examples}(a), (b),
-and (c) illustrate three representative subjects for which the
-recognition rates of our algorithm are significantly boosted
-with the new window. However, we should mention that the best
-choice of the window is problem-specific and there is not a
-simple guideline to follow. For example, although the new
-window performs better on Multi-PIE, the same window does not
-help at all on our own database, which will be introduced in
-the next section. This is because most of the training and
-testing images in our database are taken on the same day so the
-variation in hair style is very small. Hence, excluding the
-forehead part may actually result in loss of useful
-discriminative information.
-
-\subsection{Comparison to Existing Work}
-We first compare our result to the recent work
-\cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
-initial registration is obtained from manually selected outer
-eye corners. Then, a supervised hierarchical sparse coding
-model based on local image descriptors is trained, which enjoys
-certain translation invariant properties. With the same
-training and testing sets, \cite{Yang2010-CVPR} is able to
-handle the remaining misalignment and achieves state-of-the-art
-performance on the CMU Multi-PIE database.
-Table~\ref{tab:MultiPIE-recognition2} shows that our algorithm
-achieves similar or better performance on different sessions of
-Multi-PIE.
-
-% Classical algorithms
-To better examine the effectiveness of our iterative alignment
-algorithm, we next compare our result to baseline
-linear-projection-based algorithms, such as Nearest Neighbor
-(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
-Discriminant Analysis (LDA)
-\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
-PCA \cite{Turk1991-CVPR} as its performance is always below
-that of Nearest Subspace.} Since these algorithms assume
-pixel-accurate alignment, they are not expected to work well if
-the test image is not well aligned with the training. In
-Table~\ref{tab:MultiPIE-recognition}, we report the results of
-these classical algorithms with three types of testing image
-alignment: 1.\ alignment from the Viola and Jones' detector,
-2.\ alignment via manually selected outer eye
-corners,\footnote{Two manually clicked points are sufficient to
-define a similarity transformation. All of the experiments in
-this section are carried out with similarity transformations.}
-and 3.\ the output of our iterative alignment algorithm. The
-performance drop of the LDA algorithm on Multi-PIE reported
-here seems to agree with that reported already in
-\cite{Gross2008-FGR}.  All of the classical algorithms benefit
-greatly from being paired with our iterative alignment
-algorithm.
-
-\begin{table*}
-\centering
-\caption{Recognition rates on the Multi-PIE database for
-different pairings of alignment and recognition stages.}
-\centerline{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
-\hline
-\backslashbox{Rec.}{Align.}
-& \multicolumn{3}{|c|}{Face Detector}
-& \multicolumn{3}{|c|}{Manual}
-& \multicolumn{3}{|c|}{Iterative Alignment}
-\\
-\hline
-Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
-\hline
-NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
-\hline
-NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
-\hline
-LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
-\hline
-LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
-\hline
-SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition} }
-\vspace{-.1in}
-\end{table*}
-
-% LBP
-We also compare our result to Local Binary Patterns (LBP)
-\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
-able to capture fine details of facial appearance and texture.
-Due to its robustness to variations in illumination, facial
-expression, aging and other changes, LBP has achieved the
-state-of-the-art face recognition performance in the scenario
-when only one sample per person is used for training
-\cite{Tan06facerecognition}.
-In this paper, we follow the same
-steps as in \cite{Ahonen2006-PAMI} to construct an LBP
-descriptor for each training and testing sample. The $80\times
-60$ face region is first divided into a regular $10\times 10$
-grid of cells, each of size $8\times 6$ pixels. Within each
-cell, the histogram of 59 uniform binary patterns is then
-computed, where the patterns are generated by thresholding 8
-neighboring pixels in a circle of radius 2 using the central
-pixel value. Finally, the local histograms are concatenated to
-produce the global descriptor vector. As suggested in
-\cite{Ahonen2006-PAMI}, the recognition is performed using a
-nearest neighbor classifier with Chi square distance as the
-distance measure and we report the recognition rates with the
-same three types of input as before.  As shown in
-Table~\ref{tab:MultiPIE-recognition}, although LBP achieves
-competitive recognition rates given manually aligned training
-and testing samples, demonstrating its robustness to moderate
-misalignment, it still benefits from using the output of our
-iterative alignment algorithm as the input.
-In addition, like
-the other classical algorithms, the performance of LBP degrades
-dramatically if it is applied directly to the output of a face
-detector. This is notable given that LBP is often applied
-without any special alignment in practice. Finally, we
-attribute the improvement in performance of LBP over SRC in
-this experiment to its robustness to illumination components
-that cannot be linearly interpolated by the training set.
-
-In addition, although our algorithm is not designed for
-recognition when there is only a single gallery image per user,
-we compare its performance with LBP within this setting for
-completeness. For this experiment, we use the FERET dataset
-\cite{phillips1998feret}, which contains five standard
-partitions: `fa' is the gallery containing 1196 frontal images
-of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
-sets of probe images. The testing sets differ from the training
-in facial expression (`fb'), illumination (`fc'), aging (`dup1'
-) and long aging (`dup2'). In fact, except for `fb', we notice
-significant changes of illumination in all the other three test
-sets. For the training, we again crop and normalize the face
-region from each original image to an $80\times 60$ window
-using manually marked eye coordinates \cite{Deng2010-PR}. In
-Table~\ref{tab:FERET-recognition}, we report the performance of
-our algorithm on the four test sets, with input directly
-obtained from the Viola and Jones' detector. We also report the
-performance of LBP with the same three types of input as before:
-we use letters
-``$d$'', ``$m$'', and ``$i$'' to indicate face detector, manual
-alignment, and our iterative alignment algorithm, respectively.
-
-\begin{table}
-\caption{Performance on single gallery image FERET dataset}
-\vspace{-3mm} \centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate \% & fb & fc & dup1 & dup2 \\
-\hline
-$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
-\hline
-$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
-\hline
-$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
-\hline
-{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
-\hline
-\end{tabular}
-\label{tab:FERET-recognition} }
-\vspace{-12mm}
-\end{table}
-
-As expected, our algorithm does not perform well except for
-`fb', in which the illumination is similar to the training and
-the mere variation in facial expression is handled well by the
-sparse error model. For the other three test sets, our
-algorithm fails because the illumination changes and other
-variations seriously violate the assumptions of our method.
-This also explains why LBP performs worse with our iterative
-alignment algorithm, compared to manual alignment. On the other
-hand, while LBP achieves the best recognition rates given
-manually aligned training and testing samples, its performance
-degrades drastically when the input is obtained directly from
-the face detector. It is also worth noting that similar poor
-performance of LBP, as well as other descriptors, has been
-observed on the Labeled Face in the Wild (LFW) database, where
-the training is uncontrolled and limited and the input is
-directly obtained from the face detector \cite{Wolf2008-ECCV}.
-
-All of these experimental results confirm that
-both illumination and alignment need to be simultaneously handled
-well in order to achieve accurate face recognition, even when there is
-no obvious occlusion or corruption in the test.
-
-\subsection{Subject Validation}
-
-We test the algorithms' ability to reject invalid images of the
-88 subjects not appearing in the training database. As
-mentioned before, the \emph{sparsity concentration index} (SCI)
-is used as the outlier rejection rule. Given the sparse
-representation $\x$ of a test image with respect to $K$
-training classes, the SCI measures how concentrated the
-coefficients are on a single class in the dataset and is
-defined as in \cite{Wright2009-PAMI}:
-\begin{displaymath}
-\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
-\|\x\|_1 - 1}{K - 1} \in [0,1] .
-\end{displaymath}
-It is easy to see that if $\textup{SCI}(\x) = 1$, the test
-image is represented using images from one single subject
-class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
-evenly over all classes. Thus, we can choose a threshold $t \in
-[0,1]$ for the proposed method and accept a test image as valid
-if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
-invalid. We compare this classifier to classifiers based on
-thresholding the error residuals of NN, NS, LDA, and LBP.
-\begin{figure}[t]
-{\small
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
-(a) & \hspace{2mm}(b) \\
-\end{tabular}
-}
-\vspace{-2mm}
-\caption{{\bf ROC curves} for subject validation on Multi-PIE database,
-(a) for all algorithms with iterative alignment, and
-(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-2mm}
-\end{figure}
-
-Figure \ref{fig:roc-multipie} plots the receiver operating
-characteristic (ROC) curves, which are generated by sweeping
-the threshold $t$ through the entire range of possible values
-for each algorithm.\footnote{Rejecting invalid images not in
-the entire database is much more difficult than deciding if two
-face images are the same subject. Figure \ref{fig:roc-multipie}
-should not be confused with typical ROC curves for face
-similarity, e.g., \cite{PhillipsP2007}.} On the left we can see
-that the SCI based recognition approach significantly
-outperforms the other algorithms, including LBP, even when all
-algorithms are coupled with our proposed iterative alignment.
-In the right plot we again see that classical algorithms, and
-even LBP, are very sensitive to alignment.  Similar contrasts
-between our algorithm and baseline algorithms were also
-observed for SRC in \cite{Wright2009-PAMI}, though on much
-smaller datasets.
-
-\subsection{Recognition with Synthetic Random Block Occlusion}
-
-We further test the robustness of our $\ell^1$-norm based
-algorithm to synthetic occlusion. We simulate various levels of
-occlusion from 10\% to 50\% by replacing a randomly located
-block of the face image with an image of a baboon, as shown in
-Figure~\ref{fig:multipie-occ-rec}. In this experiment, to avoid
-any other factors that may contribute to extra occlusion of the
-face (such as the change of hair style), we choose illumination
-10 from Session 1\footnote{This is the same session as the
-training set.} as testing. The rest of the experimental setting
-remains unchanged. The table in
-Figure~\ref{fig:multipie-occ-rec} shows that our algorithm is
-indeed capable of handling a moderate amount of occlusion. For
-example, at 20\% occlusion, our algorithm still achieves 94.9\%
-recognition rate.
-
-\newcommand{\tempscale}{0.40}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ10.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ20.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ30.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ40.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ50.png}  \\
-\end{tabular}
-\vspace{-1mm}
-\caption{{\bf Recognition under varying level of
-random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:}
-\vspace{2mm}
-{\small
-\begin{tabular}{|c|c|c|c|c|c| }
-\hline
-Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
-\hline
-Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
-\hline
-\end{tabular}
-}
-\label{fig:multipie-occ-rec}
-\vspace{-.5in}
-\end{figure}
-
-\subsection{Recognition with Pose and Expression} We now run tests of
-our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
-Using the same training set as above, we test our algorithm on
-images in Session 2 with $15^\circ$ pose, for all 20
-illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
-algorithm on images in Session 3 with smile, again for all 20
-illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
-the performance of our method will be significantly improved if pose and expression data
-are available in the training.
-
-%\vspace{-.1in}
-\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
-system we described in Section \ref{sec:illumination}, and shown in Figure
-\ref{fig:system}, we have collected the frontal view of 109
-subjects {\em without eyeglasses} under 38 illuminations shown
-in Figure \ref{fig:illumination-patterns}. For testing our
-algorithm, we have also taken 935 images of these subjects with
-a different camera under a variety of practical conditions.
-\vspace{-.1in}
-
-\subsection{Necessity of Rear Illuminations} To see how
-training illuminations affect the performance of our algorithm
-in practice, we now compare how well a few frontal
-illuminations can linearly represent: 1. other frontal illuminations
-taken under the same laboratory conditions, and 2. typical
-indoor and outdoor illuminations. To this end, we use the face
-database acquired by our system and use 7 illuminations per
-subject as training. The illuminations are chosen to be similar
-to the 7 illuminations used in the previous experiment on
-Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
-13, 18, 21, 22\}$ shown in Figure
-\ref{fig:illumination-patterns}(b) to mimic the illuminations
- $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
-our algorithm on the remaining $24 - 7 = 17$ frontal
-illuminations for all the subjects. The recognition rate is
-$99.8\%$, nearly perfect. We also test our algorithm on 310
-indoor images and 168 outdoor images of these subjects taken
-under a variety of lighting conditions (category 1 and 2
-specified below), similar to the one shown in Figure
-\ref{fig:promo}, and the recognition rates for indoor and
-outdoor images drop down to $94.2\%$ and $89.2\%$,
-respectively. This is a strong indication that
-frontal illuminations taken under laboratory conditions
-are insufficient for representing test images under typical indoor and
-outdoor illuminations.
-\vspace{-0.1in}
-
-\renewcommand{\tempscale}{0.45}
-\begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} \\
-\end{tabular}
-\caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
-\vspace{-.1in}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG} \\
-\end{tabular}
- \caption{{\bf Representative examples of category C4}.
-Top row: successful examples with our method using overlapping blocks.
-Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
-\vspace{-.55in}
-\end{figure}
-
-\subsection{Large-Scale Test with Sufficient Training
-Illuminations} Now we use all 109 subjects and 38 illuminations
-in the training and test on 935 images taken under a variety of
-practical illuminations and conditions. We have manually partitioned the test images into four main
-categories:
-\begin{description}
-\item[C1:] 310 \emph{indoor} images of 72 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 1).
-\item[C2:] 168 \emph{outdoor} images of 48 subjects without
-    eyeglasses, frontal view
-    (Fig.~\ref{fig:examples1-3}, row 2).
-\item[C3:] 211 images of 32 subjects with \emph{eyeglasses}
-    (Fig.~\ref{fig:examples1-3}, row 3).
-\item[C4:] 246 images of 56 subjects with \emph{sunglasses}
-    (Fig.~\ref{fig:examples4}).
-\end{description}
-We apply Viola and Jones' face detector on these images and
-directly use the detected faces as the input to our algorithm.
-Table~\ref{tab:UIUC-recognition} reports the performance of our
-algorithm on each category.
-Since our focus is on face
-recognition, the errors do not include failures of the face
-detector on some of the more challenging images.
-As one can see, our algorithm achieves higher than 95\%
-recognition rates on categories 1-3. Furthermore, using the
-full set of 38 illuminations indeed improves the performance of
-our system under practical illumination conditions compared to
-only using a small subset of 7 illuminations. However, the
-performance dramatically drops when the faces are occluded by
-various types of sunglasses, which could cover up to 40\% of
-the entire face. Given the previous experimental results on
-synthetic random block occlusions, and given that the
-illuminations are more challenging, the result is not
-surprising. In the next subsection, we will show how additional
-assumptions can be used to improve the recognition performance.
-\begin{table}[h]
-\vspace{0mm}
-\centering \caption{Recognition rates on our own
-database.} \vspace{0mm}
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Test Category & C1 & C2 & C3 & C4  \\
-\hline
-\hline
-Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
-\hline
-\end{tabular}
-\label{tab:UIUC-recognition}
-\vspace{0mm}
-\end{table}
-
-\subsection{Improving the Performance with Occlusion using Overlapping Blocks}
-A traditional approach to improve the performance of face
-recognition under severe occlusion is to use subregions instead
-the entire face as a whole. This idea has been explored in many
-earlier works; see \cite{Pentland1994-CVPR, Wright2009-PAMI}
-for examples. Since in most real world cases the occlusion is contiguous, it is reasonable to argue that a minority of the
-subregions are likely to be affected by the occlusion. In this
-paper, we adopt the same idea and partition the face into four
-overlapping blocks to better handle sunglasses. This
-scheme is illustrated in Figure~\ref{fig:occ-block}. Notice
-that in this example three out of the four blocks are partially
-or almost completely occluded. In our experiment, each block is
-of size $90\times 48$ and covers about two-fifths of the entire
-face. The testing and training sets are partitioned in the same
-way. We then independently apply
-Algorithm~\ref{alg:deformable-src} and compute a sparse
-representation after registration for each block independently
-with respect to the training set. The recognition
-results for individual blocks are then aggregated by voting.
-
-\begin{figure}
-\centering
-\includegraphics[width=2.6in]{figures_pami/occ_block.png}
-\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.
-(b) The four overlapping blocks.
-(c) The sparse representation is calculated after alignment for each block independently.
-The red lines correspond to his true identity.
-(d) The true identity is successfully recovered by voting based on the SCI scores.}
-\label{fig:occ-block}
-\vspace{0mm}
-\end{figure}
-
-In this experiment, we found that the using the \emph{sparsity
-concentration index} (SCI) scores for voting achieves higher
-recognition rate than the residual measure used in Algorithm~\ref{alg:deformable-src}, on
-category 4 (sunglasses) of our database. The recognition rate
-is increased to 78.3\%, compared to 40.9\% obtained without
-this partition scheme. This is another evidence of the superior
-ability of SCI on subject validation, since a heavily occluded
-block can be regarded as an outlier for recognition and should
-be rejected while voting.
-
-However, we should point out that a major problem with this
-approach is that occlusion cannot always be expected to fall within
- any fixed partition of the face image. Therefore, the
-proposed scheme should only be viewed as an example which shows
-that the performance under occlusion can be boosted by
-leveraging local information of a face as well as global information. We
-leave the investigation of more general models (e.g., MRF \cite{ZhouZ2009}) for face
-recognition with both misalignment and occlusion as an
-interesting future work.
-
-\section{Conclusion}\label{sec:conclusion}
-Using a well-though-out combination of existing ideas
-(iterative image alignment, $\ell_1$-error function, SRC, using projectors for
-illumination), we have proposed a system for recognizing human faces
-from images taken under practical conditions that is conceptually simple, well
-motivated, and competitive with state-of-the-art recognition systems for access
-control scenarios.
-
-The system achieves extremely stable performance under
-a wide range of variations in illumination, misalignment, and even under small amounts of
-pose and occlusion. We achieve very good recognition performance on
-large-scale tests with public datasets as well as our practical face
-images, while using only frontal 2D images in the gallery and no
-explicit 3D face model.
-Our system could potentially be extended to better handle large pose
-and expression, either by incorporating training images with different poses or
-expressions or by explicitly modeling and compensating the associated deformations
-in the alignment stage.
-
-Another important direction for future
-investigation is to extend the alignment algorithm to better
-tackle contiguous occlusion. We have demonstrated that misalignment can be naturally handled within the
-sparse representation framework. More complicated models for
-spatial continuity, such as Markov random fields, have also
-been successfully integrated into the computation of a sparse
-representation of well-aligned test images
-\cite{Cevher2008-NIPS, ZhouZ2009}. A unified approach
-for face alignment and recognition in the presence of
-contiguous occlusion remains an open problem.
-
-\section*{Acknowledgment}
-This work was supported by grants NSF IIS 08-49292, NSF ECCS
-07-01676, and ONR N00014-09-1-0230. JW thanks Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
-Mathematics for discussions related to this work, and
-acknowledges support from a Microsoft Fellowship and
-the Lemelson-Illinois Student Prize. %} %\vspace{0mm}
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{42}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-\bibliographystyle{IEEEtran}
-\bibliography{faces}
-
-\input{bios/bios}
-
-\end{document}
+%\end{comment}

### 2011-04-06 faces_system.tex
Version sent to Yi.  Requires manual page merge tricks
Automating pdf creation...

@@
 
-\begin{comment}
-\appendix[L1 Minimization via Augmented Lagrange Multiplier]
-In this Appendix we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}, which is
-repeated here for convenience. It is
-not hard to see that its computational complexity is dominated
-by the two steps where the $\ell^1$-norm minimization problems
-are solved; namely Step 6 for iterative registration, and Step
-14 for global sparse representation.
-Fortunately, many fast algorithms for solving these problems have been proposed over the
-past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
-That work suggests that \emph{Augmented Lagrange Multiplier}
-(ALM) algorithms \cite{Bertsekas1982} strike a good balance
-between scalability and accuracy: as first order methods, they
-require only lightweight vector operations and matrix-vector
-multiplications at each iteration, making them preferable to
-more classical solutions such as interior point methods.
-However, compared to other first-order methods, they achieve
-higher accuracy with a fixed computational budget.
-
-We use Step 14 as an example to illustrate the ALM
-method, since solving Step 6 is very similar. Recall that in
-Step 14 the problem we are interested in is:
-\begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
-A \x + \e.
-\end{equation}
-Its corresponding augmented Lagrangian function is
-\begin{equation}
-L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
-\end{equation}
-where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
-penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
-(\x,\e,\blamda)$ by alternating between optimizing with respect
-to the primal variables $\x, \e$ and updating the dual variable
-$\blamda$, with the other fixed, as follows:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:alm}
-\end{equation}
-Although updating $\blamda$ is trivial,
-minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
-$\x$ and $\e$ could still be costly. To further reduce the
-complexity of the problem, we adopt an approach used in
-\cite{YangJ2009-pp}, called \emph{alternating direction
-method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
-over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
-\begin{equation}
-\left \{
-\begin{array}{lll}
-\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
-\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
-\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
-\end{array}
-\right .
-\label{eq:adm}
-\end{equation}
-As the objective function is convex and alternation is between two
-terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).
-
-In order to discuss the solution to the above subproblems, we
-need to define the following soft-thresholding operator for a
-vector $\x$ and a scalar $\alpha \geq 0$:
-\begin{equation}
-\mathcal{T}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
-\end{equation}
-where all the operations are performed component-wise. It is
-easy to show that the subproblem with respect to $\e$ has a
-closed-form solution given by the soft-thresholding operator:
-\begin{equation}
-\e_{k+1} = \mathcal{T}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
-\end{equation}
-To solve the subproblem associated with $\x$, we
-apply a first-order $\ell^1$-minimization method,
-called \emph{fast iterative shrinkage-threshold algorithm}
-(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
-iteratively minimize a quadratic approximation $Q(\x, \z)$ to
-$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
-carefully chosen in order to achieve a good convergence
-rate. We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
-\begin{algorithm}[t]
-\caption{\bf (Augmented Lagrange Multiplier Method for Global
-Recognition)}
-\begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
-\mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \mathcal{T}\left(\y - A\x_k +
-\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
-\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \mathcal{T}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
-\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
-\ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{small}
-\end{algorithmic}
-\label{alg:alm}
-%\vspace{-4mm}
-\end{algorithm}
-
-We have selected this algorithm because it strikes the best
-balance between speed, accuracy, and scalability for our problem out of
-many algorithms that we have tested. We refer the interested reader to
-\cite{YangA2010-pp} for a more detailed discussion of competing
-approaches.  On a Mac Pro with
-Dual-Core 2.66GHz Xeon processors and 4GB memory,
-running on our database containing images size $80\times 60$
-pixels from 109 subjects under 38 illuminations,
-our C implementation of Algorithm~\ref{alg:deformable-src} takes
-about 0.60 seconds per subject for alignment and about 2.0
-seconds for global recognition. Compared to the highly
-customized interior point method used in the conference version
-of this paper \cite{Wagner2009-CVPR}, this new algorithm is
-only slightly faster for per subject alignment. However, it is
-much simpler to implement and it achieves a
-\emph{speedup of more than a factor of 10} for global
-recognition!
-\end{comment}
+\input{appendix}
 

### 2011-05-03 faces_system.tex
Fixed double Zihan typo, and updated Hossein's bio.

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou}
-Zihan Zhou received the bachelor's degree in automation from Tsinghua
+received the bachelor's degree in automation from Tsinghua
 University, China, in 2007. Since then, he has been with University of Illinois
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]
-{bios/hossein}}]{Hossein Mobahi} received his Bachelor's and Master's degrees
-in Iran, both in Computer Engineering from Azad University (Tehran-South) in
-2003  and University of Tehran in 2005 respectively. He is currently a PhD
+{bios/hossein}}]{Hossein Mobahi} is currently a PhD
 candidate in the Computer Science Department at the University of Illinois,

### 2011-05-08 release_version/main.tex
Moving release materials into one directory

--- 2011-05-03 faces_system.tex
@@
 using SRC.}\label{fig:promo}
-\vspace{-6.9mm}
+\vspace{-5mm}
 \end{figure}
@@
 (c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
-\vspace{-8.0mm}
+\vspace{-5.0mm}
 \end{figure}
@@
 For a more detailed discussion of competing
-approaches, we refer the interested reader to \cite{YangA2010-pp} 
+approaches, we refer the interested reader to
+\cite{YangA2010-pp}.
 On a Mac Pro with
@@
 \end{tabular}
-\vspace{0mm}
-}
-\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.
-The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees.
-{\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions.
-(b) Simultaneous translation in $y$ direction and in-plane rotation.
-(c) Simultaneous translation in $y$ direction and scale variation.
-{\bf Bottom row:} (d) Translation in $x$ direction only.
-(e) Translation in $y$ direction only.
-(f) In-plane rotation only.
-(g) Scale variation only.}
-\label{fig:attraction}
-\vspace{0mm}
+\vspace{0mm} } \caption{{\bf Region of attraction.} Fraction of
+subjects for which the algorithm successfully aligns a
+synthetically perturbed test image.  The amount of translation
+is expressed as a fraction of the distance between the outer
+eye corners, and the amount of in-plane rotation in degrees.
+{\bf Top row:} (a) Simultaneous translation in $x$ and $y$
+directions. (b) Simultaneous translation in $y$ direction and
+in-plane rotation. (c) Simultaneous translation in $y$
+direction and scale variation. {\bf Bottom row:} (d)
+Translation in $x$ direction only. (e) Translation in $y$
+direction only. (f) In-plane rotation only. (g) Scale variation
+only.} \label{fig:attraction} \vspace{-4mm}
 \end{figure*}
@@
 \label{fig:multiscale}
-\vspace{-4mm}
+\vspace{-2mm}
 \end{figure}
@@
 to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
-}\label{fig:pose-alignment} \vspace{-4mm}
+}\label{fig:pose-alignment} \vspace{-8mm}
 \end{figure}
@@
 constrained only by the need to achieve a good
-SNR\footnote{Since illuminations with more pixels illuminated
+SNR,\footnote{Since illuminations with more pixels illuminated
 will have a better SNR (provided they don't saturate), there is
 an engineering tradeoff between the SNR and the number of
-training images.}, avoid saturation, and achieve
-a reasonably short acquisition time.  Two simplifying
-assumptions that we make are that every pixel is either turned
-fully on or off in every illumination, and that the illuminated
-regions do not overlap.
+training images.} avoid saturation, and achieve a reasonably
+short acquisition time.  Two simplifying assumptions that we
+make are that every pixel is either turned fully on or off in
+every illumination, and that the illuminated regions do not
+overlap.
 
@@
 \label{fig:illumination-sufficiency}
-\vspace{-4.5em}
+\vspace{-3em}
 \end{figure}
@@
 \label{fig:failed-examples}
-\vspace{-4mm}
+\vspace{-12mm}
 \end{figure}
-
 \subsection{Improving the Sampling Window}
@@
 almost all test illuminations.
-
-Meanwhile, this observation also suggests that we might be able
-to improve the performance of our method by carefully choosing
-a face region which is less affected by the above factors for
-recognition. In particular, since the forehead region is likely
-to be affected by the change of hair style, we try replacing
-the previous $80 \times 60$ canonical frame with a new
-window that better excludes the forehead. We adjust the
-resolution of the window to keep $m$ approximately constant. In addition,
-we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
-the observation that in many cases the corners
-actually contain background. An example of the new window
-is shown in Figure~\ref{fig:new-mask}.
-
-\begin{figure}
+\begin{figure}[b]
 \vspace{-1em}
@@
 \end{figure}
+Meanwhile, this observation also suggests that we might be able
+to improve the performance of our method by carefully choosing
+a face region which is less affected by the above factors for
+recognition. In particular, since the forehead region is likely
+to be affected by the change of hair style, we try replacing
+the previous $80 \times 60$ canonical frame with a new
+window that better excludes the forehead. We adjust the
+resolution of the window to keep $m$ approximately constant. In addition,
+we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
+the observation that in many cases the corners
+actually contain background. An example of the new window
+is shown in Figure~\ref{fig:new-mask}.
+
+\begin{table*}
+\centering
+\caption{Recognition rates on the Multi-PIE database for
+different pairings of alignment and recognition stages.}
+\centerline{
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
+\hline
+\backslashbox{Rec.}{Align.}
+& \multicolumn{3}{|c|}{Face Detector}
+& \multicolumn{3}{|c|}{Manual}
+& \multicolumn{3}{|c|}{Iterative Alignment}
+\\
+\hline
+Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
+\hline
+NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
+\hline
+NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
+\hline
+LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
+\hline
+LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
+\hline
+SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
+\hline
+\end{tabular}
+\label{tab:MultiPIE-recognition} }
+\vspace{-.1in}
+\end{table*}
 
@@
 \subsection{Comparison to Existing Work}
+
 We first compare our result to the recent work
@@
 
-\begin{table*}
-\centering
-\caption{Recognition rates on the Multi-PIE database for
-different pairings of alignment and recognition stages.}
-\centerline{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c| }
-\hline
-\backslashbox{Rec.}{Align.}
-& \multicolumn{3}{|c|}{Face Detector}
-& \multicolumn{3}{|c|}{Manual}
-& \multicolumn{3}{|c|}{Iterative Alignment}
-\\
-\hline
-Session $\rightarrow$	& 2		&3			&4			& 2		&3			&4			& 2		&3			&4		\\
-\hline
-NS	& 30.8\%	& 29.4\%	& 24.6\%	& 77.6\%	& 74.3\%	& 73.4\%	& 84.5\%	& 82.3\%	& 81.4\% \\
-\hline
-NN	& 26.4\%	& 24.7\%	& 21.9\%	& 67.3\%	& 66.2\%	& 62.8\%	& 73.5\%	& 69.6\%	& 69.3\% \\
-\hline
-LDA	& 5.1\%		& 5.9\%		& 4.3\%		& 49.4\%	& 44.3\%	& 47.9\%	& 91.0\%	& 89.9\%	& 88.1\% \\
-\hline
-LBP	& 39.9\%	& 38.1\%	& 33.9\%	& 93.3\%	& 91.2\%	& 92.9\%	& {\bf 95.2\%}	& {\bf 94.7\%}	& {\bf 93.5\%} \\
-\hline
-SRC	& -- & -- & -- & -- & -- & -- & 93.9\%	& 93.8\%	& 92.3\% \\
-\hline
-\end{tabular}
-\label{tab:MultiPIE-recognition} }
-\vspace{-.1in}
-\end{table*}
-
 % LBP
@@
 when only one sample per person is used for training
-\cite{Tan06facerecognition}.
-In this paper, we follow the same
+\cite{Tan06facerecognition}. In this paper, we follow the same
 steps as in \cite{Ahonen2006-PAMI} to construct an LBP
@@
 distance measure and we report the recognition rates with the
-same three types of input as before.  As shown in
-Table~\ref{tab:MultiPIE-recognition}, although LBP achieves
-competitive recognition rates given manually aligned training
-and testing samples, demonstrating its robustness to moderate
-misalignment, it still benefits from using the output of our
-iterative alignment algorithm as the input.
-In addition, like
-the other classical algorithms, the performance of LBP degrades
-dramatically if it is applied directly to the output of a face
-detector. This is notable given that LBP is often applied
-without any special alignment in practice. Finally, we
+same three types of input as before.
+
+As shown in Table~\ref{tab:MultiPIE-recognition}, although LBP
+achieves competitive recognition rates given manually aligned
+training and testing samples, demonstrating its robustness to
+moderate misalignment, it still benefits from using the output
+of our iterative alignment algorithm as the input. In addition,
+like the other classical algorithms, the performance of LBP
+degrades dramatically if it is applied directly to the output
+of a face detector. This is notable given that LBP is often
+applied without any special alignment in practice. Finally, we
 attribute the improvement in performance of LBP over SRC in
@@
 
-\begin{table}
-\caption{Performance on single gallery image FERET dataset}
-\vspace{-3mm} \centerline{
-\begin{tabular}{|c|c|c|c|c| }
-\hline
-Recognition rate \% & fb & fc & dup1 & dup2 \\
-\hline
-$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
-\hline
-$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
-\hline
-$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
-\hline
-{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
-\hline
-\end{tabular}
-\label{tab:FERET-recognition} }
-\vspace{-12mm}
-\end{table}
-
 As expected, our algorithm does not perform well except for
@@
 directly obtained from the face detector \cite{Wolf2008-ECCV}.
+\begin{table}
+\caption{Performance on single gallery image FERET dataset}
+\vspace{-3mm} \centerline{
+\begin{tabular}{|c|c|c|c|c| }
+\hline
+Recognition rate \% & fb & fc & dup1 & dup2 \\
+\hline
+$LBP_d$ & 54.8 & 10.3 & 29.8 & 19.8 \\
+\hline
+$LBP_m$ & {\bf 96.6} & {\bf 58.8} & {\bf 71.6} & {\bf 61.5} \\
+\hline
+$LBP_i$ & 94.5 & 42.8 & 46.5 & 21.1 \\
+\hline
+{Alg. 1} & 95.2 & 28.4 & 46.1 & 20.3 \\
+\hline
+\end{tabular}
+\label{tab:FERET-recognition} }
+\vspace{-12mm}
+\end{table}
 
@@
 (b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
-\vspace{-2.5em}
+\vspace{0em}
 \end{figure}
@@
 \label{fig:multipie-occ-rec}
-\vspace{-.5in}
+\vspace{-.3in}
 \end{figure}
@@
 \end{tabular}
-\label{tab:UIUC-recognition}
-\vspace{0mm}
+\label{tab:UIUC-recognition} \vspace{-5mm}
 \end{table}
@@
 \includegraphics[width=2.6in]{figures_pami/occ_block.png}
+\vspace{-2mm}
 \caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses.
@@
 \label{fig:occ-block}
-\vspace{-3em}
+\vspace{-5mm}
 \end{figure}
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou}
-received the bachelor's degree in automation from Tsinghua
-University, China, in 2007. Since then, he has been with University of Illinois
-at Urbana-Champaign, where he received the MS degree in electrical and computer
-engineering and where he is now working toward the PhD degree. During the
-summer of 2009, he was a research intern with Microsoft Research Asia, Beijing.
-His research interests include computer vision, signal processing and machine
-learning. 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou} received the bachelor's degree in automation from Tsinghua
+University, China, in 2007. Since then, he has been with
+University of Illinois at Urbana-Champaign, where he received
+the MS degree in electrical and computer engineering and where
+he is now working toward the PhD degree. During the summer of
+2009, he was a research intern with Microsoft Research Asia,
+Beijing. His research interests include computer vision, signal
+processing and machine learning.
 \end{IEEEbiography}
@@
 
-\input{appendix}
+% \input{appendix_main}
 

### 2011-05-08 release_version/supplementary.tex
Moving release materials into one directory

--- 2011-04-06 supplementary_material.tex
@@
 
-% Include other packages here, before hyperref.
+\usepackage{verbatim}
+
 \usepackage{mydefs}
@@
 
-% \usepackage[small]{caption}
-
 \newcommand{\subj}{\textup{subj to}}
 
-\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
-%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
-%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}
+\begin{document}
 
-\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}
-
-\begin{document}
-%
-% paper title
-\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
-Illumination by Sparse Representation\\
-(Supplementary Materials)}
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
-        John~Wright,~\IEEEmembership{Member,~IEEE,}
-        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
-        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
-        Hossein~Mobahi,
-        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{A.
-Wagner, A.
-Ganesh, Z.
-Zhou, and Y.
-Ma are with the Dept.
-of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign.
-H.
-Mobahi is with the Computer Science Dept.
-at the University of Illinois at Urbana-Champaign.
-J.
-Wright and Y.
-Ma are with Microsoft Research Asia.
-Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W.
-Main st.
-Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-
-\maketitle
-
-\IEEEdisplaynotcompsoctitleabstractindextext
-\vspace{-9mm} % HACK
-
-\IEEEpeerreviewmaketitle
-
-%%%%%%%%% BODY TEXT
-\section{L1 Minimization via Augmented Lagrange Multiplier}
-
-\PARstart{I}{n} these supplementary materials we discuss the computational issues related to
-the implementation of Algorithm~\ref{alg:deformable-src}, which is
+\appendix[L1 Minimization via Augmented Lagrange Multiplier]
+In this Appendix we discuss the computational issues related to
+the implementation of Algorithm 1, which is
 repeated here for convenience. It is
@@
 higher accuracy with a fixed computational budget.
-\input{deformable-src.tex}
 
@@
 \label{alg:alm}
-%\vspace{-4mm}
 \end{algorithm}
@@
 pixels from 109 subjects under 38 illuminations,
-our C implementation of Algorithm~\ref{alg:deformable-src} takes
+our C implementation of Algorithm 1 takes
 about 0.60 seconds per subject for alignment and about 2.0
@@
 
-\input{bios/bios}
-
 \end{document}

### 2011-05-08 release_version/main.tex
Deleting dead code
Got first figure to work

@@
 
-% \usepackage[small]{caption}
-
 \newcommand{\subj}{\textup{subj to}}
 
-\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
-%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
-%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}
-
-\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}
-
 \begin{document}
-%
-% paper title
+
 \title{Towards a Practical Face Recognition System: \\ Robust Alignment and
 Illumination by Sparse Representation}
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
+
 \author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
@@
 
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-
-% If you want to put a publisher's ID mark on the page
-% (can leave text blank if you just want to see how the
-% text height on the first page will be reduced by IEEE)
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
-
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%% ABSTRACT
 \IEEEcompsoctitleabstractindextext{
@@
 \IEEEdisplaynotcompsoctitleabstractindextext
-\vspace{-9mm} % HACK
+\vspace{-9mm}
 
@@
 
-%%%%%%%%% BODY TEXT
 \section{Introduction}
-%Automatic face recognition remains one of the most active areas in computer vision.
 
@@
 occlusion or corruption by solving a simple convex program.
-%For this reason, the final recognition stage of our proposed system will
-%consist of SRC, and the alignment stage that precedes it will be based on
-%similar techniques.
 
 \begin{figure}
-\centering \begin{tabular}{cc}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
-\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
-\hspace{3mm}
-\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
-\end{tabular} \caption{{\bf Effects of registration and illumination on
+\centering 
+\includegraphics{Figure_1}
+\caption{{\bf Effects of registration and illumination on
 Recognition}. In this example we identify the girl among 20 subjects, by
@@
 them.
-
-% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.
-While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.
-
-% We use tools from sparse representation
-% to align a test face image with a set of frontal training
-% images in the presence of registration error, illumination variation, and partial occlusion. The region
-% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
-% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
-% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions.
-We show
-% that our system can efficiently and effectively recognize faces under
-% a variety of realistic conditions, using only
-% frontal images under the proposed illuminations as training.
-
-%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.
-We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.
-Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.
-We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.
-We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.
 
@@
 
-%\vspace{-.1in}
 \section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
@@
 acknowledges support from a Microsoft Fellowship and
-the Lemelson-Illinois Student Prize. %} %\vspace{0mm}
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{42}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
+the Lemelson-Illinois Student Prize.
 
@@
 
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% \input{appendix_main}
-
 \end{document}

### 2011-05-08 release_version/make.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+#latexmk -pdfdvi main.tex
+#latexmk -pdfdvi supplementary.tex
+latexmk main.tex
+#latexmk -pdfdvi supplementary.tex
+

### 2011-05-08 release_version/main.tex
First 6 figures in place...

@@
 
-% Include other packages here, before hyperref.
 \usepackage{mydefs}
@@
 \centering
-{\small
-\begin{tabular}{cccc}
-\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
-\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
-\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
-\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
-\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
-\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
-(a) & (b) & (c) & (d)
-\end{tabular}
-\vspace{0mm}}
+\includegraphics{Figure_2}
 \caption{{\bf Comparing alignment of a subject wearing sunglasses by
@@
 \centering
-{\small
-\begin{tabular}{ccc}
-\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
-\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
-(a)&(b)&(c)
-\end{tabular}
-\begin{tabular}{cccc}
-\includegraphics[height=1in]{figures_pami/x_tr.png} &
-\includegraphics[height=1in]{figures_pami/y_tr.png} &
-\includegraphics[height=1in]{figures_pami/theta.png} &
-\includegraphics[height=1in]{figures_pami/scale.png}\\
-(d)&(e)&(f)& (g)
-\end{tabular}
-\vspace{0mm} } \caption{{\bf Region of attraction.} Fraction of
+\includegraphics{Figure_3}
+\caption{{\bf Region of attraction.} Fraction of
 subjects for which the algorithm successfully aligns a
@@
 \centering
-\includegraphics[width=3in]{figures_pami/multiscale.png}
+\includegraphics{Figure_4}
 \vspace{-2mm}
@@
 \centering
-{\small
-\begin{tabular}{ccccc}
-\includegraphics[height=0.5in]{figures_cvpr/5} &
-\includegraphics[height=0.5in]{figures_cvpr/7} &
-\includegraphics[height=0.5in]{figures_cvpr/09} &
-\includegraphics[height=0.5in]{figures_cvpr/11} &
-\includegraphics[height=0.5in]{figures_cvpr/13} \vspace{0mm} \\
-(a) & (b) & (c) & (d) & (e)\\
-\includegraphics[height=0.5in]{figures_cvpr/15} &
-\includegraphics[height=0.5in]{figures_cvpr/17} &
-\includegraphics[height=0.5in]{figures_cvpr/19} &
-\includegraphics[height=0.5in]{figures_cvpr/21} &
-\includegraphics[height=0.5in]{figures_cvpr/3} \vspace{0mm} \\
-(f) & (g) & (h) & (i) & (j) \vspace{0mm}
-\end{tabular}
-}
+\includegraphics{Figure_5}
 \caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
@@
 \centering
-\includegraphics[height=2in]{figures_cvpr/camera_rig.pdf}
+\includegraphics{Figure_6}
 \caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}

### 2011-05-08 release_version/main.tex
Fixed more figure references

@@
 \centering
-\begin{tabular}{cc}
-\includegraphics[height=.75in]{figures_cvpr/coverage_experiment_asplode.png} &
-\includegraphics[height=.75in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
-(a) Coverage Experiment & (b) Chosen Illumination Patterns
-\end{tabular}\vspace{0mm}
+\includegraphics{Figure_7}
 \caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
@@
 \centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
-\includegraphics[height=1.2in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
-(a) Coverage & (b) Granularity
-\end{tabular}\vspace{0mm}
+\includegraphics{Figure_8}
 \caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
@@
 \centering
-\includegraphics[width=3.2in]{figures_cvpr/training.png}
+\includegraphics[width=3.2in]{Figure_8_5}
 \vspace{-2em}
@@
 \centering
-{\small
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
-\hspace{-2mm}\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
-\includegraphics[width=\tempwidth,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \\
-\hspace{-2mm}(a) & (b) & (c) & (d) & (e) & (f) \vspace{0mm}
-\end{tabular}
-}
+\includegraphics{Figure_9}
 \caption{{\bf Representative failures from Multi-PIE}.
@@
 \centering
-{\small
-\begin{tabular}{@{}cc@{}}
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example.png} &
-\includegraphics[trim=1.9in .7in 1.9in .5in, clip, height=1.0in]{figures_pami/example_new.png} \vspace{0mm} \\
-Default window. & Proposed window. \vspace{0mm}
-\end{tabular}
-}
+\includegraphics{Figure_10}
 \caption{{\bf Choosing different sampling windows.}}
@@
 \begin{figure}[t]
-{\small
-\centering
-\begin{tabular}{@{}c@{}c@{}}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
-\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
-(a) & \hspace{2mm}(b) \\
-\end{tabular}
-}
+\includegraphics{Figure_11}
 \vspace{-2mm}

### 2011-05-08 release_version/main.tex
All images linked properly

@@
 
-\newcommand{\tempscale}{0.40}
 \begin{figure}
-\centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ10.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ20.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ30.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ40.png} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/multipie_occ/occ50.png}  \\
-\end{tabular}
+\includegraphics{Figure_11}
 \vspace{-1mm}
@@
 \centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} \\
-\end{tabular}
+\includegraphics{Figure_12}
 \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
@@
 \centering
-\begin{tabular}{@{}c@{}c@{}c@{}c@{}c@{}c@{}}
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG} &
-\includegraphics[scale=\tempscale,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG} \\
-\end{tabular}
+\includegraphics{Figure_13}
 \caption{{\bf Representative examples of category C4}.
@@
 \centering
-\includegraphics[width=2.6in]{figures_pami/occ_block.png}
+%\includegraphics[width=2.6in]{Figure_14}
+\includegraphics{Figure_14}
 \vspace{-2mm}
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Wagner}}]{Andrew Wagner}
 received his Bachelor's degree in General Engineering in 2003,
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/john}}]{John
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Wright}}]{John
 Wright} received his PhD in Electrical Engineering from the
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Arvind Ganesh}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Ganesh}}]{Arvind Ganesh}
 received his Bachelor's and Master's degrees, both in Electrical
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou} received the bachelor's degree in automation from Tsinghua
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Zhou}}]{Zihan Zhou} received the bachelor's degree in automation from Tsinghua
 University, China, in 2007. Since then, he has been with
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]
-{bios/hossein}}]{Hossein Mobahi} is currently a PhD
+{Mobahi}}]{Hossein Mobahi} is currently a PhD
 candidate in the Computer Science Department at the University of Illinois,
@@
 \biospace
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Ma}}]{Yi
 Ma} received two Bachelors degree in Automation and Applied Mathematics

### 2011-05-08 release_version/main.tex
Typesetting the same up to page 10

@@
 \centering
-\includegraphics{Figure_2}
+\includegraphics[width=3.5in]{Figure_2}
 \caption{{\bf Comparing alignment of a subject wearing sunglasses by
@@
 
-\input{deformable-src.tex}
+\begin{algorithm}[thb]
+\caption{\bf (Deformable Sparse Recovery and Classification for
+Face Recognition)} \label{alg:deformable-src}
+\begin{algorithmic}[1]
+\begin{small}
+\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
+$\y\in\Re^m$ and a deformation group $T$.
+\STATE
+{\bf for} each subject $i$,
+\STATE \hspace{3mm} $\tau^{(0)}
+\leftarrow I$.
+\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
+\STATE \hspace{6mm}
+$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
+\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
+\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
+%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
+\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
+\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
+\STATE
+\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+\tau$;
+\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
+the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
+residuals $\|\e\|_1$. \STATE Compute an average transformation
+$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
+\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
+\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
+k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
+\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
+\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
+$\ell^1$-minimization problem: \hspace{2mm}
+$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
+\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
+\end{small}
+\end{algorithmic}
+\end{algorithm}
+%\vspace{-4mm}
 
@@
 \centering
-\includegraphics{Figure_4}
+\includegraphics[width=3.4in]{Figure_4}
 \vspace{-2mm}
@@
 \centering
-\includegraphics{Figure_6}
+\includegraphics[width=2.9in]{Figure_6}
 \caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
@@
 \centering
-\includegraphics{Figure_7}
+\includegraphics[width=3.1in]{Figure_7}
 \caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.
@@
 
-\renewcommand{\tempscale}{0.45}
 \begin{figure}

### 2011-05-08 release_version/make.sh
Added Over page charge authorization form

@@ -2,5 +2,5 @@
 #latexmk -pdfdvi main.tex
 #latexmk -pdfdvi supplementary.tex
 latexmk main.tex
-#latexmk -pdfdvi supplementary.tex
+latexmk supplementary.tex
 

### 2011-05-08 release_version/make.sh
Fixed flowing on last 2 pages

@@ -1,6 +1,6 @@
 #!/bin/bash
-#latexmk -pdfdvi main.tex
-#latexmk -pdfdvi supplementary.tex
-latexmk main.tex
-latexmk supplementary.tex
+latexmk -pdfdvi main.tex
+latexmk -pdfdvi supplementary.tex
+#latexmk main.tex
+#latexmk supplementary.tex
 
